### 2023-05-17 ms.Rmd
Add manuscript skeleton

---
title: ''
shorttitle: 
leftheader: 
author: 
  - name: Matti Vuorre
    affiliation: 1
    corresponding: yes
    address: Department of Social Psychology, Tilburg University
    email: m.j.vuorre@tilburguniversity.edu
affiliation:
  - id: 1
    institution: Department of Social Psychology, Tilburg University
abstract: |

keywords: 
authornote: |
  \noindent \textbf{Working paper. Proceed with caution.}
wordcount: "`r wordcountaddin:::word_count(filename = 'ms.Rmd')`"
bibliography: references.bib
floatsintext: yes
numbersections: false
linenumbers: no
draft: no
mask: no
figurelist: no
tablelist: no
footnotelist: no
documentclass: apa7
classoption: jou
header-includes:
  - |
    \usepackage{float}
    \setlength{\parskip}{4pt}
output: 
  papaja::apa6_pdf:
    number_sections: true
    keep_tex: false
    highlight: kate
  papaja::apa6_docx: 
    number_sections: true
    keep_tex: false
---

```{r setup, include = FALSE}
# Packages
library(tidyverse)

# Output options
opts_chunk$set(
  eval = TRUE,
  cache = TRUE,
  warning = FALSE,
  error = TRUE,
  message = FALSE
)

# Plotting options
theme_set(
  theme_linedraw(base_size = 12) +
    theme(
      strip.background = element_rect(fill = "gray90", color = "gray90"),
      strip.text = element_text(color = "black", hjust = 0),
      panel.grid = element_blank()
    )
)

```

# References {-}

::: {#refs custom-style="Bibliography"}
:::

### 2023-07-13 ms.qmd
Now quarto

---
abstract: Abstract
---

```{r}
#| message: false

library(tidyverse)
```

### 2023-07-14 supplement.qmd
Content

---
subtitle: Supplementary analysis file
format:
  html:
    toc: true
    toc-depth: 2
    toc-title: Contents
    theme: materia
    mainfont: LibertinusSansRegular
    monofont: FiraCodeRegular
    code-fold: true
    # code-link: true # Slow
    code-tools: true
    header-includes: |
      <link rel="stylesheet" media="screen" href="https://fontlibrary.org//face/fira-code" type="text/css"/>
      <link rel="stylesheet" media="screen" href="https://fontlibrary.org//face/libertinus-sans" type="text/css"/> 
---

```{r document-options}
#| include: false
knitr::opts_chunk$set(
  message = FALSE,
  cache = TRUE,
  warning = FALSE,
  echo = TRUE
)
```

This document shows the output of all our supplementary analyses.

# Session info

```{r}
cat(
  "This output from",
  "\nFile:", knitr::current_input(),
  "\nCommit: ", gert::git_commit_id(),
  "\nRepo:", gert::git_remote_info()$url
)
sessionInfo()
```

### 2023-07-16 ms.qmd
Manuscript skeleton

@@
 ---
-abstract: Abstract
+abstract: Although video games receive immense attention as potential carriers of both positive and negative psychological effects, literature on games' effects remain mixed.
+One reason for this, we believe, is that the effects of games---if any---might be short lived and transient.
+To capture potential effects of games during play itself, we queried XXX players mood at the start of XXX gameplay sessions and in XXX subsequent moods.
+Critically, mood was reported in game in a modified version of a commercially available game.
+We found X Y and Z but surprisingly not XYZ.
+Implications.
+Statements.
+date: 2023-07-18
+format:
+  pdf:
+    documentclass: scrartcl
+    classoption: twocolumn
+    fontfamily: libertinus
+    fontsize: 10pt
+    papersize: a4
+    fig-align: center
+    fig-width: 8
+    pagestyle: headings
+    section-number: true
+    geometry:
+      - top=20mm
+      - left=20mm
+      - right=20mm
+      - bottom=30mm
+      - heightrounded
+    include-in-header:
+      - text: |
+          \setlength{\columnsep}{10mm}
+          \setkomafont{title}{\normalfont\bfseries}
+          \addtokomafont{title}{\normalfont}
+          \addtokomafont{section}{\normalfont}
 ---
 
-```{r}
-#| message: false
+```{r setup}
+#| include: false
 
+knitr::opts_chunk$set(
+  message = FALSE,
+  warning = FALSE,
+  cache = TRUE,
+  include = FALSE
+)
+```
+
+```{r packages}
 library(tidyverse)
 ```
+
+# Bibliography

### 2023-07-16 supplement.qmd

@@
 
-This document shows the output of all our supplementary analyses.
+This document shows the output of all our supplementary analyses including data cleaning. To get the raw data run `make data`.
+
+```{r}
+#| label: packages
+#| code-summary: R Packages
+#| cache: false
+
+library(patchwork)
+library(gt)
+library(modelr)
+library(scales)
+library(broom)
+library(ggdist)
+library(distributional)
+library(tidyverse)
+
+theme_set(
+  theme_linedraw() +
+    theme(
+      strip.background = element_rect(fill = NA, color = NA), 
+      strip.text = element_text(colour = "black", hjust = 0), 
+      panel.grid = element_blank()
+    )
+)
+```
+
+# Data cleaning
+
+We first load the relevant variables from the raw data file. We also rescale the raw mood reports to proportions.
+
+```{r}
+#| label: data-read
+#| message: false
+
+dat <- read_csv(
+  "data/study_prompt_answered.csv",
+  col_select = c(
+    pid, 
+    time = Time_utc, 
+    duration = CurrentSessionLength,
+    prompt = LastStudyPromptType, 
+    mood = response
+  )
+) |> 
+  mutate(
+    pid = factor(pid),
+    mood = mood / 1000
+  )
+dat
+```
+
+We then create a variable that identifies unique sessions, using the heuristics indicated in the code below
+
+```{r}
+#| label: clean-sessions
+#| code-fold: show
+#| include: true
+
+# Create sessions
+dat <- dat |> 
+  arrange(pid, time) |> 
+  mutate(
+    i = row_number(),
+    # Observation is in a new session if this...
+    new_session = 
+      # is the first total observation,
+      i == 1 |
+      # session duration is zero,
+      (duration == 0) | 
+      # session duration is less than previous session duration (ticker has reset),
+      (duration < lag(duration, default = 0)) | 
+      # or if more than 30 minutes since previous observation.
+      (time > lag(time) + minutes(30)),
+    # Session number is the cumulative sum of new sessions...
+    session = cumsum(new_session),
+    # ... per person
+    .by = pid
+  ) |> 
+  select(-c(i, new_session))
+dat
+```
+
+We filter out all other reports except mood ("Wellbeing"), then keep only sessions that had a "pre" game measure of mood at time zero and at least one more mood report.
+
+```{r}
+
+dat <- dat |> filter(prompt == "Wellbeing") |> select(-prompt)
+
+# Count of data with all wellbeing observations
+dat_total_wellbeing <- dat |> 
+  summarise(
+    n_total = n(),
+    n_players = length(unique(pid))
+  )
+
+dat <- dat |> 
+  mutate(
+    session_has_pre = duration[1] == 0,
+    session_has_post = sum(!is.na(mood)) > 1,
+    .by = c(pid, session)
+  )
+
+dat |> 
+  count(session_has_pre, session_has_post) |> 
+  mutate(p = n / sum(n))
+
+dat <- dat |> 
+  filter(session_has_pre, session_has_post) |> 
+  select(!contains("_has_"))
+```
+
+```{r}
+bind_rows(
+  "Total" = dat_total_wellbeing,
+  "This study" = dat |> 
+    summarise(
+      n_total = n(),
+      n_players = length(unique(pid))
+    ),
+  .id = "Data"
+)
+```
+
+The data look like in @tbl-data.
+
+```{r}
+#| label: tbl-data
+#| tbl-cap: First few rows of the data.
+
+dat |> 
+  slice(1:10) |> 
+  gt() |> 
+  fmt_datetime(time) |> 
+  cols_align("left") |> 
+  opt_stylize(6)
+```
+
+We then summarise the raw values of session durations and numbers of sessions in @fig-sessions.
+
+```{r}
+#| label: fig-sessions
+#| fig-height: 5
+#| fig-cap: '**A**.
+Raw session durations.
+**B**.
+Session durations after trimming top 1%.
+**C**.
+Raw numbers of sessions per participant.
+**D**.
+As C, but after trimming participants in the top 1% on number of sessions.'
+
+p0 <- dat |> 
+  ggplot() +
+  scale_x_continuous(
+    expand = expansion(c(0.01, 0.1)),
+  ) +
+  scale_y_continuous(
+    expand = expansion(c(0.01, 0.1)),
+  ) +
+  geom_histogram(
+    col = "dodgerblue4",
+    fill = "dodgerblue1",
+    linewidth = .25
+  )
+
+p_durations <- p0 +
+  aes(duration) +
+  labs(x = "Session duration (minutes)", y = "Observations")
+
+p_durations_99 <- p_durations %+%
+  filter(dat, duration < quantile(duration, .99))
+
+dat_sessions <- dat |> 
+  summarise(
+    sessions = max(session),
+    .by = pid
+  )
+
+p_sessions <- p0 %+% 
+  dat_sessions +
+  aes(sessions) +
+  labs(x = "Sessions", y = "Participants")
+
+p_sessions_99 <- p_sessions %+% 
+  filter(p_sessions$data, sessions < quantile(sessions, .99))
+
+(p_durations | p_durations_99) / 
+  (p_sessions | p_sessions_99) +
+  plot_annotation(tag_level = "A")
+```
+
+Based on these very long tails of implausible values (especially session duration), we trim at 99%.
+
+```{r}
+dat <- dat |> 
+  filter(duration < quantile(duration, .99)) |> 
+  right_join(
+    dat_sessions |> 
+      filter(sessions < quantile(sessions, .99)) |> 
+      select(pid)
+  )
+```
+
+Here is a scatterplot of the raw trimmed data.
+
+```{r}
+#| label: fig-scatter
+#| fig-cap: Scatterplot of mood on duration
+
+dat |> 
+  ggplot() +
+  scale_y_continuous(
+    "Mood",
+    expand = expansion(0.01)
+  ) +
+  scale_x_continuous(
+    "Session duration (minutes)",
+    expand = expansion(0.01)
+  ) +
+  aes(duration, mood) +
+  geom_point(
+    col = "dodgerblue",
+    alpha = .2,
+    size = .8
+  )
+```
+
+We then save this dataset
+
+```{r}
+dat |> 
+  write_rds("data/mood.rds")
+```
+
+# Descriptives
+
+## Demographics
+
+```{r}
+#| echo: true
+#| include: true
+
+read_csv(
+  "data/demographics.csv", 
+  col_select = c(pid, country, gender, age)
+) |> 
+  filter(pid %in% unique(dat$pid)) |> 
+  write_rds("data/demographics.rds")
+demographics <- read_rds("data/demographics.rds")
+demographics
+```
+
+```{r}
+#| label: tbl-demographics
+#| tbl-cap: Demographics
+#| tbl-subcap: 
+#|   - "Countries of origin"
+#|   - "Gender responses"
+#| layout-ncol: 2
+
+count(demographics, country, sort = TRUE) |> 
+  mutate(p = percent(n / sum(n), .1)) |> 
+  gt()
+demographics |> 
+  mutate(gender = fct_lump_min(gender, 2)) |> 
+  count(gender, sort = TRUE) |> 
+  mutate(p = percent(n / sum(n), .1)) |> 
+  gt()
+```
+
+```{r}
+#| label: fig-age
+#| fig-height: 3
+#| fig-cap: Age distribution
+
+p0 %+%
+  demographics +
+  aes(age)
+```
+
+# Modelling strategy
+
+Here's our thinking of the modelling strategy at this point. Although this is not a pre-post experiment, it helps to label the setup as such:
+
+- We rescale the original mood values from 0-1000 to 0-1 for easier interpretation
+- We model mood reports for row *i*, person *j*, and session *k* with a normal distribution censored at 0 & 1
+- We model the mean $\eta_{ijk}$ on an intercept (mood at pre-measure) and slope (mood difference btwn post and pre) with population level effects, and session and person deviations that are multivariate normal distributed
+- We also model the residual deviation: random intercepts by person and probably by session too.
+
+Here is a sketch of this analysis for one person
+
+```{r}
+#| label: fig-example
+#| fig-cap: Sketch of the modelling strategy for one participant
+#| fig-height: 3
+library(lme4)
+library(latex2exp)
+
+dat <- dat |> 
+  mutate(
+    i = row_number(),
+    i = factor(i == 1, levels = c(TRUE, FALSE), labels = c("pre", "post")),
+    .by = c(pid, session)
+  )
+
+tmp <- dat |> 
+  filter(pid == "p10004") |> 
+  mutate(session = str_glue("{pid} (session: {session})"))
+x <- lmer(
+  mood ~ 0 + i + (0 + i | session), data = tmp 
+)
+
+# Ugly code >>> no code
+newx <- tmp |> 
+  data_grid(duration = seq_range(duration, 10), i, session)
+newx$.fitted <- predict(x, newdata = newx)
+
+newy <- tmp |> 
+  data_grid(duration = seq_range(duration, 10), i)
+newy$.fitted <- predict(x, newdata = newy, re.form = NA)
+newy <- newy |> 
+  mutate(session = "p10004 (all data)")
+tmp2 <- tmp |> 
+  mutate(session = "p10004 (all data)")
+
+tmp <- bind_rows(tmp, tmp2)
+newx <- bind_rows(newx, newy)
+
+labels <- coef(x)$session |> 
+  as.data.frame() |> 
+  rownames_to_column("session") |> 
+  mutate(duration = 50)
+
+labels2 <- fixef(x) |> t() |> 
+  data.frame() |> 
+  mutate(session = "p10004 (all data)", duration = 50)
+
+labels <- bind_rows(labels, labels2)
+
+tmp |> 
+  mutate(session = fct_inorder(session)) |> 
+  ggplot() +
+  aes(duration, mood, col = i) +
+  scale_color_brewer(
+    "Pre-session measure",
+    palette = "Set1",
+    aesthetics = c("color", "fill")
+  ) +
+  scale_x_continuous(
+    "Session duration (mins)"
+  ) +
+  scale_y_continuous(
+    "Mood",
+    limits = c(0, 1)
+  ) +
+  geom_line(
+    data = newx |> mutate(session = fct_inorder(session)),
+    aes(y = .fitted)
+  ) +
+  geom_segment(
+    data = labels |> mutate(session = fct_inorder(session)),
+    aes(x = duration, xend = duration, y = ipre, yend = ipost, col = "diff"),
+    arrow = arrow(length = unit(5, "pt"), ends = "both", type = "closed")
+  ) +
+  geom_point() +
+  labs(caption = "mood ~ 1 + post + (1 + post | session)") +
+  facet_wrap("session", nrow = 1) +
+  theme(legend.position = "bottom")
+```
 

### 2023-07-17 supplement.qmd
Download data with R

@@
 
+## Download data
+
+We download data from the main PWS dataset repo:
+
+```{r}
+#| label: data-download
+path <- "data-raw/data.zip"
+if (!file.exists(path)) {
+  download.file("https://osf.io/download/j48qf/", destfile = "data-raw/data.zip")
+  dir.create("data", FALSE)
+  unzip(path, files = c("data/demographics.csv", "data/study_prompt_answered.csv"))
+}
+```
+
 We first load the relevant variables from the raw data file. We also rescale the raw mood reports to proportions.
@@
 #| label: data-read
+#| code-fold: show
 #| message: false
@@
 ```{r}
-#| label: clean-sessions
+#| label: sessions-create
 #| code-fold: show
-#| include: true
 
@@
 
-dat <- dat |> filter(prompt == "Wellbeing") |> select(-prompt)
+dat <- dat |> 
+  filter(prompt == "Wellbeing") |> 
+  select(-prompt)
 
 # Count of data with all wellbeing observations
-dat_total_wellbeing <- dat |> 
+dat_total <- dat |> 
   summarise(
@@
 
+# Include only sessions with "pre-" and "post" measure(s)
 dat <- dat |> 
-  mutate(
-    session_has_pre = duration[1] == 0,
-    session_has_post = sum(!is.na(mood)) > 1,
+  filter(
+    # Session has a wellbeing measure at time = 0
+    duration[1] == 0, 
+    # Session has at least two non-missing wellbeing measures (e.g. pre + 1 post)
+    sum(!is.na(mood)) > 1,
     .by = c(pid, session)
@@
 
-dat |> 
-  count(session_has_pre, session_has_post) |> 
-  mutate(p = n / sum(n))
-
-dat <- dat |> 
-  filter(session_has_pre, session_has_post) |> 
-  select(!contains("_has_"))
-```
-
-```{r}
-bind_rows(
-  "Total" = dat_total_wellbeing,
-  "This study" = dat |> 
+# Count of data to be used in these analyses  
+dat_filtered <- dat |> 
   summarise(
@@
     n_players = length(unique(pid))
-    ),
+  )
+
+# Summary
+bind_rows(
+  "total" = dat_total,
+  "filtered" = dat_filtered,
   .id = "Data"
@@
 
-Here is a scatterplot of the raw trimmed data.
+Here is a scatterplot of the raw data
 

### 2023-07-21 ms.qmd
Move all computation to ms.qmd

--- 2023-07-13 ms.qmd
@@
 ---
-abstract: Abstract
+abstract: "***Abstract*** Although video games receive immense attention as potential carriers of both positive and negative psychological effects, empirical evidence on games' effects remains mixed.
+One reason for this, we believe, is that the effects of games---if any---might be short lived and transient.
+To capture potential effects of games during play itself, we queried XXX players mood at the start of XXX gameplay sessions and in XXX subsequent moods.
+Critically, mood was reported in game in a modified version of a commercially available game.
+We found X Y and Z but surprisingly not XYZ.
+Implications.
+Statements."
+date: 2023-07-18
+bibliography: bibliography.bib
+format:
+  docx:
+    reference-doc: reference.docx
+  pdf:
+    keep-tex: false
+    documentclass: scrartcl
+    classoption: twocolumn
+    fontfamily: libertinus
+    fontsize: 10pt
+    papersize: a4
+    fig-align: center
+    fig-width: 8
+    section-number: true
+    geometry:
+      - top=20mm
+      - left=20mm
+      - right=20mm
+      - bottom=30mm
+    include-in-header:
+      - text: |
+          \setlength{\columnsep}{7.5mm}
+          \addtokomafont{disposition}{\normalfont}
 ---
 
-```{r}
-#| message: false
+```{r setup}
+#| include: false
 
+knitr::opts_chunk$set(
+  message = FALSE,
+  warning = FALSE,
+  cache = TRUE,
+  include = FALSE,
+  echo = FALSE
+)
+```
+
+```{r packages}
+#| cache: false
+library(scales)
+library(cmdstanr)
+library(brms)
+library(ggstance)
+library(patchwork)
 library(tidyverse)
 ```
+
+```{r options}
+#| cache: false
+
+dir.create("models", FALSE)
+options(
+  mc.cores = 8,
+  brms.backend = "cmdstanr"
+)
+theme_set(
+  theme_linedraw(base_size = 10) +
+    theme(
+      strip.background = element_rect(fill = NA, color = NA), 
+      strip.text = element_text(colour = "black", hjust = 0), 
+      axis.text = element_text(size = rel(0.75)),
+      panel.grid = element_blank()
+    )
+)
+```
+
+```{r functions}
+# Format big numbers with a comma
+number2 <- function(x, ...) {
+  number(x, ..., big.mark = ",")
+}
+# Replace extreme percentages
+percent2 <- function(x, accuracy = .1) {
+  x <- percent(x, accuracy = accuracy)
+  x <- if_else(x == "100.0%", ">99.9%", x)
+  x <- if_else(x == "0.0%", "<0.1%", x)
+  x
+}
+```
+
+```{r}
+#| label: data-download
+
+# Download, unzip data from OSF PWS database
+path <- "data-raw/data.zip"
+if (!file.exists(path)) {
+  download.file("https://osf.io/download/j48qf/", destfile = "data-raw/data.zip")
+  dir.create("data", FALSE)
+  unzip(path, files = c("data/demographics.csv", "data/study_prompt_answered.csv"))
+}
+```
+
+# Introduction
+
+## Current study
+
+Here, we ask how players' mood changes during gameplay sessions.
+
+# Methods
+
+```{r data-load}
+# Load data from csv
+dat <- read_csv(
+  "data/study_prompt_answered.csv",
+  col_select = c(
+    pid, 
+    time = Time_utc, 
+    duration = CurrentSessionLength,
+    prompt = LastStudyPromptType, 
+    mood = response
+  )
+) |> 
+  mutate(
+    pid = factor(pid),
+    mood = mood / 1000
+  )
+
+# Calculate raw data summaries
+sraw <- list()
+sraw$n_res <- dat |> 
+  count(prompt) 
+sraw$n_res_total <- number2(sum(sraw$n_res$n))
+sraw$n_res <- sraw$n_res |> 
+  mutate(n = number2(n)) |> 
+  pivot_wider(names_from = prompt, values_from = n)
+sraw$n_sub <- number2(length(unique(dat$pid)))
+
+# Create sessions using heuristics below
+dat <- dat |> 
+  arrange(pid, time) |> 
+  mutate(
+    i = row_number(),
+    # Observation is in a new session if this...
+    new_session = 
+      # is the first total observation,
+      i == 1 |
+      # session duration is zero,
+      (duration == 0) | 
+      # session duration is less than previous session duration (ticker has reset),
+      (duration < lag(duration, default = 0)) | 
+      # or if more than 30 minutes since previous observation.
+      (time > lag(time) + minutes(30)),
+    # Session number is the cumulative sum of new sessions...
+    session = cumsum(new_session),
+    # ... per person
+    .by = pid
+  ) |> 
+  select(-c(i, new_session))
+
+# Filter data: Just wellbeing responses that occurred in sessions with both "pre" and "post" measures
+dat <- dat |> 
+  filter(prompt == "Wellbeing") |> 
+  select(-prompt)
+
+# Count of data with all wellbeing observations
+dat_total <- dat |> 
+  summarise(
+    n_total = n(),
+    n_players = length(unique(pid))
+  )
+
+# Include only sessions with "pre-" and "post" measure(s)
+dat <- dat |> 
+  filter(
+    # Session has a wellbeing measure at time = 0
+    duration[1] == 0, 
+    # Session has at least two non-missing wellbeing measures (e.g. pre + 1 post)
+    sum(!is.na(mood)) > 1,
+    .by = c(pid, session)
+  )
+
+# Count of data to be used in these analyses  
+dat_filtered <- dat |> 
+  summarise(
+    n_total = n(),
+    n_players = length(unique(pid))
+  )
+
+# Summary
+bind_rows(
+  "total" = dat_total,
+  "filtered" = dat_filtered,
+  .id = "Data"
+)
+
+# Summaries of filtered data
+s <- list()
+s$n_sub <- number2(length(unique(dat$pid)))
+s$n_obs <- number2(nrow(dat))
+s$n_obs_missing <- number2(nrow(filter(dat, is.na(mood))))
+s$n_ses <- number2(nrow(count(dat, pid, session)))
+```
+
+## Participants
+
+```{r demographics-load}
+demo <- read_csv(
+  "data/demographics.csv", 
+  col_select = c(pid, country, gender, age)
+) |> 
+  filter(pid %in% unique(dat$pid))
+
+a <- quantile(demo$age, probs = c(.5, .1, .9), na.rm = TRUE)
+a <- str_glue("{a[1]} ({a[2]}, {a[3]})")
+g <- count(demo, gender, sort = TRUE) |> 
+  mutate(p = n / sum(n)) |> 
+  mutate(x = str_glue("{gender} ({number2(n)}, {percent2(p)})"))
+n_countries <- length(unique(demo$country[!is.na(demo$country)]))
+countries <- count(demo, country, sort = TRUE) |> 
+  mutate(p = n / sum(n)) |> 
+  mutate(x = str_glue("{country} ({number2(n)}, {percent2(p)})"))
+```
+
+The PowerWash Simulator case study database contains `r sraw$n_res_total` psychological item responses from `r sraw$n_sub` individuals.
+In this study, we only studied subjective well-being (mood), so did not consider the other five types of item responses in the database (enjoyment, focus, autonomy, competence, and immersion).
+Therefore, here we used data from `r s$n_sub` participants who provided `r s$n_obs` mood responses (`r s$n_obs_missing` with missing values) from `r s$n_ses` gameplay sessions.
+
+These participants median age (1st & 9th deciles) was `r a`, and four most frequent responses to gender were `r g$x[1:4]`.
+They hailed from `r n_countries`, with four most represented countries being `r countries$x[1:4]`.
+
+The study procedures were granted ethical approval by Oxford University’s Central University Research Ethics Committee (SSH_OII_CIA_21_011).
+All participants provided informed consent to participate and reported being 18 years or older.
+
+## Measures
+
+We measured affective well-being with one item.
+
+## Procedure
+
+Describe PWS.
+
+## Data analysis
+
+```{r}
+dat <- dat |> 
+  mutate(
+    post = factor(
+      row_number() > 1, 
+      levels = c(FALSE, TRUE), 
+      labels = c("0", "1")
+    ),
+    .by = c(pid, session)
+  ) |> 
+  mutate(ps = interaction(pid, session, sep = "_")) |> 
+  mutate(
+    cl = case_when(
+      mood == 0 ~ "left",
+      mood == 1 ~ "right",
+      TRUE ~ "none"
+      )
+  )
+```
+
+The focal inference in the current work were mood levels at the beginning of each session (during those subset of sessions where mood was queried at login) contrasted to mood levels during the subsequent play session.
+We modelled each mood report *i* in 1 to `r nrow(dat)` for participant *j* in 1 to `r length(unique(dat$pid))` and participant-session *k* in 1 to `r length(unique(dat$ps))` as normally distributed with the following equations:
+
+$$
+\begin{align*}
+\text{mood}_{ijk} &\sim \text{Normal}(\eta_{ijk}, \kappa_{ij}), \\
+\eta_{ijk} &= \alpha_0 + \beta_{0j} + \gamma_{0k} + \\ 
+  &\quad \ (\alpha_1 + \beta_{1j} + \gamma_{1k})\text{post}_{ijk}, \\
+log(\kappa_{ij}) &= \alpha_2 + \beta_{2j} + \alpha_3\text{post}_{ijk}, \\
+\begin{bmatrix} \alpha_0 \\ \alpha_1 \\ \alpha_2 \end{bmatrix} &\sim 
+  \text{MVN}\left(\begin{bmatrix} 0 \\ 0 \\ 0 \end{bmatrix}, 
+  \begin{bmatrix} 
+    \tau_{\alpha_0} &\rho_{{\alpha_0}{\alpha_1}} &\rho_{{\alpha_0}{\alpha_2}} \\ 
+    \rho_{{\alpha_0}{\alpha_1}} &\tau_{\alpha_1} &\rho_{{\alpha_1}{\alpha_2}} \\ 
+    \rho_{{\alpha_0}{\alpha_2}} &\rho_{{\alpha_1}{\alpha_2}} &\tau_{\alpha_2} 
+  \end{bmatrix}\right), \\
+  \begin{bmatrix} \gamma_0 \\ \gamma_1 \end{bmatrix} &\sim
+    \text{MVN}\left(\begin{bmatrix} 0 \\ 0 \end{bmatrix}, 
+    \begin{bmatrix} 
+      \tau_{\gamma_0} &\rho_{{\gamma_0}{\gamma_1}} \\ 
+      \rho_{{\gamma_0}{\gamma_1}} &\tau_{\gamma_1}
+  \end{bmatrix}\right).
+\end{align*}
+$$
+
+We analyzed the data with R [@rcoreteamLanguageEnvironmentStatistical2023] and used the brms package for models [@burknerBrmsPackageBayesian2017].
+
+```{r}
+contrasts(dat$post) <- c(-0.5, 0.5)
+model <- bf(mood | cens(cl) ~ 0 + post + (0 + post |p| pid) + (1 | ps)) +
+  lf(sigma ~ 0 + post + (1 |p| pid)) +
+  gaussian()
+
+get_prior(model, dat)
+
+fit <- brm(
+  model,
+  data = dat |> filter(pid %in% sample(unique(dat$pid), 100)),
+  silent = 0,
+  threads = 2,
+  iter = 750,
+  control = list(adapt_delta = .9),
+  file = "models/brm"
+)
+summary(fit, prior = TRUE)
+```
+
+# Results
+
+## Descriptives
+
+```{r}
+#| label: fig-descriptives
+#| include: true
+#| fig-height: 2.2
+#| fig-width: 7
+#| fig-cap: '**A**. Histogram of session durations. **B**. Summary of how many sessions each participant completed. **C**. Histogram of all mood ratings.'
+
+p0 <- dat |> 
+  ggplot() +
+  scale_x_continuous(
+    expand = expansion(c(0.01, 0.01)),
+  ) +
+  scale_y_continuous(
+    expand = expansion(c(0.001, 0.1)),
+  ) +
+  geom_histogram(
+    col = "dodgerblue4",
+    fill = "dodgerblue1",
+    linewidth = .25
+  )
+
+dat_session_durations <- dat |> 
+  filter(duration == max(duration), .by = c(pid, session))
+
+dat_session_counts <- dat |> 
+  distinct(pid, session) |> 
+  count(pid, name = "sessions")
+  summarise(
+    sessions = max(session),
+    .by = pid
+  )
+
+p_durations <- p0 %+%
+  dat_session_durations +
+  aes(duration) +
+  scale_x_log10() +
+  labs(x = "Session duration (minutes)", y = "Sessions")
+
+p_sessions <- p0 %+% 
+  dat_session_counts +
+  aes(sessions) +
+  labs(x = "Sessions", y = "Participants")
+
+p_mood <- p0 +
+  aes(mood) +
+  labs(x = "Mood", y = "Responses")
+
+(p_durations | p_sessions | p_mood) +
+  plot_annotation(tag_level = "A")
+```
+
+```{r}
+#| label: fig-data
+#| include: true
+#| fig-width: 7.2
+#| fig-height: 6
+#| fig-cap: "**A**.
+Representative data from three participants (rows) showing scatterplots of mood responses over that session's (columns) duration.
+Each session's first mood report is highlighted in red, subsequent reports are in blue.
+**B**.
+Scatterplot of person-mean mood reports at the beginning of the session (x-axis) and subsequent reports (y-axis).
+**C**.
+Histogram of differences between person-means of initial and subsequent mood reports."
+
+axislabels <- c(0, 0.5, 1.0)
+
+# Take a quasirandom sample of people who had sufficient number of sessions
+# and data per session.
+set.seed(99)
+p0 <- dat |> 
+  add_count(pid, session, name = "obs_per_session") |> 
+  filter(obs_per_session >= 4) |> 
+  mutate(session = as.numeric(as.factor(session)), .by = pid) |> 
+  mutate(session_per_person = length(unique(session)), .by = pid) |> 
+  filter(session_per_person >= 4) |> 
+  arrange(pid, session) |> 
+  filter(pid %in% sample(unique(pid), 3)) |> 
+  mutate(Session = session, Person = fct_anon(pid)) |> 
+  ggplot() +
+  aes(duration, mood, col = post) +
+  scale_color_brewer(
+    "Pre-session measure",
+    palette = "Set1",
+    aesthetics = c("color", "fill")
+  ) +
+  scale_x_continuous(
+    "Session duration (minutes)"
+  ) +
+  scale_y_continuous(
+    "Mood",
+    limits = c(0, 1),
+    breaks = axislabels,
+    labels = axislabels
+  ) +
+  geom_point(size = 1.5, alpha = .75) +
+  facet_grid(
+    Person ~ Session
+  ) +
+  theme(legend.position = "none")
+
+tmp <- dat |> 
+  summarise(
+    mood = mean(mood, na.rm = TRUE),
+    .by = c(pid, post)
+  ) |> 
+  pivot_wider(names_from = post, values_from = mood) |> 
+  mutate(delta = `1` - `0`)
+pa <- tmp |> 
+  ggplot(aes(`0`, `1`)) +
+  scale_x_continuous(
+    "First mood",
+    expand = expansion(0.01),
+    breaks = axislabels,
+    labels = axislabels
+  ) +
+  scale_y_continuous(
+    "Subsequent mood",
+    expand = expansion(0.01),
+    breaks = axislabels,
+    labels = axislabels
+  ) +
+  geom_abline(linewidth = .5, col = "dodgerblue") +
+  geom_point(
+    alpha = .25, size = 1, shape = 1
+  ) +
+  theme(aspect.ratio = 1)
+pb <- tmp |> 
+  drop_na(delta) |> 
+  ggplot(aes(delta)) +
+  scale_y_continuous(
+    "Participants",
+    expand = expansion(c(0, 0.1))
+  ) +
+  scale_x_continuous(
+    "Difference (subsequent - first mood)"
+  ) +
+  geom_histogram(binwidth = .025, col = "white") +
+  stat_summaryh(
+    fun.data = mean_cl_boot_h, 
+    aes(y = 5), 
+    col = "dodgerblue"
+  )
+(p0 / (pa | pb)) +
+  plot_layout(heights = c(6, 4)) +
+  plot_annotation(tag_level = "A")
+```
+
+```{r}
+summary(fit)
+
+hs <- c(
+  "Pre" = "post0 = 0",
+  "Post" = "post1 = 0",
+  "Difference" = "post1 - post0 = 0"
+)
+tmp <- bind_rows(
+  hypothesis(fit, hs)$hypothesis,
+  hypothesis(fit, hs, scope = "coef", group = "pid")$hypothesis
+)
+tmp |> 
+  ggplot(aes(Estimate, Hypothesis)) +
+  geom_point(
+    data = . %>% filter(is.na(Group)),
+    size = 1
+  ) +
+  geom_point(
+    data = . %>% filter(!is.na(Group)),
+    size = .25
+  )
+
+```
+
+# Bibliography
+
+::: {#refs}
+:::
+
+# Appendices / supplementary analyses / stuff

### 2023-07-21 supplement.qmd
Move all computation to ms.qmd

--- 2023-07-14 supplement.qmd
@@
 
-This document shows the output of all our supplementary analyses.
+NOTE: This is just a bit of a notebook at the moment.
+
+```{r}
+#| label: packages
+#| code-summary: R Packages
+#| cache: false
+
+library(patchwork)
+library(gt)
+library(modelr)
+library(scales)
+library(broom)
+library(ggdist)
+library(distributional)
+library(tidyverse)
+
+theme_set(
+  theme_linedraw() +
+    theme(
+      strip.background = element_rect(fill = NA, color = NA), 
+      strip.text = element_text(colour = "black", hjust = 0), 
+      panel.grid = element_blank()
+    )
+)
+```
+
+# Modelling strategy
+
+Here's our thinking of the modelling strategy at this point. Although this is not a pre-post experiment, it helps to label the setup as such:
+
+- We rescale the original mood values from 0-1000 to 0-1 for easier interpretation
+- We model mood reports for row *i*, person *j*, and session *k* with a normal distribution censored at 0 & 1
+- We model the mean $\eta_{ijk}$ on an intercept (mood at pre-measure) and slope (mood difference btwn post and pre) with population level effects, and session and person deviations that are multivariate normal distributed
+- We also model the residual deviation: random intercepts by person and probably by session too.
+
+Here is a sketch of this analysis for one person
+
+```{r}
+#| label: fig-example
+#| fig-cap: Sketch of the modelling strategy for one participant
+#| fig-height: 3
+library(lme4)
+library(latex2exp)
+
+dat <- dat |> 
+  mutate(
+    i = row_number(),
+    i = factor(i == 1, levels = c(TRUE, FALSE), labels = c("pre", "post")),
+    .by = c(pid, session)
+  )
+
+tmp <- dat |> 
+  filter(pid == "p10004") |> 
+  mutate(session = str_glue("{pid} (session: {session})"))
+x <- lmer(
+  mood ~ 0 + i + (0 + i | session), data = tmp 
+)
+
+# Ugly code >>> no code
+newx <- tmp |> 
+  data_grid(duration = seq_range(duration, 10), i, session)
+newx$.fitted <- predict(x, newdata = newx)
+
+newy <- tmp |> 
+  data_grid(duration = seq_range(duration, 10), i)
+newy$.fitted <- predict(x, newdata = newy, re.form = NA)
+newy <- newy |> 
+  mutate(session = "p10004 (all data)")
+tmp2 <- tmp |> 
+  mutate(session = "p10004 (all data)")
+
+tmp <- bind_rows(tmp, tmp2)
+newx <- bind_rows(newx, newy)
+
+labels <- coef(x)$session |> 
+  as.data.frame() |> 
+  rownames_to_column("session") |> 
+  mutate(duration = 50)
+
+labels2 <- fixef(x) |> t() |> 
+  data.frame() |> 
+  mutate(session = "p10004 (all data)", duration = 50)
+
+labels <- bind_rows(labels, labels2)
+
+tmp |> 
+  mutate(session = fct_inorder(session)) |> 
+  ggplot() +
+  aes(duration, mood, col = i) +
+  scale_color_brewer(
+    "Pre-session measure",
+    palette = "Set1",
+    aesthetics = c("color", "fill")
+  ) +
+  scale_x_continuous(
+    "Session duration (mins)"
+  ) +
+  scale_y_continuous(
+    "Mood",
+    limits = c(0, 1)
+  ) +
+  geom_line(
+    data = newx |> mutate(session = fct_inorder(session)),
+    aes(y = .fitted)
+  ) +
+  geom_segment(
+    data = labels |> mutate(session = fct_inorder(session)),
+    aes(x = duration, xend = duration, y = ipre, yend = ipost, col = "diff"),
+    arrow = arrow(length = unit(5, "pt"), ends = "both", type = "closed")
+  ) +
+  geom_point() +
+  labs(caption = "mood ~ 1 + post + (1 + post | session)") +
+  facet_wrap("session", nrow = 1) +
+  theme(legend.position = "bottom")
+```
 

### 2023-07-21 ms.qmd
Throw out prose

@@
 ---
-abstract: "***Abstract*** Although video games receive immense attention as potential carriers of both positive and negative psychological effects, empirical evidence on games' effects remains mixed.
-One reason for this, we believe, is that the effects of games---if any---might be short lived and transient.
-To capture potential effects of games during play itself, we queried XXX players mood at the start of XXX gameplay sessions and in XXX subsequent moods.
-Critically, mood was reported in game in a modified version of a commercially available game.
-We found X Y and Z but surprisingly not XYZ.
-Implications.
-Statements."
 date: 2023-07-18
@@
 ```
-
-# Introduction
-
-## Current study
-
-Here, we ask how players' mood changes during gameplay sessions.
 
@@
 We measured affective well-being with one item.
-
-## Procedure
-
-Describe PWS.
 

### 2023-07-21 ms.qmd
WIP pre-holiday changes :smile:

@@
 
-The focal inference in the current work were mood levels at the beginning of each session (during those subset of sessions where mood was queried at login) contrasted to mood levels during the subsequent play session.
-We modelled each mood report *i* in 1 to `r nrow(dat)` for participant *j* in 1 to `r length(unique(dat$pid))` and participant-session *k* in 1 to `r length(unique(dat$ps))` as normally distributed with the following equations:
+Our primary contrast of interest was between mood levels at the beginning of each session and mood levels during the subsequent play session.
+We modelled each mood report *i* in 1 to `r nrow(dat)` for participant *j* in 1 to `r length(unique(dat$pid))` and participant-session *k* in 1 to `r length(unique(dat$ps))` with a normal distribution censored at the VAS limits [0, 1].
 
@@
 ```{r}
-contrasts(dat$post) <- c(-0.5, 0.5)
+set.seed(999)
+dat_snip <- dat |> 
+  filter(pid %in% sample(unique(dat$pid), 300))
+
 model <- bf(mood | cens(cl) ~ 0 + post + (0 + post |p| pid) + (1 | ps)) +
@@
 
-get_prior(model, dat)
-
 fit <- brm(
   model,
-  data = dat |> filter(pid %in% sample(unique(dat$pid), 100)),
+  data = dat_snip,
   silent = 0,
   threads = 2,
-  iter = 750,
+  iter = 1500,
   control = list(adapt_delta = .9),
-  file = "models/brm"
+  file = "models/brm-pre-post-location-scale"
 )
@@
 #| fig-width: 7
-#| fig-cap: '**A**. Histogram of session durations. **B**. Summary of how many sessions each participant completed. **C**. Histogram of all mood ratings.'
+#| fig-cap: '**A**. Histogram of session durations (note log10 x-axis). **B**. Summary of how many sessions each participant completed. **C**. Histogram of all mood ratings.'
 
@@
   geom_histogram(
-    col = "dodgerblue4",
-    fill = "dodgerblue1",
+    col = "white",
     linewidth = .25
@@
   count(pid, name = "sessions")
-  summarise(
-    sessions = max(session),
-    .by = pid
-  )
 
@@
 #| fig-width: 7.2
-#| fig-height: 6
+#| fig-height: 5
 #| fig-cap: "**A**.
@@
 **B**.
+Horizontal histograms of all sessions' mood response for that person.
+**C**.
 Scatterplot of person-mean mood reports at the beginning of the session (x-axis) and subsequent reports (y-axis).
-**C**.
-Histogram of differences between person-means of initial and subsequent mood reports."
+**D**.
+Histograms of mood ratings at the beginning of session, during the session, and differences between person-means of first and subsequent mood reports."
 
@@
 set.seed(99)
-p0 <- dat |> 
+tmp <- dat |> 
   add_count(pid, session, name = "obs_per_session") |> 
@@
   filter(pid %in% sample(unique(pid), 3)) |> 
-  mutate(Session = session, Person = fct_anon(pid)) |> 
+  mutate(Session = str_glue("Session {session}"), Person = str_glue("Person {fct_anon(pid)}"))
+
+library(ggdist)
+p_mood_duration_sessions <- tmp |> 
   ggplot() +
-  aes(duration, mood, col = post) +
+  aes(duration / 60, mood, col = post) +
   scale_color_brewer(
@@
   scale_x_continuous(
-    "Session duration (minutes)"
+    "Session duration (hours)"
   ) +
@@
   ) +
-  geom_point(size = 1.5, alpha = .75) +
+  geom_point(size = 1.5, alpha = 1) +
   facet_grid(
-    Person ~ Session
+    Person ~ Session,
+    scales = "free_x"
   ) +
+  theme(
+    plot.margin = unit(c(8, 8, 8, 8), "pt"),
+    # strip.background.y = element_blank(),
+    # strip.text.y = element_blank()
+  )
+
+p_mood_duration_all <- tmp |> 
+  mutate(Session = "All sessions") |> 
+  ggplot() +
+  aes(y = mood, col = post, fill = post) +
+  scale_color_brewer(
+    "Pre-session measure",
+    palette = "Set1",
+    aesthetics = c("color", "fill")
+  ) +
+  scale_x_continuous(
+    "Scaled count",
+    expand = expansion(c(0.01, 0.1))
+  ) +
+  scale_y_continuous(
+    limits = c(0, 1),
+    # expand = expansion(c(0, 0.01)),
+    breaks = axislabels,
+    labels = axislabels
+  ) +
+  coord_fixed(ratio = 0.66) +
+  geom_dotplot(
+    aes(x = 0),
+    dotsize = 0.9,
+    binaxis = "y",
+    method = "histodot",
+    binwidth = 0.06,
+    stackgroups = TRUE
+  ) +
+  facet_grid(
+    rows = vars(Person), 
+    cols = vars(Session)
+  ) +
+  theme(
+    axis.text = element_blank(), 
+    axis.title.y = element_blank(), 
+    axis.ticks.y = element_blank(), 
+    strip.background.y = element_blank(),
+    strip.text.y = element_blank()
+    # plot.margin = unit(c(8, 8, 8, 8), "pt")
+  )
+
+p_mood_duration <- (p_mood_duration_sessions | p_mood_duration_all) &
+  plot_layout(widths = c(80, 20)) &
   theme(legend.position = "none")
@@
   mutate(delta = `1` - `0`)
-pa <- tmp |> 
+
+p_mood_biscatter <- tmp |> 
   ggplot(aes(`0`, `1`)) +
   scale_x_continuous(
-    "First mood",
+    "Mood at start of session",
     expand = expansion(0.01),
@@
   scale_y_continuous(
-    "Subsequent mood",
+    "Mood during session",
     expand = expansion(0.01),
@@
   ) +
-  geom_abline(linewidth = .5, col = "dodgerblue") +
   geom_point(
@@
   ) +
+  geom_abline(linewidth = .33, col = "dodgerblue") +
   theme(aspect.ratio = 1)
-pb <- tmp |> 
+p_mood_difference <- tmp |> 
   drop_na(delta) |> 
@@
     "Participants",
-    expand = expansion(c(0, 0.1))
+    expand = expansion(c(0.001, 0.1))
   ) +
@@
   ) +
+  coord_cartesian(xlim = c(-0.4, 0.6)) +
   geom_histogram(binwidth = .025, col = "white") +
@@
     fun.data = mean_cl_boot_h,
-    aes(y = 5), 
-    col = "dodgerblue"
+    aes(y = 10),
+    col = "white",
+    fill = "black",
+    shape = 21,
+    stroke = 1.5, size = 0.75
   )
-(p0 / (pa | pb)) +
+
+p_mood_difference <- tmp |> 
+  rename(
+    `Pre-session` = `0`,
+    `During session` = `1`,
+    Difference = delta
+  ) |> 
+  pivot_longer(-pid) |> 
+  mutate(name = fct_inorder(name)) |> 
+  ggplot(aes(value)) +
+  geom_histogram(bins = 30, col = "white") +
+  scale_y_continuous(
+    "Participants",
+    expand = expansion(c(0.001, 0.1))
+  ) +
+  scale_x_continuous(
+    "Mood",
+    expand = expansion(c(0.001))
+  ) +
+  facet_wrap("name", ncol = 3, scales = "free")
+
+(p_mood_duration / (p_mood_biscatter | p_mood_difference)) +
   plot_layout(heights = c(6, 4)) +
@@
 ```{r}
-summary(fit)
+#| label: tbl-avg
+#| include: true
+#| tbl-cap: Population level average mood at beginning and during session, and difference.
 
@@
 )
+
+h_avg <- hypothesis(fit, hs)$hypothesis
+h_avg |> 
+  select(` ` = Hypothesis, Estimate, CI.Lower, CI.Upper) |> 
+  mutate(across(where(is.numeric), ~percent2(., .01))) |> 
+  knitr::kable(digits = 3)
+```
+
+```{r}
+#| eval: false
+h_sub <- hypothesis(fit, hs, scope = "coef", group = "pid")$hypothesis
+
 tmp <- bind_rows(
-  hypothesis(fit, hs)$hypothesis,
-  hypothesis(fit, hs, scope = "coef", group = "pid")$hypothesis
+  h_avg,
+  h_sub
 )

### 2023-09-01 ms.qmd
Ignore output style for now

@@
 format:
-  docx:
-    reference-doc: reference.docx
-  pdf:
-    keep-tex: false
-    documentclass: scrartcl
-    classoption: twocolumn
-    fontfamily: libertinus
-    fontsize: 10pt
-    papersize: a4
-    fig-align: center
-    fig-width: 8
-    section-number: true
-    geometry:
-      - top=20mm
-      - left=20mm
-      - right=20mm
-      - bottom=30mm
-    include-in-header:
-      - text: |
-          \setlength{\columnsep}{7.5mm}
-          \addtokomafont{disposition}{\normalfont}
+  html: default
 ---

### 2023-09-01 ms.qmd
Minor code cleaning

@@
 
-```{r setup}
+```{r}
+#| label: setup
 #| include: false
@@
 
-```{r packages}
+```{r}
+#| label: packages
 #| cache: false
+
 library(scales)
@@
 
-```{r options}
+```{r}
+#| label: options
 #| cache: false
@@
 
-```{r functions}
+```{r}
+#| label: functions
+
 # Format big numbers with a comma
@@
 }
+
 # Replace extreme percentages
@@
 if (!file.exists(path)) {
-  download.file("https://osf.io/download/j48qf/", destfile = "data-raw/data.zip")
+  download.file(
+    url = "https://osf.io/download/j48qf/",
+    destfile = "data-raw/data.zip"
+  )
   dir.create("data", FALSE)
-  unzip(path, files = c("data/demographics.csv", "data/study_prompt_answered.csv"))
+  unzip(
+    zipfile = path,
+    files = c("data/demographics.csv", "data/study_prompt_answered.csv")
+  )
 }
@@
 
-```{r data-load}
-# Load data from csv
+```{r}
+#| label: data-load
+
 dat <- read_csv(
@@
       (duration == 0) |
-      # session duration is less than previous session duration (ticker has reset),
+      # session is shorter than previous session (ticker has reset),
       (duration < lag(duration, default = 0)) |
@@
 
-# Filter data: Just wellbeing responses that occurred in sessions with both "pre" and "post" measures
+# Filter data:
+# Include only wellbeing responses that occurred
+# in sessions with both "pre" and "post" measures
 dat <- dat |>
@@
     duration[1] == 0,
-    # Session has at least two non-missing wellbeing measures (e.g. pre + 1 post)
+    # Session has at least two non-missing wellbeing measures 
+    # (e.g. pre + 1 post)
     sum(!is.na(mood)) > 1,
@@
 
-```{r demographics-load}
+```{r}
+#| label: demographics-load
+
 demo <- read_csv(
@@
 ```{r}
+#| label: data-wrangle-for-analysis
+
 dat <- dat |>
@@
 ```{r}
+#| label: model-estimate
+
 set.seed(999)
@@
 ```{r}
+#| label: wip-contrasts
 #| eval: false
+
 h_sub <- hypothesis(fit, hs, scope = "coef", group = "pid")$hypothesis

### 2023-09-01 ms.qmd
Move code block up
Small model revision
Also added some verbal description

@@
 ```
-
-# Methods
 
@@
 
+# Methods
+
 ## Participants
@@
 
-Our primary contrast of interest was between mood levels at the beginning of each session and mood levels during the subsequent play session.
-We modelled each mood report *i* in 1 to `r nrow(dat)` for participant *j* in 1 to `r length(unique(dat$pid))` and participant-session *k* in 1 to `r length(unique(dat$ps))` with a normal distribution censored at the VAS limits [0, 1].
+Our primary contrast of interest was between player mood at the beginning of each session and mood during the subsequent play session.
+We implemented this contrast within a censored three-level hierarchical location-scale model: First, because mood was measured with a VAS, we specified a censored gaussian model of the responses.
+We then modelled the mean of the response distribution on indicator variables for time (pre, post), and allowed both parameters to vary randomly across players.
+In addition, we modelled random average moods over player-sessions.
+Also, because equal variances across people in such naturalistic observation seems unlikely, we specified a similar model for the response distribution's standard deviation, but only modelled random deviations across participants, and so assumed equal residual variation across sessions within players.
+
+More formally, we modelled each mood response *i* in 1 to `r s$n_obs` from participant *j* in 1 to `r s$n_sub` and participant-session *k* in 1 to `r s$n_ses` with a normal distribution censored at the VAS limits [0, 1].
+We then modelled the distribution's mean on indicator variables (0 or 1) for pre-play and during play, their random per-player deviations $\beta$, and a random deviation for player-sessions $\gamma_0$.
+We modelled the residual deviation on a global average and per-player deviations, through a log link function:
 
@@
 \begin{align*}
-\text{mood}_{ijk} &\sim \text{Normal}(\eta_{ijk}, \kappa_{ij}), \\
-\eta_{ijk} &= \alpha_0 + \beta_{0j} + \gamma_{0k} + \\ 
-  &\quad \ (\alpha_1 + \beta_{1j} + \gamma_{1k})\text{post}_{ijk}, \\
-log(\kappa_{ij}) &= \alpha_2 + \beta_{2j} + \alpha_3\text{post}_{ijk}, \\
+\text{mood}_{ijk} &\sim \text{CensNormal}(\eta_{ijk}, \kappa_{ij}, 0, 1), \\
+\eta_{ijk} &= (\alpha_0 + \beta_{0j})\text{pre}_{ijk} +
+  (\alpha_1 + \beta_{1j})\text{post}_{ijk} +
+  \gamma_{0k}, \\
+log(\kappa_{ij}) &= \alpha_2 + \beta_{2j}, \\
 \begin{bmatrix} \alpha_0 \\ \alpha_1 \\ \alpha_2 \end{bmatrix} &\sim 
@@
   \end{bmatrix}\right), \\
-  \begin{bmatrix} \gamma_0 \\ \gamma_1 \end{bmatrix} &\sim
-    \text{MVN}\left(\begin{bmatrix} 0 \\ 0 \end{bmatrix}, 
-    \begin{bmatrix} 
-      \tau_{\gamma_0} &\rho_{{\gamma_0}{\gamma_1}} \\ 
-      \rho_{{\gamma_0}{\gamma_1}} &\tau_{\gamma_1}
-  \end{bmatrix}\right).
+  \gamma_0 &\sim \text{Normal}(0, \tau_{\gamma_0})
 \end{align*}
@@
 dat_snip <- dat |>
-  filter(pid %in% sample(unique(dat$pid), 300))
+  filter(pid %in% sample(unique(dat$pid), 100))
 
 model <- bf(mood | cens(cl) ~ 0 + post + (0 + post |p| pid) + (1 | ps)) +
-  lf(sigma ~ 0 + post + (1 |p| pid)) +
+  lf(sigma ~ 1 + (1 |p| pid)) +
   gaussian()

### 2023-09-04 ms.qmd
move metadata to same file

@@
 ---
-date: 2023-07-18
+title: "Mood increases during video game play: A naturalistic observational case study of PowerWash Simulator"
+description: Computational notebook supporting our manuscript
+author:
+  - name: Matti Vuorre
+    orcid: 0000-0001-5052-066X
+    email: mjvuorre@uvt.nl
+    url: https://vuorre.netlify.app
+    affiliation:
+      - name: Tilburg University
+        url: https://www.tilburguniversity.edu/staff/m-j-vuorre
+  - name: Nick Ballou
+    orcid: 0000-0003-4126-0696
+    affiliation:
+      - name: University of Oxford
+  - name: Andrew K. Przybylski
+    orcid: 0000-0001-5547-2185
+    affiliation:
+      - name: University of Oxford
+date: 2023-09-04
 bibliography: bibliography.bib
 format:
+  docx: default
   html: default

### 2023-09-04 ms.qmd
Fix metadata

--- 2023-09-01 ms.qmd
@@
 ---
-date: 2023-07-18
+title: "Mood increases during video game play: A naturalistic observational case study of PowerWash Simulator"
+description: Computational notebook supporting our manuscript
+author:
+  - name: Matti Vuorre
+    orcid: 0000-0001-5052-066X
+    email: mjvuorre@uvt.nl
+    url: https://vuorre.netlify.app
+    affiliation:
+      - ref: a
+      - ref: b
+  - name: Nick Ballou
+    orcid: 0000-0003-4126-0696
+    affiliation:
+      - ref: b
+  - name: Andrew K. Przybylski
+    orcid: 0000-0001-5547-2185
+    affiliation:
+      - ref: b
+affiliations:
+  - id: a
+    name: Tilburg University
+  - id: b
+    name: University of Oxford
+
+date: 2023-09-04
 bibliography: bibliography.bib
 format:
+  docx: default
   html: default

### 2023-09-04 ms.qmd
Estimate full model

@@
   \end{bmatrix}\right), \\
-  \gamma_0 &\sim \text{Normal}(0, \tau_{\gamma_0})
+  \gamma_0 &\sim \text{Normal}(0, \tau_{\gamma_0}).
 \end{align*}
@@
 
-set.seed(999)
-dat_snip <- dat |>
-  filter(pid %in% sample(unique(dat$pid), 100))
-
 model <- bf(mood | cens(cl) ~ 0 + post + (0 + post |p| pid) + (1 | ps)) +
@@
   model,
-  data = dat_snip,
+  data = dat,
   silent = 0,
   threads = 2,
-  iter = 1500,
+  iter = 100,
   control = list(adapt_delta = .9),

### 2023-09-05 ms.qmd
Reparameterize model

@@
 
-We analyzed the data with R [@rcoreteamLanguageEnvironmentStatistical2023] and used the brms package for models [@burknerBrmsPackageBayesian2017].
-
 ```{r}
@@
 
-model <- bf(mood | cens(cl) ~ 0 + post + (0 + post |p| pid) + (1 | ps)) +
-  lf(sigma ~ 1 + (1 |p| pid)) +
+contrasts(dat$post) <- c(-0.5, 0.5)
+
+model <- bf(mood ~ 1 + post + (1 + post | pid) + (1 | ps)) +
   gaussian()
@@
   data = dat,
-  silent = 0,
+  silent = 0, # Required for printing mcmc progress on some systems
   threads = 2,
-  iter = 100,
-  control = list(adapt_delta = .9),
-  file = "models/brm-pre-post-location-scale"
+  iter = 500,
+  control = list(adapt_delta = .95),
+  file = "models/brm-pre-post-pid-ps"
 )
-summary(fit, prior = TRUE)
 ```
+
+```{r}
+#| label: model-convergence
+#| eval: false
+
+hist(rhat(fit))
+plot(fit)
+pairs(fit$fit, pars = variables(fit)[1:7])
+```
+
+We analyzed the data with R [@rcoreteamLanguageEnvironmentStatistical2023] and used the brms package to estimate and post-process the models [@burknerBrmsPackageBayesian2017].
+We used Stan's HMC sampler to draw `r number2(ndraws(fit))` samples from the posterior distribution, and used brms' default prior distributions on all parameters, and used numerical and graphical checks to ensure model convergence.
 

### 2023-09-05 ms.qmd
WIP updates

@@
 library(cmdstanr)
+library(posterior)
 library(brms)
@@
 In this study, we only studied subjective well-being (mood), so did not consider the other five types of item responses in the database (enjoyment, focus, autonomy, competence, and immersion).
-Therefore, here we used data from `r s$n_sub` participants who provided `r s$n_obs` mood responses (`r s$n_obs_missing` with missing values) from `r s$n_ses` gameplay sessions.
+Therefore, here we used data from `r s$n_sub` participants who provided `r s$n_obs` mood responses (`r s$n_obs_missing` with missing values, which we ignored) from `r s$n_ses` gameplay sessions.
 
@@
 
-Our primary contrast of interest was between player mood at the beginning of each session and mood during the subsequent play session.
-We implemented this contrast within a censored three-level hierarchical location-scale model: First, because mood was measured with a VAS, we specified a censored gaussian model of the responses.
-We then modelled the mean of the response distribution on indicator variables for time (pre, post), and allowed both parameters to vary randomly across players.
-In addition, we modelled random average moods over player-sessions.
-Also, because equal variances across people in such naturalistic observation seems unlikely, we specified a similar model for the response distribution's standard deviation, but only modelled random deviations across participants, and so assumed equal residual variation across sessions within players.
+Our research question concerned the difference between players’ mood at the beginning of each session (pre-play) and during the subsequent play session (during play).
+It is important to understand that this contrast does not embed a causal hypothesis: Players could begin (and end) their play ses-sions for whatever reason, and these reasons are likely to confound the pre -- during contrast.
+For example, a player might come home after a stressful day at work, and sit down to play some PWS.
+During this play session, their mood would improve because they are at home vs.
+stressful work environment, and would show as a greater during play mood compared to pre-play mood: Reasons behind starting a play session are likely to contribute to the pre -- during contrast and we are unable to disentangle those from any changes actually caused by play.
+
+We estimated this contrast within a censored three-level hierarchical regression model.
+We specified a censored gaussian model because mood was measured with a VAS.
+Ignoring the censoring inherent in VAS responses can confound changes in the mean with changes in the spread of mood, and leave the contrast susceptible to ceiling or floor effects.
+We then modelled mean mood on an intercept and contrast coded time (pre-play: -0.5; during play: 0.5), and allowed both parameters to vary randomly across players.
+In addition, we modelled random intercepts over player-sessions because over time, mood is susceptible to myriad other influences.
+Although equal variances across people in such naturalistic observation seem unlikely, we fixed this parameter to a constant for model identifiability and convergence.
 
 More formally, we modelled each mood response *i* in 1 to `r s$n_obs` from participant *j* in 1 to `r s$n_sub` and participant-session *k* in 1 to `r s$n_ses` with a normal distribution censored at the VAS limits [0, 1].
-We then modelled the distribution's mean on indicator variables (0 or 1) for pre-play and during play, their random per-player deviations $\beta$, and a random deviation for player-sessions $\gamma_0$.
-We modelled the residual deviation on a global average and per-player deviations, through a log link function:
+We then modelled the distribution's mean ($\eta$) on a population-level intercept ($\beta_0$) and contrast of time ($\beta_1$), their random per-player deviations $\alpha$, and a random intercept for player-sessions $\gamma_0$:
 
@@
 \begin{align*}
-\text{mood}_{ijk} &\sim \text{CensNormal}(\eta_{ijk}, \kappa_{ij}, 0, 1), \\
-\eta_{ijk} &= (\alpha_0 + \beta_{0j})\text{pre}_{ijk} +
+\text{mood}_{ijk} &\sim \text{CensNormal}(\eta_{ijk}, \sigma^2, 0, 1), \\
+\eta_{ijk} &= \alpha_0 + \beta_{0j} +
   (\alpha_1 + \beta_{1j})\text{post}_{ijk} +
   \gamma_{0k}, \\
-log(\kappa_{ij}) &= \alpha_2 + \beta_{2j}, \\
-\begin{bmatrix} \alpha_0 \\ \alpha_1 \\ \alpha_2 \end{bmatrix} &\sim 
-  \text{MVN}\left(\begin{bmatrix} 0 \\ 0 \\ 0 \end{bmatrix}, 
+\begin{bmatrix} \alpha_0 \\ \alpha_1 \end{bmatrix} &\sim 
+  \text{MVN}\left(\begin{bmatrix} 0 \\ 0 \end{bmatrix}, 
   \begin{bmatrix} 
-    \tau_{\alpha_0} &\rho_{{\alpha_0}{\alpha_1}} &\rho_{{\alpha_0}{\alpha_2}} \\ 
-    \rho_{{\alpha_0}{\alpha_1}} &\tau_{\alpha_1} &\rho_{{\alpha_1}{\alpha_2}} \\ 
-    \rho_{{\alpha_0}{\alpha_2}} &\rho_{{\alpha_1}{\alpha_2}} &\tau_{\alpha_2} 
+    \tau_{\alpha_0} &\rho_{{\alpha_0}{\alpha_1}} \\ 
+    \rho_{{\alpha_0}{\alpha_1}} &\tau_{\alpha_1}
   \end{bmatrix}\right), \\
@@
 ```{r}
+#| label: model-build
+
+model <- bf(mood ~ 1 + post + (1 + post |p| pid) + (1 | ps)) +
+  lf(sigma ~ 1 + (1 |p| pid)) +
+  gaussian()
+
+fit2 <- brm(
+  model,
+  data = dat,
+  silent = 0,
+  threads = 2,
+  iter = 500,
+  control = list(adapt_delta = .95),
+  file = "models/brm-pre-post-pid-ps-scale"
+)
+
+model <- bf(mood ~ 1 + post + (1 + post |p| pid)) +
+  lf(sigma ~ 1 + (1 |p| pid)) +
+  gaussian()
+
+fit3 <- brm(
+  model,
+  data = dat,
+  silent = 0,
+  threads = 2,
+  iter = 500,
+  control = list(adapt_delta = .95),
+  file = "models/brm-pre-post-pid-scale"
+)
+
+model <- bf(mood | cens(cl) ~ 1 + post + (1 + post | pid) + (1 | ps)) +
+  gaussian()
+
+fit4 <- brm(
+  model,
+  data = dat,
+  silent = 0,
+  threads = 2,
+  iter = 500,
+  control = list(adapt_delta = .95),
+  file = "models/brm-pre-post-pid-ps-censored"
+)
+fit <- fit4
+```
+
+```{r}
 #| label: model-convergence
@@
 
-hist(rhat(fit))
-plot(fit)
-pairs(fit$fit, pars = variables(fit)[1:7])
+library(posterior)
+as_draws(fit4, variable = c("b_", "sd_", "cor_"), regex = TRUE) |> 
+  summarise_draws(rhat)
+pairs(fit4$fit, pars = variables(fit)[1:7])
 ```
 
-We analyzed the data with R [@rcoreteamLanguageEnvironmentStatistical2023] and used the brms package to estimate and post-process the models [@burknerBrmsPackageBayesian2017].
-We used Stan's HMC sampler to draw `r number2(ndraws(fit))` samples from the posterior distribution, and used brms' default prior distributions on all parameters, and used numerical and graphical checks to ensure model convergence.
+We analyzed the data with R [@rcoreteamLanguageEnvironmentStatistical2023] and used the brms package to estimate, via Stan's HMC sampling procedures, and post-process the models [@@burknerBrmsPackageBayesian2017; @standevelopmentteamStanModelingLanguage2022].
+We drew `r number2(ndraws(fit))` samples from the posterior distribution using brms' default prior distributions on all parameters, and used numerical and graphical checks to ensure model convergence and adequacy.
 
@@
   dat_session_counts +
+  coord_cartesian(xlim = c(0, 15)) +
   aes(sessions) +
@@
 ```{r}
+probs <- c(.5, .1, .9)
+tmp <- list()
+tmp$a <- quantile(dat_session_durations$duration, probs = probs, na.rm = TRUE)
+tmp$b <- quantile(dat_session_counts$sessions, probs = probs, na.rm = TRUE)
+tmp$c <- quantile(dat$mood, probs = probs, na.rm = TRUE)
+tmp$d <- quantile(dat$mood[dat$post == 0], probs = probs, na.rm = TRUE)
+tmp$e <- quantile(dat$mood[dat$post == 1], probs = probs, na.rm = TRUE)
+```
+
+The median session duration was `r tmp$a` minutes [10 and 90 percentiles] (@fig-descriptives A); the median player contributed data from `r tmp$b` sessions (@fig-descriptives B), and the median mood was `r tmp$c` (@fig-descriptives C; pre-session: `r tmp$d`, during play: `r tmp$e`).
+
+```{r}
 #| label: fig-data
 #| include: true
-#| fig-width: 7.2
+#| fig-width: 8
 #| fig-height: 5
@@
   facet_grid(
-    Person ~ Session,
+    rows = vars(Person), 
+    cols = vars(Session),
     scales = "free_x"
@@
   theme(
-    plot.margin = unit(c(8, 8, 8, 8), "pt"),
-    # strip.background.y = element_blank(),
-    # strip.text.y = element_blank()
+    legend.position = "none",
+    strip.text.y = element_blank()
   )
@@
     limits = c(0, 1),
-    # expand = expansion(c(0, 0.01)),
     breaks = axislabels,
@@
   ) +
-  coord_fixed(ratio = 0.66) +
   geom_dotplot(
@@
     strip.background.y = element_blank(),
-    strip.text.y = element_blank()
-    # plot.margin = unit(c(8, 8, 8, 8), "pt")
+    legend.position = "none"
   )
 
-p_mood_duration <- (p_mood_duration_sessions | p_mood_duration_all) &
-  plot_layout(widths = c(80, 20)) &
-  theme(legend.position = "none")
+p_mood_duration <- (p_mood_duration_sessions | p_mood_duration_all) +
+  plot_layout(widths = c(5/6, 1/6))
 
@@
   scale_x_continuous(
-    "Mood at start of session",
+    "Pre-session mood",
     expand = expansion(0.01),
@@
   theme(aspect.ratio = 1)
-p_mood_difference <- tmp |> 
-  drop_na(delta) |> 
-  ggplot(aes(delta)) +
-  scale_y_continuous(
-    "Participants",
-    expand = expansion(c(0.001, 0.1))
-  ) +
-  scale_x_continuous(
-    "Difference (subsequent - first mood)"
-  ) +
-  coord_cartesian(xlim = c(-0.4, 0.6)) +
-  geom_histogram(binwidth = .025, col = "white") +
-  stat_summaryh(
-    fun.data = mean_cl_boot_h,
-    aes(y = 10),
-    col = "white",
-    fill = "black",
-    shape = 21,
-    stroke = 1.5, size = 0.75
-  )
 
@@
     `During session` = `1`,
-    Difference = delta
+    `Difference (During - Pre)` = delta
   ) |> 
@@
 #| include: true
-#| tbl-cap: Population level average mood at beginning and during session, and difference.
+#| tbl-cap: Population level average mood at beginning and during a game play session, and their difference (during - pre-play).
+(Estimate = posterior mean, CI = 95% quantile interval, pd = probability of direction.)
 
 hs <- c(
-  "Pre" = "post0 = 0",
-  "Post" = "post1 = 0",
-  "Difference" = "post1 - post0 = 0"
+  "Pre-play" = "Intercept + post1 * -.5 = 0",
+  "During play" = "Intercept + post1 * .5 = 0",
+  "Difference" = "post1 = 0"
 )
 
-h_avg <- hypothesis(fit, hs)$hypothesis
+h_avg_ <- hypothesis(fit, hs)
+h_avg <- tibble(h_avg_$samples)
+names(h_avg) <- names(hs)
 h_avg |> 
-  select(` ` = Hypothesis, Estimate, CI.Lower, CI.Upper) |> 
-  mutate(across(where(is.numeric), ~percent2(., .01))) |> 
-  knitr::kable(digits = 3)
+  summarise_draws(
+    mean = ~number(mean(.), .001), 
+    ~number2(quantile2(., probs = c(.025, .975)), .001), 
+    pd = ~percent2(Pr(. > 0))
+  ) |> 
+  mutate(CI = str_glue("[{q2.5}, {q97.5}]")) |> 
+  select(` ` = variable, Estimate = mean, CI, pd) |> 
+  knitr::kable()
 ```
 
+We then turned to our primary research question: To what extent does player mood in PWS change from pre-play to during play?
+First, we visualize the relevant data in (@fig-data): Panel A shows mood responses from three example participants' first five sessions of play.
+@fig-data B then shows histograms of those data to facilitate visual comparison of the raw data.
+We then calculated participants' mean moods at the beginning and during each session (@fig-data C): Visual inspection shows a considerable shift toward higher values for the during play means.
+@fig-data D shows histograms of the pre- and during play mood means, and their difference, and suggests that while many differences are small, overall mean moods appeared greater during play than pre-play.
+
+The model results confirmed the visual impressions described above: @tbl-avg indicates, with great certainty, that mood increased by about [`r paste0(map(h_avg_$hypothesis[3,4:5], ~number(.x, .001)), collapse = ", ")`] units.
+
 ```{r}
-#| label: wip-contrasts
+#| label: scaled-coefficients
 #| eval: false
 
-h_sub <- hypothesis(fit, hs, scope = "coef", group = "pid")$hypothesis
-
-tmp <- bind_rows(
-  h_avg,
-  h_sub
-)
-tmp |> 
-  ggplot(aes(Estimate, Hypothesis)) +
-  geom_point(
-    data = . %>% filter(is.na(Group)),
-    size = 1
-  ) +
-  geom_point(
-    data = . %>% filter(!is.na(Group)),
-    size = .25
+library(latex2exp)
+tmp_ <- as_draws_df(
+  fit, 
+  c("sd_pid__Intercept", "sd_pid__post1", "sigma", "b_post1")
+) |> 
+  mutate(
+    total_residual = sd_pid__Intercept + sd_pid__post1 + sigma,
+    z0 = b_post1 / total_residual,
+    z1 = b_post1 / sd_pid__Intercept,
+    z2 = b_post1 / sd_pid__post1,
+    z3 = b_post1 / sigma
   )
 
+tmp <- tmp_ |> 
+  summarise_draws(
+    mean = ~number(mean(.), .001), 
+    ~number2(quantile2(., probs = c(.025, .975)), .001)
+  ) |> 
+  mutate(
+    x = str_glue("{mean} [{q2.5}, {q97.5}]")
+  )
+tmp_ |> 
+  select(b_post1, starts_with("z")) |> 
+  pivot_longer(everything()) |> 
+  ggplot(aes(value, name)) +
+  stat_slabinterval(
+    .width = c(.8, .95),
+    normalize = "xy",
+    height = .5,
+    interval_size_range = c(0.5, 1),
+    fatten_point = 1
+  ) +
+  scale_x_continuous(
+    "Contrast magnitude",
+    breaks = scales::extended_breaks(7)
+  ) +
+  scale_y_discrete(
+    labels = c(
+      TeX(r"($\alpha_1$)"),
+      TeX(r"($\frac{\alpha_1}{\tau_0 + \tau_1 + \sigma}$)"),
+      TeX(r"($\frac{\alpha_1}{\tau_0}$)"),
+      TeX(r"($\frac{\alpha_1}{\tau_1}$)"),
+      TeX(r"($\frac{\alpha_1}{\sigma}$)")
+    ),
+    expand = expansion(0.1)
+  ) +
+  theme(
+    axis.title.y = element_blank()
+  )
+
+coef(fit, summary = FALSE)$pid[,,"post1"] |> 
+  as_tibble()
+
+fixef(fit, summary = FALSE)[,"post1", drop = FALSE] |> 
+  as_tibble()
 ```
+
+To further interpret the magnitude of the difference in moods between pre- and during play, we compared the contrast to the natural variations in mood as estimated by the model.
+Specifically, we compared the estimated pre - during contrast to the standard deviation of the between-person differences in the intercept ($\tau_{\alpha_0}$ = `r tmp$x[1]`), the contrast ($\tau_{\alpha_1}$ = `r tmp$x[2]`) and to the residual within-person variation ($\sigma$).
 
@@
 :::
-
-# Appendices / supplementary analyses / stuff

### 2023-09-05 ms.qmd
Fix #5

@@
 if (!file.exists(path)) {
+  dir.create("data-raw", FALSE)
   download.file(

### 2023-09-06 ms.qmd
Fix build

@@
 format:
+  pdf: default
   docx: default
@@
 
-$$
 \begin{align*}
@@
 \end{align*}
-$$
 
@@
 #| label: model-estimate
+#| cache: false
 
@@
 
-model <- bf(mood ~ 1 + post + (1 + post | pid) + (1 | ps)) +
+model <- bf(mood | cens(cl) ~ 1 + post + (1 + post | pid) + (1 | ps)) +
   gaussian()
@@
   data = dat,
-  silent = 0, # Required for printing mcmc progress on some systems
-  threads = 2,
-  iter = 500,
-  control = list(adapt_delta = .95),
-  file = "models/brm-pre-post-pid-ps"
-)
-```
-
-```{r}
-#| label: model-build
-
-model <- bf(mood ~ 1 + post + (1 + post |p| pid) + (1 | ps)) +
-  lf(sigma ~ 1 + (1 |p| pid)) +
-  gaussian()
-
-fit2 <- brm(
-  model,
-  data = dat,
   silent = 0,
   threads = 2,
-  iter = 500,
-  control = list(adapt_delta = .95),
-  file = "models/brm-pre-post-pid-ps-scale"
-)
-
-model <- bf(mood ~ 1 + post + (1 + post |p| pid)) +
-  lf(sigma ~ 1 + (1 |p| pid)) +
-  gaussian()
-
-fit3 <- brm(
-  model,
-  data = dat,
-  silent = 0,
-  threads = 2,
-  iter = 500,
-  control = list(adapt_delta = .95),
-  file = "models/brm-pre-post-pid-scale"
-)
-
-model <- bf(mood | cens(cl) ~ 1 + post + (1 + post | pid) + (1 | ps)) +
-  gaussian()
-
-fit4 <- brm(
-  model,
-  data = dat,
-  silent = 0,
-  threads = 2,
-  iter = 500,
+  iter = 2000,
   control = list(adapt_delta = .95),
@@
 )
-fit <- fit4
 ```
@@
 library(posterior)
-as_draws(fit4, variable = c("b_", "sd_", "cor_"), regex = TRUE) |> 
+as_draws(fit, variable = c("b_", "sd_", "cor_"), regex = TRUE) |> 
   summarise_draws(rhat)
-pairs(fit4$fit, pars = variables(fit)[1:7])
+pairs(fit$fit, pars = variables(fit)[1:7])
 ```

### 2023-09-06 ms.qmd
WIP changes

@@
 library(ggstance)
+library(ggdist)
 library(patchwork)
@@
 
-## Measures
-
-We measured affective well-being with one item.
-
 ## Data analysis
@@
 More formally, we modelled each mood response *i* in 1 to `r s$n_obs` from participant *j* in 1 to `r s$n_sub` and participant-session *k* in 1 to `r s$n_ses` with a normal distribution censored at the VAS limits [0, 1].
-We then modelled the distribution's mean ($\eta$) on a population-level intercept ($\beta_0$) and contrast of time ($\beta_1$), their random per-player deviations $\alpha$, and a random intercept for player-sessions $\gamma_0$:
+We then modelled the distribution's mean ($\eta$) on a population-level intercept ($\beta_0$) and contrast-coded measurement timepoint ($\beta_1$), their random per-player deviations $\alpha$, and a random intercept for player-sessions $\gamma_0$:
 
@@
 
-We analyzed the data with R [@rcoreteamLanguageEnvironmentStatistical2023] and used the brms package to estimate, via Stan's HMC sampling procedures, and post-process the models [@@burknerBrmsPackageBayesian2017; @standevelopmentteamStanModelingLanguage2022].
+We analyzed the data with R [@rcoreteamLanguageEnvironmentStatistical2023] and used the brms package to estimate, via Stan's HMC sampling procedures, and post-process the models [@burknerBrmsPackageBayesian2017; @standevelopmentteamStanModelingLanguage2022].
 We drew `r number2(ndraws(fit))` samples from the posterior distribution using brms' default prior distributions on all parameters, and used numerical and graphical checks to ensure model convergence and adequacy.
@@
 
-## Descriptives
-
 ```{r}
 #| label: fig-descriptives
-#| include: true
+#| include: false
 #| fig-height: 2.2
@@
   scale_x_log10() +
+  coord_cartesian(xlim = c(5, 900)) +
   labs(x = "Session duration (minutes)", y = "Sessions")
@@
 ```{r}
-probs <- c(.5, .1, .9)
-tmp <- list()
-tmp$a <- quantile(dat_session_durations$duration, probs = probs, na.rm = TRUE)
-tmp$b <- quantile(dat_session_counts$sessions, probs = probs, na.rm = TRUE)
-tmp$c <- quantile(dat$mood, probs = probs, na.rm = TRUE)
-tmp$d <- quantile(dat$mood[dat$post == 0], probs = probs, na.rm = TRUE)
-tmp$e <- quantile(dat$mood[dat$post == 1], probs = probs, na.rm = TRUE)
+tmp <- list(
+  dat_session_durations$duration,
+  dat_session_counts$sessions,
+  dat$mood,
+  dat$mood[dat$post == 0],
+  dat$mood[dat$post == 1]
+) |> 
+  map(~median_qi(.x, .width = 0.8, na.rm = TRUE)) |> 
+  map(~str_glue_data(.x, "{y} [{ymin}, {ymax}]"))
 ```
 
-The median session duration was `r tmp$a` minutes [10 and 90 percentiles] (@fig-descriptives A); the median player contributed data from `r tmp$b` sessions (@fig-descriptives B), and the median mood was `r tmp$c` (@fig-descriptives C; pre-session: `r tmp$d`, during play: `r tmp$e`).
+The median session duration was `r tmp[[1]]` minutes [10 and 90 percentiles]; the median player contributed data from `r tmp[[2]]` sessions, and the median mood was `r tmp[[3]]` (pre-session: `r tmp[[4]]`, during play: `r tmp[[5]]`).
 
@@
 Representative data from three participants (rows) showing scatterplots of mood responses over that session's (columns) duration.
-Each session's first mood report is highlighted in red, subsequent reports are in blue.
+Each session's first mood response is highlighted in red, subsequent reports are blue.
 **B**.
-Horizontal histograms of all sessions' mood response for that person.
+Horizontal histograms of all sessions' mood response for that person, with means and bootstrapped 95%CIs.
 **C**.
-Scatterplot of person-mean mood reports at the beginning of the session (x-axis) and subsequent reports (y-axis).
-**D**.
-Histograms of mood ratings at the beginning of session, during the session, and differences between person-means of first and subsequent mood reports."
+Histograms of session-mean (**D**.
+person-mean) moods before (top) and during (bottom) sessions.
+**E**.
+Differences in session-mean (**F**.
+player-mean) mood differences (during session - pre-play).
+**G**.
+Scatterplot of person-mean mood reports at beginning (x-axis) and during game play sessions (y-axis)."
 
@@
 
-# Take a quasirandom sample of people who had sufficient number of sessions
-# and data per session.
+# Plot: Example raw sessions
 set.seed(99)
-tmp <- dat |>
+dat_example_sessions <- dat |>
   add_count(pid, session, name = "obs_per_session") |>
@@
 
-library(ggdist)
-p_mood_duration_sessions <- tmp |> 
+p_mood_example_sessions <- dat_example_sessions |> 
   ggplot() +
@@
 
-p_mood_duration_all <- tmp |> 
+p_mood_example_all <- dat_example_sessions |> 
   mutate(Session = "All sessions") |> 
   ggplot() +
-  aes(y = mood, col = post, fill = post) +
+  aes(y = mood, x = post, col = post, fill = post) +
   scale_color_brewer(
@@
   ) +
-  scale_x_continuous(
-    "Scaled count",
-    expand = expansion(c(0.01, 0.1))
+  scale_x_discrete(
+    "Session",
+    breaks = c(0, 1),
+    labels = c("Pre", "During"),
+    expand = expansion(c(0.2, 2.4))
   ) +
@@
   ) +
-  geom_dotplot(
-    aes(x = 0),
-    dotsize = 0.9,
-    binaxis = "y",
-    method = "histodot",
-    binwidth = 0.06,
-    stackgroups = TRUE
+  stat_summary(
+    fun.data = mean_cl_boot,
+    fatten = 2
   ) +
+  geom_dots(side = "right", alpha = .33, binwidth = .075) +
   facet_grid(
@@
   theme(
-    axis.text = element_blank(), 
+    axis.text.y = element_blank(), 
     axis.title.y = element_blank(), 
     axis.ticks.y = element_blank(), 
+    axis.text.x = element_text(hjust = 0),
     strip.background.y = element_blank(),
@@
 
-p_mood_duration <- (p_mood_duration_sessions | p_mood_duration_all) +
-  plot_layout(widths = c(5/6, 1/6))
+p_mood_example <- (p_mood_example_sessions | p_mood_example_all) +
+  plot_layout(widths = c(92, 8))
 
-tmp <- dat |> 
+# Plot: Raw mood histograms at pre and during play
+p_mood_prepost_raw <- dat |> 
+  mutate(post = factor(post, labels =  c("Pre-play", "During play"))) |>
+  ggplot(aes(mood, fill = post)) +
+  scale_color_brewer(
+    "Pre-session measure",
+    palette = "Set1",
+    aesthetics = c("color", "fill")
+  ) +
+  geom_histogram(bins = 30, col = "white") +
+  scale_y_continuous(
+    "Observations",
+    expand = expansion(c(0.001, 0.1)),
+    breaks = scales::extended_breaks(5)
+  ) +
+  scale_x_continuous(
+    "Mood",
+    expand = expansion(c(0.001)),
+    breaks = axislabels
+  ) +
+  facet_wrap("post", ncol = 1, scales = "free_y") +
+  theme(legend.position = "none")
+
+# Plot: Person-session-mean mood histograms at pre and during play
+p_mood_prepost_sessions <- p_mood_prepost_raw %+%
   summarise(
-    mood = mean(mood, na.rm = TRUE),
-    .by = c(pid, post)
-  ) |> 
+    p_mood_prepost_raw$data,
+    mood = mean(mood, na.rm = TRUE), .by = c(post, pid, ps)
+  ) +
+  scale_y_continuous(
+    "Sessions",
+    expand = expansion(c(0.001, 0.1)),
+    breaks = scales::extended_breaks(5)
+  )
+
+# Plot: Person-mean mood histograms at pre and during play
+p_mood_prepost_players <- p_mood_prepost_sessions %+%
+  summarise(
+    p_mood_prepost_raw$data,
+    mood = mean(mood, na.rm = TRUE), .by = c(post, pid)
+  ) +
+  scale_y_continuous(
+    "Players",
+    expand = expansion(c(0.001, 0.1)),
+    breaks = scales::extended_breaks(5)
+  )
+
+# Plot: Difference histogram (sessions)
+p_mood_difference_sessions <- p_mood_prepost_sessions$data |> 
   pivot_wider(names_from = post, values_from = mood) |> 
-  mutate(delta = `1` - `0`)
+  ggplot(aes(`During play` - `Pre-play`)) +
+  geom_histogram(bins = 50, col = "white") +
+  geom_vline(xintercept = 0, linewidth = .33, col = "dodgerblue") +
+  scale_y_continuous(
+    "Sessions",
+    expand = expansion(c(0.001, 0.1))
+  ) +
+  scale_x_continuous(
+    "Difference",
+    expand = expansion(c(0.001))
+  ) +
+  coord_cartesian(xlim = c(-0.4, 0.6))
 
-p_mood_biscatter <- tmp |> 
-  ggplot(aes(`0`, `1`)) +
+# Plot: Difference histogram (players)
+p_mood_difference_players <- p_mood_difference_sessions %+% 
+  summarise(
+    p_mood_difference_sessions$data,
+    across(`Pre-play`:`During play`, ~mean(.x, na.rm = TRUE)),
+    .by = pid
+  ) +
+  scale_y_continuous(
+    "Players",
+    expand = expansion(c(0.001, 0.1))
+  )
+
+p_mood_difference <- (
+  (p_mood_difference_sessions + 
+     theme(
+      axis.title.x = element_blank(), 
+      axis.text.x = element_blank(),
+      axis.ticks.x = element_blank()
+    )) / 
+    p_mood_difference_players
+  )
+
+p_mood_biscatter_sessions <- p_mood_difference_sessions$data |> 
+  ggplot(aes(`Pre-play`, `During play`)) +
   scale_x_continuous(
-    "Pre-session mood",
     expand = expansion(0.01),
@@
   scale_y_continuous(
-    "Mood during session",
     expand = expansion(0.01),
@@
 
-p_mood_difference <- tmp |> 
-  rename(
-    `Pre-session` = `0`,
-    `During session` = `1`,
-    `Difference (During - Pre)` = delta
-  ) |> 
-  pivot_longer(-pid) |> 
-  mutate(name = fct_inorder(name)) |> 
-  ggplot(aes(value)) +
-  geom_histogram(bins = 30, col = "white") +
-  scale_y_continuous(
-    "Participants",
-    expand = expansion(c(0.001, 0.1))
+p_mood_biscatter_players <- p_mood_biscatter_sessions %+% 
+  p_mood_difference_players$data
+
+p_mood_biscatter <- (p_mood_biscatter_sessions / p_mood_biscatter_players)
+
+p_mood_example / 
+  (
+    p_mood_prepost_sessions | 
+      p_mood_prepost_players | 
+      p_mood_difference | 
+      p_mood_biscatter_players
   ) +
-  scale_x_continuous(
-    "Mood",
-    expand = expansion(c(0.001))
-  ) +
-  facet_wrap("name", ncol = 3, scales = "free")
+  plot_layout(heights = c(4.5, 5.5)) +
+  plot_annotation(tag_levels = "A")
+```
 
-(p_mood_duration / (p_mood_biscatter | p_mood_difference)) +
-  plot_layout(heights = c(6, 4)) +
-  plot_annotation(tag_level = "A")
-```
+We then turned to our primary research question: To what extent does player mood in PWS change from pre-play to during play?
+First, we visualized the relevant data in @fig-data: Panel A shows mood responses from three example participants' first five sessions of play.
+@fig-data B then shows histograms of those data to facilitate visual comparison of the raw data.
+We then calculated sessions' and participants' mean moods at the beginning and during each session (@fig-data C, D, and G), and their differences (E and F).
+Visual inspection suggested greater mood during play compared to the pre-play measures at both the level of sessions and players.
 
@@
 
-We then turned to our primary research question: To what extent does player mood in PWS change from pre-play to during play?
-First, we visualize the relevant data in (@fig-data): Panel A shows mood responses from three example participants' first five sessions of play.
-@fig-data B then shows histograms of those data to facilitate visual comparison of the raw data.
-We then calculated participants' mean moods at the beginning and during each session (@fig-data C): Visual inspection shows a considerable shift toward higher values for the during play means.
-@fig-data D shows histograms of the pre- and during play mood means, and their difference, and suggests that while many differences are small, overall mean moods appeared greater during play than pre-play.
-
 The model results confirmed the visual impressions described above: @tbl-avg indicates, with great certainty, that mood increased by about [`r paste0(map(h_avg_$hypothesis[3,4:5], ~number(.x, .001)), collapse = ", ")`] units.
@@
 #| label: scaled-coefficients
-#| eval: false
 
-library(latex2exp)
-tmp_ <- as_draws_df(
+posterior_pop <- as_draws_df(
   fit, 
@@
     total_residual = sd_pid__Intercept + sd_pid__post1 + sigma,
-    z0 = b_post1 / total_residual,
-    z1 = b_post1 / sd_pid__Intercept,
-    z2 = b_post1 / sd_pid__post1,
-    z3 = b_post1 / sigma
-  )
+    b_post1_scaled = b_post1 / total_residual
+  ) |> 
+  tibble() |> 
+  select(-starts_with("."))
 
-tmp <- tmp_ |> 
+posterior_pop_sum <- posterior_pop |> 
   summarise_draws(
@@
   )
-tmp_ |> 
-  select(b_post1, starts_with("z")) |> 
-  pivot_longer(everything()) |> 
-  ggplot(aes(value, name)) +
-  stat_slabinterval(
-    .width = c(.8, .95),
-    normalize = "xy",
-    height = .5,
-    interval_size_range = c(0.5, 1),
-    fatten_point = 1
-  ) +
-  scale_x_continuous(
-    "Contrast magnitude",
-    breaks = scales::extended_breaks(7)
-  ) +
-  scale_y_discrete(
-    labels = c(
-      TeX(r"($\alpha_1$)"),
-      TeX(r"($\frac{\alpha_1}{\tau_0 + \tau_1 + \sigma}$)"),
-      TeX(r"($\frac{\alpha_1}{\tau_0}$)"),
-      TeX(r"($\frac{\alpha_1}{\tau_1}$)"),
-      TeX(r"($\frac{\alpha_1}{\sigma}$)")
-    ),
-    expand = expansion(0.1)
-  ) +
-  theme(
-    axis.title.y = element_blank()
-  )
-
-coef(fit, summary = FALSE)$pid[,,"post1"] |> 
-  as_tibble()
-
-fixef(fit, summary = FALSE)[,"post1", drop = FALSE] |> 
-  as_tibble()
 ```
 
-To further interpret the magnitude of the difference in moods between pre- and during play, we compared the contrast to the natural variations in mood as estimated by the model.
-Specifically, we compared the estimated pre - during contrast to the standard deviation of the between-person differences in the intercept ($\tau_{\alpha_0}$ = `r tmp$x[1]`), the contrast ($\tau_{\alpha_1}$ = `r tmp$x[2]`) and to the residual within-person variation ($\sigma$).
+To further interpret the magnitude of the difference in moods between pre- and during play, we compared the contrast to the natural variations in mood as estimated by the model (standard deviation of the between-person intercept heterogeneity ($\tau_{\alpha_0}$ = `r posterior_pop_sum$x[1]`), between-person pre-during difference heterogeneity ($\tau_{\alpha_1}$ = `r posterior_pop_sum$x[2]`) and to the residual standard deviation ($\sigma$ = `r posterior_pop_sum$x[3]`).
+The standardized (with the total residual variation from the three above components) pre-play -- during play contrast was `r posterior_pop_sum$x[6]`, a not ignorable difference.
 

### 2023-09-07 continuous-time-draft.qmd
draft

---
title: "Continuous time"
date: now
author: Kristoffer Magnusson
format:
  html:
    code-fold: true
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  include = TRUE,
  echo = TRUE
)
```

```{r}
#| label: packages
#| cache: false
library(patchwork)
library(tidyverse)
library(lme4)
library(knitr)
library(emmeans)
```

Data cleaning copied from `ms.qmd` to create sessions and the pre/post contrast.
```{r}
#| label: options
#| cache: false
dat <- read_rds("data/session.rds")
dat2 <- dat |>
  filter(prompt == "Wellbeing") |>
  select(-prompt)

dat2 <- dat2 |>
  filter(
    # Session has a wellbeing measure at time = 0
    duration[1] == 0,
    # Session has at least two non-missing wellbeing measures 
    # (e.g. pre + 1 post)
    sum(!is.na(mood)) > 1,
    .by = c(pid, session)
  )

dat2 <- dat2 |>
  mutate(
    post = factor(
      row_number() > 1,
      levels = c(FALSE, TRUE),
      labels = c("0", "1")
    ),
    .by = c(pid, session)
  ) |>
  mutate(ps = interaction(pid, session, sep = "_")) |>
  mutate(
    cl = case_when(
      mood == 0 ~ "left",
      mood == 1 ~ "right",
      TRUE ~ "none"
    )
  )
```

## Exploring sessions
Let's look at one participant and compare the pre/post data set to the complete data.

```{r}
#| label: fig-duration
#| fig-cap: Comparing full data (top) with pre-post (bottom) data
p0 <- ggplot(
  filter(dat, 
    pid == "p6622",
    session %in% 14:16
    ),
  aes(
    time,
    session,
    group = session
  )
) + 
  geom_point(
    aes(color = factor(prompt == "Wellbeing")),
    size = 2
    ) +
  scale_x_datetime(limits = c(ymd_hms("2022-09-19 16:00:00"), ymd_hms("2022-09-20 00:00:00"))) +
  scale_y_continuous(breaks = 14:16) +
  facet_wrap(~pid)

p1 <- ggplot(
  filter(dat2, 
    pid == "p6622",
    session %in% 14:16
    ),
  aes(
    time,
    session,
    group = session
  )
) + 
  geom_point(
    aes(color = factor(post)),
    size = 2
    ) +
  scale_x_datetime(limits = c(ymd_hms("2022-09-19 16:00:00"), ymd_hms("2022-09-20 00:00:00"))) +
  scale_y_continuous(breaks = 14:16) +
  facet_wrap(~pid)
p0 / p1
```

## Concerns with the pre/post contrast.
There are issues with the pre/post contrasts.

### 1. Session duration doesn't always start at 0
As can be seen in @fig-duration, session 16 isn't included in the pre/post data, even tho the first response is a WB prompt. This is because it's duration is 6. 
```{r}
filter(dat, 
  pid == "p6622",
  session %in% 14:16
) %>% 
kable()

filter(dat2, 
  pid == "p6622",
  session %in% 14:16
) %>% 
kable()
```

### 2. Lot's of data is discarded
As can be seen in the @fig-pre-post there can be lot's of data that is not covered by the pre-post contrast. This might not be a problem as the prompts were randomized. But it's a lot of PWS gameplay that's also ignored.

```{r}
#| label: fig-pre-post
#| fig-cap: Comparing full data (top) with pre-post (bottom) data
p0 <- ggplot(
  filter(dat, 
    pid == "p10",
    session %in% 16:26
    ),
  aes(
    time,
    session,
    group = session
  )
) + 
  geom_point(
    aes(color = factor(prompt == "Wellbeing")),
    size = 2
    ) +
  scale_x_datetime(
    limits = c(
      ymd_hms("2023-02-25 00:00:00"), 
      ymd_hms("2023-02-29 00:00:00"
      )
    )
  ) +
  scale_y_continuous(breaks = 16:26)

p1 <- ggplot(
  filter(dat2, 
    pid == "p10",
    session %in% 16:26
    ),
  aes(
    time,
    session,
    group = session
  )
) + 
  geom_point(
    aes(color = factor(post)),
    size = 2
    ) +
  scale_x_datetime(
    limits = c(
      ymd_hms("2023-02-25 00:00:00"), 
      ymd_hms("2023-02-29 00:00:00")
      )
    ) +
  scale_y_continuous(breaks = 16:26)
p0 / p1
```

### 3. There are multiple post responses within a session
The "post" contrast is the average mood during a session. I know this is intentional, but I'm not sure if it's ideal. I think continuous time would be more interesting. The contrast averages over sessions that are too irregular, and will be hard to interpret. A model with continuous time within sessions would solve this.

```{r}
filter(dat, pid == "p10", session == 16) %>% 
  kable()

filter(dat2, pid == "p10", session == 16) %>% 
  kable()
```

### 4. The post contrast doesn't represent the whole session
This can also be seen in the data above. The last post measure in session 16 is at 2023-02-25 11:31:43, but the actual session continues for at least an hour longer, as can be seen in the full data: 2023-02-25 12:41:00. I'm not sure if this matter, but it feels weird that this isn't captured by the model. I guess this is related to concern #2, I think we need to somehow encode the "dose" of the PWS "treatment", and the planned missingness of the mood prompts.

### The post contrast could be replaced with continuous time
A model of the form

```r
mood ~ spline(time, knots) + (1 | session) + (1 | id)
```

would allow modeling change during session while still accounting for the highly irregular sessions.

There is also the option to scale time to 1 on a session level to more closely mimic a pre/post design. This would have the advantage of being able to include all wellbeing responses, regardless if a session starts with a wellbeing measure.

### Fit model
Reset duration within each session
```{r}
dat <- dat %>% 
  mutate(
    duration = duration - duration[1],
    hours = duration / 60,
    .by = c(pid, session)
  ) %>% 
  filter(prompt == "Wellbeing")
```

Fit linear change
```{r}
library(rms)
fit0 <- lmer(mood ~ hours + (1 | session) + (1 | pid), data = dat)
summary(fit0)
```

What's the distribution of session lengths
```{r}
median_ses_len <- dat %>% 
  group_by(pid, session) %>% 
  summarise(session_length = max(duration)) %>% 
  group_by(pid) %>% 
  summarise(median_len = median(session_length))

median(median_ses_len$median_len)

mean(median_ses_len$median_len > 60)

```

"Typical" session length is 24 minutes. Few participants typically play for more than an hour. Plot estimated trend over 0 to 2 hours.

```{r}
emm0 <- emmeans(
  fit0, 
  ~hours, 
  at = list(hours = seq(0,2, length.out = 100)),
  lmer.df = "asymptotic"
  )

p0 <- as.data.frame(emm0) %>% 
  ggplot(aes(hours, emmean)) + 
  geom_line() +
  geom_ribbon(aes(ymin = asymp.LCL, ymax = asymp.UCL), alpha = 0.25) +
  labs(x = "Session duration (hours)")
```

Let's try non-linear change using `rms::rcs` with 5 knots with default placement, which should place most knots during the first hour.
```{r}
attr(rcs(dat$hours, 5), "parms")
```

Fit model with only random intercepts.
```{r}
fit1 <- lmer(mood ~ rcs(hours, 5) + (1 | session) + (1 | pid), data = dat)
emm1 <- emmeans(
  fit1, 
  ~hours, 
  at = list(hours = seq(0,2, length.out = 100)),
  lmer.df = "asymptotic"
  )
p1 <- as.data.frame(emm1) %>% 
  ggplot(aes(hours, emmean)) + 
  geom_line() +
  geom_ribbon(aes(ymin = asymp.LCL, ymax = asymp.UCL), alpha = 0.25) +
  labs(x = "Session duration (hours)")
```

Adding `rcs(hours, 5)` as random effects would be messy, so let's fit the model with a linear random slope at the participant level. I don't think random slopes at the session level is needed? Moreover, I would expect that the exact random slopes specification have little impact on fixed effects in this case.
```{r}
path <- "models/lmm-fit2.Rds"
if (file.exists(path)) {
  fit2 <- read_rds(path)
} else {
  fit2 <- lmer(mood ~ rcs(hours, 5) + (1 | session) + (1 + hours | pid), data = dat)
  write_rds(fit2, path)
}
emm2 <- emmeans(
  fit2, 
  ~hours, 
  at = list(hours = seq(0,2, length.out = 100)),
  lmer.df = "asymptotic"
  )
p2 <- as.data.frame(emm2) %>% 
  ggplot(aes(hours, emmean)) + 
  geom_line() +
  geom_ribbon(aes(ymin = asymp.LCL, ymax = asymp.UCL), alpha = 0.25) +
  labs(x = "Session duration (hours)")
```

Plots models, ignore model 2 as it's basically identical to model 1.

```{r}
p0 / p1
```

I think the shape of the spline makes sense.

TODO:

- GAMM?
- Is censoring really needed?
- Visualize the % that have positive vs negative change over a session? or over 30 / 60 minutes.

### 2023-09-08 continuous-time-draft.qmd
re slopes YOLO

@@
 library(emmeans)
+library(splines)
 ```
@@
 
+### Session lengths
+
+What's the distribution of session lengths
+```{r}
+median_ses_len <- dat %>% 
+  group_by(pid, session) %>% 
+  mutate(
+    duration = duration - duration[1],
+  ) %>% 
+  summarise(session_length = max(duration)) %>% 
+  group_by(pid) %>% 
+  summarise(median_len = median(session_length))
+median(median_ses_len$median_len)
+mean(median_ses_len$median_len > 60)
+```
+
+"Typical" session length is 35 minutes. Few participants typically play for more than an hour. Plot estimated trend over 0 to 2 hours.
+
 ### Fit model
@@
   mutate(
+    ps = interaction(pid, session, sep = "_")
+  )  %>% 
+  filter(prompt == "Wellbeing") %>% 
+  mutate(
     duration = duration - duration[1],
@@
     .by = c(pid, session)
-  ) %>% 
-  filter(prompt == "Wellbeing")
+  ) 
 ```
@@
 library(rms)
-fit0 <- lmer(mood ~ hours + (1 | session) + (1 | pid), data = dat)
-summary(fit0)
+fit0 <- lmer(mood ~ hours + (1 | ps) + (1 | pid), data = dat)
 ```
-
-What's the distribution of session lengths
-```{r}
-median_ses_len <- dat %>% 
-  group_by(pid, session) %>% 
-  summarise(session_length = max(duration)) %>% 
-  group_by(pid) %>% 
-  summarise(median_len = median(session_length))
-
-median(median_ses_len$median_len)
-
-mean(median_ses_len$median_len > 60)
-
-```
-
-"Typical" session length is 24 minutes. Few participants typically play for more than an hour. Plot estimated trend over 0 to 2 hours.
 
@@
 ```{r}
-fit1 <- lmer(mood ~ rcs(hours, 5) + (1 | session) + (1 | pid), data = dat)
+fit1 <- lmer(mood ~ rcs(hours, 5) + (1 | ps) + (1 | pid), data = dat)
 emm1 <- emmeans(
@@
   )
+
+## for plotting knot placement
+get_knot_coords <- function(fit) {
+    knots <- attr(rcs(dat$hours, 5), "parms")
+    knots_x <- rcs(knots, knots)
+    b0 <- fixef(fit)[1]
+    b <- fixef(fit)[-1]
+    knots_y <- b0 + knots_x[,1] * b[1] + knots_x[,2] * b[2] + knots_x[,3] * b[3] + knots_x[,4] * b[4]
+    data.frame(
+        x = knots,
+        y = c(knots_y)
+    )
+}
 p1 <- as.data.frame(emm1) %>% 
@@
   geom_ribbon(aes(ymin = asymp.LCL, ymax = asymp.UCL), alpha = 0.25) +
+  geom_point(
+    data = get_knot_coords(fit1),
+    aes(x,y)
+  ) +
   labs(x = "Session duration (hours)")
@@
 } else {
-  fit2 <- lmer(mood ~ rcs(hours, 5) + (1 | session) + (1 + hours | pid), data = dat)
+  fit2 <- lmer(mood ~ rcs(hours, 5) + (1 | ps) + (1 + hours | pid), data = dat)
   write_rds(fit2, path)
@@
   geom_ribbon(aes(ymin = asymp.LCL, ymax = asymp.UCL), alpha = 0.25) +
+  # geom_point(
+  #   data = get_knot_coords(fit2),
+  #   aes(x,y)
+  # ) +
   labs(x = "Session duration (hours)")
@@
 ```{r}
-p0 / p1
+p0 / p1 / p2
 ```
@@
 I think the shape of the spline makes sense.
+
+## YOLO model with random slopes
+Here I add a random slope to each piece of the spline, i.e. 4 random slopes at the participant level. The model is fit on a random subset of 1000 pids.
+```{r}
+# include 1000 participants
+set.seed(1337)
+pids <- sample(unique(dat$pid), 1000)
+tmp <- filter(dat, pid %in% pids)
+# so predict won't complain
+tmp <- tmp %>% 
+    mutate(
+        pid = as.integer(pid),
+        ps = as.integer(ps)
+        )
+# same spline but different parameterization
+# usually less estimation problems
+spline <- ns(
+    tmp$hours, 
+    knots = c(0.12, 0.4, 1), 
+    Boundary.knots = c(0.016, 3)
+)
+colnames(spline) <- paste0("hours", 0:3)
+tmp <- cbind(tmp, spline)
+path <- "models/lmm-fit3.Rds"
+if (file.exists(path)) {
+  fit3 <- read_rds(path)
+} else {
+  fit3 <- lmer(
+    mood ~ hours0 + hours1 + hours2 + hours3 + (1 | ps) + (1 + hours0 + hours1 + hours2 + hours3 | pid), 
+    data = tmp
+  )
+  write_rds(fit3, path)
+}
+
+```
+
+I'm interested in if the spline changes based on the pre-session scores.
+Let's plot the fixed effects at -2, -1, 0 times the random intercept SD + the corresponding random slope values based on their correlation with the random intercept.
+
+```{r}
+get_trend <- function(fit, data, re_int = 0, re_s0 = 0, re_s1 = 0, re_s2 = 0, re_s3 = 0) {
+    hours <- seq(0,2, length.out = 100)
+    x <- ns(
+        hours, 
+        knots = c(0.12, 0.4, 1), 
+        Boundary.knots = c(0.016, 3)
+    )
+    b0 <- fixef(fit)[1] + re_int
+    b <- fixef(fit)[-1] 
+    y <- b0 + x[,1] * (b[1] + re_s0) + x[,2] * (b[2] + re_s1) + x[,3] * (b[3] + re_s2) + x[,4] * (b[4] + re_s3)
+    data.frame(
+        x = hours,
+        y = c(y),
+        re_int
+    )
+}
+re <- as.data.frame(VarCorr(fit3))
+pid_int <- re[2, "sdcor"]
+pid_slope0 <- re[3, "sdcor"]
+pid_slope1 <- re[4, "sdcor"]
+pid_slope2 <- re[5, "sdcor"]
+pid_slope3 <- re[6, "sdcor"]
+pid_cor0 <- re[7, "sdcor"]
+pid_cor1 <- re[8, "sdcor"]
+pid_cor2 <- re[9, "sdcor"]
+pid_cor3 <- re[10, "sdcor"]
+sd1 <- 1
+sd2 <- 2
+rbind(
+    get_trend(fit3, data = tmp),
+    get_trend(
+        fit3, 
+        data = tmp,
+        -sd1 * pid_int,
+        -sd1 * pid_cor0 * pid_slope0, 
+        -sd1 * pid_cor1 * pid_slope1, 
+        -sd1 * pid_cor2 * pid_slope2, 
+        -sd1 * pid_cor3 * pid_slope3
+    ),
+    get_trend(
+        fit3, 
+        data = tmp,
+        -sd2 * pid_int,
+        -sd2 * pid_cor0 * pid_slope0, 
+        -sd2 * pid_cor1 * pid_slope1, 
+        -sd2 * pid_cor2 * pid_slope2, 
+        -sd2 * pid_cor3 * pid_slope3
+        )
+    ) %>% 
+  ggplot(
+    aes(
+        x, 
+        y, 
+        color = factor(re_int), 
+        group = factor(re_int)
+        )
+    ) + 
+  geom_line() +
+  labs(x = "Session duration (hours)")
+```
+
+The random intercepts are only weakly correlated with each piece of the cubic spline, so the shape doesn't change much.
+
+Let's also predict some individual slopes.
+
+```{r}
+pred <- filter(tmp, pid %in% sample(unique(tmp$pid), 10))
+pred$yhat <- predict(fit3, newdata = pred)
+ggplot(
+    data = pred,
+    aes(
+        x = hours,
+        y = yhat,
+        group = ps
+    )
+) + 
+    geom_line(
+        color = "blue", 
+        linewidth = 1
+    ) +
+    geom_point(
+        aes(y = mood)
+    ) +
+    facet_wrap(~pid)
+```
+
+Zoom in on pid 9263.
+```{r}
+ggplot(
+    data = filter(pred, pid == 9263),
+    aes(
+        x = hours,
+        y = yhat,
+        group = ps
+    )
+) + 
+    geom_line(
+        aes(y = mood),
+        linetype = "dashed",
+        alpha = 0.33
+    ) +
+    geom_line(
+        color = "blue", 
+        linewidth = 2
+        ) +
+    geom_point(
+        aes(y = mood)
+    ) +
+    facet_wrap(~session)
+```
+
+Probably good enough, even tho the session effect is only modelled using a random intercept.
+
+### Predict change
+Another YOLO approach to estimate individual change during a session is to to look at the model predictions.
+
+TODO: should be based on the actual session length, not only the range of WB prompts.
+
+```{r}
+pred <- fit3@frame
+pred$yhat <- predict(fit3)
+pred %>% 
+  # compare the last and first value
+  dplyr::summarize(change = last(yhat) - yhat[1], .by = c(pid, ps)) %>%
+  # average over sessions within participants
+  dplyr::summarize(change = mean(change), .by = pid) %>% 
+  # average over participants
+  dplyr::summarize(mean(change), sd(change))
+```
+
+We can also plot the distribution of the participants average change during sessions.
+```{r}
+pred %>% 
+  dplyr::summarize(change = last(yhat) - yhat[1], .by = c(pid, ps)) %>%
+  dplyr::summarize(change = mean(change), .by = pid) %>% 
+  ggplot(aes(change)) + geom_histogram()
+```
+
+They really don't change much. Let's look at a participant out in the tails.
+```{r}
+pred %>% 
+  dplyr::summarize(change = last(yhat) - yhat[1], .by = c(pid, ps))  %>% 
+  filter(change > 0.1) %>% head
+```
+
+Zoom in on pid 23.
+```{r}
+pred2 <- filter(tmp, pid == 23)
+pred2$yhat <- predict(fit3, newdata = pred2)
+ggplot(
+    data = pred2,
+    aes(
+        x = hours,
+        y = yhat,
+        group = ps
+    )
+) + 
+    geom_line(
+        aes(y = mood),
+        linetype = "dashed",
+        alpha = 0.33
+    ) +
+    geom_line(
+        color = "blue", 
+        linewidth = 2
+        ) +
+    geom_point(
+        aes(y = mood)
+    ) +
+    facet_wrap(~session)
+```
+
+This look kinda reasonable?
 

### 2023-09-08 ms.qmd
add env variables

@@
 dir.create("models", FALSE)
+N_CORES <- as.numeric(Sys.getenv("N_CORES", unset = 1))
+N_THREADS <- as.numeric(Sys.getenv("N_THREADS", unset = 2))
+N_ITER <- as.numeric(Sys.getenv("N_THREADS", unset = 2000))
+N_SUBSET <- as.numeric(Sys.getenv("N_SUBSET"))
 options(
-  mc.cores = 8,
+  mc.cores = N_CORES,
   brms.backend = "cmdstanr"
@@
 
+if(!is.na(N_SUBSET)) {
+  pid_subset <- sample(unique(dat$pid), N_SUBSET)
+  dat <- filter(dat, pid %in% pid_subset)
+}
+
 # Count of data to be used in these analyses
@@
   silent = 0,
-  threads = 2,
-  iter = 2000,
+  threads = N_THREADS,
+  iter = N_ITER,
   control = list(adapt_delta = .95),

### 2023-09-10 docker-quarto.sh
add dockerfile
includes renv, quarto, and cmdstan.
runtime env is still TODO

@@ -1,0 +1,5 @@
+#!/bin/bash
+quarto render
+rm -rf output/*
+mv -f docs/* output
+echo "Done. Files available on host at ./docker/"

### 2023-09-11 ms.qmd
fix subset bug
avoid trying to sample when there's not enough data

typo

@@
   mutate(session = as.numeric(as.factor(session)), .by = pid) |>
-  mutate(session_per_person = length(unique(session)), .by = pid) |>
+  mutate(session_per_person = length(unique(session)), .by = pid)
+# default 
+min_sessions <- 4 
+n_filtered <- dat_example_sessions |> 
+  distinct(pid, session_per_person) |> 
   filter(session_per_person >= 4) |> 
+  nrow()
+# pick anyone if there's not 3 pids with >= 4 sessions
+if(n_filtered < 3) min_sessions <- 0
+dat_example_sessions <- dat_example_sessions |>
+ filter(session_per_person >= min_sessions) |>
   arrange(pid, session) |>

### 2023-09-11 docker-quarto.sh
set output dir

@@ -1,5 +1,5 @@
 #!/bin/bash
-quarto render
+quarto render --output-dir=docs
 rm -rf output/*
 mv -f docs/* output
 echo "Done. Files available on host at ./docker/"

### 2023-09-11 ms.qmd
Figure and heterogeneity

--- 2023-09-06 ms.qmd
@@
 
-The model results confirmed the visual impressions described above: @tbl-avg indicates, with great certainty, that mood increased by about [`r paste0(map(h_avg_$hypothesis[3,4:5], ~number(.x, .001)), collapse = ", ")`] units.
+The model results confirmed the visual impressions described above: @tbl-avg indicates, with great certainty, that mood increased by about [`r paste0(map(h_avg_$hypothesis[3,4:5], ~number(.x, .001)), collapse = ", ")`] units (@fig-coef, "Raw").
 
@@
 To further interpret the magnitude of the difference in moods between pre- and during play, we compared the contrast to the natural variations in mood as estimated by the model (standard deviation of the between-person intercept heterogeneity ($\tau_{\alpha_0}$ = `r posterior_pop_sum$x[1]`), between-person pre-during difference heterogeneity ($\tau_{\alpha_1}$ = `r posterior_pop_sum$x[2]`) and to the residual standard deviation ($\sigma$ = `r posterior_pop_sum$x[3]`).
-The standardized (with the total residual variation from the three above components) pre-play -- during play contrast was `r posterior_pop_sum$x[6]`, a not ignorable difference.
+The standardized (with the total residual variation from the three above components) pre-play -- during play contrast was `r posterior_pop_sum$x[6]` (@fig-coef, "Scaled"), a not ignorable difference.
+
+```{r}
+#| label: heterogeneity-calculate
+
+posterior_pid <- coef(fit, summary = FALSE)$pid[,,"post1"]
+posterior_pid_scaled <- apply(
+  posterior_pid, 2, 
+  function(x) x / posterior_pop$total_residual
+) |> 
+  apply(2, function(x) mean_qi(x)) |> 
+  bind_rows(.id = "pid")
+posterior_pid <- posterior_pid |> 
+  apply(2, function(x) mean_qi(x)) |> 
+  bind_rows(.id = "pid")
+
+posterior_pid_sum <- bind_rows(
+  "b_post1" = posterior_pid,
+  "b_post1_scaled" = posterior_pid_scaled,
+  .id = "name"
+)
+
+x <- posterior_pop |> 
+  mutate(
+    prop_pos = pnorm(0, b_post1, sd_pid__post1, lower.tail = FALSE)
+  ) |> 
+  mean_qi(prop_pos) |> 
+  mutate(x = str_glue("{percent(prop_pos, .1)} [{number(.lower*100, .1)}, {number(.upper*100, .1)}]")) |> 
+  pull(x)
+```
+
+We then turned our attention to between-person differences in the difference in during and pre-play mood.
+With a sample of `r s$n_sub`, it is not surprising that the population-level contrast is credibly different from zero.
+We therefore asked the subsequent question: What proportion of the population from which the sample was drawn would we expect there to be an elevation in mood during play?
+The boxplots in @fig-coef show the distribution of player-specific contrasts estimated by the model, which indicated that `r x` of players in the population experience a positive shift in mood during play, compared to their moods during login.
+
+```{r}
+#| label: fig-coef
+#| include: true
+#| fig.width: 3
+#| fig.height: 1.4
+#| fig.cap: "Difference in mood during play compared to mood at session start.
+Points and intervals indicate posterior means and 95%CIs of the population level difference; boxplots summarize the player-level estimates.
+Raw estimates are on the VAS scale, whereas Scaled estimates indicate contrasts divided by the total residual deviance of the model."
+
+posterior_pop |> 
+  select(starts_with("b_")) |> 
+  pivot_longer(everything()) |>
+  group_by(name) |> 
+  mean_qi(value, .width = 0.95) |> 
+  ggplot(aes(value, name)) +
+  scale_x_continuous(
+    "During play - pre-play contrast",
+    breaks = extended_breaks(7)
+  ) +
+  scale_y_discrete(
+    breaks = c("b_post1", "b_post1_scaled"),
+    labels = c("Raw", "Scaled")
+  ) +
+  coord_cartesian(xlim = c(-0.25, 0.65)) +
+  geom_vline(xintercept = 0, linewidth = .33, col = "dodgerblue") +
+  geom_boxploth(
+    data = posterior_pid_sum,
+    aes(x = y, group = name), 
+    outlier.shape = NA,
+    notch = TRUE,
+    size = 0.25,
+    width = .2
+  ) +
+  geom_pointrangeh(
+    aes(xmin = .lower, xmax = .upper),
+    fatten = 2.5
+  ) +
+  theme(
+    axis.title.y = element_blank()
+  )
+```
 

### 2023-09-13 continuous-time-draft.qmd
WIP GAM models

@@
   ) 
+write_rds(dat, "data/ct-dev-data.rds")
 ```
@@
 ```{r}
+#| code-fold: show
 library(rms)
@@
 ```{r}
+#| code-fold: show
 fit1 <- lmer(mood ~ rcs(hours, 5) + (1 | ps) + (1 | pid), data = dat)
+```
+
+```{r}
 emm1 <- emmeans(
@@
 Moreover, I would expect that the exact random slopes specification have little impact on fixed effects in this case.
+
 ```{r}
+#| code-fold: show
 path <- "models/lmm-fit2.Rds"
@@
 }
+```
+
+```{r}
 emm2 <- emmeans(
@@
 ```{r}
+#| cold-fold: show
+
 # include 1000 participants
@@
 
-This look kinda reasonable?
+This looks kinda reasonable?
+
+## brms
+Ignore for now.
+
+```{r}
+#| code-fold: show
+library(brms)
+options(
+  mc.cores = 4,
+  brms.backend = "cmdstanr"
+)
+fit_b0 <- brm(
+      mood ~ hours0 + hours1 + hours2 + hours3 + (1 | ps) + (1 | pid),
+      data = tmp,
+      cores = 4,
+      file = "models/brm-pre-post-ct0"
+)
+pp_check(fit_b0)
+```
+
+Censoring
+```{r}
+#| code-fold: show
+tmp <- tmp |> 
+  mutate(
+    cl = case_when(
+      mood == 0 ~ "left",
+      mood == 1 ~ "right",
+      TRUE ~ "none"
+    )
+  )
+fit_b1 <- brm(
+      mood | cens(cl) ~ hours0 + hours1 + hours2 + hours3 + (1 | ps) + (1 | pid),
+      data = tmp,
+      threads = 2,
+      iter = 1000,
+      file = "models/brm-pre-post-ct1"
+)
+pp_check(fit_b1)
+```
+
+## GAM(M)
+
+Let's try some GAMs. 
+```{r}
+library(mgcv)
+knots <- list(hours = c(0.016, 0.12, 0.4, 1, 3))
+tmp <- tmp |> 
+  mutate(
+    pid = as.factor(pid),
+    ps = as.factor(ps)
+  )
+```
+
+We'll start with a cubic spline using the same knot placement as used with the LMMs.
+```{r}
+#| code-fold: show
+fit_g_0 <- gam(
+    mood ~ s(hours, bs = "cr", k = 5), 
+    knots = knots,
+    method = "REML",
+    data = tmp,
+)
+```
+
+```{r}
+plot_gam <- function(fit, t_end = 2) {
+  emm0 <- emmeans(
+  fit, 
+  ~hours, 
+  at = list(hours = seq(0, t_end, length.out = 100))
+  )
+
+  p0 <- as.data.frame(emm0) %>% 
+    ggplot(aes(hours, emmean)) + 
+    geom_line() +
+    geom_ribbon(aes(ymin = lower.CL, ymax = upper.CL), alpha = 0.25) +
+    labs(x = "Session duration (hours)")
+  p0
+}
+plot_gam(fit_g_0)
+```
+
+we could add a random intercept for pid and session like this
+```{r}
+#| eval: false
+#| code-fold: show
+fit_g_1 <- gam(
+    mood ~ s(hours, bs = "cr", k = 5) + s(pid, bs = "re") + s(ps, bs = "re"), 
+    knots = knots,
+    data = tmp,
+)
+```
+
+The call above is super slow, so let's use `gamm4` instead.
+
+```{r}
+#| code-fold: show
+library(gamm4)
+fit_g_1_lme4 <- gamm4(
+    mood ~ s(hours, bs = "cr", k = 5), 
+    knots = knots,
+    data = tmp, 
+    random = ~ (1 | pid) + (1 | ps)
+)
+```
+
+```{r}
+pred <- data.frame(hours = seq(0,2, length.out = 100))
+pred$yhat <- predict(fit_g_1_lme4$gam, newdata = pred)
+ggplot(
+  pred, 
+  aes(hours, yhat)
+  ) + 
+    geom_line() +
+    #geom_ribbon(aes(ymin = lower.CL, ymax = upper.CL), alpha = 0.25) +
+    labs(x = "Session duration (hours)")
+```
+
+We are using the same smooth for everyone here, which is probably not what we want.
+
+### Separate smooth per participant
+Let's try using a separate smooth per pid. We only fit the model to a subset of participants to save time.
+
+```{r}
+#| code-fold: show
+set.seed(6666)
+tmp2 <- filter(tmp, pid %in% sample(unique(pid), 50)) |> 
+  filter(!is.na(mood))
+fit_g_by_lme4 <- gamm4(
+    mood ~ s(hours, by = pid, bs = "cr", k = 5), 
+    knots = knots,
+    data = tmp2, 
+    random = ~ (1 | pid) + (1 | ps)
+)
+```
+
+N.B. that only the GAM part is plotted, I'm not sure if there is a command that automagically includes the random effects when using `gamm4`.
+
+```{r}
+pred <- expand.grid(
+  hours = seq(0, 2, length.out = 100),
+  pid = unique(tmp2$pid)
+ )
+pred$yhat <- predict(fit_g_by_lme4$gam, newdata = pred)
+pred$model <- "by"
+ggplot(
+  pred, 
+  aes(
+    hours, 
+    yhat,
+    group = pid,
+    color = pid
+    )
+  ) + 
+    geom_line() +
+    #geom_ribbon(aes(ymin = lower.CL, ymax = upper.CL), alpha = 0.25) +
+    labs(x = "Session duration (hours)") +
+    theme(legend.position = "none")
+
+plot_pids <- sample(unique(tmp2$pid), 10)
+tmp2 <- filter(tmp2, !is.na(mood))
+tmp2$yhat <- predict(fit_g_by_lme4$gam)
+tmp2$model <- "by"
+ggplot(
+  filter(tmp2, pid %in% plot_pids), 
+  aes(
+    hours, 
+    yhat,
+    group = ps,
+    color = pid
+    )
+  ) + 
+    geom_line(aes(y = mood)) +
+    geom_line(color = "blue", linewidth = 1) +
+    geom_point(aes(y = mood)) +
+    #geom_ribbon(aes(ymin = lower.CL, ymax = upper.CL), alpha = 0.25) +
+    labs(x = "Session duration (hours)") +
+    facet_wrap(~pid) +
+    theme(legend.position = "none")
+
+pred_all <- filter(tmp2, pid %in% plot_pids)
+```
+
+We're fitting a separate fixed and random effects for each pid. Fitting this to the full data will be super slow?
+```{r}
+summary(fit_g_by_lme4$mer)
+```
+
+### Factor-smooth interaction 
+We can also try a factor-smooth interaction instead, this should be closer to what we want.
+
+```{r}
+#| code-fold: show
+fit_g_fs_lme4 <- gamm4(
+    mood ~ s(hours, pid, bs = "fs", xt = list(bs = "cr"), k = 5, fx = FALSE),, 
+    knots = knots,
+    data = tmp2, 
+    random = ~ (1 | pid) + (1 | ps)
+)
+```
+
+`gamm4` is not supported by `emmeans`, and I'm not sure how to plot the "marginal" smooth.
+
+```{r}
+pred <- expand.grid(
+  hours = seq(0, 2, length.out = 100),
+  pid = unique(tmp2$pid)
+ )
+pred$yhat <- predict(fit_g_fs_lme4$gam, newdata = pred)
+pred$model <- "fs"
+ggplot(
+  pred, 
+  aes(
+    hours, 
+    yhat,
+    group = pid,
+    color = pid
+    )
+  ) + 
+    geom_line() +
+    #geom_ribbon(aes(ymin = lower.CL, ymax = upper.CL), alpha = 0.25) +
+    labs(x = "Session duration (hours)") +
+    theme(legend.position = "none")
+
+tmp2$yhat <- predict(fit_g_fs_lme4$gam)
+tmp2$model <- "fs"
+ggplot(
+  filter(tmp2, pid %in% plot_pids), 
+  aes(
+    hours, 
+    yhat,
+    group = ps,
+    color = pid
+    )
+  ) + 
+    geom_line(aes(y = mood)) +
+    geom_line(color = "blue", linewidth = 1) +
+    geom_point(aes(y = mood)) +
+    lims(y = c(0, 1)) +
+    #geom_ribbon(aes(ymin = lower.CL, ymax = upper.CL), alpha = 0.25) +
+    labs(x = "Session duration (hours)") +
+    facet_wrap(~pid) +
+    theme(legend.position = "none")
+pred_all <- rbind(pred_all, filter(tmp2, pid %in% plot_pids))
+```
+
+The smooths in the plot above are all linear.
+Although, it looks pretty reasonable compared to the observed data.
+(TODO: investigate with more data.) If we fit a simple smooth to the data it's also linear.
+
+```{r}
+#| code-fold: show
+gam(
+    mood ~ s(hours, bs = "cr", k = 5), 
+    knots = knots,
+    method = "REML",
+    data = tmp2,
+) |> 
+  plot_gam()
+
+```
+
+### LMM again
+As a comparisons, let's refit the LMM with a cubic spline to the subset of data used here. And then we predict without including the random intercepts.
+
+```{r}
+#| code-fold: show
+fit3 <- lmer(
+  mood ~ hours0 + hours1 + hours2 + hours3 + (1 | ps) + (1 + hours0 + hours1 + hours2 + hours3 | pid), 
+  data = tmp2
+)
+```
+
+```{r}
+tmp2$yhat <- predict(
+  fit3, 
+  re.form = ~(0 + hours0 + hours1 + hours2 + hours3 | pid)
+)
+tmp2$model <- "lmer"
+ggplot(
+  filter(tmp2, pid %in% plot_pids), 
+  aes(
+    hours, 
+    yhat,
+    group = ps,
+    color = pid
+    )
+  ) + 
+    geom_line(aes(y = mood)) +
+    geom_line(color = "blue", linewidth = 1) +
+    geom_point(aes(y = mood)) +
+    lims(y = c(0, 1)) +
+    #geom_ribbon(aes(ymin = lower.CL, ymax = upper.CL), alpha = 0.25) +
+    labs(x = "Session duration (hours)") +
+    facet_wrap(~pid)
+
+pred_all <- rbind(pred_all, filter(tmp2, pid %in% plot_pids))
+```
+
+### Combined plot
+Plot the predictions from the 3 models in a combined figure.
+
+```{r}
+#| fig.height: 5
+#| fig.width: 7
+ggplot(
+  pred_all, 
+  aes(
+    hours, 
+    yhat,
+    group = interaction(ps, model),
+    color = model
+    )
+  ) + 
+    geom_line(data = filter(pred_all, model == "fs"), aes(y = mood, color = NA), alpha = 0.5) +
+    geom_point(data = filter(pred_all, model == "fs"), aes(y = mood, color = NA), alpha = 0.5) +
+    geom_line(linewidth = 1.5) +
+    #lims(y = c(0, 1)) +
+    #geom_ribbon(aes(ymin = lower.CL, ymax = upper.CL), alpha = 0.25) +
+    labs(x = "Session duration (hours)") +
+    facet_wrap(~pid, scales = "free_y")
+```
+
+### GAM summary
+Assuming that we wanted to fit these smooths in `brms` to model censoring, then
+
+- separate smooths using `by` is not feasible?
+- factor-smooth interaction is maybe feasible using the full data.
+  - Might be overpenalizing the smooths as they were all straight lines? 
+- the LMM is basically in between the model above when it come to "wiggliness", as would be expected
+
+I'm not sure if some if the issues is caused by how GAMs are affected by the highly irregular sessions + missing data.
+
+LMM might be _the_ model?
 

### 2023-09-14 ms.qmd
use correct env var

@@
 N_THREADS <- as.numeric(Sys.getenv("N_THREADS", unset = 2))
-N_ITER <- as.numeric(Sys.getenv("N_THREADS", unset = 2000))
+N_ITER <- as.numeric(Sys.getenv("N_ITER", unset = 2000))
 N_SUBSET <- as.numeric(Sys.getenv("N_SUBSET"))

### 2023-09-19 analyze.qmd
Separate scripts to Quarto project

--- 2023-09-11 ms.qmd
@@
 ---
-title: "Mood increases during video game play: A naturalistic observational case study of PowerWash Simulator"
-description: Computational notebook supporting our manuscript
-author:
-  - name: Matti Vuorre
-    orcid: 0000-0001-5052-066X
-    email: mjvuorre@uvt.nl
-    url: https://vuorre.netlify.app
-    affiliation:
-      - ref: a
-      - ref: b
-  - name: Nick Ballou
-    orcid: 0000-0003-4126-0696
-    affiliation:
-      - ref: b
-  - name: Andrew K. Przybylski
-    orcid: 0000-0001-5547-2185
-    affiliation:
-      - ref: b
-affiliations:
-  - id: a
-    name: Tilburg University
-  - id: b
-    name: University of Oxford
-
-date: 2023-09-04
-bibliography: bibliography.bib
-format:
-  pdf: default
-  docx: default
-  html: default
+title: Analyze
 ---
@@
   cache = TRUE,
-  include = FALSE,
-  echo = FALSE
+  include = TRUE,
+  echo = TRUE
 )
@@
 library(tidyverse)
+source("functions.R")
 ```
@@
 ```{r}
-#| label: functions
-
-# Format big numbers with a comma
-number2 <- function(x, ...) {
-  number(x, ..., big.mark = ",")
-}
-
-# Replace extreme percentages
-percent2 <- function(x, accuracy = .1) {
-  x <- percent(x, accuracy = accuracy)
-  x <- if_else(x == "100.0%", ">99.9%", x)
-  x <- if_else(x == "0.0%", "<0.1%", x)
-  x
-}
-```
-
-```{r}
-#| label: data-download
-
-# Download, unzip data from OSF PWS database
-path <- "data-raw/data.zip"
-if (!file.exists(path)) {
-  dir.create("data-raw", FALSE)
-  download.file(
-    url = "https://osf.io/download/j48qf/",
-    destfile = "data-raw/data.zip"
-  )
-  dir.create("data", FALSE)
-  unzip(
-    zipfile = path,
-    files = c("data/demographics.csv", "data/study_prompt_answered.csv")
-  )
-}
-```
-
-```{r}
-#| label: data-load
-
-dat <- read_csv(
-  "data/study_prompt_answered.csv",
-  col_select = c(
-    pid,
-    time = Time_utc,
-    duration = CurrentSessionLength,
-    prompt = LastStudyPromptType,
-    mood = response
-  )
-) |>
-  mutate(
-    pid = factor(pid),
-    mood = mood / 1000
-  )
-
-# Calculate raw data summaries
-sraw <- list()
-sraw$n_res <- dat |>
-  count(prompt)
-sraw$n_res_total <- number2(sum(sraw$n_res$n))
-sraw$n_res <- sraw$n_res |>
-  mutate(n = number2(n)) |>
-  pivot_wider(names_from = prompt, values_from = n)
-sraw$n_sub <- number2(length(unique(dat$pid)))
-
-# Create sessions using heuristics below
-dat <- dat |>
-  arrange(pid, time) |>
-  mutate(
-    i = row_number(),
-    # Observation is in a new session if this...
-    new_session =
-      # is the first total observation,
-      i == 1 |
-      # session duration is zero,
-      (duration == 0) |
-      # session is shorter than previous session (ticker has reset),
-      (duration < lag(duration, default = 0)) |
-      # or if more than 30 minutes since previous observation.
-      (time > lag(time) + minutes(30)),
-    # Session number is the cumulative sum of new sessions...
-    session = cumsum(new_session),
-    # ... per person
-    .by = pid
-  ) |>
-  select(-c(i, new_session))
-
-# Filter data:
-# Include only wellbeing responses that occurred
-# in sessions with both "pre" and "post" measures
-dat <- dat |>
-  filter(prompt == "Wellbeing") |>
-  select(-prompt)
-
-# Count of data with all wellbeing observations
-dat_total <- dat |>
-  summarise(
-    n_total = n(),
-    n_players = length(unique(pid))
-  )
-
-# Include only sessions with "pre-" and "post" measure(s)
-dat <- dat |>
-  filter(
-    # Session has a wellbeing measure at time = 0
-    duration[1] == 0,
-    # Session has at least two non-missing wellbeing measures 
-    # (e.g. pre + 1 post)
-    sum(!is.na(mood)) > 1,
-    .by = c(pid, session)
-  )
-
-if(!is.na(N_SUBSET)) {
+dat <- readRDS("data/data.rds")
+s <- readRDS("data/s.rds")
+sraw <- readRDS("data/sraw.rds")
+if (!is.na(N_SUBSET)) {
   pid_subset <- sample(unique(dat$pid), N_SUBSET)
@@
 }
-
-# Count of data to be used in these analyses
-dat_filtered <- dat |>
-  summarise(
-    n_total = n(),
-    n_players = length(unique(pid))
-  )
-
-# Summary
-bind_rows(
-  "total" = dat_total,
-  "filtered" = dat_filtered,
-  .id = "Data"
-)
-
-# Summaries of filtered data
-s <- list()
-s$n_sub <- number2(length(unique(dat$pid)))
-s$n_obs <- number2(nrow(dat))
-s$n_obs_missing <- number2(nrow(filter(dat, is.na(mood))))
-s$n_ses <- number2(nrow(count(dat, pid, session)))
 ```
@@
 
-library(posterior)
 as_draws(fit, variable = c("b_", "sd_", "cor_"), regex = TRUE) |> 

### 2023-09-19 wrangle.qmd
Separate scripts to Quarto project

--- 2023-09-11 ms.qmd
@@
 ---
-title: "Mood increases during video game play: A naturalistic observational case study of PowerWash Simulator"
-description: Computational notebook supporting our manuscript
-author:
-  - name: Matti Vuorre
-    orcid: 0000-0001-5052-066X
-    email: mjvuorre@uvt.nl
-    url: https://vuorre.netlify.app
-    affiliation:
-      - ref: a
-      - ref: b
-  - name: Nick Ballou
-    orcid: 0000-0003-4126-0696
-    affiliation:
-      - ref: b
-  - name: Andrew K. Przybylski
-    orcid: 0000-0001-5547-2185
-    affiliation:
-      - ref: b
-affiliations:
-  - id: a
-    name: Tilburg University
-  - id: b
-    name: University of Oxford
+title: Wrangle
+---
 
-date: 2023-09-04
-bibliography: bibliography.bib
-format:
-  pdf: default
-  docx: default
-  html: default
----
+This notebook wrangles the raw data on OSF to something more suitable for our current analyses.
 
@@
 
-library(scales)
-library(cmdstanr)
-library(posterior)
-library(brms)
-library(ggstance)
-library(ggdist)
-library(patchwork)
 library(tidyverse)
-```
-
-```{r}
-#| label: options
-#| cache: false
-
-dir.create("models", FALSE)
-options(
-  mc.cores = 8,
-  brms.backend = "cmdstanr"
-)
-theme_set(
-  theme_linedraw(base_size = 10) +
-    theme(
-      strip.background = element_rect(fill = NA, color = NA),
-      strip.text = element_text(colour = "black", hjust = 0),
-      axis.text = element_text(size = rel(0.75)),
-      panel.grid = element_blank()
-    )
-)
-```
-
-```{r}
-#| label: functions
-
-# Format big numbers with a comma
-number2 <- function(x, ...) {
-  number(x, ..., big.mark = ",")
-}
-
-# Replace extreme percentages
-percent2 <- function(x, accuracy = .1) {
-  x <- percent(x, accuracy = accuracy)
-  x <- if_else(x == "100.0%", ">99.9%", x)
-  x <- if_else(x == "0.0%", "<0.1%", x)
-  x
-}
+source("functions.R")
 ```
@@
 
-# Methods
-
-## Participants
-
 ```{r}
-#| label: demographics-load
-
-demo <- read_csv(
-  "data/demographics.csv",
-  col_select = c(pid, country, gender, age)
-) |>
-  filter(pid %in% unique(dat$pid))
-
-a <- quantile(demo$age, probs = c(.5, .1, .9), na.rm = TRUE)
-a <- str_glue("{a[1]} ({a[2]}, {a[3]})")
-g <- count(demo, gender, sort = TRUE) |>
-  mutate(p = n / sum(n)) |>
-  mutate(x = str_glue("{gender} ({number2(n)}, {percent2(p)})"))
-n_countries <- length(unique(demo$country[!is.na(demo$country)]))
-countries <- count(demo, country, sort = TRUE) |>
-  mutate(p = n / sum(n)) |>
-  mutate(x = str_glue("{country} ({number2(n)}, {percent2(p)})"))
+dir.create("data", FALSE)
+saveRDS(dat, "data/data.rds")
+saveRDS(s, "data/s.rds")
+saveRDS(sraw, "data/sraw.rds")
 ```
-
-The PowerWash Simulator case study database contains `r sraw$n_res_total` psychological item responses from `r sraw$n_sub` individuals.
-In this study, we only studied subjective well-being (mood), so did not consider the other five types of item responses in the database (enjoyment, focus, autonomy, competence, and immersion).
-Therefore, here we used data from `r s$n_sub` participants who provided `r s$n_obs` mood responses (`r s$n_obs_missing` with missing values, which we ignored) from `r s$n_ses` gameplay sessions.
-
-These participants median age (1st & 9th deciles) was `r a`, and four most frequent responses to gender were `r g$x[1:4]`.
-They hailed from `r n_countries`, with four most represented countries being `r countries$x[1:4]`.
-
-The study procedures were granted ethical approval by Oxford University’s Central University Research Ethics Committee (SSH_OII_CIA_21_011).
-All participants provided informed consent to participate and reported being 18 years or older.
-
-## Data analysis
-
-```{r}
-#| label: data-wrangle-for-analysis
-
-dat <- dat |>
-  mutate(
-    post = factor(
-      row_number() > 1,
-      levels = c(FALSE, TRUE),
-      labels = c("0", "1")
-    ),
-    .by = c(pid, session)
-  ) |>
-  mutate(ps = interaction(pid, session, sep = "_")) |>
-  mutate(
-    cl = case_when(
-      mood == 0 ~ "left",
-      mood == 1 ~ "right",
-      TRUE ~ "none"
-    )
-  )
-```
-
-Our research question concerned the difference between players’ mood at the beginning of each session (pre-play) and during the subsequent play session (during play).
-It is important to understand that this contrast does not embed a causal hypothesis: Players could begin (and end) their play ses-sions for whatever reason, and these reasons are likely to confound the pre -- during contrast.
-For example, a player might come home after a stressful day at work, and sit down to play some PWS.
-During this play session, their mood would improve because they are at home vs.
-stressful work environment, and would show as a greater during play mood compared to pre-play mood: Reasons behind starting a play session are likely to contribute to the pre -- during contrast and we are unable to disentangle those from any changes actually caused by play.
-
-We estimated this contrast within a censored three-level hierarchical regression model.
-We specified a censored gaussian model because mood was measured with a VAS.
-Ignoring the censoring inherent in VAS responses can confound changes in the mean with changes in the spread of mood, and leave the contrast susceptible to ceiling or floor effects.
-We then modelled mean mood on an intercept and contrast coded time (pre-play: -0.5; during play: 0.5), and allowed both parameters to vary randomly across players.
-In addition, we modelled random intercepts over player-sessions because over time, mood is susceptible to myriad other influences.
-Although equal variances across people in such naturalistic observation seem unlikely, we fixed this parameter to a constant for model identifiability and convergence.
-
-More formally, we modelled each mood response *i* in 1 to `r s$n_obs` from participant *j* in 1 to `r s$n_sub` and participant-session *k* in 1 to `r s$n_ses` with a normal distribution censored at the VAS limits [0, 1].
-We then modelled the distribution's mean ($\eta$) on a population-level intercept ($\beta_0$) and contrast-coded measurement timepoint ($\beta_1$), their random per-player deviations $\alpha$, and a random intercept for player-sessions $\gamma_0$:
-
-\begin{align*}
-\text{mood}_{ijk} &\sim \text{CensNormal}(\eta_{ijk}, \sigma^2, 0, 1), \\
-\eta_{ijk} &= \alpha_0 + \beta_{0j} +
-  (\alpha_1 + \beta_{1j})\text{post}_{ijk} +
-  \gamma_{0k}, \\
-\begin{bmatrix} \alpha_0 \\ \alpha_1 \end{bmatrix} &\sim 
-  \text{MVN}\left(\begin{bmatrix} 0 \\ 0 \end{bmatrix}, 
-  \begin{bmatrix} 
-    \tau_{\alpha_0} &\rho_{{\alpha_0}{\alpha_1}} \\ 
-    \rho_{{\alpha_0}{\alpha_1}} &\tau_{\alpha_1}
-  \end{bmatrix}\right), \\
-  \gamma_0 &\sim \text{Normal}(0, \tau_{\gamma_0}).
-\end{align*}
-
-```{r}
-#| label: model-estimate
-#| cache: false
-
-contrasts(dat$post) <- c(-0.5, 0.5)
-
-model <- bf(mood | cens(cl) ~ 1 + post + (1 + post | pid) + (1 | ps)) +
-  gaussian()
-
-fit <- brm(
-  model,
-  data = dat,
-  silent = 0,
-  threads = 2,
-  iter = 2000,
-  control = list(adapt_delta = .95),
-  file = "models/brm-pre-post-pid-ps-censored"
-)
-```
-
-```{r}
-#| label: model-convergence
-#| eval: false
-
-library(posterior)
-as_draws(fit, variable = c("b_", "sd_", "cor_"), regex = TRUE) |> 
-  summarise_draws(rhat)
-pairs(fit$fit, pars = variables(fit)[1:7])
-```
-
-We analyzed the data with R [@rcoreteamLanguageEnvironmentStatistical2023] and used the brms package to estimate, via Stan's HMC sampling procedures, and post-process the models [@burknerBrmsPackageBayesian2017; @standevelopmentteamStanModelingLanguage2022].
-We drew `r number2(ndraws(fit))` samples from the posterior distribution using brms' default prior distributions on all parameters, and used numerical and graphical checks to ensure model convergence and adequacy.
-
-# Results
-
-```{r}
-#| label: fig-descriptives
-#| include: false
-#| fig-height: 2.2
-#| fig-width: 7
-#| fig-cap: '**A**. Histogram of session durations (note log10 x-axis). **B**. Summary of how many sessions each participant completed. **C**. Histogram of all mood ratings.'
-
-p0 <- dat |>
-  ggplot() +
-  scale_x_continuous(
-    expand = expansion(c(0.01, 0.01)),
-  ) +
-  scale_y_continuous(
-    expand = expansion(c(0.001, 0.1)),
-  ) +
-  geom_histogram(
-    col = "white",
-    linewidth = .25
-  )
-
-dat_session_durations <- dat |>
-  filter(duration == max(duration), .by = c(pid, session))
-
-dat_session_counts <- dat |>
-  distinct(pid, session) |>
-  count(pid, name = "sessions")
-
-p_durations <- p0 %+%
-  dat_session_durations +
-  aes(duration) +
-  scale_x_log10() +
-  coord_cartesian(xlim = c(5, 900)) +
-  labs(x = "Session duration (minutes)", y = "Sessions")
-
-p_sessions <- p0 %+%
-  dat_session_counts +
-  coord_cartesian(xlim = c(0, 15)) +
-  aes(sessions) +
-  labs(x = "Sessions", y = "Participants")
-
-p_mood <- p0 +
-  aes(mood) +
-  labs(x = "Mood", y = "Responses")
-
-(p_durations | p_sessions | p_mood) +
-  plot_annotation(tag_level = "A")
-```
-
-```{r}
-tmp <- list(
-  dat_session_durations$duration,
-  dat_session_counts$sessions,
-  dat$mood,
-  dat$mood[dat$post == 0],
-  dat$mood[dat$post == 1]
-) |> 
-  map(~median_qi(.x, .width = 0.8, na.rm = TRUE)) |> 
-  map(~str_glue_data(.x, "{y} [{ymin}, {ymax}]"))
-```
-
-The median session duration was `r tmp[[1]]` minutes [10 and 90 percentiles]; the median player contributed data from `r tmp[[2]]` sessions, and the median mood was `r tmp[[3]]` (pre-session: `r tmp[[4]]`, during play: `r tmp[[5]]`).
-
-```{r}
-#| label: fig-data
-#| include: true
-#| fig-width: 8
-#| fig-height: 5
-#| fig-cap: "**A**.
-Representative data from three participants (rows) showing scatterplots of mood responses over that session's (columns) duration.
-Each session's first mood response is highlighted in red, subsequent reports are blue.
-**B**.
-Horizontal histograms of all sessions' mood response for that person, with means and bootstrapped 95%CIs.
-**C**.
-Histograms of session-mean (**D**.
-person-mean) moods before (top) and during (bottom) sessions.
-**E**.
-Differences in session-mean (**F**.
-player-mean) mood differences (during session - pre-play).
-**G**.
-Scatterplot of person-mean mood reports at beginning (x-axis) and during game play sessions (y-axis)."
-
-axislabels <- c(0, 0.5, 1.0)
-
-# Plot: Example raw sessions
-set.seed(99)
-dat_example_sessions <- dat |>
-  add_count(pid, session, name = "obs_per_session") |>
-  filter(obs_per_session >= 4) |>
-  mutate(session = as.numeric(as.factor(session)), .by = pid) |>
-  mutate(session_per_person = length(unique(session)), .by = pid) |>
-  filter(session_per_person >= 4) |>
-  arrange(pid, session) |>
-  filter(pid %in% sample(unique(pid), 3)) |>
-  mutate(Session = str_glue("Session {session}"), Person = str_glue("Person {fct_anon(pid)}"))
-
-p_mood_example_sessions <- dat_example_sessions |> 
-  ggplot() +
-  aes(duration / 60, mood, col = post) +
-  scale_color_brewer(
-    "Pre-session measure",
-    palette = "Set1",
-    aesthetics = c("color", "fill")
-  ) +
-  scale_x_continuous(
-    "Session duration (hours)"
-  ) +
-  scale_y_continuous(
-    "Mood",
-    limits = c(0, 1),
-    breaks = axislabels,
-    labels = axislabels
-  ) +
-  geom_point(size = 1.5, alpha = 1) +
-  facet_grid(
-    rows = vars(Person), 
-    cols = vars(Session),
-    scales = "free_x"
-  ) +
-  theme(
-    legend.position = "none",
-    strip.text.y = element_blank()
-  )
-
-p_mood_example_all <- dat_example_sessions |> 
-  mutate(Session = "All sessions") |> 
-  ggplot() +
-  aes(y = mood, x = post, col = post, fill = post) +
-  scale_color_brewer(
-    "Pre-session measure",
-    palette = "Set1",
-    aesthetics = c("color", "fill")
-  ) +
-  scale_x_discrete(
-    "Session",
-    breaks = c(0, 1),
-    labels = c("Pre", "During"),
-    expand = expansion(c(0.2, 2.4))
-  ) +
-  scale_y_continuous(
-    limits = c(0, 1),
-    breaks = axislabels,
-    labels = axislabels
-  ) +
-  stat_summary(
-    fun.data = mean_cl_boot,
-    fatten = 2
-  ) +
-  geom_dots(side = "right", alpha = .33, binwidth = .075) +
-  facet_grid(
-    rows = vars(Person), 
-    cols = vars(Session)
-  ) +
-  theme(
-    axis.text.y = element_blank(), 
-    axis.title.y = element_blank(), 
-    axis.ticks.y = element_blank(), 
-    axis.text.x = element_text(hjust = 0),
-    strip.background.y = element_blank(),
-    legend.position = "none"
-  )
-
-p_mood_example <- (p_mood_example_sessions | p_mood_example_all) +
-  plot_layout(widths = c(92, 8))
-
-# Plot: Raw mood histograms at pre and during play
-p_mood_prepost_raw <- dat |> 
-  mutate(post = factor(post, labels =  c("Pre-play", "During play"))) |>
-  ggplot(aes(mood, fill = post)) +
-  scale_color_brewer(
-    "Pre-session measure",
-    palette = "Set1",
-    aesthetics = c("color", "fill")
-  ) +
-  geom_histogram(bins = 30, col = "white") +
-  scale_y_continuous(
-    "Observations",
-    expand = expansion(c(0.001, 0.1)),
-    breaks = scales::extended_breaks(5)
-  ) +
-  scale_x_continuous(
-    "Mood",
-    expand = expansion(c(0.001)),
-    breaks = axislabels
-  ) +
-  facet_wrap("post", ncol = 1, scales = "free_y") +
-  theme(legend.position = "none")
-
-# Plot: Person-session-mean mood histograms at pre and during play
-p_mood_prepost_sessions <- p_mood_prepost_raw %+%
-  summarise(
-    p_mood_prepost_raw$data,
-    mood = mean(mood, na.rm = TRUE), .by = c(post, pid, ps)
-  ) +
-  scale_y_continuous(
-    "Sessions",
-    expand = expansion(c(0.001, 0.1)),
-    breaks = scales::extended_breaks(5)
-  )
-
-# Plot: Person-mean mood histograms at pre and during play
-p_mood_prepost_players <- p_mood_prepost_sessions %+%
-  summarise(
-    p_mood_prepost_raw$data,
-    mood = mean(mood, na.rm = TRUE), .by = c(post, pid)
-  ) +
-  scale_y_continuous(
-    "Players",
-    expand = expansion(c(0.001, 0.1)),
-    breaks = scales::extended_breaks(5)
-  )
-
-# Plot: Difference histogram (sessions)
-p_mood_difference_sessions <- p_mood_prepost_sessions$data |> 
-  pivot_wider(names_from = post, values_from = mood) |> 
-  ggplot(aes(`During play` - `Pre-play`)) +
-  geom_histogram(bins = 50, col = "white") +
-  geom_vline(xintercept = 0, linewidth = .33, col = "dodgerblue") +
-  scale_y_continuous(
-    "Sessions",
-    expand = expansion(c(0.001, 0.1))
-  ) +
-  scale_x_continuous(
-    "Difference",
-    expand = expansion(c(0.001))
-  ) +
-  coord_cartesian(xlim = c(-0.4, 0.6))
-
-# Plot: Difference histogram (players)
-p_mood_difference_players <- p_mood_difference_sessions %+% 
-  summarise(
-    p_mood_difference_sessions$data,
-    across(`Pre-play`:`During play`, ~mean(.x, na.rm = TRUE)),
-    .by = pid
-  ) +
-  scale_y_continuous(
-    "Players",
-    expand = expansion(c(0.001, 0.1))
-  )
-
-p_mood_difference <- (
-  (p_mood_difference_sessions + 
-     theme(
-      axis.title.x = element_blank(), 
-      axis.text.x = element_blank(),
-      axis.ticks.x = element_blank()
-    )) / 
-    p_mood_difference_players
-  )
-
-p_mood_biscatter_sessions <- p_mood_difference_sessions$data |> 
-  ggplot(aes(`Pre-play`, `During play`)) +
-  scale_x_continuous(
-    expand = expansion(0.01),
-    breaks = axislabels,
-    labels = axislabels
-  ) +
-  scale_y_continuous(
-    expand = expansion(0.01),
-    breaks = axislabels,
-    labels = axislabels
-  ) +
-  geom_point(
-    alpha = .25, size = 1, shape = 1
-  ) +
-  geom_abline(linewidth = .33, col = "dodgerblue") +
-  theme(aspect.ratio = 1)
-
-p_mood_biscatter_players <- p_mood_biscatter_sessions %+% 
-  p_mood_difference_players$data
-
-p_mood_biscatter <- (p_mood_biscatter_sessions / p_mood_biscatter_players)
-
-p_mood_example / 
-  (
-    p_mood_prepost_sessions | 
-      p_mood_prepost_players | 
-      p_mood_difference | 
-      p_mood_biscatter_players
-  ) +
-  plot_layout(heights = c(4.5, 5.5)) +
-  plot_annotation(tag_levels = "A")
-```
-
-We then turned to our primary research question: To what extent does player mood in PWS change from pre-play to during play?
-First, we visualized the relevant data in @fig-data: Panel A shows mood responses from three example participants' first five sessions of play.
-@fig-data B then shows histograms of those data to facilitate visual comparison of the raw data.
-We then calculated sessions' and participants' mean moods at the beginning and during each session (@fig-data C, D, and G), and their differences (E and F).
-Visual inspection suggested greater mood during play compared to the pre-play measures at both the level of sessions and players.
-
-```{r}
-#| label: tbl-avg
-#| include: true
-#| tbl-cap: Population level average mood at beginning and during a game play session, and their difference (during - pre-play).
-(Estimate = posterior mean, CI = 95% quantile interval, pd = probability of direction.)
-
-hs <- c(
-  "Pre-play" = "Intercept + post1 * -.5 = 0",
-  "During play" = "Intercept + post1 * .5 = 0",
-  "Difference" = "post1 = 0"
-)
-
-h_avg_ <- hypothesis(fit, hs)
-h_avg <- tibble(h_avg_$samples)
-names(h_avg) <- names(hs)
-h_avg |> 
-  summarise_draws(
-    mean = ~number(mean(.), .001), 
-    ~number2(quantile2(., probs = c(.025, .975)), .001), 
-    pd = ~percent2(Pr(. > 0))
-  ) |> 
-  mutate(CI = str_glue("[{q2.5}, {q97.5}]")) |> 
-  select(` ` = variable, Estimate = mean, CI, pd) |> 
-  knitr::kable()
-```
-
-The model results confirmed the visual impressions described above: @tbl-avg indicates, with great certainty, that mood increased by about [`r paste0(map(h_avg_$hypothesis[3,4:5], ~number(.x, .001)), collapse = ", ")`] units (@fig-coef, "Raw").
-
-```{r}
-#| label: scaled-coefficients
-
-posterior_pop <- as_draws_df(
-  fit, 
-  c("sd_pid__Intercept", "sd_pid__post1", "sigma", "b_post1")
-) |> 
-  mutate(
-    total_residual = sd_pid__Intercept + sd_pid__post1 + sigma,
-    b_post1_scaled = b_post1 / total_residual
-  ) |> 
-  tibble() |> 
-  select(-starts_with("."))
-
-posterior_pop_sum <- posterior_pop |> 
-  summarise_draws(
-    mean = ~number(mean(.), .001), 
-    ~number2(quantile2(., probs = c(.025, .975)), .001)
-  ) |> 
-  mutate(
-    x = str_glue("{mean} [{q2.5}, {q97.5}]")
-  )
-```
-
-To further interpret the magnitude of the difference in moods between pre- and during play, we compared the contrast to the natural variations in mood as estimated by the model (standard deviation of the between-person intercept heterogeneity ($\tau_{\alpha_0}$ = `r posterior_pop_sum$x[1]`), between-person pre-during difference heterogeneity ($\tau_{\alpha_1}$ = `r posterior_pop_sum$x[2]`) and to the residual standard deviation ($\sigma$ = `r posterior_pop_sum$x[3]`).
-The standardized (with the total residual variation from the three above components) pre-play -- during play contrast was `r posterior_pop_sum$x[6]` (@fig-coef, "Scaled"), a not ignorable difference.
-
-```{r}
-#| label: heterogeneity-calculate
-
-posterior_pid <- coef(fit, summary = FALSE)$pid[,,"post1"]
-posterior_pid_scaled <- apply(
-  posterior_pid, 2, 
-  function(x) x / posterior_pop$total_residual
-) |> 
-  apply(2, function(x) mean_qi(x)) |> 
-  bind_rows(.id = "pid")
-posterior_pid <- posterior_pid |> 
-  apply(2, function(x) mean_qi(x)) |> 
-  bind_rows(.id = "pid")
-
-posterior_pid_sum <- bind_rows(
-  "b_post1" = posterior_pid,
-  "b_post1_scaled" = posterior_pid_scaled,
-  .id = "name"
-)
-
-x <- posterior_pop |> 
-  mutate(
-    prop_pos = pnorm(0, b_post1, sd_pid__post1, lower.tail = FALSE)
-  ) |> 
-  mean_qi(prop_pos) |> 
-  mutate(x = str_glue("{percent(prop_pos, .1)} [{number(.lower*100, .1)}, {number(.upper*100, .1)}]")) |> 
-  pull(x)
-```
-
-We then turned our attention to between-person differences in the difference in during and pre-play mood.
-With a sample of `r s$n_sub`, it is not surprising that the population-level contrast is credibly different from zero.
-We therefore asked the subsequent question: What proportion of the population from which the sample was drawn would we expect there to be an elevation in mood during play?
-The boxplots in @fig-coef show the distribution of player-specific contrasts estimated by the model, which indicated that `r x` of players in the population experience a positive shift in mood during play, compared to their moods during login.
-
-```{r}
-#| label: fig-coef
-#| include: true
-#| fig.width: 3
-#| fig.height: 1.4
-#| fig.cap: "Difference in mood during play compared to mood at session start.
-Points and intervals indicate posterior means and 95%CIs of the population level difference; boxplots summarize the player-level estimates.
-Raw estimates are on the VAS scale, whereas Scaled estimates indicate contrasts divided by the total residual deviance of the model."
-
-posterior_pop |> 
-  select(starts_with("b_")) |> 
-  pivot_longer(everything()) |>
-  group_by(name) |> 
-  mean_qi(value, .width = 0.95) |> 
-  ggplot(aes(value, name)) +
-  scale_x_continuous(
-    "During play - pre-play contrast",
-    breaks = extended_breaks(7)
-  ) +
-  scale_y_discrete(
-    breaks = c("b_post1", "b_post1_scaled"),
-    labels = c("Raw", "Scaled")
-  ) +
-  coord_cartesian(xlim = c(-0.25, 0.65)) +
-  geom_vline(xintercept = 0, linewidth = .33, col = "dodgerblue") +
-  geom_boxploth(
-    data = posterior_pid_sum,
-    aes(x = y, group = name), 
-    outlier.shape = NA,
-    notch = TRUE,
-    size = 0.25,
-    width = .2
-  ) +
-  geom_pointrangeh(
-    aes(xmin = .lower, xmax = .upper),
-    fatten = 2.5
-  ) +
-  theme(
-    axis.title.y = element_blank()
-  )
-```
-
-# Bibliography
-
-::: {#refs}
-:::

### 2023-09-19 functions.R

@@ -1,0 +1,14 @@
+library(scales)
+
+# Format big numbers with a comma
+number2 <- function(x, ...) {
+  number(x, ..., big.mark = ",")
+}
+
+# Replace extreme percentages
+percent2 <- function(x, accuracy = .1) {
+  x <- percent(x, accuracy = accuracy)
+  x <- if_else(x == "100.0%", ">99.9%", x)
+  x <- if_else(x == "0.0%", "<0.1%", x)
+  x
+}

### 2023-09-19 continuous-time-draft.qmd
styling
I think

--- 2023-09-07 continuous-time-draft.qmd
@@
 p0 <- ggplot(
-  filter(dat, 
+    filter(
+        dat,
         pid == "p6622",
@@
     ) +
-  scale_x_datetime(limits = c(ymd_hms("2022-09-19 16:00:00"), ymd_hms("2022-09-20 00:00:00"))) +
+    scale_x_datetime(
+        limits = c(
+            ymd_hms("2022-09-19 16:00:00"),
+            ymd_hms("2022-09-20 00:00:00")
+        )
+    ) +
     scale_y_continuous(breaks = 14:16) +
@@
 p1 <- ggplot(
-  filter(dat2, 
+    filter(
+        dat2,
         pid == "p6622",
@@
 ```{r}
-filter(dat, 
+filter(
+    dat,
     pid == "p6622",
     session %in% 14:16
-) %>% 
+) |>
     kable()
 
-filter(dat2, 
+filter(
+    dat2,
     pid == "p6622",
     session %in% 14:16
-) %>% 
+) |>
     kable()
@@
 p0 <- ggplot(
-  filter(dat, 
+    filter(
+        dat,
         pid == "p10",
@@
             ymd_hms("2023-02-25 00:00:00"),
-      ymd_hms("2023-02-29 00:00:00"
-      )
+            ymd_hms("2023-02-29 00:00:00")
         )
@@
 p1 <- ggplot(
-  filter(dat2, 
+    filter(
+        dat2,
         pid == "p10",
@@
 ```{r}
-filter(dat, pid == "p10", session == 16) %>% 
+filter(dat, pid == "p10", session == 16) |>
     kable()
 
-filter(dat2, pid == "p10", session == 16) %>% 
+filter(dat2, pid == "p10", session == 16) |>
     kable()
@@
 
+### Session lengths
+
+What's the distribution of session lengths
+```{r}
+median_ses_len <- dat |>
+    group_by(pid, session) |>
+    mutate(
+        duration = duration - duration[1],
+    ) |>
+    summarise(session_length = max(duration)) |>
+    group_by(pid) |>
+    summarise(median_len = median(session_length))
+median(median_ses_len$median_len)
+mean(median_ses_len$median_len > 60)
+```
+
+"Typical" session length is 35 minutes. Few participants typically play for more than an hour. Plot estimated trend over 0 to 2 hours.
+
 ### Fit model
@@
 ```{r}
-dat <- dat %>% 
+dat <- dat |>
+    mutate(
+        ps = interaction(pid, session, sep = "_")
+    ) |>
+    filter(prompt == "Wellbeing") |>
     mutate(
@@
         .by = c(pid, session)
-  ) %>% 
-  filter(prompt == "Wellbeing")
+    )
+write_rds(dat, "data/ct-dev-data.rds")
 ```
@@
 ```{r}
+#| code-fold: show
 library(rms)
-fit0 <- lmer(mood ~ hours + (1 | session) + (1 | pid), data = dat)
-summary(fit0)
+fit0 <- lmer(mood ~ hours + (1 | ps) + (1 | pid), data = dat)
 ```
-
-What's the distribution of session lengths
-```{r}
-median_ses_len <- dat %>% 
-  group_by(pid, session) %>% 
-  summarise(session_length = max(duration)) %>% 
-  group_by(pid) %>% 
-  summarise(median_len = median(session_length))
-
-median(median_ses_len$median_len)
-
-mean(median_ses_len$median_len > 60)
-
-```
-
-"Typical" session length is 24 minutes. Few participants typically play for more than an hour. Plot estimated trend over 0 to 2 hours.
 
@@
     ~hours,
-  at = list(hours = seq(0,2, length.out = 100)),
+    at = list(hours = seq(0, 2, length.out = 100)),
     lmer.df = "asymptotic"
@@
 
-p0 <- as.data.frame(emm0) %>% 
+p0 <- as.data.frame(emm0) |>
     ggplot(aes(hours, emmean)) +
@@
 ```{r}
-fit1 <- lmer(mood ~ rcs(hours, 5) + (1 | session) + (1 | pid), data = dat)
+#| code-fold: show
+fit1 <- lmer(mood ~ rcs(hours, 5) + (1 | ps) + (1 | pid), data = dat)
+```
+
+```{r}
 emm1 <- emmeans(
@@
     ~hours,
-  at = list(hours = seq(0,2, length.out = 100)),
+    at = list(hours = seq(0, 2, length.out = 100)),
     lmer.df = "asymptotic"
 )
-p1 <- as.data.frame(emm1) %>% 
+
+## for plotting knot placement
+get_knot_coords <- function(fit) {
+    knots <- attr(rcs(dat$hours, 5), "parms")
+    knots_x <- rcs(knots, knots)
+    b0 <- fixef(fit)[1]
+    b <- fixef(fit)[-1]
+    knots_y <- b0 + knots_x[, 1] * b[1] + knots_x[, 2] * b[2] + knots_x[, 3] * b[3] + knots_x[, 4] * b[4]
+    data.frame(
+        x = knots,
+        y = c(knots_y)
+    )
+}
+p1 <- as.data.frame(emm1) |>
     ggplot(aes(hours, emmean)) +
@@
     geom_ribbon(aes(ymin = asymp.LCL, ymax = asymp.UCL), alpha = 0.25) +
+    geom_point(
+        data = get_knot_coords(fit1),
+        aes(x, y)
+    ) +
     labs(x = "Session duration (hours)")
@@
 } else {
-  fit2 <- lmer(mood ~ rcs(hours, 5) + (1 | session) + (1 + hours | pid), data = dat)
+    fit2 <- lmer(mood ~ rcs(hours, 5) + (1 | ps) + (1 + hours | pid), data = dat)
     write_rds(fit2, path)
 }
+```
+
+```{r}
 emm2 <- emmeans(
@@
     ~hours,
-  at = list(hours = seq(0,2, length.out = 100)),
+    at = list(hours = seq(0, 2, length.out = 100)),
     lmer.df = "asymptotic"
 )
-p2 <- as.data.frame(emm2) %>% 
+p2 <- as.data.frame(emm2) |>
     ggplot(aes(hours, emmean)) +
@@
     geom_ribbon(aes(ymin = asymp.LCL, ymax = asymp.UCL), alpha = 0.25) +
+    # geom_point(
+    #   data = get_knot_coords(fit2),
+    #   aes(x,y)
+    # ) +
     labs(x = "Session duration (hours)")
@@
 I think the shape of the spline makes sense.
+
+## YOLO model with random slopes
+Here I add a random slope to each piece of the spline, i.e. 4 random slopes at the participant level. The model is fit on a random subset of 1000 pids.
+```{r}
+#| cold-fold: show
+
+# include 1000 participants
+set.seed(1337)
+pids <- sample(unique(dat$pid), 1000)
+tmp <- filter(dat, pid %in% pids)
+# so predict won't complain
+tmp <- tmp |>
+    mutate(
+        pid = as.integer(pid),
+        ps = as.integer(ps)
+    )
+# same spline but different parameterization
+# usually less estimation problems
+spline <- ns(
+    dat$hours,
+    knots = c(0.12, 0.4, 1),
+    Boundary.knots = c(0.016, 3)
+)
+colnames(spline) <- paste0("hours", 0:3)
+tmp <- cbind(dat, spline)
+path <- "models/lmm-fit3.Rds"
+if (file.exists(path)) {
+    fit3 <- read_rds(path)
+} else {
+    fit3 <- lmer(
+        mood ~ hours0 + hours1 + hours2 + hours3 +
+         (1 | ps) + 
+         (1 + hours0 + hours1 + hours2 + hours3 | pid),
+        data = tmp
+    )
+    write_rds(fit3, path)
+}
+
+```
+
+I'm interested in if the spline changes based on the pre-session scores.
+Let's plot the fixed effects at -2, -1, 0 times the random intercept SD + the corresponding random slope values based on their correlation with the random intercept.
+
+```{r}
+get_trend <- function(fit, data, re_int = 0, re_s0 = 0, re_s1 = 0, re_s2 = 0, re_s3 = 0) {
+    hours <- seq(0, 2, length.out = 100)
+    x <- ns(
+        hours,
+        knots = c(0.12, 0.4, 1),
+        Boundary.knots = c(0.016, 3)
+    )
+    b0 <- fixef(fit)[1] + re_int
+    b <- fixef(fit)[-1]
+    y <- b0 + x[, 1] * (b[1] + re_s0) + x[, 2] * (b[2] + re_s1) + x[, 3] * (b[3] + re_s2) + x[, 4] * (b[4] + re_s3)
+    data.frame(
+        x = hours,
+        y = c(y),
+        re_int
+    )
+}
+re <- as.data.frame(VarCorr(fit3))
+pid_int <- re[2, "sdcor"]
+pid_slope0 <- re[3, "sdcor"]
+pid_slope1 <- re[4, "sdcor"]
+pid_slope2 <- re[5, "sdcor"]
+pid_slope3 <- re[6, "sdcor"]
+pid_cor0 <- re[7, "sdcor"]
+pid_cor1 <- re[8, "sdcor"]
+pid_cor2 <- re[9, "sdcor"]
+pid_cor3 <- re[10, "sdcor"]
+sd1 <- 1
+sd2 <- 2
+rbind(
+    get_trend(fit3, data = tmp),
+    get_trend(
+        fit3,
+        data = tmp,
+        -sd1 * pid_int,
+        -sd1 * pid_cor0 * pid_slope0,
+        -sd1 * pid_cor1 * pid_slope1,
+        -sd1 * pid_cor2 * pid_slope2,
+        -sd1 * pid_cor3 * pid_slope3
+    ),
+    get_trend(
+        fit3,
+        data = tmp,
+        -sd2 * pid_int,
+        -sd2 * pid_cor0 * pid_slope0,
+        -sd2 * pid_cor1 * pid_slope1,
+        -sd2 * pid_cor2 * pid_slope2,
+        -sd2 * pid_cor3 * pid_slope3
+    )
+) |>
+    ggplot(
+        aes(
+            x,
+            y,
+            color = factor(re_int),
+            group = factor(re_int)
+        )
+    ) +
+    geom_line() +
+    labs(x = "Session duration (hours)")
+```
+
+The random intercepts are only weakly correlated with each piece of the cubic spline, so the shape doesn't change much.
+
+Let's also predict some individual slopes.
+
+```{r}
+pred <- filter(tmp, pid %in% sample(unique(tmp$pid), 10))
+pred$yhat <- predict(fit3, newdata = pred)
+ggplot(
+    data = pred,
+    aes(
+        x = hours,
+        y = yhat,
+        group = ps
+    )
+) +
+    geom_line(
+        color = "blue",
+        linewidth = 1
+    ) +
+    geom_point(
+        aes(y = mood)
+    ) +
+    facet_wrap(~pid)
+```
+
+Zoom in on pid 9263.
+```{r}
+ggplot(
+    data = filter(pred, pid == 9263),
+    aes(
+        x = hours,
+        y = yhat,
+        group = ps
+    )
+) +
+    geom_line(
+        aes(y = mood),
+        linetype = "dashed",
+        alpha = 0.33
+    ) +
+    geom_line(
+        color = "blue",
+        linewidth = 2
+    ) +
+    geom_point(
+        aes(y = mood)
+    ) +
+    facet_wrap(~session)
+```
+
+Probably good enough, even tho the session effect is only modelled using a random intercept.
+
+### Predict change
+Another YOLO approach to estimate individual change during a session is to to look at the model predictions.
+
+TODO: should be based on the actual session length, not only the range of WB prompts.
+
+```{r}
+pred <- fit3@frame
+pred$yhat <- predict(fit3)
+pred |>
+    # compare the last and first value
+    dplyr::summarize(change = last(yhat) - yhat[1], .by = c(pid, ps)) |>
+    # average over sessions within participants
+    dplyr::summarize(change = mean(change), .by = pid) |>
+    # average over participants
+    dplyr::summarize(mean(change), sd(change))
+```
+
+We can also plot the distribution of the participants average change during sessions.
+```{r}
+pred |>
+    dplyr::summarize(change = last(yhat) - yhat[1], .by = c(pid, ps)) |>
+    dplyr::summarize(change = mean(change), .by = pid) |>
+    ggplot(aes(change)) +
+    geom_histogram()
+```
+
+They really don't change much. Let's look at a participant out in the tails.
+```{r}
+pred |>
+    dplyr::summarize(change = last(yhat) - yhat[1], .by = c(pid, ps)) |>
+    filter(change > 0.1) |>
+    head()
+```
+
+Zoom in on pid 23.
+```{r}
+pred2 <- filter(tmp, pid == 23)
+pred2$yhat <- predict(fit3, newdata = pred2)
+ggplot(
+    data = pred2,
+    aes(
+        x = hours,
+        y = yhat,
+        group = ps
+    )
+) +
+    geom_line(
+        aes(y = mood),
+        linetype = "dashed",
+        alpha = 0.33
+    ) +
+    geom_line(
+        color = "blue",
+        linewidth = 2
+    ) +
+    geom_point(
+        aes(y = mood)
+    ) +
+    facet_wrap(~session)
+```
+
+This looks kinda reasonable?
+
+## brms
+Ignore for now.
+
+```{r}
+#| code-fold: show
+library(brms)
+options(
+    mc.cores = 8,
+    brms.backend = "cmdstanr"
+)
+fit_b0 <- brm(
+    mood ~ hours0 + hours1 + hours2 + hours3 + (1 | ps) + (1 | pid),
+    data = tmp,
+    #threads = 2,
+    chains = 4,
+    iter = 500,
+    backend = "rstan",
+    file = "models/brm-pre-post-ct0"
+)
+pp_check(fit_b0)
+```
+
+Censoring
+```{r}
+#| code-fold: show
+tmp <- tmp |>
+    mutate(
+        cl = case_when(
+            mood == 0 ~ "left",
+            mood == 1 ~ "right",
+            TRUE ~ "none"
+        )
+    )
+fit_b1 <- brm(
+    mood | cens(cl) ~ hours0 + hours1 + hours2 + hours3 + (1 | ps) + (1 | pid),
+    data = tmp,
+    threads = 2,
+    iter = 1000,
+    file = "models/brm-pre-post-ct1"
+)
+pp_check(fit_b1)
+```
+
+## GAM(M)
+
+Let's try some GAMs. 
+```{r}
+library(mgcv)
+knots <- list(hours = c(0.016, 0.12, 0.4, 1, 3))
+tmp <- tmp |>
+    mutate(
+        pid = as.factor(pid),
+        ps = as.factor(ps)
+    )
+```
+
+We'll start with a cubic spline using the same knot placement as used with the LMMs.
+```{r}
+#| code-fold: show
+fit_g_0 <- gam(
+    mood ~ s(hours, bs = "cr", k = 5),
+    knots = knots,
+    method = "REML",
+    data = tmp,
+)
+fit_g_0 <- gam(
+    mood ~ s(hours, k = 15),
+    #knots = knots,
+    method = "REML",
+    data = filter(tmp, hours < 5),
+)
+```
+
+```{r}
+plot_gam <- function(fit, t_end = 2) {
+    emm0 <- emmeans(
+        fit,
+        ~hours,
+        at = list(hours = seq(0, t_end, length.out = 100))
+    )
+    p0 <- as.data.frame(emm0) |>
+        ggplot(aes(hours, emmean)) +
+        geom_line() +
+        geom_ribbon(aes(ymin = lower.CL, ymax = upper.CL), alpha = 0.25) +
+        labs(x = "Session duration (hours)")
+    p0
+}
+plot_gam(fit_g_0) + lims(y = c(0,1))
+```
+
+we could add a random intercept for pid and session like this
+```{r}
+#| eval: false
+#| code-fold: show
+fit_g_1 <- gam(
+    mood ~ s(hours, bs = "cr", k = 5) + s(pid, bs = "re") + s(ps, bs = "re"),
+    knots = knots,
+    data = tmp,
+)
+```
+
+The call above is super slow, so let's use `gamm4` instead.
+
+```{r}
+#| code-fold: show
+library(gamm4)
+fit_g_1_lme4 <- gamm4(
+    mood ~ s(hours, bs = "cr", k = 5),
+    knots = knots,
+    data = tmp,
+    random = ~ (1 | pid) + (1 | ps)
+)
+```
+
+```{r}
+pred <- data.frame(hours = seq(0, 2, length.out = 100))
+pred$yhat <- predict(fit_g_1_lme4$gam, newdata = pred)
+ggplot(
+    pred,
+    aes(hours, yhat)
+) +
+    geom_line() +
+    lims(y = c(0,1)) +
+    # geom_ribbon(aes(ymin = lower.CL, ymax = upper.CL), alpha = 0.25) +
+    labs(x = "Session duration (hours)")
+```
+
+We are using the same smooth for everyone here, which is probably not what we want.
+
+### Separate smooth per participant
+Let's try using a separate smooth per pid. We only fit the model to a subset of participants to save time.
+
+```{r}
+#| code-fold: show
+set.seed(6666)
+tmp2 <- filter(tmp, pid %in% sample(unique(pid), 200)) |>
+    filter(!is.na(mood), hours <= 10)
+fit_g_by_lme4 <- gamm4(
+    mood ~ s(hours, by = pid, bs = "cr", k = 5),
+    knots = knots,
+    data = tmp2,
+    random = ~ (1 | pid) + (1 | ps)
+)
+```
+
+N.B. that only the GAM part is plotted, I'm not sure if there is a command that automagically includes the random effects when using `gamm4`.
+
+```{r}
+pred <- expand.grid(
+    hours = seq(0, 2, length.out = 100),
+    pid = unique(tmp2$pid)
+)
+pred$yhat <- predict(fit_g_by_lme4$gam, newdata = pred)
+pred$model <- "by"
+ggplot(
+    pred,
+    aes(
+        hours,
+        yhat,
+        group = pid,
+        color = pid
+    )
+) +
+    geom_line() +
+    # geom_ribbon(aes(ymin = lower.CL, ymax = upper.CL), alpha = 0.25) +
+    labs(x = "Session duration (hours)") +
+    theme(legend.position = "none")
+
+plot_pids <- sample(unique(tmp2$pid), 10)
+tmp2 <- filter(tmp2, !is.na(mood))
+tmp2$yhat <- predict(fit_g_by_lme4$gam)
+tmp2$model <- "by"
+ggplot(
+    filter(tmp2, pid %in% plot_pids),
+    aes(
+        hours,
+        yhat,
+        group = ps,
+        color = pid
+    )
+) +
+    geom_line(aes(y = mood)) +
+    geom_line(color = "blue", linewidth = 1) +
+    geom_point(aes(y = mood)) +
+    # geom_ribbon(aes(ymin = lower.CL, ymax = upper.CL), alpha = 0.25) +
+    labs(x = "Session duration (hours)") +
+    facet_wrap(~pid) +
+    theme(legend.position = "none")
+
+pred_all <- filter(tmp2, pid %in% plot_pids)
+```
+
+We're fitting a separate fixed and random effects for each pid. Fitting this to the full data will be super slow?
+```{r}
+summary(fit_g_by_lme4$mer)
+```
+
+### Factor-smooth interaction 
+We can also try a factor-smooth interaction instead, this should be closer to what we want.
+
+```{r}
+#| code-fold: show
+
+# TODO: include a main effect
+fit_g_fs_lme4 <- gamm4(
+    mood ~ s(hours, bs = "cr", k = 5) + s(hours, pid, bs = "fs", xt = list(bs = "cr"), k = 5),
+    knots = knots,
+    data = tmp2,
+    random = ~ (1 | pid) + (1 | ps)
+)
+plot(fit_g_fs_lme4$gam, select = 1)
+```
+
+`gamm4` is not supported by `emmeans`, and I'm not sure how to plot the "marginal" smooth.
+
+```{r}
+pred <- expand.grid(
+    hours = seq(0, 2, length.out = 100),
+    pid = unique(tmp2$pid)
+)
+pred$yhat <- predict(fit_g_fs_lme4$gam, newdata = pred)
+pred$model <- "fs"
+ggplot(
+    pred,
+    aes(
+        hours,
+        yhat,
+        group = pid,
+        color = pid
+    )
+) +
+    geom_line() +
+    # geom_ribbon(aes(ymin = lower.CL, ymax = upper.CL), alpha = 0.25) +
+    labs(x = "Session duration (hours)") +
+    theme(legend.position = "none")
+
+tmp2$yhat <- predict(fit_g_fs_lme4$gam)
+tmp2$model <- "fs"
+ggplot(
+    filter(tmp2, pid %in% plot_pids),
+    aes(
+        hours,
+        yhat,
+        group = ps,
+        color = pid
+    )
+) +
+    geom_line(aes(y = mood)) +
+    geom_line(color = "blue", linewidth = 1) +
+    geom_point(aes(y = mood)) +
+    lims(y = c(0, 1)) +
+    # geom_ribbon(aes(ymin = lower.CL, ymax = upper.CL), alpha = 0.25) +
+    labs(x = "Session duration (hours)") +
+    facet_wrap(~pid) +
+    theme(legend.position = "none")
+pred_all <- rbind(pred_all, filter(tmp2, pid %in% plot_pids))
+```
+
+The smooths in the plot above are all linear.
+Although, it looks pretty reasonable compared to the observed data.
+(TODO: investigate with more data.) If we fit a simple smooth to the data it's also linear.
+
+```{r}
+#| code-fold: show
+gam(
+    mood ~ s(hours, bs = "cr", k = 5),
+    knots = knots,
+    method = "REML",
+    data = tmp2,
+) |>
+    plot_gam()
+
+```
+
+### LMM again
+As a comparisons, let's refit the LMM with a cubic spline to the subset of data used here. And then we predict without including the random intercepts.
+
+```{r}
+#| code-fold: show
+fit3 <- lmer(
+  mood ~ hours0 + hours1 + hours2 + hours3 + (1 | ps) + (1 + hours0 + hours1 + hours2 + hours3 | pid), 
+  data = tmp2
+)
+```
+
+```{r}
+tmp2$yhat <- predict(
+    fit3,
+    re.form = ~ (0 + hours0 + hours1 + hours2 + hours3 | pid)
+)
+tmp2$model <- "lmer"
+ggplot(
+    filter(tmp2, pid %in% plot_pids),
+    aes(
+        hours,
+        yhat,
+        group = ps,
+        color = pid
+    )
+) +
+    geom_line(aes(y = mood)) +
+    geom_line(color = "blue", linewidth = 1) +
+    geom_point(aes(y = mood)) +
+    lims(y = c(0, 1)) +
+    # geom_ribbon(aes(ymin = lower.CL, ymax = upper.CL), alpha = 0.25) +
+    labs(x = "Session duration (hours)") +
+    facet_wrap(~pid)
+
+pred_all <- rbind(pred_all, filter(tmp2, pid %in% plot_pids))
+```
+
+### Combined plot
+Plot the predictions from the 3 models in a combined figure.
+
+```{r}
+#| fig.height: 5
+#| fig.width: 7
+ggplot(
+    pred_all,
+    aes(
+        hours,
+        yhat,
+        group = interaction(ps, model),
+        color = model
+    )
+) +
+    geom_line(data = filter(pred_all, model == "fs"), aes(y = mood, color = NA), alpha = 0.5) +
+    geom_point(data = filter(pred_all, model == "fs"), aes(y = mood, color = NA), alpha = 0.5) +
+    geom_line(linewidth = 1.5) +
+    # lims(y = c(0, 1)) +
+    # geom_ribbon(aes(ymin = lower.CL, ymax = upper.CL), alpha = 0.25) +
+    labs(x = "Session duration (hours)") +
+    facet_wrap(~pid, scales = "free_y")
+```
+
+### GAM summary
+Assuming that we wanted to fit these smooths in `brms` to model censoring, then
+
+- separate smooths using `by` is not feasible?
+- factor-smooth interaction is maybe feasible using the full data.
+  - Might be overpenalizing the smooths as they were all straight lines? 
+- the LMM is basically in between the model above when it come to "wiggliness", as would be expected
+
+I'm not sure if some if the issues is caused by how GAMs are affected by the highly irregular sessions + missing data.
+
+LMM might be _the_ model?
 

### 2023-09-20 continuous-time-draft.qmd
small fixes

@@
     mutate(
-        ps = interaction(pid, session, sep = "_")
-    ) |>
-    filter(prompt == "Wellbeing") |>
-    mutate(
         duration = duration - duration[1],
@@
         .by = c(pid, session)
+    ) |> mutate(
+        ps = interaction(pid, session, sep = "_")
+    ) |>
+    filter(prompt == "Wellbeing")
+
+dat |> 
+    summarize(
+        int = duration[1], .by = ps
+        ) |> 
+    summarize(
+        median(int),
+        mean(int),
+        sd(int)
     )
+
 write_rds(dat, "data/ct-dev-data.rds")
@@
 #| cold-fold: show
-
-# include 1000 participants
 set.seed(1337)
-pids <- sample(unique(dat$pid), 1000)
-tmp <- filter(dat, pid %in% pids)
-# so predict won't complain
-tmp <- tmp |>
-    mutate(
-        pid = as.integer(pid),
-        ps = as.integer(ps)
-    )
 # same spline but different parameterization
@@
 colnames(spline) <- paste0("hours", 0:3)
-tmp <- cbind(dat, spline)
+tmp <- cbind(dat, spline) |> 
+    filter(hours <= 3)
 path <- "models/lmm-fit3.Rds"
@@
 get_trend <- function(fit, data, re_int = 0, re_s0 = 0, re_s1 = 0, re_s2 = 0, re_s3 = 0) {
-    hours <- seq(0, 2, length.out = 100)
+    hours <- seq(0, 3, length.out = 100)
     x <- ns(
@@
 ```{r}
+pred <- filter(tmp, pid == "p9263")
+pred$yhat <- predict(fit3, newdata = pred)
 ggplot(
-    data = filter(pred, pid == 9263),
+    data = pred,
     aes(
@@
 pred$yhat <- predict(fit3)
-pred |>
+ps_filtered <- pred |> 
+    summarise(n = n(), .by = ps) |> 
+    filter(n > 1)
+change <- pred |>
     # compare the last and first value
-    dplyr::summarize(change = last(yhat) - yhat[1], .by = c(pid, ps)) |>
+    filter(ps %in% ps_filtered$ps) |> 
+    summarise(change = last(yhat) - yhat[1], .by = c(pid, ps)) |>
     # average over sessions within participants
-    dplyr::summarize(change = mean(change), .by = pid) |>
+    summarise(change = median(change), .by = pid) 
+change |> 
     # average over participants
-    dplyr::summarize(mean(change), sd(change))
+    summarise(mean(change), median(change), sd(change))
 ```
@@
 ```{r}
-pred |>
-    dplyr::summarize(change = last(yhat) - yhat[1], .by = c(pid, ps)) |>
-    dplyr::summarize(change = mean(change), .by = pid) |>
+change |> 
     ggplot(aes(change)) +
@@
 ```{r}
-pred |>
-    dplyr::summarize(change = last(yhat) - yhat[1], .by = c(pid, ps)) |>
+change |> 
     filter(change > 0.1) |>
@@
 
-Zoom in on pid 23.
+Zoom in on pid p10057.
 ```{r}
-pred2 <- filter(tmp, pid == 23)
+pred2 <- filter(tmp, pid == "p10057")
 pred2$yhat <- predict(fit3, newdata = pred2)
@@
 ```{r}
+#| eval: false
+#| label: brms
 #| code-fold: show
@@
 #| code-fold: show
+#| label: brms-cens
+#| eval: false
 tmp <- tmp |>
@@
 )
-fit_g_0 <- gam(
-    mood ~ s(hours, k = 15),
-    #knots = knots,
-    method = "REML",
-    data = filter(tmp, hours < 5),
-)
 ```
@@
 }
-plot_gam(fit_g_0) + lims(y = c(0,1))
+plot_gam(fit_g_0)
 ```
@@
 ```{r}
+#| label: gamm4
 #| code-fold: show
@@
 ```{r}
+#| label: gamm4-fs
 #| code-fold: show

### 2023-09-20 wrangle.qmd
Save 'raw' data

@@
 
-# Filter data:
+# Filter data
+dat_raw <- dat
 # Include only wellbeing responses that occurred
@@
 saveRDS(dat, "data/data.rds")
+saveRDS(dat_raw, "data/data-raw.rds")
 saveRDS(s, "data/s.rds")

### 2023-09-20 analyze.qmd
Reproducible abstract

@@
 names(h_avg) <- names(hs)
-h_avg |> 
+h_avg_tab <- h_avg |> 
   summarise_draws(
@@
   mutate(CI = str_glue("[{q2.5}, {q97.5}]")) |> 
-  select(` ` = variable, Estimate = mean, CI, pd) |> 
+  select(` ` = variable, Estimate = mean, CI, pd)
+h_avg_tab |>   
   knitr::kable()
@@
 
+# Abstract
+
+```{glue}
+#| results: asis
+#| echo: false
+
+Although hundreds of millions of individuals play video games, a medium sometimes considered psychologically damaging, no firm evidence exists on how games affect players’ moods---if at all.
+One reason that current evidence remains mixed is that games are likely to affect players differently depending on the time-scale of the investigation: Do games affect players’ well-being over long periods of time, or are the effects fleeting and only evidenced in near real-time to play?
+In this case study of a single game (PowerWash Simulator), we examined {s$n_sub} players’ moods from {s$n_obs} reports provided in-game.
+We compared players’ moods before play sessions (at login) to their moods during play.
+We find that on average, players report {h_avg_tab[3, "Estimate"]} {h_avg_tab[3, "CI"]} scale percentages greater mood during play than at the beginning of a play session, and that {x} of individuals in the population are predicted to have such positive changes in mood.
+We do not know whether the current results indicate a causal effect, since the hypothetical treatment of “choosing to play” occurs prior to our first measure, nor do we know whether and how our results would generalize to other games and player populations.
+Nevertheless, current results offer a promising glimpse of the potential of games in positively contributing to player mood, and of the potential of in-game telemetry and mood measure to study games’ effects on people at a short time scale where effects are likely to exist.
+```
+
 # Bibliography

### 2023-09-21 continuous-time.qmd
WIP document
this is the real document, *_draft.qmd will be deleted.

---
title: "Continuous time"
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  include = TRUE,
  echo = TRUE
)
```

In the main analysis in [Analyze]() we focused on contrasting the mood at the beginning of a session with the average mood during session. In this document we focus on two additional research questions:

1. How does mood change during a session?
2. Is change during a session related to the participants' pretest mood?

To answer these questions time is modeled as a continuous variable.

## Setup
```{r}
#| label: packages
#| cache: false

library(scales)
library(cmdstanr)
library(posterior)
library(brms)
library(ggstance)
library(ggdist)
library(patchwork)
library(tidyverse)
library(splines)
library(emmeans)
library(lme4)
library(broom.mixed)
library(knitr)
source("functions.R")
```

```{r}
#| label: options
#| cache: false

N_CORES <- as.numeric(Sys.getenv("N_CORES", unset = 1))
N_THREADS <- as.numeric(Sys.getenv("N_THREADS", unset = 2))
N_ITER <- as.numeric(Sys.getenv("N_ITER", unset = 2000))
N_SUBSET <- as.numeric(Sys.getenv("N_SUBSET"))
options(
  mc.cores = N_CORES,
  brms.backend = "cmdstanr"
)
theme_set(
  theme_linedraw(base_size = 10) +
    theme(
      strip.background = element_rect(
        fill = NA,
        color = NA
      ),
      strip.text = element_text(
        colour = "black", hjust = 0
      ),
      axis.text = element_text(
        size = rel(0.75)
      ),
      panel.grid = element_blank()
    )
)
```

## Data
Load the data and remove observations that occurred more than 3 hours into a session.
```{r}
#| label: load-data
dat <- read_csv(
  "data/study_prompt_answered.csv",
  col_select = c(
    pid,
    time = Time_utc,
    duration = CurrentSessionLength,
    prompt = LastStudyPromptType,
    mood = response
  )
) |>
  mutate(
    pid = factor(pid),
    mood = mood / 1000
  )

# Create sessions using heuristics below
dat <- dat |>
  arrange(pid, time) |>
  mutate(
    i = row_number(),
    # Observation is in a new session if this...
    new_session =
    # is the first total observation,
      i == 1 |
        # session duration is zero,
        (duration == 0) |
        # session is shorter than previous session (ticker has reset),
        (duration < lag(duration, default = 0)) |
        # or if more than 30 minutes since previous observation.
        (time > lag(time) + minutes(30)),
    # Session number is the cumulative sum of new sessions...
    session = cumsum(new_session),
    # ... per person
    .by = pid
  ) |>
  select(-c(i, new_session))

dat <- dat |>
  mutate(
    duration = duration - duration[1],
    hours = duration / 60,
    .by = c(pid, session)
  ) |>
  mutate(
    ps = interaction(pid, session, sep = "_")
  ) |>
  filter(prompt == "Wellbeing")

dat <- dat |>
  filter(
    !is.na(mood),
    hours <= 3
  ) |>
  mutate(
    cl = case_when(
      mood == 0 ~ "left",
      mood == 1 ~ "right",
      TRUE ~ "none"
    ),
    mood100 = mood * 100
  )
```

## Main model
Define a spline that uses 4 degrees of freedom.

```{r}
#| label: define-spline

ns4 <- function(x) {
  ns(
    x,
    knots = c(0.05, 0.33, 0.7),
    Boundary.knots = c(0, 3)
  )
}
```

The main model is a LMM, we include random effects for each piece of the spline and a random intercept at the session level.

```{r}
#| label: lmm-ns4

path <- "models/lmm-ns4.Rds"
if (file.exists(path)) {
  fit <- read_rds(path)
} else {
  fit <- lmer(
    mood100 ~ ns4(hours) + (1 | ps) + (1 + ns4(hours) | pid),
    data = dat
  )
  write_rds(fit, path)
}

```

The spline is plotted in @fig-plot-curve, and we can see that there's a small but sharp increase early during a session.

```{r}
#| label: fig-plot-curve
#| fig-cap: Estimated mood during a play session

hours <- seq(0, 3, length.out = 100)
emm1 <- emmeans(
  fit,
  ~hours,
  at = list(hours = hours),
  lmer.df = "asymptotic"
)
p_curve <- as.data.frame(emm1) |>
  ggplot(aes(hours, emmean)) +
  geom_line() +
  geom_ribbon(
    aes(
      ymin = asymp.LCL,
      ymax = asymp.UCL
    ),
    alpha = 0.25
  ) +
  labs(
    x = "Session duration (hours)",
    y = "Mood (0 - 100)"
  )
p_curve
```

@fig-plot-change shows the same spline, but as a contrast where hour = 0 is the reference value.

```{r}
#| label: fig-plot-change
#| fig-cap: Estimated difference in mood over a session compared to the start of a session

d_change <- emmeans::contrast(
  emm1,
  method = "trt.vs.ctrl",
  ref = "hours0"
) |>
  confint() |>
  as.data.frame() |>
  mutate(
    hours = hours[-1]
  )

p_change <- d_change |>
  ggplot(aes(hours, estimate)) +
  geom_line() +
  geom_ribbon(
    aes(
      ymin = asymp.LCL,
      ymax = asymp.UCL
    ),
    alpha = 0.25
  ) +
  lims(y = c(0, 6)) +
  labs(x = "Session duration (hours)")
p_change
```

@tbl-change-contrast shows the estimated change from baseline after 10min, 30 min, 1 hour, 2 hours, and 3 hours.
```{r}
#| label: tbl-change-contrast
#| tbl-cap: Estimated change from baseline
emmeans(
  fit,
  ~hours,
  at = list(hours = c(0, 10 / 60, 0.5, 1, 2, 3)),
  lmer.df = "asymptotic"
) |>
  emmeans::contrast(
    method = "trt.vs.ctrl",
    ref = "hours0",
  ) |>
  confint() |>
  kable(
    digits = 2
  )
```

### Relationship between change and mood at the start of a session
We are also interested in the relationship between the initial mood and change over time. As the intercepts are correlated with the random slopes we can plot the implied splines for someone with mood levels 0, 1, and 2 SD below the average at the start of a session. The logic here is simply: if cor(intercept, slope) = 0.5, then someone with an intercept equal to -1SD is expected to have a slope equal to -0.5 * sd(slope). This is shown in @fig-intercept-sds, which tells us that someone with a lower mood at the beginning of a session is expected to have their mood increase more during a session.

Interpret this with caution.

```{r}
#| label: fig-intercept-sds
#| fig-cap: Estimated curves for participants with an intercept equal to 0, -1SD, and -2SD.
get_trend <- function(
    fit,
    data,
    sd = 0,
    intercept = TRUE) {
  re <- as.data.frame(VarCorr(fit))
  pid_int <- re[2, "sdcor"]
  pid_slope0 <- re[3, "sdcor"]
  pid_slope1 <- re[4, "sdcor"]
  pid_slope2 <- re[5, "sdcor"]
  pid_slope3 <- re[6, "sdcor"]
  pid_cor0 <- re[7, "sdcor"]
  pid_cor1 <- re[8, "sdcor"]
  pid_cor2 <- re[9, "sdcor"]
  pid_cor3 <- re[10, "sdcor"]
  re_int <- sd * pid_int
  re_s0 <- sd * pid_cor0 * pid_slope0
  re_s1 <- sd * pid_cor1 * pid_slope1
  re_s2 <- sd * pid_cor2 * pid_slope2
  re_s3 <- sd * pid_cor3 * pid_slope3
  hours <- seq(0, 3, length.out = 100)
  x <- ns4(
    hours
  )
  b0 <- fixef(fit)[1] + re_int
  b <- fixef(fit)[-1]
  y <- x[, 1] * (b[1] + re_s0) + x[, 2] * (b[2] + re_s1) + x[, 3] * (b[3] + re_s2) + x[, 4] * (b[4] + re_s3)
  if (intercept) y <- y + b0
  data.frame(
    x = hours,
    y = c(y),
    re_int
  )
}
color_scale <- list(
  scale_color_manual(
    "Intercept",
    values = c(
      "1 SD" = "#4daf4a",  
      "-1 SD" = "#e41a1c",
      "-2 SD" = "#377eb8",
      "0" = "#000000"
    )
  )
)

p_int_sd <- rbind(
get_trend(
    fit,
    data = dat,
    sd = 1
  ),
  get_trend(fit, data = dat),
  get_trend(
    fit,
    data = dat,
    sd = -1
  ),
  get_trend(
    fit,
    data = dat,
    sd = -2
  )
) |>
  mutate(
    intercept = factor(
      rep(
        c("1 SD", "0", "-1 SD", "-2 SD"),
        each = 100
      )
    )
  ) |>
  ggplot(
    aes(
      x,
      y,
      color = intercept,
      group = intercept
    )
  ) +
  geom_line() +
  labs(
    x = "Session duration (hours)",
    y = "Mood (0 - 100)"
  ) +
  color_scale

p_int_sd
```

The curves in  @fig-intercept-sds are easier to compare if we plot them without an intercept, as can be seen in @fig-no-intercept.

```{r}
#| label: fig-no-intercept
#| fig-cap: Estimated change curves for participants with an intercept equal to 0, -1SD, and -2SD.
p_int_sd_change <- rbind(
  get_trend(fit, data = dat, intercept = FALSE),
  get_trend(
    fit,
    data = dat,
    sd = -1,
    intercept = FALSE
  ),
  get_trend(
    fit,
    data = dat,
    sd = -2,
    intercept = FALSE
  )
) |>
  mutate(
    intercept = factor(
      rep(
        c("0", "-1 SD", "-2 SD"),
        each = 100
      )
    )
  ) |>
  ggplot(
    aes(
      x,
      y,
      color = intercept,
      group = intercept
    )
  ) +
  geom_line() +
  geom_ribbon(
    data = d_change,
    aes(
      x = hours,
      y = NULL,
      ymin = asymp.LCL,
      ymax = asymp.UCL,
      group = "0",
      color = NULL
    ),
    alpha = 0.15,
    show.legend = FALSE
  ) +
  labs(
    x = "Session duration (hours)",
    y = "Change in Mood (0 - 100)"
  ) +
  color_scale
p_int_sd_change
```

### Combined figured
Combined figure for manuscript(?).

```{r}
#| label: prod-figure
#| fig-width: 8
#| fig-height: 5
p_comb <- (p_curve) /
  (
    p_int_sd +
      p_int_sd_chage +
      plot_layout(guides = "collect") &
      theme(legend.position = "bottom")
  ) +
  plot_annotation(tag_levels = "A")

p_comb
ggsave(
  "figures/fig-session-change.png",
  p_comb,
  width = 8,
  height = 5.5
)
```

## Sensitivity checks
### Censoring

The main model used a censored normal distribution. A similar model would take a long time to fit here, so we'll refit the continuous time model using a censored normal distribution on a random subset of 1000 participants.

```{r}
#| label: brms-cens
set.seed(345345)
tmp <- filter(dat, pid %in% sample(unique(pid), 1000))

fit_b <- brm(
  mood | cens(cl) ~ ns4(hours) + (1 + ns4(hours) | pid) + (1 | ps),
  silent = 0,
  data = tmp,
  threads = 2,
  control = list(adapt_delta = .95),
  iter = 2000,
  prior = prior("normal(0,1)", class = "b"),
  file = "models/brms-ns4-cens"
)
```

We also re-fit the LMM to the reduced data.

```{r}
#| label: lmm-reduced

fit2 <- lmer(
  mood ~ ns4(hours) + (1 | ps) + (1 + ns4(hours) | pid),
  data = fit_b$data
)
```

If we plot the fitted spline from both models, we in @fig-cens-compare that the shape is almost identical but the censored model is shifted upwards.

```{r}
#| label: fig-cens-compare
#| fig-cap: Estimated trends with and without censoring.
d_brms <-
  emmeans(
    fit_b,
    ~hours,
    at = list(hours = seq(0, 3, length.out = 100))
  ) |>
  as.data.frame() |>
  mutate(model = "cens") |>
  select(
    hours,
    emmean,
    model,
    lower = "lower.HPD",
    upper = "upper.HPD"
  )
# SHIFT <- 0.017
SHIFT <- 0
d_lmm <-
  emmeans(
    fit2,
    ~hours,
    at = list(hours = seq(0, 3, length.out = 100))
  ) |>
  as.data.frame() |>
  select(
    hours,
    emmean,
    lower = "asymp.LCL",
    upper = "asymp.UCL"
  ) |>
  mutate(
    model = "lmm",
    emmean = emmean + SHIFT,
    upper = upper + SHIFT,
    lower = lower + SHIFT
  )
rbind(d_brms, d_lmm) |>
  ggplot(
    aes(
      hours,
      emmean,
      group = model,
      color = model
    )
  ) +
  geom_line() +
  geom_ribbon(
    aes(
      ymin = lower,
      ymax = upper,
      fill = model,
      color = NULL
    ),
    alpha = 0.25
  ) +
  labs(x = "Session duration (hours)")
```

Let's take a look at all model coefficients. @fig-model-cens-compare and @tbl-model-cens-compare show that ignoring censoring is perhaps not a huge problem.

```{r}
#| label: tidy-models
fit2_tidy <- tidy(fit2) |>
  mutate(
    model = "lmm",
    term = str_remove_all(term, "\\(|\\)")
  ) |>
  select(-statistic)
fit_b_tidy <- tidy(fit_b)
fit_b_tidy <- fit_b_tidy |>
  mutate(
    model = "cens",
    term = str_remove_all(term, "\\(|\\)")
  ) |>
  select(
    effect,
    group,
    term,
    estimate,
    std.error,
    model
  )
```

```{r}
#| label: fig-model-cens-compare
#| fig-cap: Model coefficients from an LMM with and without censoring

rbind(fit2_tidy, fit_b_tidy) |>
  ggplot(
    aes(
      y = term,
      estimate,
      color = model
    )
  ) +
  geom_point() +
  facet_wrap(
    ~effect,
    scale = "free",
    ncol = 1
  )
```

```{r}
#| label: tbl-model-cens-compare
#| tbl-cap: Model coefficients from an LMM with and without censoring

left_join(
  fit2_tidy,
  fit_b_tidy,
  by = join_by(term, group)
) |>
  mutate(
    diff = estimate.x - estimate.y,
    rel_diff = diff / estimate.y
  ) |>
  select(
    group,
    term,
    est_lmm = estimate.x,
    est_cens = estimate.y,
    diff,
    rel_diff
  ) |>
  kable(digits = 2)
```

We can also recreate @fig-intercept-sds using the Bayesian censored normal model and compare it to the LMM. As shown in @fig-bayes-int-sd, it's mostly the intercepts that are affected by the censoring.

```{r}
#| label: bayes-int-sd

get_trend_bayes <- function(sd) {
  re <- VarCorr(fit_b)
  ps <- as_draws_df(
    fit_b,
    c(
      "b_Intercept",
      "b_ns4hours1",
      "b_ns4hours2",
      "b_ns4hours3",
      "b_ns4hours4",
      "sd_pid__Intercept",
      "sd_pid__ns4hours1",
      "sd_pid__ns4hours2",
      "sd_pid__ns4hours3",
      "sd_pid__ns4hours4",
      "cor_pid__Intercept__ns4hours1",
      "cor_pid__Intercept__ns4hours2",
      "cor_pid__Intercept__ns4hours3",
      "cor_pid__Intercept__ns4hours4"
    )
  ) |>
    rename(
      pid_int = sd_pid__Intercept,
      pid_slope0 = sd_pid__ns4hours1,
      pid_slope1 = sd_pid__ns4hours2,
      pid_slope2 = sd_pid__ns4hours3,
      pid_slope3 = sd_pid__ns4hours4,
      pid_cor0 = cor_pid__Intercept__ns4hours1,
      pid_cor1 = cor_pid__Intercept__ns4hours2,
      pid_cor2 = cor_pid__Intercept__ns4hours3,
      pid_cor3 = cor_pid__Intercept__ns4hours4,
      b1 = b_ns4hours1,
      b2 = b_ns4hours2,
      b3 = b_ns4hours3,
      b4 = b_ns4hours4,
    ) |>
    mutate(
      re_int = sd * pid_int,
      b0 = b_Intercept + re_int,
      re_s0 = sd * pid_cor0 * pid_slope0,
      re_s1 = sd * pid_cor1 * pid_slope1,
      re_s2 = sd * pid_cor2 * pid_slope2,
      re_s3 = sd * pid_cor3 * pid_slope3
    )
  lapply(hours, function(hour) {
    x <- ns4(
      hour
    )
    ps |>
      mutate(
        y = b0 + x[, 1] * (b1 + re_s0) + x[, 2] * (b2 + re_s1) + x[, 3] * (b3 + re_s2) + x[, 4] * (b4 + re_s3)
      ) |>
      select(y) |>
      summarize(
        estimate = median(y),
        lwr = quantile(y, 0.025),
        upr = quantile(y, 0.975)
      ) |>
      mutate(
        hours = hour,
        re_int = mean(ps$re_int)
      )
  }) |>
    bind_rows() |>
    mutate(sd = sd)
}
res <- bind_rows(
  get_trend_bayes(0),
  get_trend_bayes(-1),
  get_trend_bayes(-2),
)
res <- res |>
  mutate(
    intercept = factor(
      sd,
      levels = c(0, -1, -2),
      labels = c("0", "-1 SD", "-2 SD")
    )
  )
# lmm for comparison
dat2 <- fit_b$data
res_lmm <- rbind(
  get_trend(fit2, data = dat2),
  get_trend(
    fit2,
    data = dat2,
    sd = -1
  ),
  get_trend(
    fit2,
    data = dat2,
    sd = -2
  )
) |>
  mutate(
    intercept = factor(
      rep(
        c("0", "-1 SD", "-2 SD"),
        each = 100
      )
    )
  ) |>
  select(
    estimate = y,
    hours = x,
    intercept
  )

```

```{r}
#| label: fig-bayes-int-sd
#| fig-cap: Bayesian censored normal model vs LMM (dashed) fit to subset of data

ggplot(
  res,
  aes(
    hours, estimate,
    group = intercept,
    color = intercept
  )
) +
  geom_line() +
  geom_line(
    data = res_lmm,
    linetype = "dashed"
  ) +
  geom_ribbon(
    aes(
      x = hours,
      y = NULL,
      ymin = lwr,
      ymax = upr,
      color = NULL
    ),
    alpha = 0.15,
    show.legend = FALSE
  ) + 
  color_scale
```

### Pretest x time continuous interaction
To check the robustness of the impact of baseline wellbeing, we also fit a separate interaction model to the subset of participants that began their session by answering a mood question. This allows us to explicitly model the interaction between the pretest values and the within-session change, without having to infer this from the multivariate normal distribution of the random effects.

```{r}
#| label: lmer-pre-interaction
dat2 <- dat |>
  filter(
    # Session has a wellbeing measure at time = 0
    duration[1] == 0,
    .by = c(pid, session)
  ) |> 
  mutate(
    pre = mood100[1],
    .by = c(pid, session)
  ) |> 
  filter(
    hours != 0
  )

path <- "models/lmm-pre-interaction.Rds"
if (file.exists(path)) {
  fit3 <- read_rds(path)
} else {
 fit3 <- lmer(
    mood100 ~ ns4(hours) * ns(pre, 5) + (1 + ns4(hours) | pid) + (1 | ps), 
    data = dat2
    )
  write_rds(fit3, path)
}
```

Plot the interaction at different pretest values. As shown in @fig-lmm-pre-interaction, this interaction model shows a pattern similar to @fig-intercept-sds where those who report lower mood at the beginning of a session also improve more. 
```{r}
#| label: fig-lmm-pre-interaction
#| fig-cap: Interaction between pretest mood and change, shown at -2 SD, -1, SD, the mean, and +1 SD.
pre_sum <- dat2 |> 
    summarise(
        pre_mean = median(pre), .by = pid
    ) |> 
    summarise(
        mean = mean(pre_mean), sd = sd(pre_mean)
    )
emm <- emmeans(
    fit3,
    ~hours + pre,
    at = list(
        hours = seq(0, 3, length.out = 100), 
        pre = c(
            pre_sum$mean - 2 * pre_sum$sd, 
            pre_sum$mean - pre_sum$sd, 
            pre_sum$mean, 
            pre_sum$mean + pre_sum$sd)
            ),
    lmer.df = "asymp"
)
p <- as.data.frame(emm) |> 
    mutate(
        pre = factor(round(pre))
    ) |> 
    ggplot(
        aes(hours, emmean, group = pre, color = pre)
    ) + 
    geom_line() +
    geom_ribbon(
        aes(ymin = asymp.LCL, ymax = asymp.UCL, fill = pre, color = NULL),
        alpha = 0.25
        ) 
p
ggsave("figures/fig-interaction-model.png", p, width = 7, height = 5)
```

(Ignore this change model --- not sure if it's worth recreating it).
```{r}
#| eval: false
#| label: fig-lmm-pre-interaction-change
emm <- emmeans(
    fit3,
    ~hours + pre,
    at = list(
        hours = seq(0, 3, length.out = 100), 
        pre = c(
            pre_sum$mean)
            ),
    lmer.df = "asymp"
)
d_change <- emmeans::contrast(
  emm,
  method = "trt.vs.ctrl",
  ref = "hours0 pre72.2106770833333"
) |>
  confint() |>
  as.data.frame() |>
  mutate(
    hours = hours[-1]
  )

p_change <- d_change |>
  ggplot(aes(hours, estimate)) +
  geom_line() +
  geom_ribbon(
    aes(
      ymin = asymp.LCL,
      ymax = asymp.UCL
    ),
    alpha = 0.25
  ) +
  labs(x = "Session duration (hours)")
p_change
```

For completeness we'll also fit the censored normal model with 1000 participants.

```{r}
#| label: cens-pre-interaction
set.seed(5454545)
tmp2 <- filter(dat2, pid %in% sample(unique(pid), 1000))

fit3_b <- brm(
  mood ~ ns4(hours) * ns(pre, 5) + (1 + ns4(hours) | pid) + (1 | ps),
  silent = 0,
  data = tmp2,
  threads = 2,
  #control = list(adapt_delta = .95),
  iter = 4000,
  file = "models/brms-ns4-interaction-cens"
)
```

Fit the LMM again to the smaller data.

```{r}
#| label: lmer-pre-interaction-small
 fit3_small <- lmer(
    mood ~ ns4(hours) * ns(pre, 5) + (1 + ns4(hours) | pid) + (1 | ps), 
    data = fit3_b$data
    )
```

We then compare the censored normal model with the LMM and recreate @fig-lmm-pre-interaction. As shown in @fig-cens-pre-interaction, there's almost no difference between the models. The LMMs CIs are slightly narrower near the intercept.
```{r}
#| label: fig-cens-pre-interaction
#| fig-cap: Censored normal vs LMM with a continuous by continuous interaction between within-session change and pretest mood.
get_emm_pre <- function(fit) {
  emmeans(
    fit,
    ~ hours + pre,
    # data = dat2,
    at = list(
      hours = seq(0, 3, length.out = 100),
      pre = c(
        pre_sum$mean - 2 * pre_sum$sd,
        pre_sum$mean - pre_sum$sd,
        pre_sum$mean,
        pre_sum$mean + pre_sum$sd
      )
    ),
    lmer.df = "asymp"
  )
}
pre_sum <- dat2 |>
  summarise(
    pre_mean = median(pre), .by = pid
  ) |>
  summarise(
    mean = mean(pre_mean), sd = sd(pre_mean)
  )
emm_cens <- get_emm_pre(fit3_b) |>
  as.data.frame() |>
  mutate(
    model = "cens"
  ) |>
  rename(
    lwr = lower.HPD,
    upr = upper.HPD
  )
emm_lmm <- get_emm_pre(fit3_small) |>
  as.data.frame() |>
  mutate(
    model = "LMM"
  ) |>
  rename(
    lwr = asymp.LCL,
    upr = asymp.UCL
  )

p <- bind_rows(emm_cens, emm_lmm) |>
  mutate(
    pre = factor(
      pre,
      levels = sort(unique(pre)),
      labels = c("-2 SD", "-1 SD", "Mean", "+1 SD")
    )
  ) |>
  {
    \(d)
    ggplot(
      data = d,
      aes(
        hours,
        emmean,
        group = interaction(pre, model),
        color = pre,
        linetype = model
      )
    ) +
      geom_line() +
      geom_ribbon(
        data = filter(d, model == "cens"),
        aes(
          ymin = lwr,
          ymax = upr,
          fill = pre,
          color = pre,
        ),
        color = NA,
        alpha = 0.25
      ) +
      geom_ribbon(
        data = filter(d, model == "LMM"),
        aes(
          ymin = lwr,
          ymax = upr,
          fill = pre,
          color = pre,
        ),
        fill = NA,
        alpha = 0.15
      ) +
      scale_color_brewer(palette = "Set1") +
      scale_fill_brewer(palette = "Set1")
  }()
p
```

### 2023-09-21 continuous-time.qmd
edit text
and some refactoring

@@
 }
-
-```
+```
+
+This model included `r number2(ngrps(fit)["pid"])` participants, with a total of `r number2(ngrps(fit)["ps"])` sessions and `r number2(nobs(fit))` observations.
 
@@
 #| fig-cap: Estimated mood during a play session
-
+#| fig-height: 3
+#| fig-width: 8
 hours <- seq(0, 3, length.out = 100)
@@
 #| fig-cap: Estimated difference in mood over a session compared to the start of a session
-
+#| fig-height: 3
+#| fig-width: 8
 d_change <- emmeans::contrast(
@@
 #| fig-cap: Estimated curves for participants with an intercept equal to 0, -1SD, and -2SD.
+#| fig-height: 3
+#| fig-width: 8
 get_trend <- function(
@@
 }
-color_scale <- list(
-  scale_color_manual(
-    "Intercept",
-    values = c(
+
+pre_labels <- c("-2 SD", "-1 SD", "Mean", "1 SD")
+pre_values <- c(-2, -1, 0, 1)
+colors <- c(
   "1 SD" = "#4daf4a",
@@
   "-2 SD" = "#377eb8",
-      "0" = "#000000"
-    )
-  )
-)
-
-p_int_sd <- rbind(
-get_trend(
-    fit,
-    data = dat,
-    sd = 1
-  ),
-  get_trend(fit, data = dat),
-  get_trend(
-    fit,
-    data = dat,
-    sd = -1
-  ),
-  get_trend(
-    fit,
-    data = dat,
-    sd = -2
-  )
-) |>
+  "Mean" = "#000000"
+)
+color_scale <- list(
+  scale_color_manual(
+    "Pre",
+    values = colors
+  ),
+    scale_fill_manual(
+    "Pre",
+    values = colors
+  )
+)
+
+p_int_sd <- map(
+  pre_values,
+  \(sd) get_trend(fit, dat, sd = sd)
+) |>
+  bind_rows() |>
   mutate(
@@
       rep(
-        c("1 SD", "0", "-1 SD", "-2 SD"),
+        pre_labels,
         each = 100
-      )
+      ),
+      levels = pre_labels
     )
@@
 #| fig-cap: Estimated change curves for participants with an intercept equal to 0, -1SD, and -2SD.
-p_int_sd_change <- rbind(
-  get_trend(fit, data = dat, intercept = FALSE),
-  get_trend(
-    fit,
-    data = dat,
-    sd = -1,
-    intercept = FALSE
-  ),
-  get_trend(
-    fit,
-    data = dat,
-    sd = -2,
-    intercept = FALSE
-  )
-) |>
+#| fig-height: 3
+#| fig-width: 8
+p_int_sd_change <- map(
+  pre_values,
+  \(sd) get_trend(fit, dat, sd = sd, intercept = FALSE)
+) |>
+  bind_rows() |>
   mutate(
@@
       rep(
-        c("0", "-1 SD", "-2 SD"),
+        pre_labels,
         each = 100
-      )
+      ),
+      levels = pre_labels
     )
@@
 
-### Combined figured
-Combined figure for manuscript(?).
-
-```{r}
-#| label: prod-figure
+```{r}
+#| label: fig-session-change
+#| echo: false
 #| fig-width: 8
@@
     p_int_sd +
-      p_int_sd_chage +
+      p_int_sd_change +
       plot_layout(guides = "collect") &
@@
 )
+```
+
+### Pretest x time continuous interaction
+Modeling the relationship between baseline wellbeing and change using random effects is limited.
+Here we'll fit a separate interaction model to the subset of participants that began their session by answering a mood question.
+This allows us to explicitly model the interaction between the pretest values and the within-session change, without having to infer this from the multivariate normal distribution of the random effects.
+This also allows us to model this as a non-linear relationship.
+
+```{r}
+#| label: lmer-pre-interaction
+dat2 <- dat |>
+  filter(
+    # Session has a wellbeing measure at time = 0
+    duration[1] == 0,
+    .by = c(pid, session)
+  ) |> 
+  mutate(
+    pre = mood100[1],
+    .by = c(pid, session)
+  ) |> 
+  filter(
+    hours != 0
+  )
+
+path <- "models/lmm-pre-interaction.Rds"
+if (file.exists(path)) {
+  fit3 <- read_rds(path)
+} else {
+ fit3 <- lmer(
+    mood100 ~ ns4(hours) * ns(pre, 5) + (1 + ns4(hours) | pid) + (1 | ps), 
+    data = dat2
+    )
+  write_rds(fit3, path)
+}
+```
+
+This model included `r number2(ngrps(fit3)["pid"])` participants, with a total of `r number2(ngrps(fit3)["ps"])` sessions and `r number2(nobs(fit3))` observations.
+
+Plot the interaction at different pretest values.
+As shown in @fig-lmm-pre-interaction, this interaction model shows a pattern similar to @fig-intercept-sds where those who report lower mood at the beginning of a session also improve more.
+```{r}
+#| label: fig-lmm-pre-interaction
+#| fig-cap: Interaction between pretest mood and change, shown at -2 SD, -1, SD, the mean, and +1 SD.
+#| fig-height: 3
+#| fig-width: 8
+pre_sum <- dat2 |>
+  summarise(
+    pre_mean = median(pre), .by = pid
+  ) |>
+  summarise(
+    mean = mean(pre_mean), sd = sd(pre_mean)
+  )
+emm <- emmeans(
+  fit3,
+  ~ hours + pre,
+  at = list(
+    hours = seq(0, 3, length.out = 100),
+    pre = pre_sum$mean + pre_values * pre_sum$sd
+  ),
+  lmer.df = "asymp"
+)
+p_pre_x <- as.data.frame(emm) |>
+  mutate(
+    pre = factor(
+      round(pre), 
+      labels = pre_labels
+      )
+  ) |>
+  ggplot(
+    aes(
+      hours,
+      emmean,
+      group = pre,
+      color = pre
+    )
+  ) +
+  geom_line() +
+  geom_ribbon(
+    aes(
+      ymin = asymp.LCL,
+      ymax = asymp.UCL,
+      fill = pre,
+      color = NULL
+    ),
+    alpha = 0.25,
+    show.legend = FALSE
+  ) + color_scale
+p_pre_x
+ggsave(
+  "figures/fig-interaction-model.png", 
+  p_pre_x, 
+  width = 7, 
+  height = 5
+)
+```
+
+Let's also calculate the within-session change for the interaction model.
+```{r}
+#| label: fig-lmm-pre-interaction-change
+#| fig-cap: Within-session change from the interaction model
+#| fig-height: 2.5
+#| fig-width: 8
+get_change_by_pre <- function(pre, label, fit, hours) {
+  emm <- emmeans(
+    fit,
+    ~ hours + pre,
+    at = list(
+      hours = hours,
+      pre = pre
+    ),
+    lmer.df = "asymp"
+  )
+  emmeans::contrast(
+    emm,
+    method = "trt.vs.ctrl",
+    ref = paste0("hours0 pre", pre)
+  ) |>
+    confint() |>
+    as.data.frame() |>
+    mutate(
+      hours = hours[-1],
+      pre = pre,
+      label = label
+    )
+}
+
+d_x_change <- data.frame(
+  pre = round(pre_sum$mean + pre_values * pre_sum$sd),
+  label = pre_labels
+) |>
+  pmap(\(pre, label) get_change_by_pre(
+    pre,
+    label,
+    fit3,
+    hours = seq(0, 3, length.out = 100)
+  )) |>
+  bind_rows()
+
+p_pre_x_change <- d_x_change |>
+  mutate(
+    label = factor(
+      label,
+      levels = pre_labels
+    )
+  ) |>
+  ggplot(
+    aes(
+      hours,
+      estimate,
+      group = label,
+      color = label,
+      fill = label
+    )
+  ) +
+  geom_line() +
+  geom_ribbon(
+    aes(
+      ymin = asymp.LCL,
+      ymax = asymp.UCL,
+      color = NULL
+    ),
+    alpha = 0.15,
+    show.legend = FALSE
+  ) +
+  geom_hline(
+    yintercept = 0,
+    linetype = "dotted"
+  ) +
+  labs(x = "Session duration (hours)") +
+  color_scale +
+  facet_wrap(~label, ncol = 4)
+p_pre_x_change
+```
+
+### Manuscript figures
+Create figures for the manuscript.
+```{r}
+#| fig-height: 4
+#| fig-width: 8
+p_tmp <- (
+  (
+    p_pre_x + 
+      facet_wrap(~pre, ncol = 4) +
+      labs(y = "Mood", x = "Hours")
+    ) / 
+    (
+      p_pre_x_change +
+        labs(y = "Change in mood", x = "Hours")
+      )
+    ) +
+   plot_annotation(tag_levels = "A") +
+  plot_layout(
+    heights = c(2, 2),
+    widths = c(8, 8),
+    guides = "collect"
+  ) & theme(legend.position = "bottom") 
+p_tmp
+```
+
+```{r}
+#| echo: false
+ggsave(
+  "figures/fig-pre-x-change.png", 
+  p_tmp, 
+  width = 8, 
+  height = 4
+  )
+```
+
+And this figure.
+
+```{r}
+#| fig-height: 4
+#| fig-width: 8
+p_tmp <- (
+  p_curve + labs(x = "Hours", y = "Mood")
+  ) / (
+  p_change + labs(x = "Hours", y = "Change in mood")
+  ) + plot_annotation(tag_levels = "A") 
+p_tmp
+```
+
+```{r}
+#| echo: false
+ggsave(
+  "figures/fig-change.png", 
+  p_tmp, 
+  width = 8, 
+  height = 4
+  )
+```
+
+And a table.
+```{r}
+#| label: tbl-pre-interaction-change
+#| tbl-cap: Estimated within-session change from the interaction model
+pre_sds <- c(1, 0, -1, -2)
+tbl_x_change <- data.frame(
+  pre = round(pre_sum$mean + pre_values * pre_sum$sd),
+  label = pre_labels
+) |>
+  pmap(\(pre, label) get_change_by_pre(
+    pre, 
+    label, 
+    fit3, 
+    hours = c(0, 10/60, 0.5, 1, 2))
+    ) |>
+  bind_rows() |> 
+  select(
+    label, 
+    pre, 
+    hours,
+    estimate, 
+    CI_lwr = asymp.LCL, 
+    CI_upr = asymp.UCL
+    )
+tbl_x_change |> 
+  kable(digits = 2)
 ```
@@
 #| label: lmm-reduced
-
 fit2 <- lmer(
@@
 #| fig-cap: Estimated trends with and without censoring.
+#| fig-height: 3
+#| fig-width: 8
 d_brms <-
@@
 }
-res <- bind_rows(
-  get_trend_bayes(0),
-  get_trend_bayes(-1),
-  get_trend_bayes(-2),
-)
-res <- res |>
+
+res <- map(
+  pre_values,
+  \(sd) get_trend_bayes(sd = sd)
+) |>
+  bind_rows() |>
   mutate(
@@
       sd,
-      levels = c(0, -1, -2),
-      labels = c("0", "-1 SD", "-2 SD")
-    )
-  )
+      levels = pre_values,
+      labels = pre_labels
+    )
+  )
+
 # lmm for comparison
 dat2 <- fit_b$data
-res_lmm <- rbind(
-  get_trend(fit2, data = dat2),
-  get_trend(
-    fit2,
-    data = dat2,
-    sd = -1
-  ),
-  get_trend(
-    fit2,
-    data = dat2,
-    sd = -2
-  )
-) |>
+res_lmm <- map(
+  pre_values,
+  \(sd) get_trend(fit2, dat2, sd = sd)
+) |>
+  bind_rows() |>
   mutate(
@@
 #| fig-cap: Bayesian censored normal model vs LMM (dashed) fit to subset of data
-
+#| fig-height: 3
+#| fig-width: 8
 ggplot(
@@
 
-### Pretest x time continuous interaction
-To check the robustness of the impact of baseline wellbeing, we also fit a separate interaction model to the subset of participants that began their session by answering a mood question.
-This allows us to explicitly model the interaction between the pretest values and the within-session change, without having to infer this from the multivariate normal distribution of the random effects.
-
-```{r}
-#| label: lmer-pre-interaction
-dat2 <- dat |>
-  filter(
-    # Session has a wellbeing measure at time = 0
-    duration[1] == 0,
-    .by = c(pid, session)
-  ) |> 
-  mutate(
-    pre = mood100[1],
-    .by = c(pid, session)
-  ) |> 
-  filter(
-    hours != 0
-  )
-
-path <- "models/lmm-pre-interaction.Rds"
-if (file.exists(path)) {
-  fit3 <- read_rds(path)
-} else {
- fit3 <- lmer(
-    mood100 ~ ns4(hours) * ns(pre, 5) + (1 + ns4(hours) | pid) + (1 | ps), 
-    data = dat2
-    )
-  write_rds(fit3, path)
-}
-```
-
-Plot the interaction at different pretest values.
-As shown in @fig-lmm-pre-interaction, this interaction model shows a pattern similar to @fig-intercept-sds where those who report lower mood at the beginning of a session also improve more.
-```{r}
-#| label: fig-lmm-pre-interaction
-#| fig-cap: Interaction between pretest mood and change, shown at -2 SD, -1, SD, the mean, and +1 SD.
-pre_sum <- dat2 |> 
-    summarise(
-        pre_mean = median(pre), .by = pid
-    ) |> 
-    summarise(
-        mean = mean(pre_mean), sd = sd(pre_mean)
-    )
-emm <- emmeans(
-    fit3,
-    ~hours + pre,
-    at = list(
-        hours = seq(0, 3, length.out = 100), 
-        pre = c(
-            pre_sum$mean - 2 * pre_sum$sd, 
-            pre_sum$mean - pre_sum$sd, 
-            pre_sum$mean, 
-            pre_sum$mean + pre_sum$sd)
-            ),
-    lmer.df = "asymp"
-)
-p <- as.data.frame(emm) |> 
-    mutate(
-        pre = factor(round(pre))
-    ) |> 
-    ggplot(
-        aes(hours, emmean, group = pre, color = pre)
-    ) + 
-    geom_line() +
-    geom_ribbon(
-        aes(ymin = asymp.LCL, ymax = asymp.UCL, fill = pre, color = NULL),
-        alpha = 0.25
-        ) 
-p
-ggsave("figures/fig-interaction-model.png", p, width = 7, height = 5)
-```
-
-(Ignore this change model --- not sure if it's worth recreating it).
-```{r}
-#| eval: false
-#| label: fig-lmm-pre-interaction-change
-emm <- emmeans(
-    fit3,
-    ~hours + pre,
-    at = list(
-        hours = seq(0, 3, length.out = 100), 
-        pre = c(
-            pre_sum$mean)
-            ),
-    lmer.df = "asymp"
-)
-d_change <- emmeans::contrast(
-  emm,
-  method = "trt.vs.ctrl",
-  ref = "hours0 pre72.2106770833333"
-) |>
-  confint() |>
-  as.data.frame() |>
-  mutate(
-    hours = hours[-1]
-  )
-
-p_change <- d_change |>
-  ggplot(aes(hours, estimate)) +
-  geom_line() +
-  geom_ribbon(
-    aes(
-      ymin = asymp.LCL,
-      ymax = asymp.UCL
-    ),
-    alpha = 0.25
-  ) +
-  labs(x = "Session duration (hours)")
-p_change
-```
-
-For completeness we'll also fit the censored normal model with 1000 participants.
+Lastly, we'll also fit the interaction model using a censored normal distribution, again with 1000 participants.
 
@@
 #| fig-cap: Censored normal vs LMM with a continuous by continuous interaction between within-session change and pretest mood.
+#| fig-height: 3
+#| fig-width: 8
 get_emm_pre <- function(fit) {
@@
 
-p <- bind_rows(emm_cens, emm_lmm) |>
+p_cens_pre_interaction <- bind_rows(emm_cens, emm_lmm) |>
   mutate(
@@
         color = NA,
-        alpha = 0.25
+        alpha = 0.25,
+        show.legend = FALSE
       ) +
@@
         fill = NA,
-        alpha = 0.15
+        alpha = 0.15,
+        show.legend = FALSE
       ) +
@@
   }()
-p
-```
+p_cens_pre_interaction
+```

### 2023-09-22 continuous-time.qmd
refactor

@@
 source("functions.R")
+source("R/fit_cached.R")
 ```
@@
 #| label: lmm-ns4
-
-path <- "models/lmm-ns4.Rds"
-if (file.exists(path)) {
-  fit <- read_rds(path)
-} else {
-  fit <- lmer(
+fit <- fit_cached(
+  "models/lmm-ns4.Rds",
+  lmer(
     mood100 ~ ns4(hours) + (1 | ps) + (1 + ns4(hours) | pid),
@@
   )
-  write_rds(fit, path)
-}
+)
 ```
@@
 #| fig-width: 8
-get_trend <- function(
-    fit,
-    data,
-    sd = 0,
-    intercept = TRUE) {
-  re <- as.data.frame(VarCorr(fit))
-  pid_int <- re[2, "sdcor"]
-  pid_slope0 <- re[3, "sdcor"]
-  pid_slope1 <- re[4, "sdcor"]
-  pid_slope2 <- re[5, "sdcor"]
-  pid_slope3 <- re[6, "sdcor"]
-  pid_cor0 <- re[7, "sdcor"]
-  pid_cor1 <- re[8, "sdcor"]
-  pid_cor2 <- re[9, "sdcor"]
-  pid_cor3 <- re[10, "sdcor"]
-  re_int <- sd * pid_int
-  re_s0 <- sd * pid_cor0 * pid_slope0
-  re_s1 <- sd * pid_cor1 * pid_slope1
-  re_s2 <- sd * pid_cor2 * pid_slope2
-  re_s3 <- sd * pid_cor3 * pid_slope3
-  hours <- seq(0, 3, length.out = 100)
-  x <- ns4(
-    hours
-  )
-  b0 <- fixef(fit)[1] + re_int
-  b <- fixef(fit)[-1]
-  y <- x[, 1] * (b[1] + re_s0) + x[, 2] * (b[2] + re_s1) + x[, 3] * (b[3] + re_s2) + x[, 4] * (b[4] + re_s3)
-  if (intercept) y <- y + b0
-  data.frame(
-    x = hours,
-    y = c(y),
-    re_int
-  )
-}
+source("R/get_trend.R")
 
 pre_labels <- c("-2 SD", "-1 SD", "Mean", "1 SD")
-pre_values <- c(-2, -1, 0, 1)
+pre_sds <- c(-2, -1, 0, 1)
 colors <- c(
@@
 p_int_sd <- map(
-  pre_values,
+  pre_sds,
   \(sd) get_trend(fit, dat, sd = sd)
@@
 p_int_sd_change <- map(
-  pre_values,
-  \(sd) get_trend(fit, dat, sd = sd, intercept = FALSE)
+  pre_sds,
+  \(sd) get_trend(
+    fit,
+    dat,
+    sd = sd,
+    intercept = FALSE
+  )
 ) |>
@@
 #| label: lmer-pre-interaction
-dat2 <- dat |>
+dat_pre <- dat |>
   filter(
@@
 
-path <- "models/lmm-pre-interaction.Rds"
-if (file.exists(path)) {
-  fit3 <- read_rds(path)
-} else {
- fit3 <- lmer(
+fit3 <- fit_cached(
+  "models/lmm-pre-interaction.Rds",
+  lmer(
     mood100 ~ ns4(hours) * ns(pre, 5) + (1 + ns4(hours) | pid) + (1 | ps), 
-    data = dat2
-    )
-  write_rds(fit3, path)
-}
+    data = dat_pre
+    )
+)
 ```
@@
 This model included `r number2(ngrps(fit3)["pid"])` participants, with a total of `r number2(ngrps(fit3)["ps"])` sessions and `r number2(nobs(fit3))` observations.
+
+Calculate pre values we are interested in plotting.
+
+```{r}
+#| label: pre-sum
+pre_sum <- dat_pre |>
+  summarise(
+    pre_mean = median(pre),
+    .by = pid
+  ) |>
+  summarise(
+    mean = mean(pre_mean),
+    sd = sd(pre_mean)
+  )
+kable(pre_sum, digits = 2)
+```
+
+```{r}
+#| label: pre-sd-values
+pre_values <- pre_sum$mean + pre_sds * pre_sum$sd
+data.frame(
+  pre = pre_values,
+  label = pre_labels
+  ) |> 
+  kable(digits = 2)
+```
 
@@
     hours = seq(0, 3, length.out = 100),
-    pre = pre_sum$mean + pre_values * pre_sum$sd
+    pre = pre_values
   ),
@@
 #| fig-width: 8
-get_change_by_pre <- function(pre, label, fit, hours) {
-  emm <- emmeans(
-    fit,
-    ~ hours + pre,
-    at = list(
-      hours = hours,
-      pre = pre
-    ),
-    lmer.df = "asymp"
-  )
-  emmeans::contrast(
-    emm,
-    method = "trt.vs.ctrl",
-    ref = paste0("hours0 pre", pre)
-  ) |>
-    confint() |>
-    as.data.frame() |>
-    mutate(
-      hours = hours[-1],
-      pre = pre,
-      label = label
-    )
-}
-
+source("R/get_change_by_pre.R")
 d_x_change <- data.frame(
-  pre = round(pre_sum$mean + pre_values * pre_sum$sd),
+  pre = round(pre_values),
   label = pre_labels
@@
 tbl_x_change <- data.frame(
-  pre = round(pre_sum$mean + pre_values * pre_sum$sd),
+  pre = round(pre_values),
   label = pre_labels
@@
 set.seed(345345)
-tmp <- filter(dat, pid %in% sample(unique(pid), 1000))
+tmp <- filter(
+  dat,
+  pid %in% sample(unique(pid), 1000)
+)
 
@@
 #| fig-cap: Model coefficients from an LMM with and without censoring
-
 rbind(fit2_tidy, fit_b_tidy) |>
@@
 #| tbl-cap: Model coefficients from an LMM with and without censoring
-
 left_join(
@@
 #| label: bayes-int-sd
-
-get_trend_bayes <- function(sd) {
-  re <- VarCorr(fit_b)
-  ps <- as_draws_df(
-    fit_b,
-    c(
-      "b_Intercept",
-      "b_ns4hours1",
-      "b_ns4hours2",
-      "b_ns4hours3",
-      "b_ns4hours4",
-      "sd_pid__Intercept",
-      "sd_pid__ns4hours1",
-      "sd_pid__ns4hours2",
-      "sd_pid__ns4hours3",
-      "sd_pid__ns4hours4",
-      "cor_pid__Intercept__ns4hours1",
-      "cor_pid__Intercept__ns4hours2",
-      "cor_pid__Intercept__ns4hours3",
-      "cor_pid__Intercept__ns4hours4"
-    )
-  ) |>
-    rename(
-      pid_int = sd_pid__Intercept,
-      pid_slope0 = sd_pid__ns4hours1,
-      pid_slope1 = sd_pid__ns4hours2,
-      pid_slope2 = sd_pid__ns4hours3,
-      pid_slope3 = sd_pid__ns4hours4,
-      pid_cor0 = cor_pid__Intercept__ns4hours1,
-      pid_cor1 = cor_pid__Intercept__ns4hours2,
-      pid_cor2 = cor_pid__Intercept__ns4hours3,
-      pid_cor3 = cor_pid__Intercept__ns4hours4,
-      b1 = b_ns4hours1,
-      b2 = b_ns4hours2,
-      b3 = b_ns4hours3,
-      b4 = b_ns4hours4,
-    ) |>
-    mutate(
-      re_int = sd * pid_int,
-      b0 = b_Intercept + re_int,
-      re_s0 = sd * pid_cor0 * pid_slope0,
-      re_s1 = sd * pid_cor1 * pid_slope1,
-      re_s2 = sd * pid_cor2 * pid_slope2,
-      re_s3 = sd * pid_cor3 * pid_slope3
-    )
-  lapply(hours, function(hour) {
-    x <- ns4(
-      hour
-    )
-    ps |>
-      mutate(
-        y = b0 + x[, 1] * (b1 + re_s0) + x[, 2] * (b2 + re_s1) + x[, 3] * (b3 + re_s2) + x[, 4] * (b4 + re_s3)
-      ) |>
-      select(y) |>
-      summarize(
-        estimate = median(y),
-        lwr = quantile(y, 0.025),
-        upr = quantile(y, 0.975)
-      ) |>
-      mutate(
-        hours = hour,
-        re_int = mean(ps$re_int)
-      )
-  }) |>
-    bind_rows() |>
-    mutate(sd = sd)
-}
-
+source("R/get_trend_bayes.R")
 res <- map(
-  pre_values,
+  pre_sds,
   \(sd) get_trend_bayes(sd = sd)
@@
       sd,
-      levels = pre_values,
+      levels = pre_sds,
       labels = pre_labels
@@
 # lmm for comparison
-dat2 <- fit_b$data
 res_lmm <- map(
-  pre_values,
-  \(sd) get_trend(fit2, dat2, sd = sd)
+  pre_sds,
+  \(sd) get_trend(fit2, fit_b$data, sd = sd)
 ) |>
@@
 set.seed(5454545)
-tmp2 <- filter(dat2, pid %in% sample(unique(pid), 1000))
+tmp_pre <- filter(dat_pre, pid %in% sample(unique(pid), 1000))
 
@@
   silent = 0,
-  data = tmp2,
+  data = tmp_pre,
   threads = 2,
@@
     ~ hours + pre,
-    # data = dat2,
     at = list(
@@
 }
-pre_sum <- dat2 |>
+pre_sum <- tmp_pre |>
   summarise(

### 2023-09-22 R/fit_cached.R

@@ -1,0 +1,18 @@
+#' Explicitly cache model
+#' 
+#' A function that will save a model as an .Rds file,
+#' and reload the file without having to fit the model
+#' on the next run.
+#' 
+#' @param path string; path where the model is saved.
+#' Should end with .rds.
+#' @param model a model call to fit
+fit_cached <- function(path, model) {
+  if (file.exists(path)) {
+    fit <- read_rds(path)
+  } else {
+    fit <- model
+    write_rds(fit, path)
+  }
+  fit
+}

### 2023-09-22 R/get_change_by_pre.R

@@ -1,0 +1,29 @@
+#' Within-session change by pretest values
+#' 
+#' @param pre the baseline mood value to calculate the change for.
+#' @param label indicated the label to use for pre value.
+#' @param fit the model object.
+#' @param hours the range of hours to calculate the vales for
+get_change_by_pre <- function(pre, label, fit, hours) {
+  emm <- emmeans(
+    fit,
+    ~ hours + pre,
+    at = list(
+      hours = hours,
+      pre = pre
+    ),
+    lmer.df = "asymp"
+  )
+  emmeans::contrast(
+    emm,
+    method = "trt.vs.ctrl",
+    ref = paste0("hours0 pre", pre)
+  ) |>
+    confint() |>
+    as.data.frame() |>
+    mutate(
+      hours = hours[-1],
+      pre = pre,
+      label = label
+    )
+}

### 2023-09-22 R/get_trend_bayes.R

@@ -1,0 +1,74 @@
+#' Get trend for different random intercept values from brms.
+#' @param sd integer to scale the random intercept SD by.
+#' @details 
+#' The model object is hard coded to be `fit_b`.
+#' 
+#' The curve is calculated over all posterior samples, and
+#' summaries are based on medians and 2.5% and 97.5% percentile
+#' intervals.
+get_trend_bayes <- function(sd) {
+  re <- VarCorr(fit_b)
+  ps <- as_draws_df(
+    fit_b,
+    c(
+      "b_Intercept",
+      "b_ns4hours1",
+      "b_ns4hours2",
+      "b_ns4hours3",
+      "b_ns4hours4",
+      "sd_pid__Intercept",
+      "sd_pid__ns4hours1",
+      "sd_pid__ns4hours2",
+      "sd_pid__ns4hours3",
+      "sd_pid__ns4hours4",
+      "cor_pid__Intercept__ns4hours1",
+      "cor_pid__Intercept__ns4hours2",
+      "cor_pid__Intercept__ns4hours3",
+      "cor_pid__Intercept__ns4hours4"
+    )
+  ) |>
+    rename(
+      pid_int = sd_pid__Intercept,
+      pid_slope0 = sd_pid__ns4hours1,
+      pid_slope1 = sd_pid__ns4hours2,
+      pid_slope2 = sd_pid__ns4hours3,
+      pid_slope3 = sd_pid__ns4hours4,
+      pid_cor0 = cor_pid__Intercept__ns4hours1,
+      pid_cor1 = cor_pid__Intercept__ns4hours2,
+      pid_cor2 = cor_pid__Intercept__ns4hours3,
+      pid_cor3 = cor_pid__Intercept__ns4hours4,
+      b1 = b_ns4hours1,
+      b2 = b_ns4hours2,
+      b3 = b_ns4hours3,
+      b4 = b_ns4hours4,
+    ) |>
+    mutate(
+      re_int = sd * pid_int,
+      b0 = b_Intercept + re_int,
+      re_s0 = sd * pid_cor0 * pid_slope0,
+      re_s1 = sd * pid_cor1 * pid_slope1,
+      re_s2 = sd * pid_cor2 * pid_slope2,
+      re_s3 = sd * pid_cor3 * pid_slope3
+    )
+  lapply(hours, function(hour) {
+    x <- ns4(
+      hour
+    )
+    ps |>
+      mutate(
[…]

### 2023-09-22 continuous-time.qmd
show quantiles

@@
 ## Main model
-Define a spline that uses 4 degrees of freedom.
+Define a spline that uses 4 degrees of freedom. We base this on the quantiles of `hours`.
+
+```{r}
+#| label: spline-percentiles
+quantile(dat$hours)
+```
+
+Not all participants contributed an equal number of observations. We'll calculate the max session length per participant.
+
+```{r}
+#| label: tbl-session-max-prop
+#| tbl-cap: Proportion of participants that had at least on session last longer than gt_* hours
+dat |> 
+  summarize(
+    max = max(hours), .by = c(pid)
+  ) |> 
+  summarise(
+    gt_0.33 = mean(max > 0.33),
+    gt_0.71 = mean(max > 0.716),
+    gt_1 = mean(max > 1),
+    gt_2 = mean(max > 2),
+  ) |> 
+  kable(digits = 2)
+```
+
+@tbl-session-max-prop shows that 66% of the participant have at least one session that lasted for over 0.71 hours. So we will base our knot placement on these quantiles.
 
@@
 #| label: define-spline
-
 ns4 <- function(x) {

### 2023-09-22 continuous-time.qmd
tweak figures

@@
         size = rel(0.75)
-      ),
-      panel.grid = element_blank()
+      )
+    )
+)
+scale_x_hours <- list(
+  scale_x_continuous(
+  breaks = c(0, 15/60, 0.5, 45/60, 1, 2, 3), 
+  labels = c(0, "15m", "30m", "45m", "1h", "2h", "3h")
   )
@@
     y = "Mood (0 - 100)"
-  )
+  ) +
+  scale_x_hours
 p_curve
@@
   lims(y = c(0, 6)) +
-  labs(x = "Session duration (hours)")
+  labs(x = "Session duration (hours)") +
+  scale_x_hours
 p_change
@@
   ) +
-  color_scale
+  color_scale +
+  scale_x_hours
 
@@
   ) +
-  color_scale
+  color_scale +
+  scale_x_hours
 p_int_sd_change
@@
 #| fig-width: 8
-pre_sum <- dat2 |>
-  summarise(
-    pre_mean = median(pre), .by = pid
-  ) |>
-  summarise(
-    mean = mean(pre_mean), sd = sd(pre_mean)
-  )
 emm <- emmeans(
@@
     show.legend = FALSE
-  ) + color_scale
+  ) +
+  color_scale +
+  scale_x_hours
+
 p_pre_x
@@
   color_scale +
+  scale_x_hours +
   facet_wrap(~label, ncol = 4) 
@@
 ```{r}
+#| label: fig-ms-interaction
 #| fig-height: 4
 #| fig-width: 8
+
+scale_x_hours2 <- list(
+  scale_x_continuous(
+  breaks = c(0, 15/60, 0.5, 45/60, 1, 2, 3), 
+  labels = c(0, "", "30m", "", "1h", "2h", "3h")
+  )
+)
 p_tmp <- (
@@
       facet_wrap(~pre, ncol = 4) +
-      labs(y = "Mood", x = "Hours")
+      scale_x_hours2 +
+      labs(y = "Mood", x = NULL) +
+      theme(plot.margin = margin(0, 0, 0, 0, "pt"))
     ) / 
@@
       p_pre_x_change +
-        labs(y = "Change in mood", x = "Hours")
+        scale_x_hours2 +
+        labs(y = "Change", x = "Duration") +
+        theme(plot.margin = margin(0, 0, 0, 0, "pt"))
       ) 
@@
 ```{r}
+#| label: ggsave-fig-ms-interaction
 #| echo: false
@@
   p_tmp, 
-  width = 8, 
-  height = 4
+  width = 7, 
+  height = 3.5
   )
@@
 ```{r}
+#| label: fig-ms-no-interaction
 #| fig-height: 4
@@
 p_tmp <- (
-  p_curve + labs(x = "Hours", y = "Mood")
+  p_curve + labs(x = NULL, y = "Mood")
   ) / (
-  p_change + labs(x = "Hours", y = "Change in mood")
+  p_change + labs(x = "Duration", y = "Change")
   ) + plot_annotation(tag_levels = "A") 
@@
 ```{r}
+#| label: ggsave-fig-ms-no-interaction
 #| echo: false
@@
   p_tmp, 
-  width = 8, 
-  height = 4
+  width = 6, 
+  height = 3
   )

### 2023-09-25 continuous-time.qmd
more sensitivity analyses

@@
 ```
+
+### Functional form of `hours` and `pre`
+Here we'll investigate the impact of the assumed functional form of `hours` and `pre`.
+As can be seen in @fig-lmm-pre-interaction-sens-pre, that impact of the assumed functional form of `pre` has very little impact on the results.
+
+```{r}
+#| label: fig-lmm-pre-interaction-sens-pre
+#| fig-cap: Estimated trends for various functional forms of `pre`
+tmp_models <- list()
+tmp_models[["linear"]] <- fit_cached(
+  "models/lmm-pre-interaction-sens-pre-linear.Rds",
+  lmer(
+    mood100 ~ ns4(hours) * pre + (1 + ns4(hours) | pid) + (1 | ps),
+    data = dat_pre
+  )
+)
+tmp_models[["ns2"]] <-
+  fit_cached(
+    "models/lmm-pre-interaction-sens-pre-ns2.Rds",
+    lmer(
+      mood100 ~ ns4(hours) * ns(pre, 2) + (1 + ns4(hours) | pid) + (1 | ps),
+      data = dat_pre,
+    )
+  )
+tmp_models[["ns3"]] <-
+  fit_cached(
+    "models/lmm-pre-interaction-sens-pre-ns3.Rds",
+    lmer(
+      mood100 ~ ns4(hours) * ns(pre, 3) + (1 + ns4(hours) | pid) + (1 | ps),
+      data = dat_pre,
+    )
+  )
+tmp_models[["ns4"]] <-
+  fit_cached(
+    "models/lmm-pre-interaction-sens-pre-ns4.Rds",
+    lmer(
+      mood100 ~ ns4(hours) * ns(pre, 4) + (1 + ns4(hours) | pid) + (1 | ps),
+      data = dat_pre,
+    )
+  )
+tmp_models[["ns5"]] <- fit3
+
+get_trend_by_model <- function(fit, model_label) {
+  emm <- emmeans(
+    fit,
+    ~ hours + pre,
+    at = list(
+      hours = seq(0, 3, length.out = 100),
+      pre = pre_values
+    ),
+    lmer.df = "asymp"
+  )
+  as.data.frame(emm) |>
+    mutate(
+      pre = factor(
+        round(pre),
+        labels = pre_labels,
+      ),
+      model = model_label
+    )
+}
+tmp_p <- map2(
+  tmp_models,
+  names(tmp_models),
+  \(fit, lab) get_trend_by_model(fit, lab)
+) |>
+  bind_rows()
+p_pre_x_sens <- tmp_p |>
+  ggplot(
+    aes(
+      hours,
+      emmean,
+      group = interaction(model, pre),
+      color = model,
+      linetype = pre
+    )
+  ) +
+  geom_line() +
+  geom_ribbon(
+    aes(
+      ymin = asymp.LCL,
+      ymax = asymp.UCL,
+      fill = model,
+      color = NULL
+    ),
+    alpha = 0.10,
+    show.legend = FALSE
+  ) +
+  scale_color_brewer(palette = "Set1") +
+  scale_fill_brewer(palette = "Set1") +
+  scale_x_hours
+p_pre_x_sens
+```
+
+Next, we'll lok at `hours`. In @fig-lmm-pre-interaction-sens-hours we see that the results aren't fundamentally different using splines with less degrees of freedom, or even a linear term.
+
+```{r}
+#| label: fig-lmm-pre-interaction-sens-hours
+#| fig-cap: Estimated trends for various functional forms of `hours`  
+tmp_models2 <- list()
+tmp_models2[["linear"]] <- fit_cached(
+  "models/lmm-pre-interaction-sens-hours-linear.Rds",
+  lmer(
+    mood100 ~ hours * ns(pre, 5) + (1 + hours | pid) + (1 | ps),
+    data = dat_pre
+  )
+)
+tmp_models2[["ns2"]] <-
+  fit_cached(
+    "models/lmm-pre-interaction-sens-hours-ns2.Rds",
+    lmer(
+      mood100 ~ ns(hours, 2) * ns(pre, 5) + (1 + ns(hours, 2) | pid) + (1 | ps),
+      data = dat_pre,
+    )
+  )
+tmp_models2[["ns3"]] <- fit_cached(
+  "models/lmm-pre-interaction-sens-hours-n3.Rds",
+  lmer(
+    mood100 ~ ns(hours, 3) * ns(pre, 5) + (1 + ns(hours, 3) | pid) + (1 | ps),
+    data = dat_pre,
+  )
+)
+tmp_models2[["poly2"]] <- fit_cached(
+  "models/lmm-pre-interaction-sens-hours-poly2.Rds",
+  lmer(
+    mood100 ~ poly(hours, 2) * ns(pre, 5) + (1 + poly(hours, 2) | pid) + (1 | ps),
+    data = dat_pre,
+  )
+)
+tmp_models2[["ns4"]] <- fit3
+tmp_p <- map2(
+  tmp_models2,
+  names(tmp_models2),
+  \(fit, lab) get_trend_by_model(fit, lab)
+) |>
+  bind_rows()
+p_pre_x_sens %+% tmp_p 
+```

### 2023-09-25 continuous-time.qmd
cleaning

@@
 library(brms)
-library(ggstance)
-library(ggdist)
 library(patchwork)
@@
 #| label: load-data
-dat <- read_csv(
-  "data/study_prompt_answered.csv",
-  col_select = c(
-    pid,
-    time = Time_utc,
-    duration = CurrentSessionLength,
-    prompt = LastStudyPromptType,
-    mood = response
-  )
-) |>
-  mutate(
-    pid = factor(pid),
-    mood = mood / 1000
-  )
-
-# Create sessions using heuristics below
-dat <- dat |>
-  arrange(pid, time) |>
-  mutate(
-    i = row_number(),
-    # Observation is in a new session if this...
-    new_session =
-    # is the first total observation,
-      i == 1 |
-        # session duration is zero,
-        (duration == 0) |
-        # session is shorter than previous session (ticker has reset),
-        (duration < lag(duration, default = 0)) |
-        # or if more than 30 minutes since previous observation.
-        (time > lag(time) + minutes(30)),
-    # Session number is the cumulative sum of new sessions...
-    session = cumsum(new_session),
-    # ... per person
-    .by = pid
-  ) |>
-  select(-c(i, new_session))
-
+dat <- read_rds("data/data-raw.rds")
+if (!is.na(N_SUBSET)) {
+  pid_subset <- sample(unique(dat$pid), N_SUBSET)
+  dat <- filter(dat, pid %in% pid_subset)
+}
 dat <- dat |>
@@
 dat |> 
-  summarize(
+  summarise(
     max = max(hours), .by = c(pid)
@@
 #| label: define-spline
+#| code-fold: show
 ns4 <- function(x) {
@@
 #| label: lmm-ns4
+#| code-fold: show
 fit <- fit_cached(
@@
 
-```{r}
-#| label: fig-session-change
-#| echo: false
-#| fig-width: 8
-#| fig-height: 5
-p_comb <- (p_curve) /
-  (
-    p_int_sd +
-      p_int_sd_change +
-      plot_layout(guides = "collect") &
-      theme(legend.position = "bottom")
-  ) +
-  plot_annotation(tag_levels = "A")
-
-p_comb
-ggsave(
-  "figures/fig-session-change.png",
-  p_comb,
-  width = 8,
-  height = 5.5
-)
-```
-
 ### Pretest x time continuous interaction
@@
 #| label: lmer-pre-interaction
+#| code-fold: show
 dat_pre <- dat |>
@@
 ```{r}
-#| label: pre-sum
+#| label: tbl-pre-sum
+#| tbl-cap: Average mood levels the start of a session
 pre_sum <- dat_pre |>
@@
 ```{r}
-#| label: pre-sd-values
+#| label: tbl-pre-sd-values
+#| tbl-cap: Distribution of average mood levels at the start of a session
 pre_values <- pre_sum$mean + pre_sds * pre_sum$sd
@@
 #| fig-width: 8
+#| fig-cap: Figure included in manuscript
 
@@
 #| fig-width: 8
+#| fig-cap: Figure shown in manuscript
 p_tmp <- (
@@
 
-The main model used a censored normal distribution.
-A similar model would take a long time to fit here, so we'll refit the continuous time model using a censored normal distribution on a random subset of 1000 participants.
+```{r}
+#| label: n-pids-sample
+#| echo: false
+n_pids_sample <- case_when(
+  is.na(N_SUBSET) ~ 1000,
+  !is.na(N_SUBSET) ~ min(N_SUBSET, 1000)
+)
+```
+
+The main discrete time model used a censored normal distribution.
+A similar model would take a long time to fit here, so we'll refit the continuous time model using a censored normal distribution on a random subset of `r n_pids_sample` participants.
 
@@
   dat,
-  pid %in% sample(unique(pid), 1000)
+  pid %in% sample(unique(pid), n_pids_sample)
 )
@@
 
-Lastly, we'll also fit the interaction model using a censored normal distribution, again with 1000 participants.
+Lastly, we'll also fit the interaction model using a censored normal distribution, again with `r n_pids_sample` participants.
 
@@
 set.seed(5454545)
-tmp_pre <- filter(dat_pre, pid %in% sample(unique(pid), 1000))
+tmp_pre <- filter(dat_pre, pid %in% sample(unique(pid), n_pids_sample))
 
@@
 
-Next, we'll lok at `hours`. In @fig-lmm-pre-interaction-sens-hours we see that the results aren't fundamentally different using splines with less degrees of freedom, or even a linear term.
+Next, we'll look at `hours`. In @fig-lmm-pre-interaction-sens-hours we see that the results aren't fundamentally different using splines with less degrees of freedom, or even a linear term.
 

### 2023-09-25 analyze.qmd
move file

@@
 library(tidyverse)
-source("functions.R")
+source("R/functions.R")
 ```

### 2023-09-25 continuous-time.qmd
move file

@@
 library(knitr)
-source("functions.R")
+source("R/functions.R")
 source("R/fit_cached.R")

### 2023-09-25 continuous-time.qmd
deps

--- 2023-09-25 continuous-time.qmd
@@
 library(knitr)
-source("functions.R")
+library(rstanarm)
+source("R/functions.R")
 source("R/fit_cached.R")

### 2023-09-25 continuous-time.qmd
subset data

@@
 dat <- read_rds("data/data-raw.rds")
-if (!is.na(N_SUBSET)) {
-  pid_subset <- sample(unique(dat$pid), N_SUBSET)
-  dat <- filter(dat, pid %in% pid_subset)
-}
 dat <- dat |>
@@
   )
+if (!is.na(N_SUBSET)) {
+  pid_subset <- sample(unique(dat$pid), N_SUBSET)
+  dat <- filter(dat, pid %in% pid_subset) |> 
+    droplevels()
+}
 ```
@@
     hours != 0
-  )
+  ) |> 
+  droplevels()
 
@@
 #| label: brms-cens
+#| results: hide
 set.seed(345345)
@@
 
-If we plot the fitted spline from both models, we in @fig-cens-compare that the shape is almost identical but the censored model is shifted upwards.
+If we plot the fitted spline from both models, we see in @fig-cens-compare that the shape is almost identical but the censored model is shifted upwards.
 
@@
 
-Lastly, we'll also fit the interaction model using a censored normal distribution, again with `r n_pids_sample` participants.
+Lastly, we'll also fit the interaction model using a censored normal distribution, with `r min(n_pids_sample, nlevels(dat_pre$pid))` participants.
 
@@
 #| label: cens-pre-interaction
+#| results: hide
 set.seed(5454545)
-tmp_pre <- filter(dat_pre, pid %in% sample(unique(pid), n_pids_sample))
+tmp_pre <- filter(
+    dat_pre,
+    pid %in% sample(
+        unique(pid),
+        # avoid taking sample larger than population 
+        min(
+            n_pids_sample,
+            nlevels(pid)
+        )
+    )
+)
 

### 2023-09-25 R/get_trend.R
Continuous time analysis (#7)
add continuous time analyses

---------

Co-authored-by: Kristoffer Magnusson <1629860+rpsychologist@users.noreply.github.com>

@@ -1,0 +1,46 @@
+#' Get trend for different random intercept values
+#' 
+#' sdfsdf
+#'
+#' @param fit lme4 model object.
+#' @param data the data.
+#' @param sd integer to scale the random intercept SD by.
+#' @param intercept boolean; indicates whether to include the intercept in the
+#' response calculation.
+#' 
+#' @details
+#' This only works with the model `mood ~ ns4(hours) + (1 + ns4(hours) | pid)`
+get_trend <- function(
+    fit,
+    data,
+    sd = 0,
+    intercept = TRUE) {
+  re <- as.data.frame(VarCorr(fit))
+  pid_int <- re[2, "sdcor"]
+  pid_slope0 <- re[3, "sdcor"]
+  pid_slope1 <- re[4, "sdcor"]
+  pid_slope2 <- re[5, "sdcor"]
+  pid_slope3 <- re[6, "sdcor"]
+  pid_cor0 <- re[7, "sdcor"]
+  pid_cor1 <- re[8, "sdcor"]
+  pid_cor2 <- re[9, "sdcor"]
+  pid_cor3 <- re[10, "sdcor"]
+  re_int <- sd * pid_int
+  re_s0 <- sd * pid_cor0 * pid_slope0
+  re_s1 <- sd * pid_cor1 * pid_slope1
+  re_s2 <- sd * pid_cor2 * pid_slope2
+  re_s3 <- sd * pid_cor3 * pid_slope3
+  hours <- seq(0, 3, length.out = 100)
+  x <- ns4(
+    hours
+  )
+  b0 <- fixef(fit)[1] + re_int
+  b <- fixef(fit)[-1]
+  y <- x[, 1] * (b[1] + re_s0) + x[, 2] * (b[2] + re_s1) + x[, 3] * (b[3] + re_s2) + x[, 4] * (b[4] + re_s3)
+  if (intercept) y <- y + b0
+  data.frame(
+    x = hours,
+    y = c(y),
+    re_int
+  )
+}

### 2023-09-25 continuous-time.qmd
use percentiles

@@
 #| tbl-cap: Distribution of average mood levels at the start of a session
-pre_values <- pre_sum$mean + pre_sds * pre_sum$sd
+#pre_values <- pre_sum$mean + pre_sds * pre_sum$sd
+
+pre <- dat_pre |> 
+  summarise(pre = pre[1], .by = ps) |> 
+  pull(pre)
+pre_values <- quantile(pre, c(0.05, .25, 0.5, 0.75))
+pre_labels <- c("5%", "25%", "Median", "75%")
+colors <- c(
+  "5%" = "#4daf4a",
+  "25%" = "#e41a1c",
+  "75%" = "#377eb8",
+  "Median" = "#000000"
+)
+color_scale_x <- list(
+  scale_color_manual(
+    "Pre",
+    values = colors
+  ),
+    scale_fill_manual(
+    "Pre",
+    values = colors
+  )
+)
 data.frame(
@@
   ) +
-  color_scale +
+  color_scale_x +
   scale_x_hours
@@
   labs(x = "Session duration (hours)") +
-  color_scale +
+  color_scale_x +
   scale_x_hours +

### 2023-09-26 wrangle.qmd
Continuous time analysis (#7)
add continuous time analyses

---------

Co-authored-by: Kristoffer Magnusson <1629860+rpsychologist@users.noreply.github.com>
Add some text

@@
 
+Load some packages:
+
 ```{r}
@@
 
+library(patchwork)
 library(tidyverse)
-source("functions.R")
+source("R/functions.R")
 ```
+
+## Raw data
+
+Download data from OSF:
 
@@
 ```
+
+Load data and rename and rescale variables:
 
@@
 
+## Save data
+
 ```{r}
@@
 saveRDS(dat, "data/data.rds")
-saveRDS(dat_raw, "data/data-raw.rds")
-saveRDS(s, "data/s.rds")
-saveRDS(sraw, "data/sraw.rds")
 ```

### 2023-09-26 wrangle.qmd
Redo sessions and filtering
We now use original sessions and truncate sessions at 5hrs

@@
   )
-) |>
+)
+dat <- dat |> 
   mutate(
     pid = factor(pid),
-    mood = mood / 1000
+    # Mood to 0-100 scale
+    mood = mood / 10, 
+    # hours indicates session duration in hours
+    hours = duration / 60,
+    .keep = "unused"
   )
+```
 
-# Calculate raw data summaries
-sraw <- list()
-sraw$n_res <- dat |>
-  count(prompt)
-sraw$n_res_total <- number2(sum(sraw$n_res$n))
-sraw$n_res <- sraw$n_res |>
-  mutate(n = number2(n)) |>
-  pivot_wider(names_from = prompt, values_from = n)
-sraw$n_sub <- number2(length(unique(dat$pid)))
+## Sessions
 
-# Create sessions using heuristics below
+Mood responses were recorded with a session length indicator, which was reset to 0 when players started the game.
+We therefore create a session indicator that increments from 1 whenever the observation's session duration is less than the previous one.
+
+```{r}
+#| label: session-indicator-create
 dat <- dat |> 
+  # filter(hours <= 5) |> 
   arrange(pid, time) |> 
   mutate(
-    i = row_number(),
-    # Observation is in a new session if this...
-    new_session =
-      # is the first total observation,
-      i == 1 |
-      # session duration is zero,
-      (duration == 0) |
-      # session is shorter than previous session (ticker has reset),
-      (duration < lag(duration, default = 0)) |
-      # or if more than 30 minutes since previous observation.
-      (time > lag(time) + minutes(30)),
-    # Session number is the cumulative sum of new sessions...
+    new_session = hours < lag(hours, default = 999),
     session = cumsum(new_session),
-    # ... per person
     .by = pid
   ) |> 
-  select(-c(i, new_session))
+  select(-c(new_session)) |> 
+  mutate(ps = paste(pid, session, sep = "_"))
 
-# Filter data
-dat_raw <- dat
+str_glue(
+  "Raw data\n",
+  "Players: {number2(length(unique(dat$pid)))}\n",
+  "Sessions: {number2(length(unique(dat$ps)))}\n",
+  "Observations: {number2(nrow(dat))}\n"
+)
+```
+
+Roughly 10% of sessions should have a mood prompt at login.
+However, many sessions are short and only have the login prompt when just the response data is considered.
+Therefore, in the current data the proportion will be somewhat greater than 10%.
+
+```{r}
+# What proportion of sessions have a login mood prompt?
+dat |> 
+  mutate(
+    has_login_prompt = any(hours == 0 & prompt == "Wellbeing"),
+    .by = ps
+  ) |> 
+  distinct(ps, has_login_prompt) |> 
+  summarise(mean(has_login_prompt))
+```
+
+## Response types
+
+We then filter out all but the mood (Wellbeing) item responses
+
+```{r}
+#| label: mood-filter
+
 # Include only wellbeing responses that occurred
@@
 
-# Count of data with all wellbeing observations
-dat_total <- dat |>
-  summarise(
-    n_total = n(),
-    n_players = length(unique(pid))
+str_glue(
+  "Mood-only data\n",
+  "Players: {number2(length(unique(dat$pid)))}\n",
+  "Sessions: {number2(length(unique(dat$ps)))}\n",
+  "Observations: {number2(nrow(dat))}\n"
 )
+```
 
-# Include only sessions with "pre-" and "post" measure(s)
+## EDA / outliers / exclusions
+
+At this point the data table contains all in-game mood responses of all sessions. It might be reasonable to do some filtering of e.g. long sessions.
+
+```{r}
+#| label: fig-session-durations
+#| fig-height: 3
+#| fig-cap: Summaries of session durations. All x-axes are truncated at 10 hours.
+
+p_durations <- dat |> 
+  filter(hours == max(hours), .by = ps) |> 
+  ggplot(aes(hours)) +
+  scale_x_continuous(
+    "Session duration",
+    expand = expansion(c(0.05, 0.05))
+  ) +
+  scale_y_continuous(
+    "Sessions",
+    expand = expansion(c(0, 0.05))
+  ) +
+  coord_cartesian(xlim = c(0, 10)) +
+  geom_histogram(binwidth = 0.33)
+
+p_mean_durations <- dat |> 
+  filter(hours == max(hours), .by = ps) |> 
+  summarise(hours = mean(hours), n = n(), .by = pid) |> 
+  ggplot(aes(hours)) +
+  scale_x_continuous(
+    "Mean session duration",
+    expand = expansion(c(0.05, 0.05))
+  ) +
+  scale_y_continuous(
+    "Players",
+    expand = expansion(c(0, 0.05))
+  ) +
+  coord_cartesian(xlim = c(0, 10)) +
+  geom_histogram(binwidth = 0.33)
+
+p_ecdf_durations <- dat |> 
+  filter(hours == max(hours), .by = c(session, pid)) |> 
+  ggplot(aes(hours)) +
+  scale_x_continuous(
+    "Session duration",
+    expand = expansion(c(0.05, 0.05))
+  ) +
+  scale_y_continuous(
+    "Cumulative proportion",
+    expand = expansion(0.01),
+    breaks = (0:10)/10
+  ) +
+  stat_ecdf() +
+  coord_cartesian(xlim = c(0, 10))
+
+p_durations | p_mean_durations | p_ecdf_durations
+
 dat <- dat |> 
-  filter(
-    # Session has a wellbeing measure at time = 0
-    duration[1] == 0,
-    # Session has at least two non-missing wellbeing measures 
-    # (e.g. pre + 1 post)
-    sum(!is.na(mood)) > 1,
-    .by = c(pid, session)
+  filter(hours <= 5)
+
+str_glue(
+  "Mood-only 5<h session data\n",
+  "Players: {number2(length(unique(dat$pid)))}\n",
+  "Sessions: {number2(length(unique(dat$ps)))}\n",
+  "Observations: {number2(nrow(dat))}\n"
 )
+```
 
-# Count of data to be used in these analyses
-dat_filtered <- dat |>
-  summarise(
-    n_total = n(),
-    n_players = length(unique(pid))
+As with many things concerning numbers, we are happy with 5. Finally we just drop observations where participants skipped the prompt
+
+```{r}
+#| label: drop-missing-data
+
+dat <- dat |> 
+  drop_na()
+
+str_glue(
+  "Mood-only 5<h session non-missing data\n",
+  "Players: {number2(length(unique(dat$pid)))}\n",
+  "Sessions: {number2(length(unique(dat$ps)))}\n",
+  "Observations: {number2(nrow(dat))}\n"
 )
-
-# Summary
-bind_rows(
-  "total" = dat_total,
-  "filtered" = dat_filtered,
-  .id = "Data"
-)
-
-# Summaries of filtered data
-s <- list()
-s$n_sub <- number2(length(unique(dat$pid)))
-s$n_obs <- number2(nrow(dat))
-s$n_obs_missing <- number2(nrow(filter(dat, is.na(mood))))
-s$n_ses <- number2(nrow(count(dat, pid, session)))
 ```

### 2023-09-28 continuous-time.qmd
fix #9

--- 2023-09-25 continuous-time.qmd
@@
 #| label: load-data
-dat <- read_rds("data/data-raw.rds")
-dat <- dat |>
-  mutate(
-    duration = duration - duration[1],
-    hours = duration / 60,
-    .by = c(pid, session)
-  ) |>
-  mutate(
-    ps = interaction(pid, session, sep = "_")
-  ) |>
-  filter(prompt == "Wellbeing")
+dat <- read_rds("data/data.rds")
 
@@
       mood == 0 ~ "left",
-      mood == 1 ~ "right",
+      mood == 100 ~ "right",
       TRUE ~ "none"
     ),
-    mood100 = mood * 100
   )
@@
 #| label: tbl-session-max-prop
-#| tbl-cap: Proportion of participants that had at least on session last longer than gt_* hours
+#| tbl-cap: Proportion of participants that had at least one session last longer than gt_* hours
 dat |>
   summarise(
-    max = max(hours), .by = c(pid)
+    max = max(hours),
+    .by = c(pid)
   ) |>
   summarise(
-    gt_0.33 = mean(max > 0.33),
-    gt_0.71 = mean(max > 0.716),
-    gt_1 = mean(max > 1),
+    gt_0.18 = mean(max > 0.18),
+    gt_0.51 = mean(max > 0.51),
+    gt_1.3 = mean(max > 1.3),
     gt_2 = mean(max > 2),
@@
 
-@tbl-session-max-prop shows that 66% of the participant have at least one session that lasted for over 0.71 hours. So we will base our knot placement on these quantiles.
+@tbl-session-max-prop shows that 85% of the participant have at least one session that lasted for over 0.51 hours, and 64% for over 1.3 hours. So we will base our knot placement on these quantiles.
 
@@
     x,
-    knots = c(0.05, 0.33, 0.7),
+    knots = c(0.18, 0.5, 1.3),
     Boundary.knots = c(0, 3)
@@
   lmer(
-    mood100 ~ ns4(hours) + (1 | ps) + (1 + ns4(hours) | pid),
+    mood ~ ns4(hours) + (1 | ps) + (1 + ns4(hours) | pid),
     data = dat
@@
   ) +
-  lims(y = c(0, 6)) +
+  lims(y = c(0, 10)) +
   labs(x = "Session duration (hours)") +
@@
 
-@tbl-change-contrast shows the estimated change from baseline after 10min, 30 min, 1 hour, 2 hours, and 3 hours.
+@tbl-change-contrast shows the estimated change from baseline after 15min, 30 min, 1 hour, 2 hours, and 3 hours.
 ```{r}
@@
   ~hours,
-  at = list(hours = c(0, 10 / 60, 0.5, 1, 2, 3)),
+  at = list(hours = c(0, 15 / 60, 0.5, 1, 2, 3)),
   lmer.df = "asymptotic"
@@
     # Session has a wellbeing measure at time = 0
-    duration[1] == 0,
+    hours[1] == 0,
     .by = c(pid, session)
@@
   mutate(
-    pre = mood100[1],
+    pre = mood[1],
     .by = c(pid, session)
@@
   filter(
-    hours != 0
+    hours > 0
   ) |> 
@@
   lmer(
-    mood100 ~ ns4(hours) * ns(pre, 5) + (1 + ns4(hours) | pid) + (1 | ps),
+    mood ~ ns4(hours) * ns(pre, 5) + (1 + ns4(hours) | pid) + (1 | ps),
     data = dat_pre
@@
 
-Calculate pre values we are interested in plotting.
+Next we'll calculate the pre values we are interested in plotting.
 
@@
 
+However, baseline mood is not normally distributed, so we'll use percentiles instead. 
 ```{r}
@@
 #| tbl-cap: Distribution of average mood levels at the start of a session
-pre_values <- pre_sum$mean + pre_sds * pre_sum$sd
+#pre_values <- pre_sum$mean + pre_sds * pre_sum$sd
+
+pre <- dat_pre |> 
+  summarise(pre = pre[1], .by = ps) |> 
+  pull(pre)
+pre_values <- quantile(pre, c(0.05, .25, 0.5, 0.75))
+pre_labels_x <- c("5th", "25th", "Median", "75th")
+colors <- c(
+  "5th" = "#4daf4a",
+  "25th" = "#e41a1c",
+  "75th" = "#377eb8",
+  "Median" = "#000000"
+)
+color_scale_x <- list(
+  scale_color_manual(
+    "Pre",
+    values = colors
+  ),
+    scale_fill_manual(
+    "Pre",
+    values = colors
+  )
+)
 data.frame(
   pre = pre_values,
-  label = pre_labels
+  label = pre_labels_x
   ) |> 
@@
   ),
-  lmer.df = "asymp"
+  lmer.df = "asymptotic"
 )
@@
       round(pre), 
-      labels = pre_labels
+      labels = pre_labels_x
       )
@@
   pre = round(pre_values),
-  label = pre_labels
+  label = pre_labels_x
 ) |>
@@
       label,
-      levels = pre_labels
+      levels = pre_labels_x
     )
@@
     fit3, 
-    hours = c(0, 10/60, 0.5, 1, 2))
+    hours = c(0, 15/60, 0.5, 1, 2))
     ) |>
@@
   pid %in% sample(unique(pid), n_pids_sample)
-)
+) |> 
+  mutate(mood = mood / 100)
 
@@
   )
-# SHIFT <- 0.017
-SHIFT <- 0
 d_lmm <-
@@
     ~hours,
-    at = list(hours = seq(0, 3, length.out = 100))
+    at = list(hours = seq(0, 3, length.out = 100)),
+    lmer.df = "asymptotic"
   ) |>
@@
     model = "lmm",
-    emmean = emmean + SHIFT,
-    upper = upper + SHIFT,
-    lower = lower + SHIFT
+    emmean = emmean,
+    upper = upper,
+    lower = lower
   )
@@
 
-We then compare the censored normal model with the LMM and recreate @fig-lmm-pre-interaction.
-As shown in @fig-cens-pre-interaction, there's almost no difference between the models.
-The LMMs CIs are slightly narrower near the intercept.
+We then compare the censored normal model with the LMM and recreate @fig-lmm-pre-interaction. As shown in @fig-cens-pre-interaction, there's almost no difference between the models.
 ```{r}
@@
   lmer(
-    mood100 ~ ns4(hours) * pre + (1 + ns4(hours) | pid) + (1 | ps),
+    mood ~ ns4(hours) * pre + (1 + ns4(hours) | pid) + (1 | ps),
     data = dat_pre
@@
     lmer(
-      mood100 ~ ns4(hours) * ns(pre, 2) + (1 + ns4(hours) | pid) + (1 | ps),
+      mood ~ ns4(hours) * ns(pre, 2) + (1 + ns4(hours) | pid) + (1 | ps),
       data = dat_pre,
@@
     lmer(
-      mood100 ~ ns4(hours) * ns(pre, 3) + (1 + ns4(hours) | pid) + (1 | ps),
+      mood ~ ns4(hours) * ns(pre, 3) + (1 + ns4(hours) | pid) + (1 | ps),
       data = dat_pre,
@@
     lmer(
-      mood100 ~ ns4(hours) * ns(pre, 4) + (1 + ns4(hours) | pid) + (1 | ps),
+      mood ~ ns4(hours) * ns(pre, 4) + (1 + ns4(hours) | pid) + (1 | ps),
       data = dat_pre,
@@
         round(pre),
-        labels = pre_labels,
+        labels = pre_labels_x,
       ),
@@
   lmer(
-    mood100 ~ hours * ns(pre, 5) + (1 + hours | pid) + (1 | ps),
+    mood ~ hours * ns(pre, 5) + (1 + hours | pid) + (1 | ps),
     data = dat_pre
@@
     lmer(
-      mood100 ~ ns(hours, 2) * ns(pre, 5) + (1 + ns(hours, 2) | pid) + (1 | ps),
+      mood ~ ns(hours, 2) * ns(pre, 5) + (1 + ns(hours, 2) | pid) + (1 | ps),
       data = dat_pre,
@@
   lmer(
-    mood100 ~ ns(hours, 3) * ns(pre, 5) + (1 + ns(hours, 3) | pid) + (1 | ps),
+    mood ~ ns(hours, 3) * ns(pre, 5) + (1 + ns(hours, 3) | pid) + (1 | ps),
     data = dat_pre,
@@
   lmer(
-    mood100 ~ poly(hours, 2) * ns(pre, 5) + (1 + poly(hours, 2) | pid) + (1 | ps),
+    mood ~ poly(hours, 2) * ns(pre, 5) + (1 + poly(hours, 2) | pid) + (1 | ps),
     data = dat_pre,

### 2023-10-06 analyze.qmd
Fix #8
Cross reference sections (#11)
WIP

--- 2023-09-19 analyze.qmd
@@
----
-title: Analyze
----
+# Analyze {#sec-analyze}
 
@@
 library(tidyverse)
-source("functions.R")
-```
-
-```{r}
-#| label: options
-#| cache: false
-
+source("R/functions.R")
+source("R/fit_cached.R")
+source("R/common.R")
 dir.create("models", FALSE)
-N_CORES <- as.numeric(Sys.getenv("N_CORES", unset = 1))
-N_THREADS <- as.numeric(Sys.getenv("N_THREADS", unset = 2))
-N_ITER <- as.numeric(Sys.getenv("N_ITER", unset = 2000))
-N_SUBSET <- as.numeric(Sys.getenv("N_SUBSET"))
-options(
-  mc.cores = N_CORES,
-  brms.backend = "cmdstanr"
-)
-theme_set(
-  theme_linedraw(base_size = 10) +
-    theme(
-      strip.background = element_rect(fill = NA, color = NA),
-      strip.text = element_text(colour = "black", hjust = 0),
-      axis.text = element_text(size = rel(0.75)),
-      panel.grid = element_blank()
-    )
-)
 ```
@@
 dat <- readRDS("data/data.rds")
-s <- readRDS("data/s.rds")
-sraw <- readRDS("data/sraw.rds")
+
 if (!is.na(N_SUBSET)) {
@@
 a <- str_glue("{a[1]} ({a[2]}, {a[3]})")
-g <- count(demo, gender, sort = TRUE) |>
+```
+
+- N participants: `r number2(length(unique(dat$pid)))`
+- N sessions: `r number2(length(unique(dat$ps)))`
+- age: `r a`
+- gender:
+
+```{r}
+#| label: demographics-gender
+
+count(demo, gender, sort = TRUE) |>
   mutate(p = n / sum(n)) |>
   mutate(x = str_glue("{gender} ({number2(n)}, {percent2(p)})"))
+```
+
+- countries:
+```{r}
+#| label: demographics-countries
+
 n_countries <- length(unique(demo$country[!is.na(demo$country)]))
-countries <- count(demo, country, sort = TRUE) |>
+count(demo, country, sort = TRUE) |>
   mutate(p = n / sum(n)) |>
@@
 
-The PowerWash Simulator case study database contains `r sraw$n_res_total` psychological item responses from `r sraw$n_sub` individuals.
-In this study, we only studied subjective well-being (mood), so did not consider the other five types of item responses in the database (enjoyment, focus, autonomy, competence, and immersion).
-Therefore, here we used data from `r s$n_sub` participants who provided `r s$n_obs` mood responses (`r s$n_obs_missing` with missing values, which we ignored) from `r s$n_ses` gameplay sessions.
-
-These participants median age (1st & 9th deciles) was `r a`, and four most frequent responses to gender were `r g$x[1:4]`.
-They hailed from `r n_countries`, with four most represented countries being `r countries$x[1:4]`.
-
-The study procedures were granted ethical approval by Oxford University’s Central University Research Ethics Committee (SSH_OII_CIA_21_011).
-All participants provided informed consent to participate and reported being 18 years or older.
-
-## Data analysis
-
-```{r}
-#| label: data-wrangle-for-analysis
-
-dat <- dat |>
-  mutate(
-    post = factor(
-      row_number() > 1,
-      levels = c(FALSE, TRUE),
-      labels = c("0", "1")
-    ),
-    .by = c(pid, session)
-  ) |>
-  mutate(ps = interaction(pid, session, sep = "_")) |>
-  mutate(
-    cl = case_when(
-      mood == 0 ~ "left",
-      mood == 1 ~ "right",
-      TRUE ~ "none"
-    )
-  )
-```
-
-Our research question concerned the difference between players’ mood at the beginning of each session (pre-play) and during the subsequent play session (during play).
-It is important to understand that this contrast does not embed a causal hypothesis: Players could begin (and end) their play ses-sions for whatever reason, and these reasons are likely to confound the pre -- during contrast.
-For example, a player might come home after a stressful day at work, and sit down to play some PWS.
-During this play session, their mood would improve because they are at home vs.
-stressful work environment, and would show as a greater during play mood compared to pre-play mood: Reasons behind starting a play session are likely to contribute to the pre -- during contrast and we are unable to disentangle those from any changes actually caused by play.
-
-We estimated this contrast within a censored three-level hierarchical regression model.
-We specified a censored gaussian model because mood was measured with a VAS.
-Ignoring the censoring inherent in VAS responses can confound changes in the mean with changes in the spread of mood, and leave the contrast susceptible to ceiling or floor effects.
-We then modelled mean mood on an intercept and contrast coded time (pre-play: -0.5; during play: 0.5), and allowed both parameters to vary randomly across players.
-In addition, we modelled random intercepts over player-sessions because over time, mood is susceptible to myriad other influences.
-Although equal variances across people in such naturalistic observation seem unlikely, we fixed this parameter to a constant for model identifiability and convergence.
-
-More formally, we modelled each mood response *i* in 1 to `r s$n_obs` from participant *j* in 1 to `r s$n_sub` and participant-session *k* in 1 to `r s$n_ses` with a normal distribution censored at the VAS limits [0, 1].
-We then modelled the distribution's mean ($\eta$) on a population-level intercept ($\beta_0$) and contrast-coded measurement timepoint ($\beta_1$), their random per-player deviations $\alpha$, and a random intercept for player-sessions $\gamma_0$:
-
-\begin{align*}
-\text{mood}_{ijk} &\sim \text{CensNormal}(\eta_{ijk}, \sigma^2, 0, 1), \\
-\eta_{ijk} &= \alpha_0 + \beta_{0j} +
-  (\alpha_1 + \beta_{1j})\text{post}_{ijk} +
-  \gamma_{0k}, \\
-\begin{bmatrix} \alpha_0 \\ \alpha_1 \end{bmatrix} &\sim 
-  \text{MVN}\left(\begin{bmatrix} 0 \\ 0 \end{bmatrix}, 
-  \begin{bmatrix} 
-    \tau_{\alpha_0} &\rho_{{\alpha_0}{\alpha_1}} \\ 
-    \rho_{{\alpha_0}{\alpha_1}} &\tau_{\alpha_1}
-  \end{bmatrix}\right), \\
-  \gamma_0 &\sim \text{Normal}(0, \tau_{\gamma_0}).
-\end{align*}
-
-```{r}
-#| label: model-estimate
-#| cache: false
-
-contrasts(dat$post) <- c(-0.5, 0.5)
-
-model <- bf(mood | cens(cl) ~ 1 + post + (1 + post | pid) + (1 | ps)) +
-  gaussian()
-
-fit <- brm(
-  model,
-  data = dat,
-  silent = 0,
-  threads = N_THREADS,
-  iter = N_ITER,
-  control = list(adapt_delta = .95),
-  file = "models/brm-pre-post-pid-ps-censored"
-)
-```
-
-```{r}
-#| label: model-convergence
-#| eval: false
-
-as_draws(fit, variable = c("b_", "sd_", "cor_"), regex = TRUE) |> 
-  summarise_draws(rhat)
-pairs(fit$fit, pars = variables(fit)[1:7])
-```
-
-We analyzed the data with R [@rcoreteamLanguageEnvironmentStatistical2023] and used the brms package to estimate, via Stan's HMC sampling procedures, and post-process the models [@burknerBrmsPackageBayesian2017; @standevelopmentteamStanModelingLanguage2022].
-We drew `r number2(ndraws(fit))` samples from the posterior distribution using brms' default prior distributions on all parameters, and used numerical and graphical checks to ensure model convergence and adequacy.
-
-# Results
+## Descriptives
 
@@
 #| label: fig-descriptives
-#| include: false
 #| fig-height: 2.2
@@
   scale_y_continuous(
-    expand = expansion(c(0.001, 0.1)),
+    expand = expansion(c(0.001, 0.05)),
   ) +
@@
     col = "white",
-    linewidth = .25
+    bins = 50,
+    linewidth = .25,
+    boundary = 0
   )
@@
 dat_session_durations <- dat |>
-  filter(duration == max(duration), .by = c(pid, session))
+  filter(hours == max(hours), .by = c(pid, session))
 
@@
   dat_session_durations +
-  aes(duration) +
-  scale_x_log10() +
-  coord_cartesian(xlim = c(5, 900)) +
-  labs(x = "Session duration (minutes)", y = "Sessions")
+  aes(hours) +
+  labs(x = "Session duration (hours)", y = "Sessions")
 
@@
   dat_session_counts +
-  coord_cartesian(xlim = c(0, 15)) +
+  coord_cartesian(xlim = c(0, 75)) +
   aes(sessions) +
@@
 (p_durations | p_sessions | p_mood) +
-  plot_annotation(tag_level = "A")
+  plot_annotation(tag_levels = "A")
 ```
@@
 tmp <- list(
-  dat_session_durations$duration,
+  dat_session_durations$hours,
   dat_session_counts$sessions,
@@
   map(~median_qi(.x, .width = 0.8, na.rm = TRUE)) |> 
-  map(~str_glue_data(.x, "{y} [{ymin}, {ymax}]"))
+  map(~mutate(.x, across(c(y, ymin, ymax), ~number2(., .01))) |> str_glue_data("{y} [{ymin}, {ymax}]"))
 ```
 
-The median session duration was `r tmp[[1]]` minutes [10 and 90 percentiles]; the median player contributed data from `r tmp[[2]]` sessions, and the median mood was `r tmp[[3]]` (pre-session: `r tmp[[4]]`, during play: `r tmp[[5]]`).
+The median session duration was `r tmp[[1]]` hours [10 and 90 percentiles]; the median player contributed data from `r tmp[[2]]` sessions, and the median mood was `r tmp[[3]]` (pre-session: `r tmp[[4]]`, during play: `r tmp[[5]]`).
 
@@
 #| fig-width: 8
-#| fig-height: 5
+#| fig-height: 4.6
 #| fig-cap: "**A**.
@@
 
-axislabels <- c(0, 0.5, 1.0)
-
 # Plot: Example raw sessions
@@
   mutate(session_per_person = length(unique(session)), .by = pid)
+
 # default 
-min_sessions <- 4 
+min_sessions <- 8
 n_filtered <- dat_example_sessions |> 
   distinct(pid, session_per_person) |> 
-  filter(session_per_person >= 4) |> 
+  filter(session_per_person >= 8) |> 
   nrow()
+
 # pick anyone if there's not 3 pids with >= 4 sessions
@@
 
-p_mood_example_sessions <- dat_example_sessions |> 
+p_mood_example <- dat_example_sessions |> 
+  filter(session <= 8) |> 
   ggplot() +
-  aes(duration / 60, mood, col = post) +
+  aes(hours, mood, col = post) +
   scale_color_brewer(
@@
     limits = c(0, 1),
-    breaks = axislabels,
-    labels = axislabels
   ) +
@@
     cols = vars(Session),
-    scales = "free_x"
+    scales = "fixed"
   ) +
@@
   )
-
-p_mood_example_all <- dat_example_sessions |> 
-  mutate(Session = "All sessions") |> 
-  ggplot() +
-  aes(y = mood, x = post, col = post, fill = post) +
-  scale_color_brewer(
-    "Pre-session measure",
-    palette = "Set1",
-    aesthetics = c("color", "fill")
-  ) +
-  scale_x_discrete(
-    "Session",
-    breaks = c(0, 1),
-    labels = c("Pre", "During"),
-    expand = expansion(c(0.2, 2.4))
-  ) +
-  scale_y_continuous(
-    limits = c(0, 1),
-    breaks = axislabels,
-    labels = axislabels
-  ) +
-  stat_summary(
-    fun.data = mean_cl_boot,
-    fatten = 2
-  ) +
-  geom_dots(side = "right", alpha = .33, binwidth = .075) +
-  facet_grid(
-    rows = vars(Person), 
-    cols = vars(Session)
-  ) +
-  theme(
-    axis.text.y = element_blank(), 
-    axis.title.y = element_blank(), 
-    axis.ticks.y = element_blank(), 
-    axis.text.x = element_text(hjust = 0),
-    strip.background.y = element_blank(),
-    legend.position = "none"
-  )
-
-p_mood_example <- (p_mood_example_sessions | p_mood_example_all) +
-  plot_layout(widths = c(92, 8))
 
@@
     "Mood",
-    expand = expansion(c(0.001)),
-    breaks = axislabels
+    expand = expansion(c(0.001))
   ) +
@@
   pivot_wider(names_from = post, values_from = mood) |> 
-  ggplot(aes(`During play` - `Pre-play`)) +
+  mutate(Difference = `During play` - `Pre-play`) |> 
+  drop_na(Difference) |> 
+  ggplot(aes(Difference)) +
   geom_histogram(bins = 50, col = "white") +
@@
   scale_x_continuous(
-    "Difference",
     expand = expansion(c(0.001))
   ) +
-  coord_cartesian(xlim = c(-0.4, 0.6))
+  coord_cartesian(xlim = c(-.4, .6))
 
 # Plot: Difference histogram (players)
-p_mood_difference_players <- p_mood_difference_sessions %+% 
-  summarise(
-    p_mood_difference_sessions$data,
-    across(`Pre-play`:`During play`, ~mean(.x, na.rm = TRUE)),
-    .by = pid
-  ) +
+p_mood_difference_players <- p_mood_prepost_players$data |> 
+  pivot_wider(names_from = post, values_from = mood) |> 
+  mutate(Difference = `During play` - `Pre-play`) |> 
+  drop_na(Difference) |> 
+  ggplot(aes(Difference)) +
+  geom_histogram(bins = 50, col = "white") +
+  geom_vline(xintercept = 0, linewidth = .33, col = "dodgerblue") +
   scale_y_continuous(
@@
     expand = expansion(c(0.001, 0.1))
-  )
+  ) +
+  scale_x_continuous(
+    expand = expansion(c(0.001))
+  ) +
+  coord_cartesian(xlim = c(-.4, .6))
 
@@
   scale_x_continuous(
-    expand = expansion(0.01),
-    breaks = axislabels,
-    labels = axislabels
+    expand = expansion(0.01)
   ) +
   scale_y_continuous(
-    expand = expansion(0.01),
-    breaks = axislabels,
-    labels = axislabels
+    expand = expansion(0.01)
   ) +
@@
 
-p_mood_biscatter_players <- p_mood_biscatter_sessions %+% 
-  p_mood_difference_players$data
+p_mood_biscatter_players <- p_mood_difference_players$data |> 
+  ggplot(aes(`Pre-play`, `During play`)) +
+  scale_x_continuous(
+    expand = expansion(0.01)
+  ) +
+  scale_y_continuous(
+    expand = expansion(0.01)
+  ) +
+  geom_point(
+    alpha = .25, size = 1, shape = 1
+  ) +
+  geom_abline(linewidth = .33, col = "dodgerblue") +
+  theme(aspect.ratio = 1)
 
@@
   ) +
-  plot_layout(heights = c(4.5, 5.5)) +
+  plot_layout(heights = c(5, 5)) +
   plot_annotation(tag_levels = "A")
@@
 
-We then turned to our primary research question: To what extent does player mood in PWS change from pre-play to during play?
-First, we visualized the relevant data in @fig-data: Panel A shows mood responses from three example participants' first five sessions of play.
-@fig-data B then shows histograms of those data to facilitate visual comparison of the raw data.
-We then calculated sessions' and participants' mean moods at the beginning and during each session (@fig-data C, D, and G), and their differences (E and F).
-Visual inspection suggested greater mood during play compared to the pre-play measures at both the level of sessions and players.
+## Model
+
+Our research question concerned the difference between players’ mood at the beginning of each session (pre-play) and during the subsequent play session (during play).
+It is important to understand that this contrast does not embed a causal hypothesis: Players could begin (and end) their play ses-sions for whatever reason, and these reasons are likely to confound the pre -- during contrast.
+For example, a player might come home after a stressful day at work, and sit down to play some PWS.
+During this play session, their mood would improve because they are at home vs.
+stressful work environment, and would show as a greater during play mood compared to pre-play mood: Reasons behind starting a play session are likely to contribute to the pre -- during contrast and we are unable to disentangle those from any changes actually caused by play.
+
+We estimated this contrast within a censored three-level hierarchical regression model.
+We specified a censored gaussian model because mood was measured with a VAS.
+Ignoring the censoring inherent in VAS responses can confound changes in the mean with changes in the spread of mood, and leave the contrast susceptible to ceiling or floor effects.
+We then modelled mean mood on an intercept and contrast coded time (pre-play: -0.5; during play: 0.5), and allowed both parameters to vary randomly across players.
+In addition, we modelled random intercepts over player-sessions because over time, mood is susceptible to myriad other influences.
+Although equal variances across people in such naturalistic observation seem unlikely, we fixed this parameter to a constant for model identifiability and convergence.
+
+More formally, we modelled each mood response *i* in 1 to `r number2(nrow(dat))` from participant *j* in 1 to `r number2(length(unique(dat$pid)))` and participant-session *k* in 1 to `r number2(length(unique(dat$ps)))` with a normal distribution censored at the VAS limits [0, 1].
+We then modelled the distribution's mean ($\eta$) on a population-level intercept ($\beta_0$) and contrast-coded measurement timepoint ($\beta_1$), their random player ($\alpha$) and player-session ($\gamma$) deviations,
+
+$$
+\begin{align*}
+\text{mood}_{ijk} &\sim \text{CensNormal}(\eta_{ijk}, \sigma^2, 0, 1), \\
+\eta_{ijk} &= \alpha_0 + \beta_{0j} + \gamma_{0k} +
+  (\alpha_1 + \beta_{1j})\text{post}_{ijk}, \\
+\begin{bmatrix} \alpha_0 \\ \alpha_1 \end{bmatrix} &\sim 
+  \text{MVN}\left(\begin{bmatrix} 0 \\ 0 \end{bmatrix}, 
+  \begin{bmatrix} 
+    \tau_{\alpha_0} &\rho_{{\alpha_0}{\alpha_1}} \\ 
+    \rho_{{\alpha_0}{\alpha_1}} &\tau_{\alpha_1}
+  \end{bmatrix}\right), \\
+\begin{bmatrix} \gamma_0 \end{bmatrix} &\sim 
+  \text{Normal}\left(\begin{bmatrix} 0 \end{bmatrix}, 
+  \begin{bmatrix} 
+    \tau_{\gamma_0}
+  \end{bmatrix}\right).
+\end{align*}
+$$
+
+```{r}
+#| label: model-estimate-brms
+#| cache: false
+
+model <- bf(
+  mood | cens(cl) ~ 
+    1 + post + 
+    (1 + post | pid) + 
+    (1 | ps)
+) +
+  gaussian()
+
+fit <- brm(
+  model,
+  data = dat,
+  silent = 0,
+  iter = N_ITER,
+  control = list(adapt_delta = .95),
+  file = "models/brm-prepost-pid-ps-censored"
+)
+```
+
+```{r}
+#| label: fig-convergence
+#| fig-cap: Convergence diagnostic plot showing bivariate scatterplots of the model's population-level parameters' posterior draws.
+
+pairs(
+  fit, 
+  variable = c("b_", "sd_", "cor_"), 
+  regex = TRUE,
+  off_diag_args = list(size = .33, shape = 1)
+  )
+```
+
+## Population level effects
 
@@
   mutate(CI = str_glue("[{q2.5}, {q97.5}]")) |> 
-  select(` ` = variable, Estimate = mean, CI, pd) |> 
+  select(` ` = variable, Estimate = mean, CI, pd)
+h_avg_tab |>   
   knitr::kable()
 ```
-
-The model results confirmed the visual impressions described above: @tbl-avg indicates, with great certainty, that mood increased by about [`r paste0(map(h_avg_$hypothesis[3,4:5], ~number(.x, .001)), collapse = ", ")`] units (@fig-coef, "Raw").
 
@@
   fit, 
-  c("sd_pid__Intercept", "sd_pid__post1", "sigma", "b_post1")
+  c("sd_", "sigma", "b_post1"),
+  regex = TRUE
 ) |> 
   mutate(
-    total_residual = sd_pid__Intercept + sd_pid__post1 + sigma,
-    b_post1_scaled = b_post1 / total_residual
+    tr = sd_pid__Intercept + sd_pid__post1 + sd_ps__Intercept + sigma,
+    b_post1_z = b_post1 / tr,
+    prop_pos = pnorm(0, b_post1, sd_pid__post1, lower.tail = FALSE)
   ) |> 
@@
   select(-starts_with("."))
-
-posterior_pop_sum <- posterior_pop |> 
-  summarise_draws(
-    mean = ~number(mean(.), .001), 
-    ~number2(quantile2(., probs = c(.025, .975)), .001)
+posterior_pop |> 
+  summarize_draws(
+    mean, ~quantile2(., probs = c(0.025, 0.975))
   ) |> 
-  mutate(
-    x = str_glue("{mean} [{q2.5}, {q97.5}]")
-  )
+  knitr::kable(digits = 3)
 ```
 
-To further interpret the magnitude of the difference in moods between pre- and during play, we compared the contrast to the natural variations in mood as estimated by the model (standard deviation of the between-person intercept heterogeneity ($\tau_{\alpha_0}$ = `r posterior_pop_sum$x[1]`), between-person pre-during difference heterogeneity ($\tau_{\alpha_1}$ = `r posterior_pop_sum$x[2]`) and to the residual standard deviation ($\sigma$ = `r posterior_pop_sum$x[3]`).
-The standardized (with the total residual variation from the three above components) pre-play -- during play contrast was `r posterior_pop_sum$x[6]` (@fig-coef, "Scaled"), a not ignorable difference.
+## Heterogeneity
 
@@
   posterior_pid, 2, 
-  function(x) x / posterior_pop$total_residual
+  function(x) x / posterior_pop$tr
 ) |> 
@@
   "b_post1" = posterior_pid,
-  "b_post1_scaled" = posterior_pid_scaled,
+  "b_post1_z" = posterior_pid_scaled,
   .id = "name"
-)
-
-x <- posterior_pop |> 
-  mutate(
-    prop_pos = pnorm(0, b_post1, sd_pid__post1, lower.tail = FALSE)
-  ) |> 
-  mean_qi(prop_pos) |> 
-  mutate(x = str_glue("{percent(prop_pos, .1)} [{number(.lower*100, .1)}, {number(.upper*100, .1)}]")) |> 
-  pull(x)
+) |> tibble()
 ```
-
-We then turned our attention to between-person differences in the difference in during and pre-play mood.
-With a sample of `r s$n_sub`, it is not surprising that the population-level contrast is credibly different from zero.
-We therefore asked the subsequent question: What proportion of the population from which the sample was drawn would we expect there to be an elevation in mood during play?
-The boxplots in @fig-coef show the distribution of player-specific contrasts estimated by the model, which indicated that `r x` of players in the population experience a positive shift in mood during play, compared to their moods during login.
 
@@
 #| include: true
-#| fig.width: 3
-#| fig.height: 1.4
-#| fig.cap: "Difference in mood during play compared to mood at session start.
-Points and intervals indicate posterior means and 95%CIs of the population level difference; boxplots summarize the player-level estimates.
-Raw estimates are on the VAS scale, whereas Scaled estimates indicate contrasts divided by the total residual deviance of the model."
+#| fig.width: 4
+#| fig.height: 2.6
+#| fig.cap: Histograms of person-specific contrasts on the raw scaled and standardized scale.
 
-posterior_pop |> 
-  select(starts_with("b_")) |> 
-  pivot_longer(everything()) |>
-  group_by(name) |> 
-  mean_qi(value, .width = 0.95) |> 
-  ggplot(aes(value, name)) +
+posterior_pid_sum |> 
+  ggplot(aes(y, name)) +
+  stat_slab(
+    density = "histogram", 
+    breaks = 100, 
+    slab_color = "white", 
+    outline_bars = TRUE, 
+    linewidth = .25
+  ) +
+  geom_vline(xintercept = 0, linewidth = .33, col = "dodgerblue") +
   scale_x_continuous(
@@
   scale_y_discrete(
-    breaks = c("b_post1", "b_post1_scaled"),
-    labels = c("Raw", "Scaled")
-  ) +
-  coord_cartesian(xlim = c(-0.25, 0.65)) +
-  geom_vline(xintercept = 0, linewidth = .33, col = "dodgerblue") +
-  geom_boxploth(
-    data = posterior_pid_sum,
-    aes(x = y, group = name), 
-    outlier.shape = NA,
-    notch = TRUE,
-    size = 0.25,
-    width = .2
-  ) +
-  geom_pointrangeh(
-    aes(xmin = .lower, xmax = .upper),
-    fatten = 2.5
+    breaks = c("b_post1", "b_post1_z"),
+    labels = c("Raw", "Scaled"),
+    expand = expansion(c(0.05, 0.01))
   ) +
@@
     axis.title.y = element_blank()
-  )
+  ) +
+  coord_cartesian(ylim = c(1, 2.5), xlim = c(-0.2, 0.4))
 ```
 
-# Bibliography
-
-::: {#refs}
-:::
+```{r}
+p_mood_difference_players$data |> 
+  left_join(posterior_pid) |> 
+  left_join(count(dat, pid)) |> 
+  ggplot(aes(Difference, y, alpha = n, size = n)) +
+  scale_x_continuous(
+    "Difference"
+  ) +
+  scale_y_continuous(
+    "Modelled difference"
+  ) +
+  coord_cartesian(
+    xlim = c(-1, 1), ylim = c(-.5, .5)
+  ) +
+  geom_point() +
+  geom_abline() +
+  theme(aspect.ratio = 1)
+```

### 2023-10-06 continuous-time.qmd
Cross reference sections (#11)
WIP

@@
----
-title: "Continuous time"
----
+# Continuous time {#sec-continuous-time}
 
@@
 
-In the main analysis in [Analyze]() we focused on contrasting the mood at the beginning of a session with the average mood during session.
+In the main analysis in @sec-analyze we focused on contrasting the mood at the beginning of a session with the average mood during session.
 In this document we focus on two additional research questions:
@@
 source("R/fit_cached.R")
-```
-
-```{r}
-#| label: options
-#| cache: false
-
-N_CORES <- as.numeric(Sys.getenv("N_CORES", unset = 1))
-N_THREADS <- as.numeric(Sys.getenv("N_THREADS", unset = 2))
-N_ITER <- as.numeric(Sys.getenv("N_ITER", unset = 2000))
-N_SUBSET <- as.numeric(Sys.getenv("N_SUBSET"))
-options(
-  mc.cores = N_CORES,
-  brms.backend = "cmdstanr"
-)
-theme_set(
-  theme_linedraw(base_size = 10) +
-    theme(
-      strip.background = element_rect(
-        fill = NA,
-        color = NA
-      ),
-      strip.text = element_text(
-        colour = "black", hjust = 0
-      ),
-      axis.text = element_text(
-        size = rel(0.75)
-      )
-    )
-)
-scale_x_hours <- list(
-  scale_x_continuous(
-  breaks = c(0, 15/60, 0.5, 45/60, 1, 2, 3), 
-  labels = c(0, "15m", "30m", "45m", "1h", "2h", "3h")
-  )
-)
+source("R/common.R")
 ```
@@
 ## Data
-Load the data and remove observations that occurred more than 3 hours into a session.
+Load the data.
 ```{r}
 #| label: load-data
+
 dat <- read_rds("data/data.rds")
 
-dat <- dat |>
-  filter(
-    !is.na(mood),
-    hours <= 3
-  ) |>
-  mutate(
-    cl = case_when(
-      mood == 0 ~ "left",
-      mood == 100 ~ "right",
-      TRUE ~ "none"
-    ),
-  )
 if (!is.na(N_SUBSET)) {
@@
 #| fig-width: 8
+
 hours <- seq(0, 3, length.out = 100)
@@
     x = "Session duration (hours)",
-    y = "Mood (0 - 100)"
+    y = "Mood (0 - 1)"
   ) +
@@
 #| fig-width: 8
+
 d_change <- emmeans::contrast(
@@
   ggplot(aes(hours, estimate)) +
+  scale_y_continuous(
+    expand = expansion(c(0.1, 0.1))
+  ) +
   geom_line() +
@@
   ) +
-  lims(y = c(0, 10)) +
   labs(x = "Session duration (hours)") +
@@
   kable(
-    digits = 2
+    digits = 3
   )
@@
     pre = factor(
-      round(pre), 
+      pre,
       labels = pre_labels_x
@@
 d_x_change <- data.frame(
-  pre = round(pre_values),
+  pre = pre_values,
   label = pre_labels_x
@@
 #| label: fig-ms-interaction
-#| fig-height: 4
-#| fig-width: 8
+#| fig-height: 3.5
+#| fig-width: 7
 #| fig-cap: Figure included in manuscript
@@
       scale_x_hours2 +
-      labs(y = "Mood", x = NULL) +
-      theme(plot.margin = margin(0, 0, 0, 0, "pt"))
+      labs(y = "Mood") +
+      theme(axis.title.x = element_blank())
   ) / 
@@
         labs(y = "Change", x = "Duration") +
-        theme(plot.margin = margin(0, 0, 0, 0, "pt"))
-      ) 
-    ) +
-   plot_annotation(tag_levels = "A") +
-  plot_layout(
-    heights = c(2, 2),
-    widths = c(8, 8),
-    guides = "collect"
-  ) & theme(legend.position = "bottom") 
+        theme()
+    ) 
+) +
+  plot_annotation(tag_levels = "A") &
+  theme(legend.position = "none") 
 p_tmp
@@
 #| echo: false
+
 ggsave(
@@
 #| label: fig-ms-no-interaction
-#| fig-height: 4
-#| fig-width: 8
+#| fig-height: 3
+#| fig-width: 6
 #| fig-cap: Figure shown in manuscript
+
 p_tmp <- (
@@
       pre = factor(
-        round(pre),
+        pre,
         labels = pre_labels_x,

### 2023-10-06 wrangle.qmd
Cross reference sections (#11)
WIP

@@
----
-title: Wrangle
----
+# Wrangle {#sec-wrangle}
 
@@
 source("R/functions.R")
+source("R/common.R")
 ```
@@
     pid = factor(pid),
-    # Mood to 0-100 scale
-    mood = mood / 10, 
+    # Mood to 0-1 scale
+    mood = mood / 1000, 
     # hours indicates session duration in hours
@@
 #| label: session-indicator-create
+
 dat <- dat |> 
-  # filter(hours <= 5) |> 
   arrange(pid, time) |> 
@@
 ```{r}
+#| label: check-login-prompt-rate
+
 # What proportion of sessions have a login mood prompt?
@@
 
+Other stuff
+
+```{r}
+dat <- dat |>
+  mutate(
+    post = factor(
+      row_number() > 1,
+      levels = c(FALSE, TRUE),
+      labels = c("0", "1")
+    ),
+    .by = c(pid, session)
+  ) |>
+  # Censoring indicator
+  mutate(
+    cl = case_when(
+      mood == 0 ~ "left",
+      mood == 1 ~ "right",
+      TRUE ~ "none"
+    )
+  )
+
+contrasts(dat$post) <- c(-0.5, 0.5)
+```
+
 ## Save data
@@
 ```{r}
+#| label: data-save
+
 dir.create("data", FALSE)

### 2023-10-06 R/common.R

@@ -1,0 +1,36 @@
+N_CORES <- as.numeric(Sys.getenv("N_CORES", unset = 1))
+N_THREADS <- as.numeric(Sys.getenv("N_THREADS", unset = 2))
+N_ITER <- as.numeric(Sys.getenv("N_ITER", unset = 2000))
+N_SUBSET <- as.numeric(Sys.getenv("N_SUBSET"))
+
+options(
+  mc.cores = N_CORES,
+  brms.backend = "cmdstanr",
+  brms.threads = N_THREADS
+)
+
+theme_set(
+  theme_linedraw(base_size = 10) +
+    theme(
+      strip.background = element_rect(
+        fill = NA,
+        color = NA
+      ),
+      strip.text = element_text(
+        colour = "black",
+        hjust = 0
+      ),
+      axis.text = element_text(
+        size = rel(0.75)
+      ),
+      panel.grid = element_blank(),
+      plot.tag.position = c(0, 1)
+    )
+)
+
+scale_x_hours <- list(
+  scale_x_continuous(
+    breaks = c(0, 15/60, 0.5, 45/60, 1, 2, 3),
+    labels = c(0, "15m", "30m", "45m", "1h", "2h", "3h")
+  )
+)

### 2023-10-18 ms.Rmd
Notebooks to reproducible manuscript.

@@
 ---
-title: ''
-shorttitle: 
-leftheader: 
+title: "Affective Uplift During Video Game Play: A Naturalistic Case Study"
+shorttitle: Affective Uplift During Play
+leftheader: Affective Uplift During Play
 author: 
   - name: Matti Vuorre
-    affiliation: 1
     corresponding: yes
-    address: Department of Social Psychology, Tilburg University
-    email: m.j.vuorre@tilburguniversity.edu
+    affiliation: "1,2*"
+    address: Tilburg University
+    email: mjvuorre@uvt.nl
+  - name: Nick Ballou
+    affiliation: "2*"
+  - name: Thomas Hakman
+    affiliation: "2"
+  - name: Kristoffer Magnusson
+    affiliation: "2,3"
+  - name: Andrew K. Przybylski
+    affiliation: "2"
+    address: Oxford Internet Institute
+    email: andy.przybylski@oii.ox.ac.uk
 affiliation:
   - id: 1
-    institution: Department of Social Psychology, Tilburg University
-abstract: |
+    institution: Tilburg School of Social and Behavioral Sciences, Tilburg University
+  - id: 2
+    institution: Oxford Internet Institute, University of Oxford
+  - id: 3
+    institution: Centre for Psychiatry Research, Department of Clinical Neuroscience, Karolinska Institutet, & Stockholm Health Care Services, Region Stockholm, Sweden
+authornote: |
+  \noindent \textbf{This pre-print is not yet peer-reviewed.} 
+  *MV and NB contributed equally to this work.
 
-keywords: 
-authornote: |
-  \noindent \textbf{Working paper. Proceed with caution.}
-wordcount: "`r wordcountaddin:::word_count(filename = 'ms.Rmd')`"
+wordcount: '`r scales::number(wordcountaddin:::word_count("ms.Rmd"), big.mark = ",")`'
 bibliography: references.bib
+keywords: "video games, well-being, mood, play behavior, telemetry"
+csl: apa.csl
 floatsintext: yes
-numbersections: false
 linenumbers: no
@@
 classoption: jou
-header-includes:
-  - |
-    \usepackage{float}
-    \setlength{\parskip}{4pt}
 output: 
   papaja::apa6_pdf:
-    number_sections: true
-    keep_tex: false
-    highlight: kate
+    number_sections: false
+    keep_tex: true
   papaja::apa6_docx: 
-    number_sections: true
-    keep_tex: false
+    number_sections: false
+editor_options: 
+  chunk_output_type: console
 ---
 
-```{r setup, include = FALSE}
-# Packages
+```{r}
+#| label: prepare
+#| cache: false
+
+library(scales)
+library(cmdstanr)
+library(splines)
+library(emmeans)
+library(posterior)
+library(lme4)
+library(brms)
+library(gt)
+library(ggstance)
+library(ggdist)
+library(patchwork)
 library(tidyverse)
+source("R/functions.R")
+source("R/common.R")
 
-# Output options
-opts_chunk$set(
-  eval = TRUE,
-  cache = TRUE,
-  warning = FALSE,
-  error = TRUE,
-  message = FALSE
+dir.create("models", FALSE)
+```
+
+# Introduction
+
+How do video games affect players' well-being?
+Games are often studied for their potential in catalyzing psychological change over timescales spanning from weeks to months [e.g., effects on school performance, depression, or life satisfaction, @sauterSocialContextGaming2021; @vuorreTimeSpentPlaying2022], and the surrounding public debate has typically focused on play's far-reaching consequences on players' mental health, social attitudes, or cognitive development [@fergusonDoesSexualizationVideo2022; @hilgardOverestimationActionGameTraining2019; @mathurFindingCommonGround2019].
+In stark contrast, typical play appears to be motivated by short-term goals, such as wanting to unwind after a long day, escape to a pleasant non-reality in the moment, or engage in uplifting social interaction over periods of hours [@bourgonjonPlayersPerspectivesPositive2016; @kahnTrojanPlayerTypology2015; @stensengAreThereTwo2021].
+Such short-term dynamics between play and affect can exist but need not necessarily accumulate into long-term impacts.
+For example, games might provide relief, relaxation, and brief improvements in mood over several hours [@riegerEatingGhostsUnderlying2015; @russonielloEffectivenessCasualVideo2009; @tyackRestorativePlayVideogames2020], after which the effects taper out as individuals return to their baseline moods.
+
+Understanding whether and when games' short-term effects emerge is critical for establishing games' potential for mood-related interventions, as well as for building a theoretical foundation for repeated short-term gaming experiences' long-term effects on mental health.
+At present, however, the prevalence and magnitude of these short-term effects remains poorly understood.
+Some existing evidence suggests that games can provide short-term boosts to well-being ["mood repair"; @bowmanMoodGameSelective2015; @tyackRestorativePlayVideogames2020], possibly to a greater extent than non-interactive media such as videos [@riegerEatingGhostsUnderlying2015].
+On the other hand, games might affect players negatively: Frustrating gaming experiences, for example, can lead to negative consequences such as immediate post-play aggression [@przybylskiCompetenceimpedingElectronicGames2014].
+
+Despite the above examples, the validity and generalizability of research on games' short-term affective effects remains limited by three challenges.
+First, a substantial portion of gameplay research has relied on artificial stimuli; games created or substantially modified by academic researchers.
+While such customized games allow for greater experimental control, they are unlikely to reflect actual games' rich complexity [@mcmahanConsiderationsUseCommercial2011].
+This issue of the limited ecological validity and generalizability of research stimuli (games) limits current inferences about popularly played games' psychological effects.
+
+The second challenge is providing an ecologically valid *context* for play.
+Research participants typically play games in (online or physical) labs that do not resemble the natural contexts of play, such as when, with whom, and why people choose to play.
+In lab settings, research participants play to satisfy study requirements, rather than the intrinsic motivations that typically lead them to play.
+While beneficial to clarifying causal inferences, the extrinsically motivated play behaviors necessary in lab studies might relate differently to well-being than intrinsically motivated naturally occurring play [@bruhlmannMotivationalProfilingLeague2020; @howardStudentMotivationAssociated2021].
+Therefore, results from such studies are less likely to accurately generalize to how games are played in the real world.
+
+The third challenge concerns the timescale of effects: How quickly do potential effects emerge, and how long are they sustained?
+For example, some studies indicate that by the end of a half-hour game session, players may exhibit changes in stress [@russonielloEffectivenessCasualVideo2009], aggressive affect [@przybylskiCompetenceimpedingElectronicGames2014], and vitality [@tyackRestorativePlayVideogames2020].
+When and how video games' effects evolve during the initial half-hour remains unclear and difficult to study because researchers are typically unable to ask questions at a sufficient temporal resolution, with notable exceptions of @boweyPredictingBeliefsNPC2021 and @frommelGatheringSelfReportData2021, who used non-player characters to ask questions directly within the game.
+However, they did not enquire about well-being, leaving it unclear when and how the affective dimensions of play change on short timescales.
+
+Here, we aim to address these three challenges to better understand how real play in natural contexts might predict mood on short timescales.
+Specifically, we examined an intensive longitudinal dataset from the popular commercially available game PowerWash Simulator [PWS; @vuorreIntensiveLongitudinalDataset2023], which includes mood questions embedded in the game itself, to ask three questions: First, to what extent does mood change from immediately before video game play to during play?
+Second, hyow heterogeneous are these changes in the population of similar players?
+And third, how do changes in mood develop over the course of a gaming session?
+
+## Methods
+
+<!-- Data wrangling code -->
+
+```{r}
+#| label: data-get
+
+# Download, unzip data from OSF PWS database
+path <- "data-raw/data.zip"
+if (!file.exists(path)) {
+  dir.create("data-raw", FALSE)
+  download.file(
+    url = "https://osf.io/download/j48qf/",
+    destfile = "data-raw/data.zip"
   )
+  dir.create("data", FALSE)
+  unzip(
+    zipfile = path,
+    files = c(
+      "data/demographics.csv",
+      "data/study_prompt_answered.csv"
+    )
+  )
+}
 
-# Plotting options
-theme_set(
-  theme_linedraw(base_size = 12) +
-    theme(
-      strip.background = element_rect(fill = "gray90", color = "gray90"),
-      strip.text = element_text(color = "black", hjust = 0),
-      panel.grid = element_blank()
+# Load data to R session
+dat <- read_csv(
+  "data/study_prompt_answered.csv",
+  col_select = c(
+    pid,
+    time = Time_utc,
+    duration = CurrentSessionLength,
+    prompt = LastStudyPromptType,
+    mood = response
   )
@@
 
+# Convert types and units
+dat <- dat |>
+  mutate(
+    pid = factor(pid),
+    # Mood to 0-1 scale
+    mood = mood / 1000,
+    # hours indicates session duration in hours
+    hours = duration / 60,
+    .keep = "unused"
+  )
 ```
 
-# References {-}
+```{r}
+#| label: session-indicator-create
+#| include: true
 
-::: {#refs custom-style="Bibliography"}
+dat <- dat |>
+  arrange(pid, time) |>
+  mutate(
+    new_session = hours < lag(hours, default = 999),
+    session = cumsum(new_session),
+    .by = pid
+  ) |>
+  select(-c(new_session)) |>
+  mutate(ps = paste(pid, session, sep = "_"))
+
+dat_sum <- dat |>
+  summarize(
+    when = "raw",
+    n_pid = length(unique(pid)),
+    n_session = length(unique(ps)),
+    n_obs = n()
+  )
+```
+
+```{r}
+#| label: mood-filter
+
+# Include only wellbeing responses that occurred
+# in sessions with both "pre" and "post" measures
+dat <- dat |>
+  filter(prompt == "Wellbeing") |>
+  select(-prompt)
+
+dat_sum <- dat |>
+  summarize(
+    when = "mood",
+    n_pid = length(unique(pid)),
+    n_session = length(unique(ps)),
+    n_obs = n()
+  ) |>
+  bind_rows(dat_sum)
+```
+
+```{r}
+#| label: fig-session-durations
+#| include: false
+#| fig-height: 3
+#| fig.cap: Summaries of session durations. X-axes are truncated at 10 hours.
+
+p_durations <- dat |>
+  filter(hours == max(hours), .by = ps) |>
+  ggplot(aes(hours)) +
+  scale_x_continuous(
+    "Session duration",
+    expand = expansion(c(0.05, 0.05))
+  ) +
+  scale_y_continuous(
+    "Sessions",
+    expand = expansion(c(0, 0.05))
+  ) +
+  coord_cartesian(xlim = c(0, 10)) +
+  geom_histogram(binwidth = 0.33)
+
+p_mean_durations <- dat |>
+  filter(hours == max(hours), .by = ps) |>
+  summarise(hours = mean(hours), n = n(), .by = pid) |>
+  ggplot(aes(hours)) +
+  scale_x_continuous(
+    "Mean session duration",
+    expand = expansion(c(0.05, 0.05))
+  ) +
+  scale_y_continuous(
+    "Players",
+    expand = expansion(c(0, 0.05))
+  ) +
+  coord_cartesian(xlim = c(0, 10)) +
+  geom_histogram(binwidth = 0.33)
+
+p_ecdf_durations <- dat |>
+  filter(hours == max(hours), .by = c(session, pid)) |>
+  ggplot(aes(hours)) +
+  scale_x_continuous(
+    "Session duration",
+    expand = expansion(c(0.05, 0.05))
+  ) +
+  scale_y_continuous(
+    "Cumulative proportion",
+    expand = expansion(0.01),
+    breaks = (0:10) / 10
+  ) +
+  stat_ecdf() +
+  coord_cartesian(xlim = c(0, 10))
+
+p_durations | p_mean_durations | p_ecdf_durations
+
+dat <- dat |>
+  filter(hours <= 5)
+
+dat_sum <- dat |>
+  summarize(
+    when = "5h",
+    n_pid = length(unique(pid)),
+    n_session = length(unique(ps)),
+    n_obs = n()
+  ) |>
+  bind_rows(dat_sum)
+```
+
+```{r}
+#| label: wrangle-more
+#| include: false
+
+# Drop rows with missing mood responses
+dat <- dat |>
+  drop_na(mood) |>
+  mutate(pid = fct_drop(pid))
+
+# Create indicators for analyses
+dat <- dat |>
+  mutate(
+    post = factor(
+      row_number() > 1,
+      levels = c(FALSE, TRUE),
+      labels = c("0", "1")
+    ),
+    .by = c(pid, session)
+  ) |>
+  # Censoring indicator
+  mutate(
+    cl = case_when(
+      mood == 0 ~ "left",
+      mood == 1 ~ "right",
+      TRUE ~ "none"
+    )
+  )
+
+dat_sum <- dat |>
+  summarize(
+    when = "non-na",
+    n_pid = length(unique(pid)),
+    n_session = length(unique(ps)),
+    n_obs = n()
+  ) |>
+  bind_rows(dat_sum)
+
+# How many sessions had pre-mood & post-mood
+dat |>
+  mutate(
+    has_pre = any(hours == 0),
+    has_post = any(hours != 0),
+    has_both = has_pre & has_post,
+    .by = ps
+  ) |>
+  distinct(ps, has_pre, has_post, has_both) |>
+  summarise(
+    n_pre = number2(sum(has_pre)),
+    n_post = number2(sum(has_post)),
+    n_both = number2(sum(has_both)),
+    p_pre = percent2(mean(has_pre)),
+    p_post = percent2(mean(has_post)),
+    p_both = percent2(mean(has_both))
+  )
+
+# Use contrast coding with unit difference
+contrasts(dat$post) <- c(-0.5, 0.5)
+
+# Write data for any supplementary notebooks
+dir.create("data", FALSE)
+write_rds("data/data.rds")
+```
+
+<!-- End code -->
+
+In this study, we analyzed data from a large open dataset on PowerWash Simulator (PWS) play and psychological experiences [@vuorreIntensiveLongitudinalDataset2023].
+The data was collected in a research edition of PWS that recorded gameplay events, game status records, and responses to psychological survey items.
+We developed the research edition of PWS in collaboration with PWS's developer, FuturLab, who made it freely available on Steam to anyone who owned the original game (£19.99 on 2023-09-20).
+From the players' perspective, the research edition was nearly identical to that of the main game with the addition of in-game pop-ups that inquired about psychological states during play.
+
+### PowerWash Simulator
+
+PWS is a first-person simulation game developed by FuturLab.
+In the game, players run a small power washing business and take jobs from a variety of clients in different locations in the form of levels.
+The core mechanic of PWS is aiming and using a pressure washer to remove dirt from various objects and levels, ranging from Ferris wheels to skateparks.
+Progression happens sequentially through a career mode in which the player earns credits for cleaning objects and completing cleaning jobs.
+These credits can be used to upgrade the pressure washer to increase its range and effectiveness, as well as to purchase cosmetic modifications for the washer or avatar.
+The game offers a multiplayer mode which was disabled in the research version.
+
+Critically, in addition to regular gameplay, the research edition surfaced psychological survey items to the player during play sessions.
+These survey items were integrated into the game as pop-ups using the existing in-game character dialogue system and delivered by a newly created character called "The Researchers" making them both conversational and part of the game lore, ensuring minimal disruption to the play experience.
+The maximum number of questions per hour was six, with a window of at least five minutes in between pop-ups.
+In addition, at the beginning of each play session, at player login, there was a 10% probability that the player was asked a question about their mood before starting play.
+Furthermore, players were also given the option to self-report mood in the main menu once every 30 minutes, but we excluded those menu reports in this manuscript.
+
+### Participants
+
+```{r}
+#| label: demographics
+
+demo <- read_csv(
+  "data/demographics.csv",
+  col_select = c(pid, country, gender, age)
+) |>
+  filter(pid %in% unique(dat$pid))
+
+age <- median_qi(demo, age, .width = .8, na.rm = TRUE)
+
+age <- str_glue("{age[1]} ({age[2]}, {age[3]}; 1st and 9th deciles)")
+
+gender <- count(demo, gender, sort = TRUE) |>
+  mutate(p = n / sum(n)) |>
+  mutate(x = str_glue("{gender} ({number2(n)}, {percent2(p)})")) |>
+  slice(1:4) |>
+  pull(x) |>
+  paste(collapse = ", ")
+
+n_country <- length(unique(demo$country[!is.na(demo$country)]))
+country <- count(demo, country, sort = TRUE) |>
+  mutate(p = n / sum(n)) |>
+  mutate(x = str_glue("{country} ({number2(n)}, {percent2(p)})")) |>
+  slice(1:4) |>
+  pull(x) |>
+  knitr::combine_words()
+```
+
+The characteristics of the full sample of 11,080 players in the PWS dataset are described in @vuorreIntensiveLongitudinalDataset2023; here, we describe the subset of data relevant to our questions (see Data analysis below).
+All participants were over 18 years old, provided informed consent, answered at least one mood question, and did not request their data to be deleted.
+The median age was `r age`, and the four most frequent gender responses were `r gender`.
+Participants played in `r n_country` countries, with the `r country` being the most represented.
+Recruitment happened in multiple waves through multiple avenues inside and outside of the game [@vuorreIntensiveLongitudinalDataset2023].
+Study participation was incentivized through cosmetic in-game rewards (e.g.
+item skins).
+For every 12 questions answered, players could unlock a reward, of which five were available.
+These rewards could only be unlocked in the research version but were usable in both the research and main versions of PWS.
+
+The study procedures were granted ethical approval by Oxford University's Central University Research Ethics Committee (SSH_OII_CIA_21_011).
+
+### Measures
+
+We measured mood with a single item: "How are you feeling right now?" [@killingsworthWanderingMindUnhappy2010].
+Participants responded using a visual analogue scale (VAS) with endpoints "Very bad" and "Very good" that recorded 1000 possible values, which we rescaled to the unit interval (0-1) for this study.
+Consequently, our results can also be interpreted on the "[proportion] of maximum possible" scale [POMP; @cohenProblemUnitsCircumstance1999].
+While well-being is often studied with multi-item scales to differentiate between dimensions of positive and negative affect, the frequent probing of mental states in this study required a minimally intrusive instrument that would interrupt the participants' play experience as little as possible.
+Moreover, such single-item assessments have previously been validated and are recommended for intensive longitudinal studies [@songExaminingConcurrentPredictive2022].
+
+### Data analysis
+
+For the analyses reported here, we used a subset of the data in @vuorreIntensiveLongitudinalDataset2023 that was relevant to our questions.
+The full dataset contains `r number2(filter(dat_sum, when=="raw")$n_obs)` in-game survey responses, but here we ignored the enjoyment, focus, autonomy, competence, and immersion items and focused on the `r number2(filter(dat_sum, when=="mood")$n_obs)` mood responses from `r number2(filter(dat_sum, when=="mood")$n_session)` sessions and `r number2(filter(dat_sum, when=="mood")$n_pid)` players.
+We then excluded sessions longer than 5 hours in duration (`r percent2(1 - filter(dat_sum, when=="5h")$n_session / filter(dat_sum, when=="mood")$n_session)`) and dropped all responses with missing values (`r percent2(1 - filter(dat_sum, when=="non-na")$n_session / filter(dat_sum, when=="5h")$n_session)`).
+We made these decisions to reduce the complexity of our anticipated models, and under the belief that very long sessions are likely to be qualitatively different, and very rare, compared to typically shorter sessions.
+Our final dataset consisted of `r number2(filter(dat_sum, when=="non-na")$n_obs)` mood responses from `r number2(filter(dat_sum, when=="non-na")$n_session)` sessions and `r number2(filter(dat_sum, when=="non-na")$n_pid)` players.
+
+Our first and main research question concerned the difference between players' moods at the beginning of each session (pre-play) and during the subsequent play session (during play).
+This contrast does not represent a causal hypothesis: Players could begin (and end) their play sessions for whatever reason, and these reasons are likely to confound the pre – during contrast.
+For example, a player might come home after a stressful day at work and then play PWS.
+Coming home from a stressful work environment might then cause the person to both (1) choose to play, and (2) experience an elevated mood, in which case we would be in error if we attributed play itself the position of causal antecedent of any potential mood consequences.
+Generally, reasons for starting to play are likely to contribute to the pre-during contrast and we are unable to disentangle those from any changes specifically caused by play.
+
+We estimated this contrast within a three-level hierarchical regression model that nested observations within sessions, and sessions within players.
+We specified a censored Gaussian model of mood because a VAS necessarily limits response options at the lower and upper ends.
+Ignoring censoring would leave the contrast susceptible to ceiling or floor effects and might confound changes in the mood distribution's location with changes in its scale.
+We then modelled mean mood on an intercept and contrast coded time (pre-play: -0.5; during play: 0.5) and allowed both parameters to vary randomly across players.
+In addition, we modelled random intercepts over player sessions.
+Although equal residual variances across people in natural observation seem unlikely, we estimated only one residual deviation parameter to limit model complexity.
+
+We analyzed the data with R and used the brms package to estimate, via Stan's HMC sampling algorithm, and post-process the models [@burknerBrmsPackageBayesian2017; @rcoreteamLanguageEnvironmentStatistical2023; @standevelopmentteamStanModelingLanguage2021].
+We drew `r number2(N_ITER*4/2)` samples from the model's posterior distribution using brms' default prior distributions on all parameters and used numerical and graphical checks to ensure model convergence and adequacy.
+
+```{r}
+#| label: model-1-brms
+
+model <- bf(
+  mood | cens(cl) ~
+    1 + post +
+    (1 + post | pid) +
+    (1 | ps)
+) +
+  gaussian()
+
+fit1 <- brm(
+  model,
+  data = dat,
+  silent = 0,
+  iter = N_ITER,
+  control = list(adapt_delta = .95),
+  file = "models/brm-prepost-pid-ps-censored"
+)
+```
+
+```{r}
+#| label: fig-convergence
+#| fig.cap: Convergence diagnostic plot showing bivariate scatterplots of the model's population-level parameters' posterior draws.
+#| eval: false
+
+pairs(
+  fit1,
+  variable = c("b_", "sd_", "cor_"),
+  regex = TRUE,
+  off_diag_args = list(size = .33, shape = 1)
+)
+```
+
+## Results
+
+```{r}
+#| label: fig-descriptives
+#| fig-height: 2.2
+#| fig-width: 7
+#| include: false
+#| fig.cap: 'A. Histogram of session durations (note log10 x-axis). B. Summary of how many sessions each participant completed. C. Histogram of all mood ratings.'
+
+p0 <- dat |>
+  ggplot() +
+  scale_x_continuous(
+    expand = expansion(c(0.01, 0.01)),
+  ) +
+  scale_y_continuous(
+    expand = expansion(c(0.001, 0.05)),
+  ) +
+  geom_histogram(
+    col = "white",
+    bins = 50,
+    linewidth = .25,
+    boundary = 0
+  )
+
+dat_session_durations <- dat |>
+  filter(hours == max(hours), .by = c(pid, session))
+
+dat_session_counts <- dat |>
+  distinct(pid, session) |>
+  count(pid, name = "sessions")
+
+p_durations <- p0 %+%
+  dat_session_durations +
+  aes(hours) +
+  labs(x = "Session duration (hours)", y = "Sessions")
+
+p_sessions <- p0 %+%
+  dat_session_counts +
+  coord_cartesian(xlim = c(0, 75)) +
+  aes(sessions) +
+  labs(x = "Sessions", y = "Participants")
+
+p_mood <- p0 +
+  aes(mood) +
+  labs(x = "Mood", y = "Responses")
+
+(p_durations | p_sessions | p_mood) +
+  plot_annotation(tag_levels = "A")
+```
+
+```{r}
+#| include: false
+tmp <- list(
+  dat_session_durations$hours,
+  dat_session_counts$sessions,
+  dat$mood,
+  dat$mood[dat$post == 0],
+  dat$mood[dat$post == 1]
+) |>
+  map(
+    ~ median_qi(.x, .width = 0.8, na.rm = TRUE)
+  ) |>
+  map(
+    ~ mutate(
+      .x,
+      across(
+        c(y, ymin, ymax),
+        ~ number2(., .01)
+      )
+    ) |>
+      str_glue_data("{y} [{ymin}, {ymax}]")
+  )
+tmp
+```
+
+The median session duration was `r tmp[[1]]` hours [10 and 90 percentiles]; the median player contributed data from `r tmp[[2]]` sessions, and the median mood was `r tmp[[3]]` (pre-session: `r tmp[[4]]`, during play: `r tmp[[5]]`).
+We illustrate these basic features of the data in Figure \@ref(fig:fig-data).
+
+### RQ1: Mood changes from pre- to during play
+
+```{r}
+#| label: fig-data
+#| include: true
+#| fig-width: 8
+#| fig-height: 5
+#| fig.env: "figure*"
+#| fig.cap: "A.
+Scatterplots of three participants' (rows) mood responses (pre-play: red; during play: blue) over eight sessions' (columns) durations.
+B.
+Histograms of session-mean (C.
+person-mean) moods before (top) and during (bottom) play sessions.
+D.
+Differences in session-mean (E.
+player-mean) mood differences (during session - pre-play).
+F.
+Scatterplot of person-mean mood reports at the beginning (x-axis) and during gameplay sessions (y-axis).
+Identity line is shown in green, and an exploratory GAM regression line is shown in blue."
+
+# Plot: Example raw sessions
+set.seed(99)
+dat_example_sessions <- dat |>
+  add_count(pid, session, name = "obs_per_session") |>
+  filter(obs_per_session >= 4) |>
+  mutate(session = as.numeric(as.factor(session)), .by = pid) |>
+  mutate(session_per_person = length(unique(session)), .by = pid)
+
+# default
+min_sessions <- 8
+n_filtered <- dat_example_sessions |>
+  distinct(pid, session_per_person) |>
+  filter(session_per_person >= 8) |>
+  nrow()
+
+# pick anyone if there's not 3 pids with >= 4 sessions
+if (n_filtered < 3) min_sessions <- 0
+dat_example_sessions <- dat_example_sessions |>
+  filter(session_per_person >= min_sessions) |>
+  arrange(pid, session) |>
+  filter(pid %in% sample(unique(pid), 3)) |>
+  mutate(
+    Session = str_glue("Session {session}"), 
+    Person = str_glue("Person {fct_anon(pid)}")
+  )
+
+p_mood_example <- dat_example_sessions |>
+  filter(session <= 8) |>
+  ggplot() +
+  aes(hours, mood, col = post) +
+  scale_color_brewer(
+    "Pre-session measure",
+    palette = "Set1",
+    aesthetics = c("color", "fill")
+  ) +
+  scale_x_continuous(
+    "Session duration (hours)"
+  ) +
+  scale_y_continuous(
+    "Mood",
+    limits = c(0, 1),
+  ) +
+  geom_point(size = 1.5, alpha = 1) +
+  facet_grid(
+    rows = vars(Person),
+    cols = vars(Session),
+    scales = "fixed"
+  ) +
+  theme(
+    legend.position = "none",
+    strip.text.y = element_blank()
+  )
+
+# Plot: Raw mood histograms at pre and during play
+p_mood_prepost_raw <- dat |>
+  mutate(post = factor(post, labels = c("Pre-play", "During play"))) |>
+  ggplot(aes(mood, fill = post)) +
+  scale_color_brewer(
+    "Pre-session measure",
+    palette = "Set1",
+    aesthetics = c("color", "fill")
+  ) +
+  geom_histogram(bins = 30, col = "white") +
+  scale_y_continuous(
+    "Observations",
+    expand = expansion(c(0.001, 0.1)),
+    breaks = scales::extended_breaks(5)
+  ) +
+  scale_x_continuous(
+    "Mood",
+    expand = expansion(c(0.001))
+  ) +
+  facet_wrap("post", ncol = 1, scales = "free_y") +
+  theme(legend.position = "none")
+
+# Plot: Person-session-mean mood histograms at pre and during play
+p_mood_prepost_sessions <- p_mood_prepost_raw %+%
+  summarise(
+    p_mood_prepost_raw$data,
+    mood = mean(mood, na.rm = TRUE), .by = c(post, pid, ps)
+  ) +
+  scale_y_continuous(
+    "Sessions",
+    expand = expansion(c(0.001, 0.1)),
+    breaks = scales::extended_breaks(5)
+  )
+
+# Plot: Person-mean mood histograms at pre and during play
+p_mood_prepost_players <- p_mood_prepost_sessions %+%
+  summarise(
+    p_mood_prepost_raw$data,
+    mood = mean(mood, na.rm = TRUE), .by = c(post, pid)
+  ) +
+  scale_y_continuous(
+    "Players",
+    expand = expansion(c(0.001, 0.1)),
+    breaks = scales::extended_breaks(5)
+  )
+
+# Plot: Difference histogram (sessions)
+p_mood_difference_sessions <- p_mood_prepost_sessions$data |>
+  pivot_wider(names_from = post, values_from = mood) |>
+  mutate(Difference = `During play` - `Pre-play`) |>
+  drop_na(Difference) |>
+  ggplot(aes(Difference)) +
+  geom_histogram(bins = 50, col = "white") +
+  geom_vline(xintercept = 0, linewidth = .5, col = "#2ca25f") +
+  scale_y_continuous(
+    "Sessions",
+    expand = expansion(c(0.001, 0.1))
+  ) +
+  scale_x_continuous(
+    expand = expansion(c(0.001))
+  ) +
+  coord_cartesian(xlim = c(-.4, .6))
+
+# Plot: Difference histogram (players)
+p_mood_difference_players <- p_mood_prepost_players$data |>
+  pivot_wider(names_from = post, values_from = mood) |>
+  mutate(Difference = `During play` - `Pre-play`) |>
+  drop_na(Difference) |>
+  ggplot(aes(Difference)) +
+  geom_histogram(bins = 50, col = "white") +
+  geom_vline(xintercept = 0, linewidth = .5, col = "#2ca25f") +
+  scale_y_continuous(
+    "Players",
+    expand = expansion(c(0.001, 0.1))
+  ) +
+  scale_x_continuous(
+    expand = expansion(c(0.001))
+  ) +
+  coord_cartesian(xlim = c(-.4, .6))
+
+p_mood_difference <- (
+  (p_mood_difference_sessions +
+     theme(
+       axis.title.x = element_blank(),
+       axis.text.x = element_blank(),
+       axis.ticks.x = element_blank()
+     )) /
+    p_mood_difference_players
+)
+
+p_mood_biscatter_sessions <- p_mood_difference_sessions$data |>
+  ggplot(aes(`Pre-play`, `During play`)) +
+  scale_x_continuous(
+    expand = expansion(0.01)
+  ) +
+  scale_y_continuous(
+    expand = expansion(0.01)
+  ) +
+  geom_point(
+    alpha = .2, size = 0.33, shape = 1
+  ) +
+  geom_abline(linewidth = .5, col = "#2ca25f") +
+  geom_smooth(
+    method = "gam",
+    se = FALSE,
+    linewidth = .75,
+    col = "dodgerblue"
+  ) +
+  theme(aspect.ratio = 1)
+
+p_mood_biscatter_players <- p_mood_difference_players$data |>
+  ggplot(aes(`Pre-play`, `During play`)) +
+  scale_x_continuous(
+    expand = expansion(0.01)
+  ) +
+  scale_y_continuous(
+    expand = expansion(0.01)
+  ) +
+  geom_point(
+    alpha = .2, size = 0.33, shape = 1
+  ) +
+  geom_abline(linewidth = .5, col = "#2ca25f") +
+  geom_smooth(
+    method = "gam",
+    se = FALSE,
+    linewidth = .75,
+    col = "dodgerblue"
+  ) +
+  theme(aspect.ratio = 1)
+
+p_mood_example /
+  (
+    p_mood_prepost_sessions |
+      p_mood_prepost_players |
+      p_mood_difference |
+      p_mood_biscatter_players
+  ) +
+  plot_layout(heights = c(5, 5)) +
+  plot_annotation(tag_levels = "A")
+```
+
+We first focused on our primary research question: To what extent do PWS players' mood change from pre-play to during play?
+We visualized the relevant data in Figure \@ref(fig:fig-data): Panel A shows mood responses from three example participants' first eight sessions of play.
+Figure \@ref(fig:fig-data) B (C) then shows histograms of all sessions' (players') aggregated pre- and during play moods to facilitate visual comparison of the raw data.
+We show the differences in these aggregated moods in Figure \@ref(fig:fig-data) D (sessions) and E (players).
+Moreover, Figure \@ref(fig:fig-data) F shows the difference in player-mean pre- and during play moods across different values of pre-play moods.
+Overall, these figures suggested small increases in mood from pre- to during play, but also that there were broad distributions of this difference over sessions and players, and that the difference was greater for lower pre-play moods (Panel F).
+
+```{r}
+#| label: tbl-avg
+#| include: true
+#| tbl-cap: Summaries of key population-level estimates
+
+fit_tbl <- as_draws_df(
+  fit1,
+  c("b_", "sd_", "sigma"),
+  regex = TRUE
+) |>
+  transmute(
+    `Pre-play` = b_Intercept + b_post1 * -0.5,
+    `During play` = b_Intercept + b_post1 * 0.5,
+    Difference = b_post1,
+    `Difference (scaled)` = Difference /
+      (sd_pid__Intercept + sd_pid__post1 + sd_ps__Intercept + sigma),
+    `(SD) Difference` = sd_pid__post1,
+    `Positive shifts` = pnorm(0, b_post1, sd_pid__post1, lower.tail = FALSE)
+  ) |>
+  summarise_draws(
+    mean = ~ mean(.),
+    ~ quantile2(., probs = c(.025, .975))
+  ) |>
+  mutate(
+    res = str_glue("{number2(mean, .001)} [{number2(q2.5, .001)}, {number2(q97.5, .001)}]"),
+    resp = str_glue("{percent2(mean, .1)} [{percent2(q2.5, .1)}, {percent2(q97.5, .1)}]"),
+  )
+
+# Show this one as a percentage
+fit_tbl[6, "res"] <- fit_tbl[6, "resp"]
+
+fit_tbl |> 
+  mutate(Variable = variable, Estimate = res, .keep = "none") |> 
+  papaja::apa_table(
+    span_text_columns = FALSE, 
+    caption = "Summaries of the hierarchical model's key population-level estimates.",
+    note = "Numbers indicate posterior means and 95%CIs. Difference (scaled) is the standardized during play--pre-play difference."
+  )
+```
+
+We then turned to the model's results regarding (differences) in players' moods.
+They confirmed the visual impressions described above: Table \@ref(tab:tbl-avg) indicates that the average PWS player experiences a `r filter(fit_tbl, variable=="Difference")$res` unit increase in mood during PWS play, on a VAS from 0 to 1.
+We also interpreted this difference in a different light by dividing it by the total random variation estimated by the model.
+This standardized pre-play – during play contrast was `r filter(fit_tbl, variable=="Difference (scaled)")$res`.
+
+### RQ2: Heterogeneity in mood changes
+
+Although the above quantities indicated that the average player's mood did increase from before to during play, they say little about the robustness of this difference in the population of similar players.
+We therefore next turned to our second research question: How heterogeneous are mood shifts in the population of similar PWS players?
+First, the standard deviation of the person-specific mood differences was `r filter(fit_tbl, variable=="(SD) Difference")$res`.
+In comparison to the average person's estimated difference (`r filter(fit_tbl, variable=="Difference")$res`), that quantity indicated a moderate degree of heterogeneity between individuals.
+More specifically, the model-predicted proportion of individuals expected to experience positive changes from pre- to during play was `r filter(fit_tbl, variable=="Positive shifts")$resp`.
+
+In sum, the results from our model contrasting pre- and during play moods indicated small increases in mood during play, and that those changes were robust across people.
+
+### RQ3: Time course of mood changes during play
+
+The above analysis provides an easily interpretable contrast between during-play moods and moods just before play.
+However, it does not address the time course of moods *within* the sessions.
+We therefore next turned to our third question: How do (changes in) players' moods evolve during gameplay sessions?
+To answer, we used time as a continuous predictor and allowed mood changes during sessions to be non-linear by estimating a piecewise cubic spline with 4 degrees of freedom.
+Just like the main model, this was a three-level hierarchical model, with random intercepts at the session and participant levels, and random participant slopes for each piece of the spline.
+Moreover, in a separate model, we also examined how within-session change related to mood before play by including pretest values as a covariate and modeled the hour-by-pretest continuous interaction.
+We also modeled pretest mood with a cubic spline to allow the relationship to be non-linear.
+
+```{r}
+#| label: model-2-3-lmer
+
+fit2 <- fit_cached(
+  "models/lmm-ns4.Rds",
+  lmer(
+    mood ~ ns4(hours) + (1 | ps) + (1 + ns4(hours) | pid),
+    data = dat
+  )
+)
+
+dat_pre <- dat |>
+  filter(
+    # Session has a wellbeing measure at time = 0
+    hours[1] == 0,
+    .by = c(pid, session)
+  ) |>
+  mutate(
+    pre = mood[1],
+    .by = c(pid, session)
+  ) |>
+  filter(
+    hours > 0
+  ) |>
+  droplevels()
+
+fit3 <- fit_cached(
+  "models/lmm-pre-interaction.Rds",
+  lmer(
+    mood ~ ns4(hours) * ns(pre, 5) + (1 + ns4(hours) | pid) + (1 | ps),
+    data = dat_pre
+  )
+)
+```
+
+The main model without an interaction with pre-play mood included all mood responses, sessions, and participants as above.
+However, the interaction model required each session to have a pre-play mood measure, which led to `r number2(lme4::ngrps(fit3)["pid"])` players, `r number2(lme4::ngrps(fit3)["ps"])` sessions, and `r number2(nobs(fit3))` observations included in that model.
+
+We chose not to use model censoring in these models due to the increased computational cost.
+However, we performed sensitivity analyses with and without censoring on a reduced data set (1000 participants), which indicated nearly identical results.
+At worst, ignoring censoring resulted in slightly different intercepts; that is, the whole curve was shifted up or down.
+
+```{r}
+#| label: fig-ct1
+#| include: true
+#| fig.cap: "Estimated (changes) in mood as a function of session duration.
+Top: Average mood during a gaming session.
+Bottom: Change in mood during a session compared to mood at the beginning of a session.
+Gray ribbons indicate 95\\% confidence bands.
+We truncated the x-axis at three hours for this figure."
+#| fig-height: 3
+#| fig-width: 5
+
+XMAX <- 3
+XOUT <- 100
+hours <- seq(0, XMAX, length.out = XOUT)
+emm2_mood <- emmeans(
+  fit2,
+  ~hours,
+  at = list(hours = hours),
+  lmer.df = "asymptotic"
+)
+
+emm2_diff <- contrast(
+  emm2_mood,
+  method = "trt.vs.ctrl",
+  ref = "hours0"
+) |>
+  confint() |>
+  as.data.frame() |>
+  mutate(
+    hours = hours[-1],
+    res = str_glue(
+      "{number2(estimate, .001)} ",
+      "[{number2(asymp.LCL, .001)}, {number2(asymp.UCL, .001)}]"
+    )
+  )
+
+bind_rows(
+  "Mood" = as_tibble(emm2_mood),
+  "Difference" = as_tibble(emm2_diff) |>
+    rename(emmean = estimate),
+  .id = "x"
+) |>
+  mutate(x = fct_rev(x)) |>
+  ggplot(aes(hours, emmean)) +
+  scale_y_continuous(
+    "Value",
+    breaks = extended_breaks(7)
+  ) +
+  geom_hline(
+    data = tibble(
+      x = c("Mood", "Difference") |> fct_rev(),
+      y = c(NaN, 0)
+    ),
+    aes(yintercept = y),
+    linewidth = .5, col = "#2ca25f"
+  ) +
+  geom_line() +
+  geom_ribbon(
+    aes(
+      ymin = asymp.LCL,
+      ymax = asymp.UCL
+    ),
+    alpha = 0.25
+  ) +
+  scale_x_continuous(
+    "Session duration (hours)",
+    breaks = c(0, 0.5, 1, 2, 3, 4, 5),
+    labels = c("0m", "30m", "1h", "2h", "3h", "4h", "5h"),
+    expand = expansion(0.01)
+  ) +
+  facet_wrap(
+    "x",
+    ncol = 1,
+    scales = "free_y",
+    strip.position = "left"
+  ) +
+  theme(
+    axis.title.y = element_blank(),
+    strip.placement = "outside",
+    strip.text = element_text(size = rel(1), hjust = 0.5)
+  )
+```
+
+```{r}
+#| label: fig-ct2
+#| fig.env: "figure*"
+#| include: true
+#| fig.cap: "Estimated (changes) in mood as a function of session duration and pre-play mood.
+Top.
+Mood for the average player during a gaming session with pre-play mood at 5th, 25th, 50th, and 75th percentiles (columns).
+Ribbons indicate 95\\% confidence.
+Bottom.
+Same as above but with change in mood on the y-axis."
+#| fig-height: 3
+#| fig-width: 8
+
+pre <- dat_pre |>
+  summarise(pre = pre[1], .by = ps) |>
+  pull(pre)
+pre_values <- quantile(pre, c(0.05, .25, 0.5, 0.75))
+pre_labels <- c("5th", "25th", "Median", "75th")
+
+emm3_mood <- emmeans(
+  fit3,
+  ~ hours + pre,
+  at = list(
+    hours = hours,
+    pre = pre_values
+  ),
+  lmer.df = "asymptotic"
+)
+
+emm3_diff <- emm3_mood |>
+  contrast(
+    method = "trt.vs.ctrl",
+    ref = "hours0",
+    by = "pre"
+  ) |>
+  confint() |>
+  mutate(
+    hours = rep(hours[-1], length(pre_values)),
+    emmean = estimate,
+    .keep = "unused"
+  )
+
+bind_rows(
+  "Mood" = as_tibble(emm3_mood),
+  "Difference" = as_tibble(emm3_diff),
+  .id = "x"
+) |>
+  mutate(x = fct_rev(x)) |>
+  mutate(pre = factor(pre, levels = pre_values, labels = pre_labels)) |>
+  ggplot(aes(hours, emmean)) +
+  scale_y_continuous(
+    "Value",
+    breaks = extended_breaks(7)
+  ) +
+  geom_hline(
+    data = tibble(
+      x = c("Mood", "Difference") |> fct_rev(),
+      y = c(NaN, 0)
+    ),
+    aes(yintercept = y),
+    linewidth = .5, col = "#2ca25f"
+  ) +
+  geom_line() +
+  geom_ribbon(
+    aes(
+      ymin = asymp.LCL,
+      ymax = asymp.UCL
+    ),
+    alpha = 0.25
+  ) +
+  scale_x_continuous(
+    "Session duration (hours)",
+    breaks = c(0, 0.5, 1, 2, 3, 4, 5),
+    labels = c("0", "30m", "1h", "2h", "3h", "4h", "5h"),
+    expand = expansion(0.01)
+  ) +
+  facet_grid(
+    x ~ pre,
+    scales = "free_y",
+    switch = "y",
+  ) +
+  theme(
+    axis.title.y = element_blank(),
+    strip.placement = "outside",
+    strip.text = element_text(size = rel(1), hjust = 0.5)
+  )
+```
+
+This continuous time analysis added three important nuances to the simpler pre-during play contrast presented above.
+First, Figure \@ref(fig:fig-ct1) shows how the average mood increased during a session, suggesting a small but sharp uplift early during a session, and slightly greater in magnitude to the pre-during contrast.
+Second, the bulk of this increase occurred early in play sessions, with an increase of `r slice(emm2_diff, floor(XOUT/XMAX*0.25))$res` units for the average player after 15 minutes of play.
+Third, the rate and shape of change depended on the participants' initial mood levels.
+Figure \@ref(fig:fig-ct2) shows (changes in) estimated mood over a typical session based on different percentiles of pre-play mood, where the lower percentiles (5th percentile of pre-play mood = `r number2(pre_values[1], .01)`, and 25th = `r number2(pre_values[2], .01)`) showed greater uplift in moods during a session compared to median or greater pre-play mood levels.
+
+---
+abstract: Do video games affect players' well-being?
+In this case study, we examined `r number2(nrow(dat))` intensive longitudinal in-game mood reports from `r number2(length(unique(dat$ps)))` play sessions of `r number2(length(unique(dat$pid)))` players of the popular game PowerWash Simulator.
+We compared players' moods at the beginning of play session with their moods during play, and found that the average player reported  `r filter(fit_tbl, variable=="Difference")$res` visual analog scale (VAS; 0-1) units greater mood during than at the beginning of play sessions.
+Moreover, we predict that `r filter(fit_tbl, variable=="Positive shifts")$resp` of similar players experience this affective uplift during play, and that the bulk of it happens during the first 15 minutes of play.
+We do not know whether these results indicate causal effects or to what extent they generalize to other games or player populations.
+Yet, these results based on in-game subjective reports from players of a popular commercially available game suggest good external validity, and as such offer a promising glimpse of the scientific value of transparent industry-academia collaborations in understanding the psychological roles of popular digital entertainment.
+---
+
+## Discussion
+
+The current study corroborates what qualitative research and reports from video game players around the world have long suggested: People feel good playing games.
+Specifically, we find that playing a popular commercial video game, PowerWash Simulator, is linked with a small improvement in mood, that this improvement is experienced by `r filter(fit_tbl, variable=="Positive shifts")$resp` of players, and that the bulk of the improvement occurs during the first 15 minutes of play.
+
+Although the overall magnitude of the estimated change was small (`r filter(fit_tbl, variable=="Difference")$res`), it nevertheless is comparable to mood shifts following other frequent voluntary activities.
+For example, experience sampling studies of US [@killingsworthWanderingMindUnhappy2010] and Korean [@choiTakingStockHappiness2017] adults indicated that watching television (+2% change compared to an individual's average across all activities), reading (+2%), and shopping (+3%) were associated with smaller mood shifts than we observed in association with PWS.
+On the other hand, those studies indicated that listening to music (+7–9%), eating or cooking (+8%), taking a walk (+9%), visiting an urban green space [+9–10%; @tostNeuralCorrelatesIndividual2019], exercising (+11%), dating (26.5%), sexual intercourse (+28%), or taking a trip (+30.5%) correlate with greater shifts in well-being.
+
+Another interpretation of our result is that although the estimated PWS-play associated mood uplift is small, it is potentially large enough to be subjectively perceived.
+@anvariUsingAnchorbasedMethods2021 suggest that, on average, people are able to subjectively perceive a change of 2% in well-being on the related PANAS scale.
+Since our estimate of `r filter(fit_tbl, variable=="Difference")$res` is greater, we tentatively suggest that gaming, on average, is associated with mood uplifts that are large enough to be consciously felt by players.
+
+Moreover, it is possible that the uplift we report is an underestimate of the true effect of playing games.
+The estimated average pre-play mood in our study (`r filter(fit_tbl, variable=="Pre-play")$res`) was comparatively greater than average moods reported in previous experience sampling studies using the same measure [58%, @choiTakingStockHappiness2017; 65%, @killingsworthWanderingMindUnhappy2010].
+Assuming measurement equivalence, we might then hypothesize reasons for these between-sample differences: It is possible that our sample felt, on average, more positive than those other samples of US and Korean adults.
+Alternatively, it is possible that players in our study experienced an anticipatory mood benefit from playing video games.
+We believe the second of these is more likely: Anticipating an opportunity to play, and everything associated with the opportunity to play, is likely to immediately impact the player's mood, before play starts.
+
+By examining questionnaire items embedded in the game interface during natural play sessions, our study is the first to examine changes in mood during play sessions in a minute-by-minute resolution.
+These data indicate that the majority of the observed rise in affective well-being during play sessions occurs during the first 15 minutes of play, and that mood stabilizes for the following few hours without return to pre-play levels.
+These results suggest that, at least for certain kinds of games and keeping in mind our caveats regarding causality, many players might benefit from interspersing short play sessions into their day or throughout their leisure time.
+These patterns are already commonly seen in casual or idle games [@cuttingBusyDoingNothing2019], which share certain features with lower-demand simulation games such as PowerWash Simulator.
+
+Our results highlight at least three important unanswered questions.
+First, what are the mechanisms driving mood changes associated with video game play?
+Second, how long do mood changes last after gameplay?
+Third, under what conditions do the short-term "mood repair" effects of games accumulate and contribute to long-term increases in well-being?
+
+Current theories provide a few candidates regarding mood change mechanisms.
+Mood management theory [MMT; @zillmannMoodManagementCommunication1988] posits that players choose specific entertainment media as part of a (conscious or subconscious) strategy to regulate valence and arousal and to increase hedonic well-being.
+Specifically, MMT predicts that when players feel more negative emotions, they are more likely to select games with positive hedonic valence (i.e., games that are generally positive in tone), and that when players are in high arousal states such as anger stress, they are more likely to choose calm and relaxing games with.
+
+PowerWash Simulator provides an ideal case study for mood management theory, in that it is typically described as a low cognitive effort game: Players are known to play semi-idly with only partial focus, or even while consuming other media.
+As one reviewer writes, "When my brain is tired, or I'm not in the mood to compete or struggle in any way, I grab a power washer" [@marshallPowerWashSimulatorMy2021].
+Accordingly, PWS may be especially well-suited for players who are emotionally or cognitively depleted, such as those with persistently low mood [@helsbyBandwidthComesGoes2023]---something that is partially supported by our findings that sessions with lower initial moods were those with larger uplifts during play.
+Although outside the scope of this study, we believe that the additional telemetry in the PWS dataset [@vuorreIntensiveLongitudinalDataset2023] would be valuable in addressing more nuanced questions regarding MMT.
+
+Another candidate derives from self-determination theory [@ryanSelfdeterminationTheoryBasic2017].
+At least one laboratory study has shown that the experience of basic psychological need satisfaction during media consumption moderates the change from pre–post play well-being [@tyackRestorativePlayVideogames2020].
+Another found that need satisfaction in two commercial games, Animal Crossing: New Horizons and Plants vs Zombies: Battle for Neighborville, was independently associated with well-being, regardless of the amount of playtime [@johannesVideoGamePlay2021].
+It may be the case, then, that experiences with greater feelings of mastery and effectiveness (competence satisfaction) or control and volition (autonomy satisfaction) would result in larger mood increases during PWS play.
+While we did not test these causal mechanisms between play and well-being in the current study, we anticipate several theory-driven detailed investigations of the PWS dataset [@vuorreIntensiveLongitudinalDataset2023] that might do so.
+
+Second, we do not know whether the mood uplift associated with PWS play lingers or disappears after the gaming session ends.
+Answering this question would afford a better understanding of the motivations behind play: To what extent do people play games for their immediate hedonic benefits in contrast to the emotional and cognitive state players find themselves in after stopping?
+The difference between these two experiences might be an important differentiator between, for example, maladaptive *escaping from* and adaptive *escaping to* [@stensengAreThereTwo2021].
+To answer this question, however, in-game telemetry data will need to be complemented with out-of-game psychological measures assessing well-being in the period after play.
+We anticipate that experience sampling methods, which are common in social media research [@aalbersCaughtMomentAre2022; @verbeijExperienceSamplingSelfreports2022] but have yet to make meaningful inroads in video games research, would be particularly valuable in addressing this question.
+
+Finally, what does this result tell us about the long-term impacts of video game play on well-being?
+Observational studies relating well-being to time spent playing games over 2 weeks [@vuorreTimeSpentPlaying2022], 1 month [@sibillaHarmoniousObsessiveInvolvement2021], 6 months [@lemmensPsychosocialCausesConsequences2011; @weinsteinProspectiveStudyMotivational2017] and 1 year [@kowertPsychosocialCausesConsequences2015] suggest null or practically inconsequential relationships between play and well-being indicators such as affect and psychological health.
+Thus, we must square the current evidence for a short-term uplift in mood with existing evidence suggesting little to no meaningful long-term relations, and explicitly articulate why they would differ [@anvariNotAllEffects2023].
+Against that juxtaposition, we believe our findings are most consistent with the notion that gaming---for most people---is a recovery activity that helps to manage day-to-day stresses and mood fluctuations, without necessarily having substantial long-term impacts.
+The majority of players have several options for activities in their environment that would have comparable effects on their well-being.
+These activities are thus "exchangeable" and serve the same short-term goals without consequences to people's long-term adjustment.
+
+Moreover, while we have focused on discussing mood changes for the average player, our results also indicated moderate between-person heterogeneity in mood changes during play.
+Future studies might benefit from examining variations in mood shifts across types of players (and play), rather than focusing solely on the general gaming population, which necessarily varies widely in the games they play, how they play them, and their psychological characteristics.
+For example, disordered gaming is one type of play linked with negative effects [@karhulahtiPhenomenologicalStrandsGaming2022], and playing during difficult life circumstances is one linked with positive effects [@iacovidesRoleGamingDifficult2019].
+We believe that understanding the multiverse of play, including temporal patterns, social experiences, in-game behaviors and events, players' personalities, its antecedents, and consequences warrants continued research efforts coordinated across and beyond academia.
+
+### Limitations
+
+This study was not an experiment, nor did we employ methods required for rigorous causal inference from observational data, and therefore our results regarding the causal effects of video game play on mood are tentative at best.
+Without a control condition, we have nothing to compare our results to: We cannot say if the changes in mood observed during PWS play would have occurred with other games, non-game activities, or indeed no activity at all.
+For example, our results might indicate mood increases that are accounted for by starting a period of leisure time, and not play specifically.
+Future work should consider the use of randomized controlled trials to evaluate the effect of playing PowerWash simulator or other games compared to other leisure activities or therapeutic interventions.
+
+As always, the results here are likely to generalize in some ways but not in others.
+Our sample is somewhat representative with regard to gender and covers a wide age range roughly in line with the age demographics of general US adult video game players [@engelstatterVideoGamesBecome2022], suggesting that our results may generalize to other adult PWS players from Western countries.
+They are less likely to generalize to younger players and those from non-Western countries.
+The sample may further suffer from self-selection bias: Given that players voluntarily chose to download the PWS research edition, it is plausible that people who felt more positively toward the game (and more positively while playing it) were more likely to opt into the study.
+Finally, we studied just one commercial video game with a feature set that is very different from today's most commonly-played PC games [which at the time of writing include, for example, Minecraft, Fortnite, and Baldur's Gate 3, but only one other simulation game, The Sims 4; @newzooMostPopularPC2023].
+
+### Conclusion
+
+By investigating player experiences during natural play of a popular and commercially available game, we found strong evidence for a small positive change in mood over the course of a play session.
+Our findings invite further research into the mechanisms governing who experiences the larger impacts of video game play on mood, which likely includes both psychological factors and in-game behavior.
+
+## Author notes
+
+### Data and Material Availability
+
+The data and code we used to analyze it are available on GitHub (<https://github.com/digital-wellbeing/pws-prepost>) and archived at Zenodo ().
+Outputs of the computational notebooks are at ().
+For the full PWS dataset, see @vuorreIntensiveLongitudinalDataset2023.
+
+### Author contributions
+
+Conceptualization: MV, AKP\
+Data Curation: MV, KM\
+Formal Analysis: MV, KM\
+Funding Acquisition: AKP, MV, KM\
+Methodology: MV, KM\
+Project Administration: MV, AKP\
+Resources: AKP, MV\
+Supervision: AKP, MV\
+Validation: KM, MV\
+Visualization: MV, KM\
+Writing – Original Draft: NB, MV, TH, AKP, KM\
+Writing – Review & Editing: MV, NB, AKP, KM\
+
+### Declaration of Conflicting Interests
+
+The data used in this article derives from @vuorreIntensiveLongitudinalDataset2023, which is co-authored by three members of the current research team (MV, KM, AKP) and a FuturLab employee (JB) who collaboratively created the dataset.
+FuturLab had no role in the conceptualization, conduct, or publication of the research presented here.
+The authors perceive no other conflicts of interest with respect to the research, authorship, and/or publication of this article.
+
+### Funding
+
+This research was supported by Huo Family Foundation and the ESRC (ES/W012626/1). KM was supported by Forskningsrådet för hälsa, arbetsliv och välfärd (2021- 01284).
+
+# References
+
+::: {#refs}
 :::

### 2023-10-18 R/common.R

@@ -27,10 +27,3 @@ theme_set(
       plot.tag.position = c(0, 1)
     )
 )
-
-scale_x_hours <- list(
-  scale_x_continuous(
-    breaks = c(0, 15/60, 0.5, 45/60, 1, 2, 3),
-    labels = c(0, "15m", "30m", "45m", "1h", "2h", "3h")
-  )
-)

### 2023-10-18 R/fit_cached.R

@@ -1,18 +1,0 @@
-#' Explicitly cache model
-#' 
-#' A function that will save a model as an .Rds file,
-#' and reload the file without having to fit the model
-#' on the next run.
-#' 
-#' @param path string; path where the model is saved.
-#' Should end with .rds.
-#' @param model a model call to fit
-fit_cached <- function(path, model) {
-  if (file.exists(path)) {
-    fit <- read_rds(path)
-  } else {
-    fit <- model
-    write_rds(fit, path)
-  }
-  fit
-}

### 2023-10-18 R/functions.R

@@ -12,3 +12,192 @@ percent2 <- function(x, accuracy = .1) {
   x <- if_else(x == "0.0%", "<0.1%", x)
   x
 }
+
+#
+ns4 <- function(x) {
+  ns(
+    x,
+    knots = c(0.18, 0.5, 1.3),
+    Boundary.knots = c(0, 3)
+  )
+}
+
+scale_x_hours <- list(
+  scale_x_continuous(
+    "Session duration (hours)",
+    breaks = c(0, 15/60, 0.5, 45/60, 1, 2, 3, 4, 5),
+    labels = c(0, "15m", "30m", "45m", "1h", "2h", "3h", "4h", "5h"),
+    expand = expansion(0.01)
+  )
+)
+
+#' Explicitly cache model
+#'
+#' A function that will save a model as an .Rds file,
+#' and reload the file without having to fit the model
+#' on the next run.
+#'
+#' @param path string; path where the model is saved.
+#' Should end with .rds.
+#' @param model a model call to fit
+fit_cached <- function(path, model) {
+  if (file.exists(path)) {
+    fit <- read_rds(path)
+  } else {
+    fit <- model
+    write_rds(fit, path)
+  }
+  fit
+}
+
+#' Within-session change by pretest values
+#'
+#' @param pre the baseline mood value to calculate the change for.
+#' @param label indicated the label to use for pre value.
+#' @param fit the model object.
+#' @param hours the range of hours to calculate the vales for
+get_change_by_pre <- function(pre, label, fit, hours) {
+  emm <- emmeans(
+    fit,
+    ~ hours + pre,
+    at = list(
+      hours = hours,
+      pre = pre
+    ),
+    lmer.df = "asymp"
+  )
+  emmeans::contrast(
+    emm,
+    method = "trt.vs.ctrl",
+    ref = paste0("hours0 pre", pre)
+  ) |>
+    confint() |>
+    as.data.frame() |>
+    mutate(
+      hours = hours[-1],
+      pre = pre,
+      label = label
+    )
+}
+
+#' Get trend for different random intercept values from brms.
+#' @param sd integer to scale the random intercept SD by.
+#' @details
[…]

### 2023-10-18 R/get_change_by_pre.R

@@ -1,29 +1,0 @@
-#' Within-session change by pretest values
-#' 
-#' @param pre the baseline mood value to calculate the change for.
-#' @param label indicated the label to use for pre value.
-#' @param fit the model object.
-#' @param hours the range of hours to calculate the vales for
-get_change_by_pre <- function(pre, label, fit, hours) {
-  emm <- emmeans(
-    fit,
-    ~ hours + pre,
-    at = list(
-      hours = hours,
-      pre = pre
-    ),
-    lmer.df = "asymp"
-  )
-  emmeans::contrast(
-    emm,
-    method = "trt.vs.ctrl",
-    ref = paste0("hours0 pre", pre)
-  ) |>
-    confint() |>
-    as.data.frame() |>
-    mutate(
-      hours = hours[-1],
-      pre = pre,
-      label = label
-    )
-}

### 2023-10-18 R/get_trend.R

@@ -1,46 +1,0 @@
-#' Get trend for different random intercept values
-#' 
-#' sdfsdf
-#'
-#' @param fit lme4 model object.
-#' @param data the data.
-#' @param sd integer to scale the random intercept SD by.
-#' @param intercept boolean; indicates whether to include the intercept in the
-#' response calculation.
-#' 
-#' @details
-#' This only works with the model `mood ~ ns4(hours) + (1 + ns4(hours) | pid)`
-get_trend <- function(
-    fit,
-    data,
-    sd = 0,
-    intercept = TRUE) {
-  re <- as.data.frame(VarCorr(fit))
-  pid_int <- re[2, "sdcor"]
-  pid_slope0 <- re[3, "sdcor"]
-  pid_slope1 <- re[4, "sdcor"]
-  pid_slope2 <- re[5, "sdcor"]
-  pid_slope3 <- re[6, "sdcor"]
-  pid_cor0 <- re[7, "sdcor"]
-  pid_cor1 <- re[8, "sdcor"]
-  pid_cor2 <- re[9, "sdcor"]
-  pid_cor3 <- re[10, "sdcor"]
-  re_int <- sd * pid_int
-  re_s0 <- sd * pid_cor0 * pid_slope0
-  re_s1 <- sd * pid_cor1 * pid_slope1
-  re_s2 <- sd * pid_cor2 * pid_slope2
-  re_s3 <- sd * pid_cor3 * pid_slope3
-  hours <- seq(0, 3, length.out = 100)
-  x <- ns4(
-    hours
-  )
-  b0 <- fixef(fit)[1] + re_int
-  b <- fixef(fit)[-1]
-  y <- x[, 1] * (b[1] + re_s0) + x[, 2] * (b[2] + re_s1) + x[, 3] * (b[3] + re_s2) + x[, 4] * (b[4] + re_s3)
-  if (intercept) y <- y + b0
-  data.frame(
-    x = hours,
-    y = c(y),
-    re_int
-  )
-}

### 2023-10-18 R/get_trend_bayes.R

@@ -1,74 +1,0 @@
-#' Get trend for different random intercept values from brms.
-#' @param sd integer to scale the random intercept SD by.
-#' @details 
-#' The model object is hard coded to be `fit_b`.
-#' 
-#' The curve is calculated over all posterior samples, and
-#' summaries are based on medians and 2.5% and 97.5% percentile
-#' intervals.
-get_trend_bayes <- function(sd) {
-  re <- VarCorr(fit_b)
-  ps <- as_draws_df(
-    fit_b,
-    c(
-      "b_Intercept",
-      "b_ns4hours1",
-      "b_ns4hours2",
-      "b_ns4hours3",
-      "b_ns4hours4",
-      "sd_pid__Intercept",
-      "sd_pid__ns4hours1",
-      "sd_pid__ns4hours2",
-      "sd_pid__ns4hours3",
-      "sd_pid__ns4hours4",
-      "cor_pid__Intercept__ns4hours1",
-      "cor_pid__Intercept__ns4hours2",
-      "cor_pid__Intercept__ns4hours3",
-      "cor_pid__Intercept__ns4hours4"
-    )
-  ) |>
-    rename(
-      pid_int = sd_pid__Intercept,
-      pid_slope0 = sd_pid__ns4hours1,
-      pid_slope1 = sd_pid__ns4hours2,
-      pid_slope2 = sd_pid__ns4hours3,
-      pid_slope3 = sd_pid__ns4hours4,
-      pid_cor0 = cor_pid__Intercept__ns4hours1,
-      pid_cor1 = cor_pid__Intercept__ns4hours2,
-      pid_cor2 = cor_pid__Intercept__ns4hours3,
-      pid_cor3 = cor_pid__Intercept__ns4hours4,
-      b1 = b_ns4hours1,
-      b2 = b_ns4hours2,
-      b3 = b_ns4hours3,
-      b4 = b_ns4hours4,
-    ) |>
-    mutate(
-      re_int = sd * pid_int,
-      b0 = b_Intercept + re_int,
-      re_s0 = sd * pid_cor0 * pid_slope0,
-      re_s1 = sd * pid_cor1 * pid_slope1,
-      re_s2 = sd * pid_cor2 * pid_slope2,
-      re_s3 = sd * pid_cor3 * pid_slope3
-    )
-  lapply(hours, function(hour) {
-    x <- ns4(
-      hour
-    )
-    ps |>
-      mutate(
[…]

### 2023-10-18 continuous-time.qmd
First draft (#2)
* Download data with R

* Small edits

* Init bibliography

* gitignore Word output

* Move all computation to ms.qmd

* Throw out prose

* WIP pre-holiday changes :smile:

* Update renv

* Ignore output style for now

* Minor code cleaning

* Move code block up

* Small model revision

Also added some verbal description

* rm old file

* move metadata to same file

* Fix metadata

* Estimate full model

* Reparameterize model

* WIP updates

* Fix #5

* Fix build

* WIP changes

* dockerize (#6)

Dockerize the project

Closes #3

* Figure and heterogeneity

* use correct env

@@
 
-In the main analysis in @sec-analyze we focused on contrasting the mood at the beginning of a session with the average mood during session.
-In this document we focus on two additional research questions:
+In the main analysis we focused on contrasting the mood at the beginning of a session with the average mood during session. In this document we focus on two additional research questions:
 
@@
 source("R/functions.R")
-source("R/fit_cached.R")
 source("R/common.R")
@@
 #| fig-width: 8
-source("R/get_trend.R")
 
@@
 p_pre_x
-ggsave(
-  "figures/fig-interaction-model.png", 
-  p_pre_x, 
-  width = 7, 
-  height = 5
-)
 ```
@@
 #| fig-width: 8
-source("R/get_change_by_pre.R")
+
 d_x_change <- data.frame(
@@
 
-```{r}
-#| label: ggsave-fig-ms-interaction
-#| echo: false
-
-ggsave(
-  "figures/fig-pre-x-change.png", 
-  p_tmp, 
-  width = 7, 
-  height = 3.5
-  )
-```
-
 And this figure.
@@
 p_tmp
-```
-
-```{r}
-#| label: ggsave-fig-ms-no-interaction
-#| echo: false
-
-ggsave(
-  "figures/fig-change.png", 
-  p_tmp, 
-  width = 6, 
-  height = 3
-  )
 ```
@@
 #| label: bayes-int-sd
-source("R/get_trend_bayes.R")
+
 res <- map(

### 2023-10-18 R/common.R

@@ -1,0 +1,29 @@
+N_CORES <- as.numeric(Sys.getenv("N_CORES", unset = 1))
+N_THREADS <- as.numeric(Sys.getenv("N_THREADS", unset = 2))
+N_ITER <- as.numeric(Sys.getenv("N_ITER", unset = 2000))
+N_SUBSET <- as.numeric(Sys.getenv("N_SUBSET"))
+
+options(
+  mc.cores = N_CORES,
+  brms.backend = "cmdstanr",
+  brms.threads = N_THREADS
+)
+
+theme_set(
+  theme_linedraw(base_size = 10) +
+    theme(
+      strip.background = element_rect(
+        fill = NA,
+        color = NA
+      ),
+      strip.text = element_text(
+        colour = "black",
+        hjust = 0
+      ),
+      axis.text = element_text(
+        size = rel(0.75)
+      ),
+      panel.grid = element_blank(),
+      plot.tag.position = c(0, 1)
+    )
+)

### 2023-10-18 R/functions.R

@@ -1,0 +1,203 @@
+library(scales)
+
+# Format big numbers with a comma
+number2 <- function(x, ...) {
+  number(x, ..., big.mark = ",")
+}
+
+# Replace extreme percentages
+percent2 <- function(x, accuracy = .1) {
+  x <- percent(x, accuracy = accuracy)
+  x <- if_else(x == "100.0%", ">99.9%", x)
+  x <- if_else(x == "0.0%", "<0.1%", x)
+  x
+}
+
+#
+ns4 <- function(x) {
+  ns(
+    x,
+    knots = c(0.18, 0.5, 1.3),
+    Boundary.knots = c(0, 3)
+  )
+}
+
+scale_x_hours <- list(
+  scale_x_continuous(
+    "Session duration (hours)",
+    breaks = c(0, 15/60, 0.5, 45/60, 1, 2, 3, 4, 5),
+    labels = c(0, "15m", "30m", "45m", "1h", "2h", "3h", "4h", "5h"),
+    expand = expansion(0.01)
+  )
+)
+
+#' Explicitly cache model
+#'
+#' A function that will save a model as an .Rds file,
+#' and reload the file without having to fit the model
+#' on the next run.
+#'
+#' @param path string; path where the model is saved.
+#' Should end with .rds.
+#' @param model a model call to fit
+fit_cached <- function(path, model) {
+  if (file.exists(path)) {
+    fit <- read_rds(path)
+  } else {
+    fit <- model
+    write_rds(fit, path)
+  }
+  fit
+}
+
+#' Within-session change by pretest values
+#'
+#' @param pre the baseline mood value to calculate the change for.
+#' @param label indicated the label to use for pre value.
+#' @param fit the model object.
+#' @param hours the range of hours to calculate the vales for
+get_change_by_pre <- function(pre, label, fit, hours) {
+  emm <- emmeans(
+    fit,
+    ~ hours + pre,
+    at = list(
+      hours = hours,
+      pre = pre
+    ),
+    lmer.df = "asymp"
+  )
+  emmeans::contrast(
+    emm,
+    method = "trt.vs.ctrl",
+    ref = paste0("hours0 pre", pre)
+  ) |>
+    confint() |>
+    as.data.frame() |>
+    mutate(
[…]

### 2023-10-19 ms.Rmd
Remove Quarto
Submission formatting

@@
 draft: no
-mask: no
+mask: yes
 figurelist: no
@@
 documentclass: apa7
-classoption: jou
+classoption: doc
 output: 
@@
 
-# Introduction
-
 How do video games affect players' well-being?
@@
 At present, however, the prevalence and magnitude of these short-term effects remains poorly understood.
-Some existing evidence suggests that games can provide short-term boosts to well-being ["mood repair"; @bowmanMoodGameSelective2015; @tyackRestorativePlayVideogames2020], possibly to a greater extent than non-interactive media such as videos [@riegerEatingGhostsUnderlying2015].
+Some existing evidence suggests that games can provide short-term boosts to well-being ["mood repair", @bowmanMoodGameSelective2015; @tyackRestorativePlayVideogames2020], possibly to a greater extent than non-interactive media such as videos [@riegerEatingGhostsUnderlying2015].
 On the other hand, games might affect players negatively: Frustrating gaming experiences, for example, can lead to negative consequences such as immediate post-play aggression [@przybylskiCompetenceimpedingElectronicGames2014].
@@
 Here, we aim to address these three challenges to better understand how real play in natural contexts might predict mood on short timescales.
-Specifically, we examined an intensive longitudinal dataset from the popular commercially available game PowerWash Simulator [PWS; @vuorreIntensiveLongitudinalDataset2023], which includes mood questions embedded in the game itself, to ask three questions: First, to what extent does mood change from immediately before video game play to during play?
+Specifically, we examined an intensive longitudinal dataset from the popular commercially available game PowerWash Simulator [PWS, @vuorreIntensiveLongitudinalDataset2023], which includes mood questions embedded in the game itself, to ask three questions: First, to what extent does mood change from immediately before video game play to during play?
 Second, hyow heterogeneous are these changes in the population of similar players?
@@
 dir.create("data", FALSE)
-write_rds("data/data.rds")
+write_rds(dat, "data/data.rds")
 ```
@@
 Participants responded using a visual analogue scale (VAS) with endpoints "Very bad" and "Very good" that recorded 1000 possible values, which we rescaled to the unit interval (0-1) for this study.
-Consequently, our results can also be interpreted on the "[proportion] of maximum possible" scale [POMP; @cohenProblemUnitsCircumstance1999].
+Consequently, our results can also be interpreted on the "[proportion] of maximum possible" scale [POMP, @cohenProblemUnitsCircumstance1999].
 While well-being is often studied with multi-item scales to differentiate between dimensions of positive and negative affect, the frequent probing of mental states in this study required a minimally intrusive instrument that would interrupt the participants' play experience as little as possible.
@@
 #| fig-height: 3
-#| fig-width: 5
+#| fig-width: 8
 
@@
 For example, experience sampling studies of US [@killingsworthWanderingMindUnhappy2010] and Korean [@choiTakingStockHappiness2017] adults indicated that watching television (+2% change compared to an individual's average across all activities), reading (+2%), and shopping (+3%) were associated with smaller mood shifts than we observed in association with PWS.
-On the other hand, those studies indicated that listening to music (+7–9%), eating or cooking (+8%), taking a walk (+9%), visiting an urban green space [+9–10%; @tostNeuralCorrelatesIndividual2019], exercising (+11%), dating (26.5%), sexual intercourse (+28%), or taking a trip (+30.5%) correlate with greater shifts in well-being.
+On the other hand, those studies indicated that listening to music (+7–9%), eating or cooking (+8%), taking a walk (+9%), visiting an urban green space [+9–10%, @tostNeuralCorrelatesIndividual2019], exercising (+11%), dating (26.5%), sexual intercourse (+28%), or taking a trip (+30.5%) correlate with greater shifts in well-being.
 
@@
 Current theories provide a few candidates regarding mood change mechanisms.
-Mood management theory [MMT; @zillmannMoodManagementCommunication1988] posits that players choose specific entertainment media as part of a (conscious or subconscious) strategy to regulate valence and arousal and to increase hedonic well-being.
+Mood management theory [MMT, @zillmannMoodManagementCommunication1988] posits that players choose specific entertainment media as part of a (conscious or subconscious) strategy to regulate valence and arousal and to increase hedonic well-being.
 Specifically, MMT predicts that when players feel more negative emotions, they are more likely to select games with positive hedonic valence (i.e., games that are generally positive in tone), and that when players are in high arousal states such as anger stress, they are more likely to choose calm and relaxing games with.
@@
 The sample may further suffer from self-selection bias: Given that players voluntarily chose to download the PWS research edition, it is plausible that people who felt more positively toward the game (and more positively while playing it) were more likely to opt into the study.
-Finally, we studied just one commercial video game with a feature set that is very different from today's most commonly-played PC games [which at the time of writing include, for example, Minecraft, Fortnite, and Baldur's Gate 3, but only one other simulation game, The Sims 4; @newzooMostPopularPC2023].
+Finally, we studied just one commercial video game with a feature set that is very different from today's most commonly-played PC games [which at the time of writing include, for example, Minecraft, Fortnite, and Baldur's Gate 3, but only one other simulation game, The Sims 4, @newzooMostPopularPC2023].
 
@@
 
-## Author notes
-
-### Data and Material Availability
-
-The data and code we used to analyze it are available on GitHub (<https://github.com/digital-wellbeing/pws-prepost>) and archived at Zenodo ().
-Outputs of the computational notebooks are at ().
-For the full PWS dataset, see @vuorreIntensiveLongitudinalDataset2023.
-
-### Author contributions
-
+# References
+
+::: {#refs}
+:::
+
+# Contributions
+
+\noindent
 Conceptualization: MV, AKP\
@@
 Writing – Original Draft: NB, MV, TH, AKP, KM\
-Writing – Review & Editing: MV, NB, AKP, KM\
-
-### Declaration of Conflicting Interests
+Writing – Review & Editing: MV, NB, TH, AKP, KM\
+
+# Funding information
+
+This research was supported by Huo Family Foundation and the ESRC (ES/W012626/1). KM was supported by Forskningsrådet för hälsa, arbetsliv och välfärd (2021- 01284).
+
+# Competing interests
 
@@
 
-### Funding
-
-This research was supported by Huo Family Foundation and the ESRC (ES/W012626/1). KM was supported by Forskningsrådet för hälsa, arbetsliv och välfärd (2021- 01284).
-
-# References
-
-::: {#refs}
-:::
+# Data accessibility statement
+
+The data and code we used to analyze it are available on GitHub (<https://github.com/digital-wellbeing/pws-prepost>) and archived at Zenodo ().
+Outputs of the computational notebooks are at ().
+For the full PWS dataset, see @vuorreIntensiveLongitudinalDataset2023.
+
+# Figure titles and legends
+
+## Figure 1.
+
+A.
+Scatterplots of three participants' (rows) mood responses (pre-play: red; during play: blue) over eight sessions' (columns) durations.
+B.
+Histograms of session-mean (C.
+person-mean) moods before (top) and during (bottom) play sessions.
+D.
+Differences in session-mean (E.
+player-mean) mood differences (during session - pre-play).
+F.
+Scatterplot of person-mean mood reports at the beginning (x-axis) and during gameplay sessions (y-axis).
+Identity line is shown in green, and an exploratory GAM regression line is shown in blue.
+
+## Figure 2.
+
+Estimated (changes) in mood as a function of session duration.
+Top: Average mood during a gaming session.
+Bottom: Change in mood during a session compared to mood at the beginning of a session.
+Gray ribbons indicate 95\\% confidence bands.
+We truncated the x-axis at three hours for this figure.
+
+## Figure 3.
+
+Estimated (changes) in mood as a function of session duration and pre-play mood.
+Top.
+Mood for the average player during a gaming session with pre-play mood at 5th, 25th, 50th, and 75th percentiles (columns).
+Ribbons indicate 95\\% confidence.
+Bottom.
+Same as above but with change in mood on the y-axis.

### 2023-10-19 ms.Rmd
Add Zenodo link

@@
 
-The data and code we used to analyze it are available on GitHub (<https://github.com/digital-wellbeing/pws-prepost>) and archived at Zenodo ().
-Outputs of the computational notebooks are at ().
+The data and code we used to analyze it are available on GitHub (<https://github.com/digital-wellbeing/pws-prepost>) and archived at Zenodo (<https://zenodo.org/doi/10.5281/zenodo.10021014>).
 For the full PWS dataset, see @vuorreIntensiveLongitudinalDataset2023.
@@
 Bottom: Change in mood during a session compared to mood at the beginning of a session.
-Gray ribbons indicate 95\\% confidence bands.
+Gray ribbons indicate 95\% confidence bands.
 We truncated the x-axis at three hours for this figure.
@@
 Mood for the average player during a gaming session with pre-play mood at 5th, 25th, 50th, and 75th percentiles (columns).
-Ribbons indicate 95\\% confidence.
+Ribbons indicate 95\% confidence.
 Bottom.

### 2023-10-19 ms.Rmd
Typos etc

@@
   \noindent \textbf{This pre-print is not yet peer-reviewed.}
+
   *MV and NB contributed equally to this work.
@@
 draft: no
-mask: yes
+mask: no
 figurelist: no
@@
 documentclass: apa7
-classoption: doc
+classoption: jou
 output: 
@@
 #| fig-height: 3
-#| fig-width: 8
+#| fig-width: 5
 
@@
 
-# References
-
-::: {#refs}
-:::
-
-# Contributions
-
-\noindent
+# Data and Material Availability
+
+The data and code we used to analyze it are available on GitHub (<https://github.com/digital-wellbeing/pws-prepost>) and archived at Zenodo (<https://zenodo.org/doi/10.5281/zenodo.10021014>).
+
+# Author contributions
+
 Conceptualization: MV, AKP\
@@
 
-# Funding information
-
-This research was supported by Huo Family Foundation and the ESRC (ES/W012626/1). KM was supported by Forskningsrådet för hälsa, arbetsliv och välfärd (2021- 01284).
-
-# Competing interests
+# Declaration of Conflicting Interests
 
@@
 
-# Data accessibility statement
-
-The data and code we used to analyze it are available on GitHub (<https://github.com/digital-wellbeing/pws-prepost>) and archived at Zenodo (<https://zenodo.org/doi/10.5281/zenodo.10021014>).
-For the full PWS dataset, see @vuorreIntensiveLongitudinalDataset2023.
-
-# Figure titles and legends
-
-## Figure 1.
-
-A.
-Scatterplots of three participants' (rows) mood responses (pre-play: red; during play: blue) over eight sessions' (columns) durations.
-B.
-Histograms of session-mean (C.
-person-mean) moods before (top) and during (bottom) play sessions.
-D.
-Differences in session-mean (E.
-player-mean) mood differences (during session - pre-play).
-F.
-Scatterplot of person-mean mood reports at the beginning (x-axis) and during gameplay sessions (y-axis).
-Identity line is shown in green, and an exploratory GAM regression line is shown in blue.
-
-## Figure 2.
-
-Estimated (changes) in mood as a function of session duration.
-Top: Average mood during a gaming session.
-Bottom: Change in mood during a session compared to mood at the beginning of a session.
-Gray ribbons indicate 95\% confidence bands.
-We truncated the x-axis at three hours for this figure.
-
-## Figure 3.
-
-Estimated (changes) in mood as a function of session duration and pre-play mood.
-Top.
-Mood for the average player during a gaming session with pre-play mood at 5th, 25th, 50th, and 75th percentiles (columns).
-Ribbons indicate 95\% confidence.
-Bottom.
-Same as above but with change in mood on the y-axis.
+# Funding
+
+This research was supported by Huo Family Foundation and the ESRC (ES/W012626/1). KM was supported by Forskningsrådet för hälsa, arbetsliv och välfärd (2021- 01284).
+
+# References
+
+::: {#refs}
+:::

### 2023-10-27 ms.Rmd
Adjust figures

@@
 dir.create("models", FALSE)
+
+knitr::opts_chunk$set(
+  fig.width = 8,
+  fig.asp = 0.618
+)
 ```
@@
 #| include: false
-#| fig-height: 3
+#| fig.asp: 0.5
 #| fig.cap: Summaries of session durations. X-axes are truncated at 10 hours.
@@
 #| fig-height: 2.2
-#| fig-width: 7
 #| include: false
@@
 #| include: true
-#| fig-width: 8
-#| fig-height: 5
 #| fig.env: "figure*"
@@
     limits = c(0, 1),
+    breaks = c(0, 0.25, 0.5, 0.75, 1.0),
+    labels = c("0", "", "0.5", "", "1")
   ) +
@@
     "Observations",
-    expand = expansion(c(0.001, 0.1)),
+    expand = expansion(c(0.01, 0.1)),
     breaks = scales::extended_breaks(5)
@@
     "Mood",
-    expand = expansion(c(0.001))
+    expand = expansion(c(0.01))
   ) +
@@
     "Sessions",
-    expand = expansion(c(0.001, 0.1)),
+    expand = expansion(c(0.01, 0.1)),
     breaks = scales::extended_breaks(5)
@@
     "Players",
-    expand = expansion(c(0.001, 0.1)),
+    expand = expansion(c(0.01, 0.1)),
     breaks = scales::extended_breaks(5)
@@
     "Sessions",
-    expand = expansion(c(0.001, 0.1))
+    expand = expansion(c(0.01, 0.1))
   ) +
   scale_x_continuous(
-    expand = expansion(c(0.001))
+    expand = expansion(c(0.01))
   ) +
@@
     "Players",
-    expand = expansion(c(0.001, 0.1))
+    expand = expansion(c(0.01, 0.1))
   ) +
   scale_x_continuous(
-    expand = expansion(c(0.001))
+    expand = expansion(c(0.01))
   ) +
@@
 #| include: true
+#| fig.width: 4
 #| fig.cap: "Estimated (changes) in mood as a function of session duration.
@@
 We truncated the x-axis at three hours for this figure."
-#| fig-height: 3
-#| fig-width: 5
 
@@
 #| include: true
+#| fig.asp: 0.5
 #| fig.cap: "Estimated (changes) in mood as a function of session duration and pre-play mood.
@@
 Same as above but with change in mood on the y-axis."
-#| fig-height: 3
-#| fig-width: 8
 

### 2024-02-19 ms.Rmd
Improve environment variable use

@@
 )
+
+# mcmc settings
+MAX_CORES <- as.numeric(Sys.getenv("N_CORES"), 4)
+N_ITER <- as.numeric(Sys.getenv("N_ITER"), 2000)
+options(
+  brms.backend = Sys.getenv("BRMS_BACKEND", "rstan"), 
+  brms.threads = as.numeric(Sys.getenv("N_THREADS"), 1),
+  mc.cores = MAX_CORES
+)
 ```

### 2024-02-19 ms.Rmd
Revision draft 1

@@
 
-wordcount: '`r scales::number(wordcountaddin:::word_count("ms.Rmd"), big.mark = ",")`'
+wordcount: "4,802"
 bibliography: references.bib
@@
 output: 
+  papaja::apa6_docx: 
+    number_sections: false
   papaja::apa6_pdf:
@@
     keep_tex: true
-  papaja::apa6_docx: 
-    number_sections: false
 editor_options: 
@@
 
+# Output and temporary files
 dir.create("models", FALSE)
-
+dir.create("data", FALSE)
+
+# Document options
 knitr::opts_chunk$set(
+  warning = FALSE,
+  message = FALSE,
   fig.width = 8,
@@
 Specifically, we examined an intensive longitudinal dataset from the popular commercially available game PowerWash Simulator [PWS, @vuorreIntensiveLongitudinalDataset2023], which includes mood questions embedded in the game itself, to ask three questions: First, to what extent does mood change from immediately before video game play to during play?
-Second, hyow heterogeneous are these changes in the population of similar players?
+Second, how heterogeneous are these changes in the population of similar players?
 And third, how do changes in mood develop over the course of a gaming session?
@@
 # Write data for any supplementary notebooks
-dir.create("data", FALSE)
 write_rds(dat, "data/data.rds")
@@
 
+After downloading the PWS research edition and starting the game for the first time, but before entering the game menu, participants gave informed consent, confirmed that they were 18 years old or older, and answered optional demographic questions.
 The characteristics of the full sample of 11,080 players in the PWS dataset are described in @vuorreIntensiveLongitudinalDataset2023; here, we describe the subset of data relevant to our questions (see Data analysis below).
@@
 Our first and main research question concerned the difference between players' moods at the beginning of each session (pre-play) and during the subsequent play session (during play).
-This contrast does not represent a causal hypothesis: Players could begin (and end) their play sessions for whatever reason, and these reasons are likely to confound the pre – during contrast.
+This contrast does not represent a causal hypothesis (see Limitations, below): Players could begin (and end) their play sessions for whatever reason, and these reasons are likely to confound the pre – during contrast.
 For example, a player might come home after a stressful day at work and then play PWS.
@@
 
+```{r}
+#| include: false
+
+# How many sessions per player
+dat |> 
+  count(pid, session) |> 
+  count(pid) |> 
+  summarise(median(n))
+
+# How many data per session
+dat |> 
+  count(ps) |> 
+  summarise(median(n))
+```
+
 We estimated this contrast within a three-level hierarchical regression model that nested observations within sessions, and sessions within players.
-We specified a censored Gaussian model of mood because a VAS necessarily limits response options at the lower and upper ends.
+We decided this three-level hierarchy as most appropriate, because individuals typically contributed data over many sessions (the median player contributed five sessions' data), and sessions typically had multiple observations (the median session included two observations).
+More formally, then, we modelled the mood report of the $i^{th}$ observation, of the $j^{th}$ person's $k^{th}$ session as censored-normal distributed with a common variance using the following equations
+
+\begin{align*}
+\text{mood}_{ijk} &\sim \text{CensNorm}^{[0, 1]}(\beta_{0jk} + \beta_{1j}\text{during}_{ijk}, \sigma^2), \\
+\beta_{0j} &= \gamma_0 + u_{0j} + v_{0k}, \\
+\beta_{1j} &= \gamma_1 + u_{1j}, \\
+\begin{bmatrix}
+  u_{0j} \\ u_{1j}
+\end{bmatrix} &\sim \text{MVN}\left(
+  \begin{bmatrix}
+    0 \\ 0
+  \end{bmatrix},
+  \begin{pmatrix}
+    \tau_{0} \ & \\ 
+    \rho_{01} \ &\tau_{1}
+  \end{pmatrix}
+\right), \\
+v_{0k} &\sim \text{Normal}\left(0, \kappa_{0}\right)
+\end{align*}
+
+We specified a censored (at 0 and 1) Gaussian model of mood because a VAS necessarily limits response options at the lower and upper ends.
 Ignoring censoring would leave the contrast susceptible to ceiling or floor effects and might confound changes in the mood distribution's location with changes in its scale.
-We then modelled mean mood on an intercept and contrast coded time (pre-play: -0.5; during play: 0.5) and allowed both parameters to vary randomly across players.
+We then modelled mean mood on an intercept and a coefficient of *during* play (coded as pre-play: -0.5; during play: 0.5) and allowed both parameters to vary randomly across players ($u_{0j}$ and $u_{1j}$).
+Thus, to answer RQ2, we could examine $\tau_1$, describing the variability of individuals' mood changes around the mean mood change ($\gamma_1$).
 In addition, we modelled random intercepts over player sessions.
@@
 We analyzed the data with R and used the brms package to estimate, via Stan's HMC sampling algorithm, and post-process the models [@burknerBrmsPackageBayesian2017; @rcoreteamLanguageEnvironmentStatistical2023; @standevelopmentteamStanModelingLanguage2021].
+These probabilistic methods are especially helpful for complex models where some variance parameters might be small---as we anticipated here for the session-level variances.
 We drew `r number2(N_ITER*4/2)` samples from the model's posterior distribution using brms' default prior distributions on all parameters and used numerical and graphical checks to ensure model convergence and adequacy.
@@
 
-Although the above quantities indicated that the average player's mood did increase from before to during play, they say little about the robustness of this difference in the population of similar players.
+Above, we estimated that the average player's mood increased by approximately `r filter(fit_tbl, variable=="Difference")$res` units (on a 0-1 scale) from the beginning of the session to during play.
+However, that number does not indicate how representative this "average player" is.
+In other words, we do not know how variable this mood increase is likely to be in the population of similar players.
 We therefore next turned to our second research question: How heterogeneous are mood shifts in the population of similar PWS players?
-First, the standard deviation of the person-specific mood differences was `r filter(fit_tbl, variable=="(SD) Difference")$res`.
-In comparison to the average person's estimated difference (`r filter(fit_tbl, variable=="Difference")$res`), that quantity indicated a moderate degree of heterogeneity between individuals.
-More specifically, the model-predicted proportion of individuals expected to experience positive changes from pre- to during play was `r filter(fit_tbl, variable=="Positive shifts")$resp`.
-
-In sum, the results from our model contrasting pre- and during play moods indicated small increases in mood during play, and that those changes were robust across people.
+As a first approximation to an answer, we looked at the model's standard deviation of the person-specific mood increases.
+It was `r filter(fit_tbl, variable=="(SD) Difference")$res`.
+In comparison to the average person's estimated difference, that quantity indicated a moderate degree of heterogeneity between individuals.
+To give a more concrete quantity describing heterogeneity in this mood uplift, we then calculated the model-estimated proportion of individuals in this population who are expected to experience positive mood changes from pre- to during play.
+This proportion was `r filter(fit_tbl, variable=="Positive shifts")$resp`: Nearly three quarters of individuals are predicted to experience mood lifts during PWS play.
+
+In sum, the results from our model contrasting pre- and during play moods indicated small increases in mood during play, and that those changes were somewhat robust across people.
 
@@
 We therefore next turned to our third question: How do (changes in) players' moods evolve during gameplay sessions?
-To answer, we used time as a continuous predictor and allowed mood changes during sessions to be non-linear by estimating a piecewise cubic spline with 4 degrees of freedom.
+To answer, we used time (hours) as a continuous predictor and allowed mood changes during sessions to be non-linear by estimating a piecewise cubic spline with 4 degrees of freedom using the R package lme4 [@lme4].
 Just like the main model, this was a three-level hierarchical model, with random intercepts at the session and participant levels, and random participant slopes for each piece of the spline.
@@
 The sample may further suffer from self-selection bias: Given that players voluntarily chose to download the PWS research edition, it is plausible that people who felt more positively toward the game (and more positively while playing it) were more likely to opt into the study.
+
 Finally, we studied just one commercial video game with a feature set that is very different from today's most commonly-played PC games [which at the time of writing include, for example, Minecraft, Fortnite, and Baldur's Gate 3, but only one other simulation game, The Sims 4, @newzooMostPopularPC2023].
+The fact that we studied only one game---and one that is not likely representative of today's most commonly-played games---suggests caution in generalizing from our findings to other games.
 

### 2024-02-28 ms.Rmd
MS revised after reviwe

@@
     corresponding: yes
-    affiliation: "1,2*"
+    affiliation: "1,2"
     address: Tilburg University
@@
   - name: Nick Ballou
-    affiliation: "2*"
+    affiliation: "2"
   - name: Thomas Hakman
@@
 
-  *MV and NB contributed equally to this work.
-
 wordcount: "4,802"
@@
 output: 
-  papaja::apa6_docx: 
-    number_sections: false
   papaja::apa6_pdf:
@@
     keep_tex: true
+  papaja::apa6_docx: 
+    number_sections: false
 editor_options: 
@@
 )
-
-# mcmc settings
-MAX_CORES <- as.numeric(Sys.getenv("N_CORES"), 4)
-N_ITER <- as.numeric(Sys.getenv("N_ITER"), 2000)
-options(
-  brms.backend = Sys.getenv("BRMS_BACKEND", "rstan"), 
-  brms.threads = as.numeric(Sys.getenv("N_THREADS"), 1),
-  mc.cores = MAX_CORES
-)
 ```
@@
 Understanding whether and when games' short-term effects emerge is critical for establishing games' potential for mood-related interventions, as well as for building a theoretical foundation for repeated short-term gaming experiences' long-term effects on mental health.
-At present, however, the prevalence and magnitude of these short-term effects remains poorly understood.
-Some existing evidence suggests that games can provide short-term boosts to well-being ["mood repair", @bowmanMoodGameSelective2015; @tyackRestorativePlayVideogames2020], possibly to a greater extent than non-interactive media such as videos [@riegerEatingGhostsUnderlying2015].
-On the other hand, games might affect players negatively: Frustrating gaming experiences, for example, can lead to negative consequences such as immediate post-play aggression [@przybylskiCompetenceimpedingElectronicGames2014].
-
+Substantial existing evidence suggests that games can provide short-term boosts to well-being [@bowmanMoodGameSelective2015; @tyackRestorativePlayVideogames2020], possibly to a greater extent than non-interactive media such as videos [@riegerEatingGhostsUnderlying2015].
+Much of that work took place under the "mood repair" and "mood management" labels [@zillmannMoodManagementCommunication1988], which describe how media might support users in balancing internal states following unpleasant feelings, possibly through addressing basic psychological needs [@BallouDeterding2023Basic; @ReineckeEtAl2012characterizing; @riegerEatingGhostsUnderlying2015; @TamboriniEtAl2011Media; @Tyack2019Need].
+On the other hand, games might also affect players negatively: Frustrating gaming experiences, for example, can lead to negative consequences such as immediate post-play aggression [@przybylskiCompetenceimpedingElectronicGames2014].
+
+At present, however, the prevalence and magnitude of these short-term effects remain poorly understood.
 Despite the above examples, the validity and generalizability of research on games' short-term affective effects remains limited by three challenges.
-First, a substantial portion of gameplay research has relied on artificial stimuli; games created or substantially modified by academic researchers.
+First, a substantial portion of gameplay research has relied on artificial stimuli; games created or substantially modified by academic researchers [@bowmanMoodGameSelective2015].
 While such customized games allow for greater experimental control, they are unlikely to reflect actual games' rich complexity [@mcmahanConsiderationsUseCommercial2011].
@@
 The second challenge is providing an ecologically valid *context* for play.
-Research participants typically play games in (online or physical) labs that do not resemble the natural contexts of play, such as when, with whom, and why people choose to play.
+Research participants typically play games in (online or physical) labs that do not resemble the natural contexts of play, such as when, with whom, and why people choose to play [@tyackRestorativePlayVideogames2020].
 In lab settings, research participants play to satisfy study requirements, rather than the intrinsic motivations that typically lead them to play.
@@
 In this study, we analyzed data from a large open dataset on PowerWash Simulator (PWS) play and psychological experiences [@vuorreIntensiveLongitudinalDataset2023].
-The data was collected in a research edition of PWS that recorded gameplay events, game status records, and responses to psychological survey items.
+The data was collected in a research edition of PWS that recorded gameplay events, game status records, participant demographics, and responses to psychological survey items.
 We developed the research edition of PWS in collaboration with PWS's developer, FuturLab, who made it freely available on Steam to anyone who owned the original game (£19.99 on 2023-09-20).
@@
 
-Although the overall magnitude of the estimated change was small (`r filter(fit_tbl, variable=="Difference")$res`), it nevertheless is comparable to mood shifts following other frequent voluntary activities.
+Although the overall magnitude of the estimated change was small considering the scale range (`r filter(fit_tbl, variable=="Difference")$res`), comparative evidence from other frequent voluntary activities indicates that it might still be of meaningful magnitude.
 For example, experience sampling studies of US [@killingsworthWanderingMindUnhappy2010] and Korean [@choiTakingStockHappiness2017] adults indicated that watching television (+2% change compared to an individual's average across all activities), reading (+2%), and shopping (+3%) were associated with smaller mood shifts than we observed in association with PWS.
@@
 @anvariUsingAnchorbasedMethods2021 suggest that, on average, people are able to subjectively perceive a change of 2% in well-being on the related PANAS scale.
-Since our estimate of `r filter(fit_tbl, variable=="Difference")$res` is greater, we tentatively suggest that gaming, on average, is associated with mood uplifts that are large enough to be consciously felt by players.
+Since our estimate of `r filter(fit_tbl, variable=="Difference")$res` is greater, we tentatively suggest that gaming, on average, is associated with mood uplifts that are large enough to be consciously experienced by players.
+This might hold especially for people who started the session with lower pre-play mood (Figure \@ref(fig:fig-ct2)), a finding that aligns with literature indicating that people in particularly low moods will selectively seek out media with potential to uplift and balance mood [@ReineckeEtAl2012characterizing].
 
@@
 
+We believe the mood uplift observed here is significant for three reasons: First, it compares in magnitude to associations with other activities commonly considered as mood-enhancing.
+Second, the uplift is greater for individuals who started with a lower mood.
+Third, anticipatory effects, whereby mood increases before play in anticipation of this rewarding activity, are likely and would negatively bias the uplift observed here as an estimate of the overall effect.
+Nevertheless, we aimed to provide information about plausible mood-uplift magnitudes, whose interpretation likely depends on their context in future studies and applications.
+
 By examining questionnaire items embedded in the game interface during natural play sessions, our study is the first to examine changes in mood during play sessions in a minute-by-minute resolution.
@@
 Current theories provide a few candidates regarding mood change mechanisms.
-Mood management theory [MMT, @zillmannMoodManagementCommunication1988] posits that players choose specific entertainment media as part of a (conscious or subconscious) strategy to regulate valence and arousal and to increase hedonic well-being.
+Mood management theory [MMT, @zillmannMoodManagementCommunication1988] posits that people choose media partly as strategies to regulate valence and arousal and to increase hedonic well-being.
 Specifically, MMT predicts that when players feel more negative emotions, they are more likely to select games with positive hedonic valence (i.e., games that are generally positive in tone), and that when players are in high arousal states such as anger stress, they are more likely to choose calm and relaxing games with.
@@
 Accordingly, PWS may be especially well-suited for players who are emotionally or cognitively depleted, such as those with persistently low mood [@helsbyBandwidthComesGoes2023]---something that is partially supported by our findings that sessions with lower initial moods were those with larger uplifts during play.
+We think the consistency of this uplift---with over 70% of people predicted to experience uplifts---means that PWS and similar games might be especially effective for stabilizing noxious moods rather than enhancing neutral ones.
 Although outside the scope of this study, we believe that the additional telemetry in the PWS dataset [@vuorreIntensiveLongitudinalDataset2023] would be valuable in addressing more nuanced questions regarding MMT.
 
-Another candidate derives from self-determination theory [@ryanSelfdeterminationTheoryBasic2017].
+Another candidate derives from Self-determination theory [@ryanSelfdeterminationTheoryBasic2017].
 At least one laboratory study has shown that the experience of basic psychological need satisfaction during media consumption moderates the change from pre–post play well-being [@tyackRestorativePlayVideogames2020].
@@
 Second, we do not know whether the mood uplift associated with PWS play lingers or disappears after the gaming session ends.
-Answering this question would afford a better understanding of the motivations behind play: To what extent do people play games for their immediate hedonic benefits in contrast to the emotional and cognitive state players find themselves in after stopping?
+PWS might function not only as a mood-enhancer, but also as a stabilizer.
+Considering the generally high pre-play moods observed here, when compared to previous studies [@choiTakingStockHappiness2017; @killingsworthWanderingMindUnhappy2010], PWS player might indeed aim to sustain rather than reach an above-average mood.
+Of course, data on mood before, during, and after play is needed to answer this question, whose answer would afford a better understanding of the motivations behind play: To what extent do people play games for their immediate hedonic benefits in contrast to the emotional and cognitive state players find themselves in after stopping?
 The difference between these two experiences might be an important differentiator between, for example, maladaptive *escaping from* and adaptive *escaping to* [@stensengAreThereTwo2021].

### 2024-04-04 ms.Rmd
Fixes to reproducibility

@@
   )
-```
-
-```{r}
-#| label: session-indicator-create
-#| include: true
-
+
+# Create session indicators
 dat <- dat |>
@@
   )
-```
-
-```{r}
-#| label: mood-filter
 
@@
 
+# Analyze a subset of participants if required 
+# (for testing etc.; be careful)
+dat <- dat |> 
+  filter(
+    pid %in%
+      sample(
+        unique(dat$pid), 
+        size = length(unique(dat$pid)) * N_SUBSET_PROPORTION
+      )
+  )
+
+# Calculate basic data summaries
 dat_sum <- dat |>

### 2024-04-04 ms.Rmd
Data path in env

@@
 
+library(papaja)
 library(scales)
@@
 library(brms)
-library(gt)
-library(ggstance)
 library(ggdist)
@@
 knitr::opts_chunk$set(
+  cache = TRUE,
   warning = FALSE,
@@
 # Download, unzip data from OSF PWS database
-path <- "data-raw/data.zip"
-if (!file.exists(path)) {
-  dir.create("data-raw", FALSE)
+PWS_DATA_PATH <- Sys.getenv("PWS_DATA_PATH", unset = "data-raw/data.zip")
+if (!file.exists(PWS_DATA_PATH)) {
+  dir.create(
+    dirname(PWS_DATA_PATH), 
+    showWarnings = FALSE, 
+    recursive = TRUE
+  )
   download.file(
     url = "https://osf.io/download/j48qf/",
-    destfile = "data-raw/data.zip"
+    destfile = PWS_DATA_PATH
   )
@@
   unzip(
-    zipfile = path,
+    zipfile = PWS_DATA_PATH,
     files = c(
@@
 
-# Analyze a subset of participants if required 
-# (for testing etc.; be careful)
-dat <- dat |> 
-  filter(
-    pid %in%
-      sample(
-        unique(dat$pid), 
-        size = length(unique(dat$pid)) * N_SUBSET_PROPORTION
-      )
-  )
-
 # Calculate basic data summaries
@@
 contrasts(dat$post) <- c(-0.5, 0.5)
+
+# Analyze a subset of participants if required 
+# (for testing etc.; be careful)
+dat <- dat |> 
+  filter(
+    pid %in%
+      sample(
+        unique(dat$pid), 
+        size = length(unique(dat$pid)) * N_SUBSET_PROPORTION
+      )
+  )
 

### 2024-04-04 ms.Rmd
Uncompress if not yet

@@
 
-# Download, unzip data from OSF PWS database
+# Download data from OSF PWS database
 PWS_DATA_PATH <- Sys.getenv("PWS_DATA_PATH", unset = "data-raw/data.zip")
@@
   )
+}
+
+# unzip required data tables
+if (!file.exists("data/demographics.csv")) {
   dir.create("data", FALSE)

### 2024-04-04 ms.Rmd
Fix reproducibility issues (#12)
* Fixes to reproducibility

* Data path in env

* Uncompress if not yet

* Correctly load env var

* Indicate phony targets

@@
 # (for testing etc.; be careful)
+N_SUBSET_PROPORTION <- as.numeric(Sys.getenv("N_SUBSET_PROPORTION", unset = 1))
 dat <- dat |> 

### 2024-04-04 R/common.R

@@ -1,11 +1,12 @@
 N_CORES <- as.numeric(Sys.getenv("N_CORES", unset = 1))
-N_THREADS <- as.numeric(Sys.getenv("N_THREADS", unset = 2))
 N_ITER <- as.numeric(Sys.getenv("N_ITER", unset = 2000))
-N_SUBSET <- as.numeric(Sys.getenv("N_SUBSET"))
+BRMS_BACKEND <- Sys.getenv("N_SUBSET", unset = "rstan")
+N_THREADS <- as.numeric(Sys.getenv("N_THREADS", unset = 2))
+N_SUBSET_PROPORTION <- as.numeric(Sys.getenv("N_SUBSET", unset = 1))
 
 options(
   mc.cores = N_CORES,
-  brms.backend = "cmdstanr",
+  brms.backend = BRMS_BACKEND,
   brms.threads = N_THREADS
 )
 
